\documentclass{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/hong/1_NQBH/reference/bib.bib}
\usepackage[vietnamese,english]{babel}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\small \textsc{Sect.} ~\thesection}
\rhead{\small \nouppercase{\leftmark}}
\renewcommand{\sectionmark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Series: A Personal Journey to Realms of Knowledge}
\author{\selectlanguage{vietnamese} Nguyễn Quản Bá Hồng\footnote{Independent Researcher, Ben Tre City, Vietnam\\e-mail: \texttt{nguyenquanbahong@gmail.com}}}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
	
\end{abstract}

\textbf{Keywords.}
\tableofcontents
\vspace{5mm}

%------------------------------------------------------------------------------%

\selectlanguage{vietnamese}

\begin{center}
	\textit{Dedicated to my high-school mathematics teachers, Thầy \textsc{Liêm}, Thầy \textsc{Huynh}, Thầy \textsc{Hải}, Cô \textsc{An}, Cô \textsc{Hạnh}.}
	
	\textit{In the memory of, especially, Thầy \textsc{Nguyễn Văn Quí}.}
\end{center}

\section{A List of Journeys}
This series, \textit{A Personal Journey to Knowledge}, consists of the following writings:
\begin{enumerate}
	\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/NQBH_elementary_mathematics.pdf}{\textit{Some Topics in Elementary Mathematics: Problems, Theories, Applications, \& Bridges to Advanced Mathematics}}. Mar 2022--now.
	
	Also the following items, structured according to grades in Vietnamese Secondary \& High schools:
	\begin{itemize}
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/grade_6/NQBH_elementary_math_grade_6.pdf}{\textit{Elementary Mathematics}\texttt{/}\textit{Grade 6}}. Mar 2022--now.
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/grade_7/NQBH_elementary_math_grade_7.pdf}{\textit{Elementary Mathematics}\texttt{/}\textit{Grade 7}}. 2022--now. \texttt{[not initialized yet]}
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/grade_8/NQBH_elementary_math_grade_8.pdf}{\textit{Elementary Mathematics}\texttt{/}\textit{Grade 8}}. 2022--now. \texttt{[not initialized yet]}
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/grade_9/NQBH_elementary_math_grade_9.pdf}{\textit{Elementary Mathematics}\texttt{/}\textit{Grade 9}}. 2022--now. \texttt{[not initialized yet]}
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/grade_10/NQBH_elementary_math_grade_10.pdf}{\textit{Elementary Mathematics}\texttt{/}\textit{Grade 10}}. 2022--now. \texttt{[not initialized yet]}
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/grade_11/NQBH_elementary_math_grade_11.pdf}{\textit{Elementary Mathematics}\texttt{/}\textit{Grade 11}}.2022--now. \texttt{[not initialized yet]}
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/elementary_mathematics/grade_12/NQBH_elementary_math_grade_12.pdf}{\textit{Elementary Mathematics}\texttt{/}\textit{Grade 12}}. 2022--now. \texttt{[not initialized yet]}
	\end{itemize}
	\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/tree/master/advanced_mathematics}{\textit{Advanced Mathematics}}. Mar 2022--now. \texttt{[not initialized yet]}
	\begin{itemize}
		\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/advanced_mathematics/mathematical_optimization/NQBH_mathematical_optimization.pdf}{\textit{Some Topics in Mathematical Optimization}}. Mar 2022--now. \texttt{just initialize}
	\end{itemize}
	\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/psychology/NQBH_a_personal_journey_to_psychology.pdf}{\textit{A Personal Journey to Psychology: The Way I Perceive}}. Mar 2022--now. \texttt{[still writing]}
	\item Nguyễn Quản Bá Hồng. \href{https://github.com/NQBH/hobby/blob/master/philosophy/NQBH_a_personal_journey_to_philosophy.pdf}{\textit{A Personal Journey to Philosophy}}. Mar 2022--now. \texttt{just initialize}
\end{enumerate}
\textit{What an arrogant freak!} I can hear that silently from you. But for me, this is totally fine, at least for a monomaniac.

\section{Quick Notes}
\selectlanguage{vietnamese}
Đa số những nhà Toán học giỏi thường là những bậc thầy về ngôn từ -- mình tin là như vậy.

\appendix

\printbibliography[heading=bibintoc]
	
\end{document}