\documentclass{article}
\usepackage[backend=biber,natbib=true,style=alphabetic]{biblatex}
\addbibresource{/home/nqbh/reference/bib.bib}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=1cm,right=1cm,top=5mm,bottom=5mm,footskip=4mm]{geometry}
\def\labelitemii{$\circ$}

\title{The Science of Living}
\author{Alfred Adler}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\vspace{5mm}
\textbf{The Science of Living.} ``Originally published in 1930 \textit{The Science of Living} looks at Individual Psychology as a science. Adler discusses the various elements of Individual Psychology \& its application to everyday life: including the inferiority complex, the superiority complex \& other social aspects, such as, love \& marriage, sex \& sexuality, children \& their education. This is an important book in the history of psychoanalysis \& Alderian therapy.''

\section*{A Note on the Author \& His Work}
``\textsc{Dr. Alfred Adler}'s work in psychology, while it is scientific \& general in method, is essentially the study of the separate personalities we are, \& is therefore called Individual Psychology. Concrete, particular, unique human beings are the subjects of this psychology, \& it can only be truly learned from the men, women \& children we meet.

The supreme importance of this contribution to modern psychology is due to the manner in which it reveals how all the activities of the soul are drawn together into the service of the individual, how all his faculties \& strivings are related to 1 end. We are enabled by this to enter into the ideals, the difficulties, the efforts \& discouragements of our fellow-men, in such a way that we may obtain a whole \& living picture of each as a personality. In this co-ordinating idea, something like finality is achieved, though we must understand it as finality of foundation. There has never before been a method so rigorous \& yet adaptable for following the fluctuations of that most fluid, variable \& elusive of all realities, the individual human soul.

Since Adler regards not only science but even intelligence itself as the result of the communal efforts of humanity, we shall find his consciousness of his own unique contribution more than usually tempered by recognition of his collaborators, both past \& contemporary. It will therefore be useful to consider Adler's relation to the movement called Psycho-analysis, \& 1st of all to recall, however briefly, the philosophic impulses which inspired the psycho-analytic movement as a whole.

The conception of the Unconscious as vital memory -- biological memory -- is a common to modern psychology as a whole. But Freud, from the 1st a specialist in hysteria, took the memories of success or failure in the sexual life, as of the 1st -- \& almost the only -- importance. Jung, a psychiatrist of genius, has tried to widen this distressingly narrow view, by seeking to reveal the super-individual or racial memories which, he believes, have as much power as the sexual \& a higher kind of value for life.

It was left to Alfred Adler, a physician of wide \& general experience, to unite the conception of the Unconscious more firmly with biological reality. A man of the original school of psycho-analysts, he had done much work by that method of analyzing memories out of their coagulated emotional state into clearness \& objectivity. But he showed that the whole scheme of memory is different in every individual. Individuals do not form their unconscious memories all around the same central motive -- not all around sexuality, e.g., . In every individual we find an individual way of selecting its experiences from all possible experience. What is the principle of that selectivity? Adler has answered that it is, fundamentally, the organic consciousness of a \textit{need}, of some specific inferiority which has to be compensated. It is as though every soul had consciousness of its whole physical reality, \& were concentrated, with sleepless insistence, upon achieving compensation for the defects in it.

Thus the whole life of the small man, e.g., , would be interpretable as a struggle to achieve immediate greatness in some way, \& that of a deaf man to obtain a compensation for not hearing. It is not so simple as that, of course, for a system of defects may give rise to a constellation of guiding ideas, \& also in human life we have to deal with imaginary inferiorities \& fantastic strivings, but even here the principle is the same.

The sexual life, far from controlling all activities, fits perfectly into the frame of those more important strivings, for it is pre-eminently under the control of emotion, \& emotion is moulded by the entire vital history. Thus a Freudian analysis gives a true account of the sexual \textit{consequences} of a given life-line, but it is a true \textit{diagnosis} only in that sense.

Psychology becomes now for the 1st time rooted in biology. The tendencies of the soul, \& the mind's development, are seen to be controlled from the 1st by the effort to compensate for organic defects or for positions of inferiority. Everything that is exceptional or individual in the disposition of an organic being originates in this way. The principle is common to man \& animal, probably even to the vegetable kingdom also; \& the special endowments of species are to be taken as arising from experience of defects \& inferiorities in relation to their environment, which has been successfully compensated by activity, growth \& structure.

There is nothing new in the idea of compensation as a biological principle, for it has been long known that the body will over-develop certain parts in compensation for the injury of others. If 1 kidney ceases to function, e.g., the other develops abnormally until it does the work of both; if the heart springs a leak in a valve, the whole organ grows larger to allow for its loss of efficiency, \& when nervous tissue is destroyed, adjacent tissue of another kind endeavors to take on the nerve-function. The compensatory developments of the whole organism to meet the exigencies of any special work or exertion are too numerous \& well known to need illustration. But it is Dr. Adler who has 1st transferred this principle bodily to psychology as a fundamental idea, \& demonstrated the part it plays in the soul \& intellect.

Adler recommends the study of Individual Psychology not only to doctors, but generally to laymen \& especially to teachers. Culture in psychology has become a general necessity, \& must be firmly advocated in the teeth of popular opposition to it, which is founded upon the notion that modern psychology requires an unhealthy concentration of the mind upon cases of disease \& misery. It is true that the literature of psychoanalysis has revealed the most central \& the most universal evils in modern society. But it is not now a question of contemplating our errors, it is necessary that we should learn by them. We have been trying to live as though the soul of man were not a reality, as though we could build up a civilized life in defiance of psychic truths. What Adler proposes is not the universal study of psycho-pathology, but the practical reform of society \& culture in accordance with a positive \& scientific psychology to which he has contributed the 1st principles. But this is impossible if we are too much afraid of the truth. The clearer consciousness of right aims in life, which is indispensable to us, cannot be gained without a deeper understanding also of the mistakes in which we are involved. We may not desire to know ugly facts, but the more truly we are aware of life, the more clearly we perceive the real errors which frustrate it, much as the concentration of a light gives definition to the shadows.

A positive psychology, useful for human life, cannot be derived from the psychic phenomena alone, still less from pathological manifestations. It requires also a regulative principle, \& Adler has not shrunk from this necessity, by recognizing, as if it were of absolute metaphysical validity, the logic of our communal life in the world.

Recognizing this principle, we must proceed to estimate the psychology of the individual in relation too it. The way in which an individual's inner life is related to the communal being is distinguishable in 3 ``life-attitudes,'' as they are called -- his general reactions to society, to work \& to love.

By their feeling towards society as a whole -- to any other \& to tell others -- man \& women may know how much social courage they possess. The feeling of inferiority is always manifested in a sense of fear or uncertainty in the presence of society, whether its outward expression is 1 of timidity or defiance, reserve or over-anxiety. All feelings of innate suspicion or hostility, of an undefined caution \& desire for some concealment, when such feelings affect the individual in social relations generally, evince the same tendency to withdraw from reality, which inhibits self-affirmation. The ideal, or rather normal, attitude to society is an unstrained \& unconsidered assumption of human equality unchanged by any inequalities of position. Social courage depends upon this feeling of secure membership of the human family, a feeling which depends upon the harmony of one's own life. By the tone of his feeling towards his neighbors, his township \& nation \& to other nationalities, \& even by his reactions when he reads of all these things in his newspaper, a man may infer how securely his own soul is grounded in itself.

The attitude towards work is closely dependent upon this self-security in society. In the occupation by which a man earns his share in social goods \& privileges, he has to face the logic of social needs. If he has too great a sense of weakness or division from society, it will make him unable to believe that his worth will ever be recognized, \& he will not even work for recognition: instead, he will play for safety, \& work for money or advantage only, suppressing his own valuation of what is the truest service he can render. He will always be afraid to supply or demand the best, for fear it may not pay. Or he may be always seeking for some quiet backwater of the economic life, where he can do something just as he likes himself, without proper consideration of either usefulness or profit. In both cases it is not only society that suffers by not getting the best service: the individual who has not attained his proper social significance is also deeply dissatisfied. The modern world is full of men, both successful \& unsuccessful in a worldly sense, who are in open conflict with their occupation. They do not believe in it, \& they blame social \& economic conditions with some real justice; but it is also a fact that they have often had too little courage to fight for the best value in their economic function. They were afraid to claim the right to give what they genuinely believed in, or else they felt disdainful of the service society really needed of them. Hence they pursued their gain in an individualistic or even furtive spirit. We must, of course, recognize that so much is wrong in the organization of society, that, besides the possibility of making mistakes of judgment, the individual who is determined to render real social service has often to face heavy opposition. But it is precisely that sense of struggle to give his best which the individual needs no less than society benefits by it. One cannot love a vocation which does not afford some experience of victory over difficulties, \& not merely of compromise with them.

It is the 3rd of these life-attitudes -- the attitude to love -- which determines the course of the erotic life. Where the 2 previous life-attitudes, to society \& to work, have been rightly adjusted, this last comes right by itself. Where it is distorted \& wrong it cannot be improved by itself apart from the others. Although we can think how to improve the social relations \& the occupations, a concentration of thought upon the individual sex-problem is almost sure to make it worse. For this is far more the sphere of results than of course. A soul that is defeated in ordinary social life, or thwarted in its occupation, acts in the sex-life as though it were trying to obtain compensation for the kinds of expression of which it fails in their proper spheres. This is actually the best way in which we can understand all sexual vagaries, whether they isolate the individual, degrade the sexual partner or in any way distort the instinct. The friendships of an individual also are integral with the love-life as a whole; not, as the 1st psycho-analysts imagined, because friendship is a sublimation of sexual attraction, but the other way about. Sexual compulsion -- sex as an insubordinate psychic factor -- is an abnormal substitute for the vitalizing intimacy of useful friendships, \& homosexuality is always the consequence of incapability for love.

The meaning \& value which we give to sensations are also united closely with the erotic life, as many good poets have testified. The quality of our feeling for Nature, our response to the beauty of sea \& land, \& to significances of form \& sound \& color, as well as our confidence in scenes of storm \& gloom, are all involved with our integrity as lovers. The aesthetic life, with all it means to art \& culture, is thus ultimately derived, through individuals, from social courage \& intelligent usefulness.

We ought not to regard the communal feeling as something to be created with difficulty. It is as natural \& inherent as egoism itself, \& indeed as a principle of life it has priority. We have not to create, but only to liberate, it where it is repressed. It is the saving principle of life as we experience it. If anyone thinks that the services of 'busmen, railwaymen \& milkmen would be rendered as well as they are without the presence of very much instinctive communal feeling he must be suspected of a highly neurotic scheme of apperception. What inhibits it is, to speak bluntly, the enormous vanity of the human soul, which is, moreover, so subtle that no professional psychologist before Adler had been able to demonstrate it, though a few artists had divined its omnipresence. All unsuspected as it often is, the ambition of many a minor journalist or shop-assistant, to say nothing of the great ones of the world, would be enough to bring about the fall of an archangel. Every feeling of inferiority that has embittered his contact with life has fed the imagination of greatness with another god-like assumption until, in many cases, the fantasy has become so inflated as to demand not even supremacy in this world for its appeasement, but the creation of a new world altogether, \& to be the god of it. This revelation of the depth of human nature is verified, not so strikingly from the study of cases of practical ambition, however Napoleonic, as from those of passive resistance, procrastination, \& malingering, for it is these which show most clearly that an individual who feels painfully unable to dominate the real world will refuse to co-operate with it, at whatever disadvantage to himself, partly in order to tyrannize over a narrower sphere, \& partly even from an irrational feeling that the real world, without his divine assistance, will some day crumble \& shrink to his own diminished measure.\footnote{In case this should seem an exaggeration, we may recall the fact that nearly all the narrowest kind of sects, religious or secular, have a belief in world-catastrophe: the world from which they have withdrawn, \& which they despair of converting, is to be brought to destruction, \& only a remnant will survive, who will be of their own persuasion.}

The question is thus raised, how should we act, knowing this tendency to inordinate vanity in the human soul, \& that we dare not merely add to that vanity by assuming ourselves to be miraculous exceptions? Adler's reply is that we should preserve a certain attitude to all our experience, which he calls the attitude of ``half-\&-half.'' Our conception of normal behavior should be to allow the world or society, or the person with whom we are confronted, to be somehow in the right equally with ourselves. We should not depreciate either ourselves or our environment; but, assuming that each is 1-half in the right, affirm the reality of ourselves \& others equally. This applies not only to contacts with other souls, but to our mental reactions towards rainy weather, holidays or comforts that we cannot afford, \& even to the omnibus we have just missed.

Rightly understood, this is not an ideal of difficult \& distasteful humility. It is in reality a tremendous assumption of worth, to claim exactly equal reality \& omnipotence with the whole of the rest of creation, in whatever particular manifestation we may be meeting with it. To claim less than this is a false humility, for what results from any contact we make does in fact depend for half its reality upon the way in which we make it. The individual should affirm his part in everything which occurs to him, as his own half of it.

This is often a particularly difficult counsel to keep in relation to the occupation. In their business, people face more naked realities than are usually allowed to appear in social life; \& it is often almost impossible to allow equal validity to one's own aims \& to the conditions of a disorganized world. To do so, means the admission that conditions, just such as they are, are one's real problem -- \&, indeed, one's proper sphere of action. The division of labor, logical \& useful as it is in itself, has given opportunity for human megalomania to create entirely false inequalities, distinctions \& injustices, so that we live in an economic disorder which will hardly hold together. To such crazy conditions, the best of men often find it difficult to oppose themselves with perseverance, equally grating its reality \& working for its reform. They are tempted to acquiesce in disorder by some inner subterfuge, or to devote themselves to superficial remedies which evade the real problem; \& sometimes they treat their work-life as an unavoidable contamination by things inherently squalid, quite unaware that such an attitude makes them conceited, haughty \&, in a profound sense, unscrupulous. It occurs to very few that the right way would be to make alliance on human grounds with others in the same predicament \& profession, to assert its proper dignity as a social service \& improve it; but this is the only way in which the individual can really be reconciled with his economic function. Many of those who complain most about the conditions prevailing in their work are doing nothing whatever to reorganize it as a function of human life, \& never think of attacking the anarchic individualism which is its ruin. We derive it from Individual Psychology, as a categorical imperative, that every man's duty is to work to make his profession, whatever it may be, into a brotherhood, a friendship, a social unity with a powerful morale of co-operation, \& that if a man does not want to do this his own psychological state is precarious. It is true that now, in many professions, the task that this presents is terribly difficult. It is all the more essential that the effort should be made towards integration. For a man's work will never liberate the forces of his psyche unless he is striving, in a large sense, to make it the expression of his whole being, \& his idea of his profession must be not only an executive in which he has independence of action, but also a legislative in which he has some authority of direction. In a man's business life that half-\&-half valuation leads equally to recognition of reality \& to struggle with it by the only realistic method, which is necessarily co-operative.

The pedagogic principles of Individual Psychology, infallible as far as they go, are useless without this practical work of social organization. What has been written above of an individual's duty in his occupation applis in a large sense to his entire social function. A person's function includes active membership of his nation \& of humanity, to say nothing of his family. There is a certain parliament which rises for no vacation, \& to whose decisions all elected assemblies must in the end defer. It meets in schools, markets, \& everywhere on sea \& land, for it is the Parliament of Man, in which every word or look exchanged, whether of courtesy or recrimination, of wisdom or folly, has its measure of importance in the affairs of the race. It is everyone's interest to make this wide assembly more united \& its discussion more intelligible, for none of us has any real human existence except by reflection from it. When its conclaves are peaceful, all our lives are heightened in tone, health \& wealth accrues \& arts \& education flourish; when its conversation is reversed \& suspicious, work fails, men starve \& children languish. In the heat of its dissensions we perish by the million. All its decrees, by which we live or die, \& grow or decay, are rooted in our individual attitudes towards man, woman \& child in every relation of life.

When we face, objectively, this fact of the relation of all souls \& their mutual responsibility, what are we to think of the inner confusion of the neurotic? Is it not simply a narrowing of the sphere of interest, an over-concentration upon certain personal or subjective gains? The neurotic soul is the result of treating the rest of humanity as though its life \& aims were altogether of less importance than one's own, \& thus losing interest in any larger life. Paradoxically, it often happens that a neurotic has very large schemes of saving himself \& others. He is intelligent enough to try to compensate his real sense of isolation \& impotence in the human assembly, by a fantasy of exaggerated importance \& beneficent activity. He may want to reform education, to abolish war, to establish universal brotherhood or create a new culture, \& even plans or joins societies with these aims. He is defeated in such aims, of course, by the unreality of his contact with others \& with life as a whole. It is as though he had taken a standpoint outside of life altogether \& were trying to direct it by some unexplained magic.

Modern city life especially, with its intellectualism, gives unlimited scope for the neurotic thus to compensate his real unsociability with imaginary messianism, \& the result is the disintegration of a people full of saviors who are not on speaking terms.

What is needed, of course, is something very different. It is not that the individual should renounce messianism; for it is a fact that a share of responsibility for the whole future of the race is his alone. It is only necessary that he should take a reasonable view of his power to save society, correctly viewed from his own standpoint: he must become able to regard his immediate personal relations \& his occupations as if \textit{they} were of world-importance, for in fact they are so, being the only world-meaning an individual has. When they are chaotic or wrong, it is because we do not, in day-to-day experience, treat them as things of universal meaning. We sometimes treat them as important, no doubt, but generally in a personal sense only.

This tendency of the modern soul, to narrow the sphere of interest, both practically \& ideally, is most difficult to subdue, because it is reinforced by the scheme of apperception. For that reason an individual alone cannot do it, excepting only in rare cases. He needs conference with other minds, \& an entirely new kind of conference. A resolution to treat one's immediate surroundings \& daily activities as if they were the supreme significance of life brings an individual immediately into conflict with internal resistances of his own, \& often with external difficulties also, which he cannot at once understand \& which no others could rightly estimate unless they were making the same experiment. Hence, the practice of Individual Psychology demands that its students should submit themselves to mutual scrutiny, each one to be estimated by the others as a whole personality. This practice, striking at the root of the false individualism which is the basis of all neurosis, is naturally very difficult to initiate. Upon its success, however, depends the whole future of psycho-analysis as an influence in life at large, outside of clinics \& consulting rooms.

In Vienna the work of such groups has already made itself felt in education. The co-operation it has established between teachers \& medical practitioners has revolutionized the work of certain schools, \& established an equality between teachers \& pupils \& between pupils themselves, which has cured many children of criminal tendencies, dullness \& laziness. Abolition of competition \& the cultivation of encouragement have been found to liberate the energy of both pupils \& teachers. These changes are already affecting the surrounding family life, which comes into question immediately the child is psychologically considered. Education, though naturally the 1st, is not the only sphere of life which out to be invaded by the activity of such groups. Business \& political circles, which experience the deadlock of modern life most acutely, need to be vitalized with knowledge of human nature, which they have forgotten how to recognize.

It is for this work of releasing a new energy for daily life \& its reformation, that Alfred Adler has founded the International Society for Individual Psychology. The culture of human behavior which this work has begun already to propagate might well be mistaken for an almost platitudinous ethics, but for 2 things -- its practical results, \& the background of scientific method out of which it is appearing. In his realistic grasp of the social nature of the individual's problem \& his inexorable demonstration of the unity of health \& harmonious behavior, Adler resembles no one so much as the great Chinese thinkers. If the occidental world is not too far gone to make use of his service, he may well come to be known as the Confucius of the West. -- \textsc{Phillipe Mairet}'' -- \cite[pp. 9--30]{Adler2013}

%------------------------------------------------------------------------------%

\section{The Science of Living}
``Only a science which is directly related to life, said the great philosopher William James, is really a science. It might also be said that in a science which is directly related to life theory \& practice become almost inseparable. The science of life, precisely because it models itself directly on the movement of life, becomes a science of living. These considerations apply with special force to the science of Individual Psychology. Individual Psychology tries to see individual lives as a whole \& regards each single reaction, each movement \& impulse as an articulated part of an individual attitude towards life. Such a science is of necessity oriented in a practical sense, for which the aid of knowledge we can correct \& alter our attitudes. Individual Psychology is thus \textit{prophetic} in a double sense: not only does it predict what will happen, but, like the prophet Jonah, it predicts what \textit{will} happen in order that it should \textit{not} happen.

The science of Individual Psychology developed out of the effort to understand that mysterious creative power of life -- that power which expresses itself in the desire to develop, to strive \& to achieve -- \& even to compensate for defeats in 1 direction by striving for success in another. This power is \textit{teleological} -- it expresses itself in the striving after a goal, \& in this striving every bodily \& psychic movement is made to co-operate. It is thus absurd to study bodily movements \& mental conditions abstractly without relation to an individual whole. It is absurd, e.g., that in criminal psychology we should pay so much more attention to the crime than to the criminal. It is the criminal, not the crime that counts, \& no matter how much we contemplate the criminal act we shall never understand its criminality unless we see it as an episode in the life of a particular individual. The same outward act may be criminal in 1 case \& not criminal in another. The important thing is to understand the individual context -- the goal of an individual's life which marks the line of direction for all his acts \& movements. This goal enables us to understand the hidden meaning behind the various separate acts -- we see them as parts of a whole. Vice versa when we study the parts -- provided we study them as parts of a whole -- we get a better sense of the whole.

In the author's own case the interest in psychology developed out of the practice of medicine. The practice of medicine provided the teleological or purposive viewpoint which is necessary for the understanding of psychological facts. In medicine we see all organs striving to develop towards definite goals. They have definite forms which they achieve upon maturity. Moreover, in cases where there are organic defects we always find nature making special efforts to overcome the deficiency, or else to compensate for it by developing another organ to take over the functions of the defective one. Life always seeks to continue, \& the life force never yields to external obstacles without a struggle.

Now the movement of the psyche is analogous to the movement of organic life. In each mind there is the conception of a goal or ideal to get beyond the present state, \& to overcome the present deficiencies \& difficulties by postulating a concrete aim for the future. By means of this concrete aim or goal the individual can think \& feel himself superior to the difficulties of the present because he has in mind his success of the future. Without the sense of a goal individual activity would cease to have anything meaning.

All evidence points to the fact that the fixing of this goal -- giving it a concrete form -- must take place early in life, during the formative period of childhood. A kind of prototype or model of a matured personality begins to develop at this time. We can imagine how the process takes place. A child, being weak, feels inferior \& finds itself in a situation which it cannot bear. Hence it strives to develop, \& it strives to develop along a line of direction fixed by the goal which it chooses for itself. The material used for development at this stage is less important than the goal which decides the line of direction. How this goal is fixed it is difficult to say, but it is obvious that such a goal exists \& that it dominates the child's every movement. Little is indeed understood about powers, impulses, reasons, abilities or disabilities at this early period. As yet there is really no key, for the direction is definitely established only after the child has fixed its goal. Only when we see the direction in which a life is tending can we guess what steps will be taken in the future.

When the prototype -- that early personality which embodies the goal -- is formed, the line of direction is established \& the individual becomes definitely oriented. It is this fact which enables us to predict what will happen later in life. The individual's apperceptions are from then on bound to fall into a groove established by the line of direction. The child will not perceive given situations as they actually exist, but according to a personal scheme of apperception -- that is to say, he will perceive situations under the prejudice of his own interests.

An interesting fact that has been discovered in this connection is that children with organic defects connect all their experiences with the function of the defective organ. E.g., a child having stomach trouble shows an abnormal interest in eating, while one with defective eyesight is more preoccupied with things visible. This preoccupation is in keeping with the private scheme of apperception which we have said characterizes all persons. It might be suggested, therefore, that in order to find out where a child's interest lies we need only to ascertain which organ is defective. But things do not work out quite so simply. The child does not experience the fact of organ inferiority in the way that an external observer sees it, but as modified by its own scheme of apperception. Hence while the fact of organ inferiority counts as an element in the child's scheme of appreciation, the external observation of the inferiority does not necessarily give the cue to the scheme of apperception.

The child is steeped in a scheme of relativity, \& in this he is indeed like the rest of us -- none of us is blessed with the knowledge of the absolute truth. Even our science is not blessed with absolute truth. It is based on common sense, which is to say that it is ever changing \& that it is content gradually to replace big mistakes by smaller ones. We all make mistakes, but the important thing is that we can correct them. Such correction is easier at the time of the formation of the prototype. \& when we do not correct them at that time, we may correct the mistakes later on by recalling the whole situation of that period. Thus if we are confronted with the task of treating a neurotic patient, our problem is to discover, not the ordinary mistakes he makes in later life, but the very fundamental mistakes made early in his life in the course of the constitution of his prototype. If we discover these mistakes, it is possible to correct them by appropriate treatment.

In the light of Individual Psychology the problem of inheritance thus decreases in importance. It is not what one has inherited that is important, but what one does with his inheritance in the early years -- that is to say, the prototype that is built up in the childhood environment. Heredity is of course responsible for inherited organic defects, but our problem there is simply to relieve the particular difficulty \& place the child in a favorable situation. As a matter of fact we have even a great advantage here, inasmuch as when we see the defect we know how to act accordingly. Oftentimes a healthy child without any inherited defects may fare worse through malnutrition or through any of the many errors in upbringing.

In the case of children born with imperfect organs it is the psychological situation which is all-important. Because these children are placed in a more difficult situation they show marked indications of an exaggerated feeling of inferiority. At the time the prototype is being formed they are already more interested in themselves than in others, \& they tend to continue that way later on in life. Organic inferiority is not the only cause of mistakes in the prototype: other situations may also cause the same mistakes -- the situations of pampered \& hated children, e.g. We shall have occasion later on to describe these situations more in detail \& to present actual case histories illustrating the 3 situations which are particularly unfavorable, that of children with imperfect organs, that of petted children, \& that of hated children. For the present it is sufficient to note that these children grow up handicapped \& that they constantly fear attacks inasmuch as they have grown up in an environment in which they never learned independence.

It is necessary to understand the social interest from the very upset since it is the most important part of our education, of our treatment \& of our cure. Only such persons as are courageous, self-confident \& at home in the world can benefit both by the difficulties \& by the advantages of life. They are never afraid. They know that there are difficulties, but they also know that they can overcome them. They are prepared for all the problems of life, which are invariably social problems. From a human standpoint it is necessary to be prepared for social behavior. The 3 types of children we have mentioned develop a prototype with a lesser degree of social interest. They have not the mental attitude which is conducive to the accomplishment of what is necessary in life or to the solution of its difficulties. Feeling defeated, the prototype has a mistaken attitude towards the problems of life \& tends to develop the personality on the useless side of life. On the other hand our task in treating such patients is to develop behavior on the useful side \& to establish in general a useful attitude towards life \& society.

Lack of social interest is equivalent to being oriented towards the useless side of life. The individuals who lack social interest are those who make up the groups of problem children, criminals, insane persons, \& drunkards. Our problem in their case is to find means to influence them to go back to the useful side of life \& to make them interested in others. In this way it may be said that our so-called Individual Psychology is actually a social psychology.

After the social interest, our next task is to find out the difficulties that confront the individual in his development. This task is somewhat more confusing at 1st glance, but it is in reality not very complicated. We know that every petted child becomes a hated child. Our civilization is such that neither society nor the family wishes to continue the pampering process indefinitely. A pampered child is very soon confronted with life's problems. In school he finds himself in a new social institution, with a new social problem. He does not want to write or play with his fellows, for his experience has not prepared him for the communal life of the school. In fact his experiences as lived through at the prototype stage make him afraid of such situations \& make him look for more pampering. Now the characteristics of such an individual are not inherited -- far from it -- for we can deduce them from a knowledge of the nature of his prototype \& his goal. Because he has the particular characteristics conducive to his moving in the direction of his goal, it is not possible for him to have characteristics that would tend in any other direction.

The next step in the science of living lies in the study of the feelings. Not only does the axis line, the line of direction posited by the goal, affect individual characteristics, physical movements, expressions \& general outward symptoms, but it dominates the life of the feelings as well. It is a remarkable thing that individuals always try to justify their attitudes by feelings. Thus if a man wants to do good work, we will find this idea magnified \& dominating his whole emotional life. We can conclude that the feelings always agree with the individual's viewpoint of his task: they strengthen the individual in his bent for activity. We always do that which we would do even without the feelings, \& the feelings are simply an accompaniment to our acts.

We can see this fact quite clearly in dreams, the discovery of whose purpose was perhaps 1 of the latest achievements of Individual Psychology. Every dream has of course a purpose, although this was never clearly understood until now. The purpose of a dream -- expressed in general \& not specific terms -- is to create a certain movement of feeling or emotion, which movement of emotion in turn furthers the movement of the dream. It is an interesting commentary on the old idea that a dream is always a deception. We dream in the way that we would like to behave. Dreams are an emotional rehearsal of plans \& attitudes for waking behavior -- a rehearsal, however, in which the actual play may never come off. In this sense dreams are deceptive -- the emotional imagination gives us the thrill of action without the action.

This characteristic of dreams is also found in our waking life. We always have a strong inclination to deceive ourselves emotionally -- we always want to persuade ourselves to go the way of our prototypes as they were formed in the 4th or 5th year of life.

The analysis of the prototype is next in order in our scheme of science. As we have said, at 4 or 5 the prototype is already built up, \& so we have to look for impressions made n the child before or at that time. These impressions can be quite varied, far more varied than we imagine from a normal adult's point of view. 1 of the most common influences on a child's mind is the feeling of suppression brought about by a father's or mother's excessive punishment or abuse. This influence makes the child strive for release, \& sometimes this is expressed in an attitude of psychological exclusion. Thus we find that some girls having high-tempered fathers have prototypes that exclude men because they are high-tempered. Or boys suppressed by severe mothers may exclude women. This excluding attitude may of course be variously expressed: e.g., the child may become bashful, or on the other hand, he may become perverted sexually (which is simply another way of excluding women). Such perversions are not inherited, but arise from the environment surrounding the child in these years.

The early mistakes of the child are costly. \& despite this fact the child receives little guidance. Parents do not know or will not confess to the child the results of their experiences, \& the child must thus follow his own line.

Curiously enough we will find that no 2 children, even those born in the same family, grow up in the same situation. Even within the same family the atmosphere that surrounds each individual child is quite particular. Thus the 1st child has notoriously a different set of a circumstances from the other children. The 1st child is at 1st alone \& is thus the center of attention. Once the 2nd child is born, he finds himself dethroned \& he does not like the change of situation. In fact it is quite a tragedy in his life that he has been in power \& is so no longer. This sense of tragedy goes into the formation of his prototype \& will crop out in his adult characteristics. As a matter of fact case histories show that such children always suffer downfall.

Another intra-family difference of environment is to be found in the different treatments accorded to boys \& to girls. The usual case is for boys to be overvalued \& the girls to be treated as if they could not accomplish anything. These girls will grow up always hesitating \& in doubt. Throughout life they will hesitate too much, always remaining under the impression that only men are really able to accomplish anything.

The position of the 2nd child is also characteristic \& individual. He is in an entirely different position from that of the 1st child, inasmuch as for him there is always a pace-maker, moving along parallel with him. Usually the 2nd child overcomes his pace-maker, \& if we look for the cause we shall find simply that the older child was annoyed by having such a competitor \& that the annoyance in the end affected his position in the family. The older child becomes frightened by the competition \& does not do so well. He sinks more \& more in the estimation of his parents, who begin to appreciate the 2nd child. On the other hand the 2nd child is always confronted by the pace-maker, \& he is thus always in a race. All his characteristics will reflect this peculiar position in the family constellation. He shows rebellion \& does not recognize power or authority.

History \& legend recount numerous incidents of powerful youngest children. Joseph is a case in point: he wanted to overcome all the others. The fact that a younger brother was born into the family unknown to him years after he left home obviously does not alter the situation. His position was that of the youngest. We find also the same description in all the fairy tales, in which the youngest child plays the leading role. We can see how these characteristics actually originate in early childhood \& cannot be changed until the insight of the individual has increased. In order to readjust a child you must make him understand what happened in his 1st childhood. He must be made to understand that his prototype is erroneously influencing all the situations in his life.

A valuable tool for understanding the prototype \& hence the nature of the individual is the study of old remembrances. All our knowledge \& observation force us to the conclusion that our remembrances belong to the prototype. An illustration will make our point clear. Consider a child of the 1st type, one with imperfect organs -- with a weak stomach, let us say. If he remembers having seen something or heard something it will probably in some way concern eatables. Or take a child that is left-handed: his left-handedness will likewise affect his viewpoint. A person may tell you about his mother who pampered him, or about the birth of a younger child. He may tell you how he was beaten, if he had a high-tempered father, or how he was attacked if he was a hated child at school. All such indications are very valuable provided we learn the art of reading their significance.

The art of understanding old remembrances involves a very high power of sympathy, a power to identify oneself with the child in his childhood situation. It is only by such power of sympathy that we are able to understand the intimate significance in a child's life of the advent of a younger child in the family, or the impression made on a child's mind by the abuse of a high-tempered father.

\& while we are on the subject it cannot be overemphasized that nothing is gained by punishing, admonishing \& preaching. Nothing is accomplished when neither the child nor the adult knows on which point the change has to be made. When the child does not understand, he becomes slyer \& more cowardly. His prototype, however, cannot be changed by such punishment \& preaching. It cannot be changed by mere experience of life, for the experience of life is already in accordance with the individual's personal scheme of apperception. It is only when we get at the basic personality that we accomplish any changes.

If we observe a family with badly developed children, we shall see that though they all seem to be intelligent (in the sense that if you ask a question they give the right answer), yet when we look for symptoms \& expressions, they have a great feeling of inferiority. Intelligence of course is not necessarily common sense. The children have an entirely personal -- what we might term, a private -- mental attitude of the sort that one finds among neurotic persons. In a compulsion neurosis, e.g., the patient realizes the futility of always counting windows but cannot stop. One interested in useful things would never act this way. Private understanding \& language are also characteristic of the insane. The insane never speak in the language of common sense, which represents the height of social interest.

If we contrast the judgment of common sense with private judgment, we shall find that the judgment of common sense is usually nearly right. By common sense we distinguish between good \& bad, \& while in a complicated situation we usually make mistakes, the mistakes tend to correct themselves through the very movement of common sense. But those who are always looking out for their own private interests cannot distinguish between right \& wrong as readily as others. In fact they rather betray their inability, inasmuch as all their movements are transparent to the observer.

Consider e.g. the commission of crimes. If we inquire about the intelligence, the understanding \& the motive of a criminal, we shall find that the criminal always looks upon his crimes as both clever \& heroic. He believes that he has achieved a goal of superiority -- namely, that he has become more clever than the police \& is able to overcome others. He is thus a hero in his own mind, \& does not see that his actions indicate something quite different, something very far from heroic. His lack of social interest, which puts his activity on the useless side of life, is connected with a lack of courage, with cowardice, but he does not know this. Those who turn to the useless side of things are often afraid of darkness \& isolation; they wish to be with others. This is cowardice \& should be labeled as such. Indeed the best way to stop crime would be to convince everybody that crime is nothing but an expression of cowardice.

It is well known that some criminals when they approach the age of 30 will take a job, marry \& become good citizens in later life. What happens? Consider a burglar. How can a 30-year old burglar compete with a 20-year old burglar? The latter is cleverer \& more powerful. Moreover, at the age of 30 the criminal is forced to live differently from the way he lived before. As a result the profession of crime no longer pays the criminal \& he finds it convenient to retire.

Another fact to be borne in mind in connection with criminals is that if we increase the punishments, so far from frightening the individual criminal, we merely help to increase his belief that he is a hero. We must not forget that the criminal lives in a self-centered world, a world in which one will never find true courage, self-confidence, communal sense, or understanding of common values. It is not possible for such persons to join a society. Neurotics seldom start a club, \& it is an impossible feat for persons suffering from agoraphobia or for insane persons. Problem children or persons who commit suicide never make friends -- a fact for which the reason is never given. There is a reason, however: they never make friends because the early life took a self-centered direction. Their prototypes were oriented towards false goals \& followed lines of direction on the useless side of life.

Let us now consider the program which Individual Psychology offers for the education \& training of neurotic persons -- neurotic children, criminals, \& persons who are drunkards \& wish to escape by such means from the useful side of life.

In order to understand easily \& quickly what is wrong, we begin by asking at what time the trouble originated. Usually the blame is laid on some new situation. But this is a mistake, for before this actual occurrence, our patient -- so we shall find upon investigation -- had not been well prepared for the situation. So long as he was in a favorable situation the mistakes of his prototype were not apparent, for each new situation is in the nature of an experiment to which he reacts according to the scheme of apperception created by his prototype. His responses are not mere reactions, they are creative \& consistent with his goal, which is dominant throughout his life. Experience taught us early in our studies of Individual Psychology that we might exclude the importance of inheritance, as well as the importance of an isolated part. We see that the prototype answers experiences in accordance with its own scheme of apperception. \& it is this scheme of apperception that we must work upon in order to produce any results.

This sums up the approach of Individual Psychology which has been developed in the last 25 years. As one may see, Individual Psychology has traveled a long way in a new direction. There are many psychologies \& psychiatries in existence. 1 psychologist takes 1 direction, another another direction, \& no one believes that the others are right. Perhaps the reader, too, should not rely on belief \& faith. Let him compare. He will see that we cannot agree with what is called ``drive'' psychology (McDougall represents this tendency best in America), because in their ``drives'' too big a place is set aside for inherited tendencies. Similarly we cannot agree with the ``conditioning'' \& ``reactions'' of Behaviorism. It is useless to construct the fate \& character of an individual out of ``drives'' \& ``reactions'' unless we understand the goal to which such movements are directed. Neither of these psychologies thinks in terms of individual goals.

It is true that when the word ``goal'' is mentioned, the reader is likely to have a hazy impression. The idea needs to be concretized. Now in the last analysis to have a goal is to aspire to be like God. But to be like God is of course the ultimate goal -- the goal of goals, if we may use the term. Educators should be cautious in attempting to educate themselves \& their children to be like God. As a matter of fact we find that the child in his development substitutes a more concrete \& immediate goal. Children look for the strongest person in their environment \& make him their model or their goal. It may be the father, or perhaps the mother, for we find that even a boy may be influenced to imitate his mother if she seems the strongest person. Later on they want to be coachmen because they believe the coachman is the strongest person.

When children 1st conceive such a goal they behave, feel \& dress like the coachman \& take on all the characteristics consistent with the goal. But let the policeman lift a finger, \& the coachman becomes nothing $\ldots$ Later on the ideal may become the doctor or the teacher. For the teacher can punish the child \& thus he arouses his respect as a strong person.

The child has a choice of concrete symbols in selecting his goal, \& we find that the goal he chooses is really an index of his social interests. A boy, asked what he wanted to be in later life, said, ``I want to be a hangman.'' This displays a lack of social interest. The boy wished to be the master of life \& death -- a role which belongs to God. He wished to be more powerful than society, \& he was thus headed for the useless life. The goal of being a doctor is also fashioned around the God-like desire of being master of life \& death, but here the goal is realized through social service.'' -- \cite[pp. 31--55]{Adler2013}

%------------------------------------------------------------------------------%

\section{The Inferiority Complex}
``The use of the terms ``consciousness'' \& ``unconsciousness'' to designate distinctive factors is incorrect in the practice of Individual Psychology. Consciousness \& unconsciousness move together in the same direction \& are not contradictions, as is so often believed. What is more, there is no definite line of demarcation between them. It is merely a question of discovering the purpose of their joint movement. It is impossible to decide on what is conscious \& what is not until the whole connection has been obtained. This connection is revealed in the prototype, that pattern of life which we analyzed in the last chapter.

A case history will serve to illustrate the intimate connection between conscious \& unconscious life. A married man, 40 years old, suffered from 1 anxiety -- a desire to jump out of the window. He was always struggling against this desire, but aside from this he was quite well. He had friends, a good position, \& lived with his wife happily. His case is inexplicable except in terms of the collaboration of consciousness \& unconsciousness. Consciously he had the feeling that he must jump out of a window. Nonetheless he lived on, \& in fact he never even attempted to jump out of a window. The reason for this is that there was another side to his life, a side in which a struggle against his desire to commit suicide played an important part. As a result of the collaboration of this unconscious side of his being with his consciousness, he came out victorious. In fact in his ``style of life'' -- to use a term about which we shall have more to say in a later chapter -- he was a conqueror who had attained the goal of superiority. The reader might ask how could this man feel superior when he had this conscious tendency \textit{to commit suicide?} The answer is that there was something in his being that was fighting his battle against his suicidal tendency. It is his success in this battle that made him a conqueror \& a superior being. Objectively his struggle for superiority was conditioned by his own weakness, as is very often the case with persons who in 1 way or another feel inferior. But the important thing is that in his own private battle his striving for superiority, his striving to live \& to conquer, came out ahead of his sense of inferiority \& desire to die -- \& this despite the fact that the latter was expressed in his conscious life \& the former in his unconscious life.

Let us see if the development of this man's prototype bears out our theory. Let us analyze his childhood remembrances. At an early age, we learn, he had trouble at school. He did not like other boys \& wanted to run away from them. Nonetheless he collected all his powers to stay \& face them. In other words we can already perceive an effort on his part to overcome his weakness. He faced his problem \& conquered.

If we analyze our patient's character, we shall see that his 1 aim in life was to overcome fear \& anxiety. In this aim his conscious ideas cooperated with his unconscious ones \& formed a unity. Now a person who does not see the human being as a unity might believe that this patient was not superior \& was not successful. He might think him to be only an ambitious person, one who wanted to struggle \& fight but who was at bottom a coward. Such a view would be erroneous, however, since it would not take into consideration all the facts in the case \& interpret them with reference to the unity of a human life. Our whole psychology, our whole understanding or striving to understand individuals would be futile \& useless if we could not be sure that the human being is a unity. If we presupposed 2 sides without relation to one another it would be impossible to see life as a complete entity.

In addition to regarding an individual's life as a unity, we must also take it together with its context of social relations. Thus children when 1st born are weak, \& their weakness makes it necessary for other persons to care for them. Now the style or the pattern of a child's life cannot be understood without reference to the persons who look after him \& who make up for his inferiority. The child has interlocking relations with the mother \& family which could never be understood if we confined our analysis to the periphery of the child's physical being in space. The individuality of the child cuts across his physical individuality, it involves a whole context of social relations.

What applies to the child applies also, to a certain extent, to men as a whole. The weakness which is responsible for the child's living in a family group is paralleled by the weakness which drives men to live in society. All persons feel inadequate in certain situations. They feel overwhelmed by the difficulties of life \& are incapable of meeting them single-handed. Hence 1 of the strongest tendencies in man has been to form groups in order that he may live as a member of a society \& not as an isolated individual. This social life has without doubt been a great help to him in overcoming his feeling of inadequacy \& inferiority. We know that this is the case with animals, where the weaker species always live in groups in order that their combined powers might help to meet their individual needs. Thus a herd of buffaloes can defend themselves against wolves. 1 buffalo alone would find this impossible, but in a group they stick their heads together \& fight with their feet until they are saved. On the other hand, gorillas, lions \& tigers can live isolated because nature has given them the means of self-protection. A human being has not their great strength, their claws, nor their teeth, \& so cannot live apart. Thus we find that the beginning of social life lies in the weakness of the individual.

Because of this fact we cannot expect to find that the abilities \& faculties of all human begins in society are equal. But a society that is rightly adjusted will not be behindhand in supporting the abilities of the individuals who compose it. This is an important point to grasp, since otherwise we would be led to suppose that individuals have to be judged entirely on their inherited abilities. As a matter of fact an individual who might be deficient in certain faculties if he lived in an isolated condition could well compensate for his lacks in a rightly organized society.

Let us suppose that our individual insufficiencies are inherited. It then becomes the aim of psychology to train people to live well with others, in order to help decrease the effect of their natural disabilities. The history of social progress tells the story of how men co-operated in order to overcome deficiencies \& lacks. Everybody knows that language is  a social invention, but few people realize that individual deficiency was the mother o that invention. This truth, however, is illustrated in the early behavior of children. When their desires are not being satisfied, they want to gain attention \& they try to do so by some sort of language. But if a child should not need to gain attention, he would not try to speak at all. This is the case in the 1st few months, when the child's mother supplies everything that the child wishes before it speaks. There are cases on record of children who did not speak until 6 years of age because it was never necessary for them to do so. The same truth is illustrated in the case of a particular child of deaf \& dumb parents. When he fell \& hurt himself he cried, but he cried without noise. He knew that noise would be useless as his parents could not hear him. Therefore he made the appearance of crying in order to gain the attention of his parents, but it was noiseless.

We see therefore that we must always look at the whole social context of the facts we study. We must look at the social environment in order to understand the particular ``goal of superiority'' an individual chooses. We must look at the social situation, too, in order to understand a particular maladjustment. Thus many persons are mal-adjusted because they find it impossible to make the normal contact with others by means of language. The stammerer is a case in point. If we examine the stammerer we shall see that since the beginning of his life he was never socially well adjusted. He did not want to join in activities, \& he did not want friends or comrades. His language development needed association with others, but he did not want to associate. Therefore his stammering continued. There are really 2 tendencies in stammerers -- one to associate with others, \& another that makes them seek isolation for themselves.

Later in life, among adult persons not living a social life, we find that they cannot speak in public \& have a tendency to stage fright. This is because they regard their audiences as enemies. They have a feeling of inferiority when confronted by a seemingly hostile \& dominating audience. The fact is that only when a person trusts himself \& his audience can he speak well, \& only then will he not have stage fright.

The feeling of inferiority \& the problem of social training are thus intimately connected. Just as the feeling of inferiority arises from a social maladjustment, so social training is the basic method by which we can all overcome our feelings of inferiority.

There is a direct connection between social training \& common sense. When we say that people solve their difficulties by common sense, we have in mind the pooled intelligence of the social group. On the other hand, as we indicated in the last chapter, persons who act with a private language \& a private understanding manifest an abnormality. The insane, the neurotics \& the criminals are of this type. We find that certain things are not interesting to them -- people, institutions, the social norms make no appeal to them. \& yet it is through these things that the road to their salvation lies.

In working with such persons our task is to make social facts appeal to them. Nervous persons always feel justified if they show good will. But more than good will is needed. We must teach them that it is what they actually accomplish, what they actually give, that matters in society.

While the feeling of inferiority \& the striving for superiority are universal, it would be a mistake to regard this fact as indicating that all men are equal. Inferiority \& superiority are the general conditions which govern the behavior of men, but besides these conditions there are differences in bodily strength, in health, \& in environment. For that reason different mistakes are made by individuals in the same given conditions. If we examine children we shall see that there is no one absolutely fixed \& right manner for them to respond. They respond in their individual ways. They strive towards a better style of life, but they all strive in their own way, making their own mistakes \& their own type of approximations to success.

Let us analyze some of the variations \& peculiarities of individuals. Let us take, e.g., left-handed children. There are children who may never known that they are left-handed because they have been so carefully trained in the use of the right hand. At 1st they are clumsy \& imperfect with the right hand, \& they are scolded, criticized \& derided. It is an error to deride, but both hands should be trained. A left-handed child can be recognized in the cradle because his left hand moves more than his right. In later life he may feel that he is burdened because of the imperfection of his right hand. On the other hand, he often develops a greatest interest in his right hand \& arm, which interest is manifested, e.g., in drawing, writing, etc. In fact it is not surprising to find that later in life such a child is better trained than a normal child. Because he has had to get interested, he has gotten up earlier, so to speak, \& thus his imperfection has led him to more careful training. This is often a great advantage in developing artistic talent \& ability. A child in such a position is usually ambitious \& fights to overcome his limitations. Sometimes, however, if the struggle is a serious one, he may become envious or jealous of others \& thus develop a greater feeling of inferiority which is more difficult to overcome than in normal cases. Through constant struggling a child may become a fighting child or a fighting adult, always striving with the fixed idea in mind that he ought not to be clumsy \& deficient. Such an individual is more burdened than others.

Children strive, make mistakes, \& develop in various ways according to the prototypes they formed in the 1st 4 or 5 years of life. The goal of each is different. 1 child may want to be a painter, while another may wish himself out of this world where he is a misfit. We may know how he can overcome his imperfection, but he does not know it, \& all too often the facts are not explained to him in the right way.

Many children have imperfect eyes, ears, lungs or stomachs, \& we find their interest stimulated in the direction of the imperfection. A curious instance of this is revealed in the case of a man who suffered from attacks of asthma only when he came home at night from the office. He was a man of 45, married, \& with a good position. He was asked why the attacks always occurred after he came home from the office. He explained, ``You see, my wife is very materialistic \& I am idealistic, hence we do not agree. When I come home I would like to be quiet, to enjoy myself at home, but my wife wants to go into society \& so she complains about remaining at home. Whereupon I get into a bad temper \& start to suffocate.''

Why did this man suffocate: why did he not vomit? The fact is he was only being true to his prototype. It seems that as a child he had to be bandaged for some weakness \& this tight binding affected his breathing \& made him very uncomfortable. He had a maid servant, however, who liked him \& would sit beside him \& console him. All her interest was in him \& not in herself. She thus gave him the impression that he would always be amused \& consoled. When he was 4 years old the nurse went away to a wedding \& he accompanied her to the station crying very bitterly. After the nurse had left he said to his mother, ``The world has no more interest for me now that my nurse has gone away.''

Hence we see him in manhood as in the years of his prototype, looking for an ideal person who would always amuse him \& console him \& be interested in him alone. The trouble was not too little air but the fact that he was not being amused \& consoled at all times. Naturally, to find a person who will always amuse you is not easy. He always wanted to rule the whole situation \& to a certain degree it helped him when he succeeded. Thus when he took to suffocating, his wife stopped wanting to go to the theater or into society. He had then obtained his ``goal of superiority.''

Consciously this man was always right \& proper, but in his mind he has the desire to be the conqueror. He wanted to make his wife what he called idealistic instead of materialistic. We should suspect such a man of motives different from those on the surface $\ldots$

We often see children with imperfect eyes take more of an interest in visual things. They develop a keen faculty in this way. We see Gustav Freitag, a great poet who had poor, astigmatic eyes, accomplishing much. Poets  \& painter often have trouble with their eyes. But this in itself often creates greater interest. Freitag said about himself: ``Because my eyes were different from those of other people, it seems that I was compelled to use \& train my fantasy. I do not know that this has helped me to be a great writer, but in any case as a result of my eyesight it has come about that I can see better in fantasy than others in reality.''

If we examine the personalities of geniuses we shall often find poor eyes or some other deficiency. In the history of all ages even the gods have had some deficiency such as blindness in 1 or both eyes. The fact that there are geniuses who though nearly blind are yet able to understand better than others the differences in lines, shadows \& colors shows what can be done with afflicted children if their problems are properly understood.

Some people are more interested in eatables than others. Because of this they are always discussing what they can \& what they cannot eat. Usually such persons have had a hard time at the beginning of life in the matter of eating \& so have developed more interest in it than others. They had probably been told constantly by a watchful mother what they could \& could not eat. Such persons have to train to overcome the imperfections of their stomachs, \& they become vitally interested in what they will have for lunch, dinner or breakfast. As a result of their constant thought about eating they sometimes develop the art of cookery or become experts on questions of diet.

At times, however, a weakness of the stomach or the intestines causes people to look for a substitute for eating. Sometimes this substitute is money, \& such persons become miserly or great money-making bankers. They often strive extremely hard to collect money, training themselves for this purpose day \& night. They never stop thinking of their business, -- a fact which may sometimes give them a great advantage over others in similar walks of life. \& it is interesting to note that we often hear of rich men suffering from stomach trouble.

Let us remind ourselves at this point of the connection frequently made between body \& mind. A given defect does not always lead to the same result. There is no necessary cause \& effect relation between a physical imperfection \& a bad style of life. For the physical imperfection we can often give good treatment in the form of right nutrition \& thereby partly obviate the physical situation. But it is not the physical defect which causes the bad results: it is the patient's attitude which is responsible. That is why for the individual psychologist mere physical defects or exclusive physical causality does not exist, but only mistaken attitudes towards physical situations. Also that is why the individual psychologist seeks to foster a striving against the feeling of inferiority during the development of the prototype.

Sometimes we see a person impatient because he cannot wait to overcome difficulties. Whenever we see persons constantly in motion, with strong tempers \& passions, we can always conclude that they are persons with a great feeling of inferiority. A person who knows he can overcome his difficulties will not be impatient. On the other hand he may not always accomplish what is necessary. Arrogant, impertinent, fighting children also indicate a great feeling of inferiority. It is our task in their case to look for the reasons -- for the difficulties they have -- in order to prescribe the treatment. We should never criticize or \textit{punish} mistakes in the style of life of the prototype.

We can recognize these prototype traits among children in very peculiar ways -- in their unusual interests, in their scheming \& striving to surpass others, \& in building toward the goal of superiority. There is a type that does not trust himself in movement \& expression. He prefers to exclude others as far as possible. He prefers not to go where he is confronted with new situations but to stay in the little circle in which he feels sure. In school, in life, in society, in marriage he does the same. He is always hoping to accomplish much in his little place in order to arrive at a goal of superiority. We find this trait among many human beings. They all forget that to accomplish results, one must be prepared to meet all situations. Everything must be faced. If one eliminates certain situations \& certain persons, one has only private intelligence to justify oneself, \& this is not enough. One needs all the renovating winds of social contact \& common sense.

If a philosopher wants to accomplish his work, he cannot always go to lunch or dinner with others, for he needs to be alone for long periods of time in order to collect his ideas \& use the right method. But later on he must grow through contact with society. This contact is an important part of his development. \& so when we meet with such a person we must remember his 2 requirements. We must remember, too, that he can be useful or useless \& should therefore look carefully for the difference between useful \& useless behavior.

The key to the entire social process is to be found in the fact that persons are always striving to find a situation in which they excel. Thus children who have a great feeling of inferiority want to exclude stronger children \& play with weaker children whom they can rule \& domineer. This is an abnormal \& pathological expression of the feeling of inferiority, for it is important to realize that it is not the sense of inferiority which matters but the degree \& character of it.

The abnormal feeling of inferiority has acquired the name of ``inferiority complex.'' But complex is not the correct word for this feeling of inferiority that permeates the whole personality. It is more than a complex, it is almost a disease whose ravages vary under different circumstances. Thus we sometimes do not notice the feeling of inferiority when a person is on his job because he feels sure of his work. On the other hand he may not be sure of himself in society or in his relations with the opposite sex, \& in  this way we are able to discover his true psychological situation.

We notice mistakes in a greater degree in a tense or difficult situation. It is in the difficult or new situation that the prototype appears rightly, \& in fact the difficult situation is nearly always the new one. That is why, as we said in the 1st chapter, the expression of the degree of social interest appears in a new social situation.

If we put a child to school we may observe his social interest there just as in general social life. We can see whether he mixes with his fellows or avoids them. If we see hyperactive, sly, clever children, we must look into their minds to find the reasons. \& if we see some go forward only conditionally or hesitatingly, we must be on the lookout for the same characteristics to be revealed later on in society, life \& marriage.

We always meet persons who say, ``I would do this in this way,'' ``I would take that job,'' ``I would fight that man, $\ldots$ but $\ldots$!'' All such statements are a sign of a great feeling of inferiority, \& in fact if we read them this way we get a new light on certain emotions, such as doubt. We recognize that a person in doubt usually remains in doubt \& accomplishes nothing. However, when a person says ``I won't,'' he will probably act accordingly.

The psychologist, if he looks closely can often see contradictions in men. Such contradictions may be considered as a sign of a feeling of inferiority. But we must also observe the movements of a person who constitutes our problem on hand. Thus, his approach, his way of meeting people, may be poor, \& we must observe if he comes towards persons with a hesitating step \& bodily attitude. This hesitation will often be expressed in other situations of life. There are many persons who take 1 step forward \& 1 backward -- a sign of a great feeling of inferiority.

Our whole task is to train such persons away from their hesitating attitude. The proper treatment for such persons is to encourage them -- never to discourage them. We must make them understand that they are capable of facing difficulties \& solving the problems of life. This is the only way to build self-confidence, \& this is the only way the feeling of inferiority should be treated.'' -- \cite[pp. 56--77]{Adler2013}

%------------------------------------------------------------------------------%

\section{The Superiority Complex}
``In the last chapter we discussed the inferiority complex \& its relation to the general feeling of inferiority which all of us share \& struggle against. Now we have to turn to the inverse topic, the superiority complex.

We have seen how every symptom of an individual's life is expressed in a movement -- in a progress. Thus the symptom may be said to have a past \& a future. Now the future is tied up with our striving \& with our goal, while the past represents the state of inferiority or inadequacy which we are trying to overcome. That is why in an inferiority complex we are interested in the beginning, while in a superiority complex we are more interested in the continuity, in the progression of the movement itself. Moreover, the 2 complexes are naturally related. We should not be astonished if in the cases where we see an inferiority complex we find a superiority complex more or less hidden. On the other hand, if we inquire into a superiority complex \& study its continuity, we can always find a more or less hidden inferiority complex.

We muse bear in mind of course that the word complex as attached to inferiority \& superiority merely represents an exaggerated condition of the sense of inferiority \& the striving for superiority. If we look at things this way it takes away the apparent paradox of 2 contradictory tendencies, the inferiority complex \& the superiority complex, existing in the same individual. For it is obvious that as normal sentiments the striving for superiority \& the feeling of inferiority are naturally complementary. We should not strive to be superior \& to succeed if we did not feel a certain lack in our present condition. Now inasmuch as the so-called complexes develop out of the natural sentiments, there is no more contradiction in them than in the sentiments.

The striving for superiority never ceases. It constitutes in fact the mind, the psyche of the individual. As we have said, life is the attainment of a goal or form, \& it is the striving for superiority which sets the attainment of form into motion. It is like a stream which drags along all the material it can find. If we look at lazy children \& see their lack of activity, their lack of interest in anything, we should say that they do not seem to be moving. But nonetheless we find in them a desire to be superior, a desire which makes them say, ``If I were not so lazy, I could be president.'' They are moving \& striving conditionally, so to speak. They hold a high opinion of themselves \& take the view that they could accomplish much on the useful side of life, if $\ldots$! This is lying, of course -- it's fiction, but as we all know, mankind is very often satisfied with fiction. \& this is especially true of persons who lack courage. They content themselves quite well with fiction. They do not feel very strong \& so they always make detours -- they always want to escape difficulties. Through this escape, through this avoiding of battle they get a feeling of being much stronger \& cleverer than they really are.

We see children who start stealing suffering from the feeling of superiority. They believe they are deceiving others; that others do not know they are stealing. Thus they are richer with little effort. This same feeling is very pronounced among criminals who have the idea that they are superior heroes.

We have already spoken of this trait from another aspect as a manifestation of private intelligence. It is not common or social sense. If a murderer thinks himself a hero, it is a private idea. He is lacking in courage since he wants to arrange matters so that he escapes the solution of the problems of life. Criminality is thus the result of a superiority complex \& not the expression of fundamental \& original viciousness.

We see similar symptoms appearing among neurotic persons. E.g., they suffer from sleeplessness \& so are not strong enough next day to comply with the demands of their occupations. Because of sleeplessness they feel that they cannot be required to work because they are not equal to doing what they could accomplish. They lament, ``What could I not do if I could only get my sleep!''

We see this also among depressed persons suffering from anxiety. Their anxiety makes them tyrants over others. In fact they use their anxiety to rule others, for they must always have people with them, they must be accompanied wherever they go, etc. The companions are made to live their lives in accordance with the demands of the depressed person.

Melancholy \& insane persons are always the center of attention in the family. In them we see the power wielded by the inferiority complex. They complain that they feel weak \& are losing weight, etc., but nonetheless they are the strongest of all. They dominate healthy persons. This fact should not surprise us, for in our culture weakness can be quite strong \& powerful. (In fact if we were to ask ourselves who is the strongest person in our culture, the logical answer would be, the baby. The baby rules \& cannot be dominated.)

Let us study the connection between the superiority complex \& inferiority. Let us take e.g. a problem child with a superiority complex -- a child that is impertinent, arrogant \& pugnacious. We shall find that he always wants to appear greater than he really is. We all know how children with temper tantrums want to control others by getting a sudden attack. Why are they so impatient? Because they are not sure they are strong enough to attain their goal. They feel inferior. We will always discover in fighting, aggressive children an inferiority complex \& a desire to overcome it. It is as if they were trying to lift themselves on their toes in order to appear greater \& to gain by this easy method success, pride \& superiority.

We have to find methods of treatment for such children. They act that way because they do not see the coherence of life. They do not see the natural order of things. We should not censure them because they do not want to see it, for if we confront them with the question, they will always insist that they do not feel inferior but superior. We must therefore in a friendly manner explain to them our point of view \& get them gradually to understand.

If a person is a show-off it is only because he feels inferior, because he does not feel strong enough to compete with others on the useful side of life. That is why he stays on the useless ride. He is not in harmony with society. He is not socially adjusted, \& he does not know how to solve the social problems of life. \& so we always find a struggle between him \& his parents \& teachers during his childhood. In such cases the situation must be understood \& also made understandable to the children.

We see the same combination of inferiority \& superiority complexes in neurotic illnesses. The neurotic frequently expresses his superiority complex but does not see his inferiority complex. The case history of a compulsion neurotic is very illuminating in this regard. There was a young girl in close association with an elder sister who was very charming \& much esteemed. This fact is significant at the outset, for if 1 person out of a family is more outstanding than the others, the latter will suffer. This is always so, whether the favored individual be the father, 1 of the children, or the mother. A very difficult situation is created for the other members of the family, \& sometimes they feel they cannot bear it.

Now we will find among these other children that they all have an inferiority complex \& are striving toward a superiority complex. So long as they are interested not only in themselves but in others, they will solve their problems of life satisfactorily. But if their inferiority complex is clearly marked, they find themselves living, as it were, in an enemy country -- always looking out for their own interests rather than for those of others, \& thus not having the right amount of communal sense. They approach the social questions of life with a feeling that is not conducive to their solution. \& so, seeking relief, they go over to the useless side of life. We know that  this is not really relief, but it seems like relief not to solve questions but to be supported by others. They are like beggars, who are being supported by others \& who feel comfortable neurotically exploiting their weakness.

It seems to be a trait of human nature that when individuals -- both children \& adults -- feel weak, they cease to be interested socially but strive for superiority. They want to solve the problems of life in such a way as to obtain personal superiority without any admixture of social interest. As long as a person strives for superiority \& tempers it with social interest, he is on the useful side of life \& can accomplish good. But if he lacks social interest, he is not really prepared for the solution of the problems of life. In this category should be put, as we have already said, the problem children, the insane, the criminals, those who commit suicide, etc.

Now this girl of whom we started to speak grew up outside of a favorable circle \& felt herself restricted. If she had been socially interested, \& had understood what we understand, she could have developed along another line. She began to study to be a musician, but she was always at such tension, due to the inferiority complex caused by always thinking of her preferred sister, that she was blocked here too. When she was 20 her sister married \& so she began to look for marriage in order to compete with her sister. In this way she was getting in deeper, \& drifting more \& more from the healthy, useful side of life. She developed the idea that she was a bad, bad girl, \& possessed magic power which could send a person to hell.

We see this magic power as a superiority complex but she on the other hand complained, just as we sometimes hear rich men complain of how bad their fate is to be rich men. Not only did she feel that she had the god-like power of sending people to hell, but at times she got the impression that she could \& ought to save these people. Of course both of these claims were ridiculous, but by means of this system of fiction she assured herself of possessing a power that was higher than her preferred sister's. She could overcome her sister only by this game. \& so she complained that she had this power, for the more she complained about it the more plausible it was that she actually possessed it. If she had laughed about it, the claim of power would have been questionable. Only by complaining could she feel happy with her lot. We see here how a superiority complex may sometimes be hidden, not recognized as present, yet existing in fact as a compensation for the inferiority complex.

The older sister -- of whom we shall now speak -- was very much favored, for at 1 time she was the only child, much pampered, \& the center of attention in the family. 3 years later there arrived a younger sister, which fact changed the whole situation for the older girl. Formerly she had always been alone, the center of attention. Now she was suddenly thrown out of this position. As a result she became a fighting child. But there can be fighting only where there are weaker companions. A fighting child is not really courageous -- he fights only against weaker persons. If the environment is strong, then instead of becoming pugnacious, a child becomes peevish, or depressed, \& is likely to be less appreciated in the home circle for this reason.

In such cases the older child feels she is not as dearly loved as before, \& she sees the manifestations of the changed attitude as a confirmation of her view. She considers her mother the most guilty inasmuch as it is she who has brought this other girl into the home. Thus we can understand her directing attacks against her mother.

The baby, on the other hand, has to be watched, observed, pampered as all babies are, \& is thus in a favorable position. Therefore she does not need to exert herself, does not need to fight. She develops as a very sweet, very soft \& very much beloved creature -- the center of the family. Sometimes virtue in the form of obedience may conquer!

Now let us examine \& see if this sweetness, softness \& kindness was on the useful side of life or not. We may presuppose that she was so amenable \& tractable only because she was so pampered. But our civilization does not regard pampered children with favor. Sometimes the father realizes this \& wants to end this state of affairs. Sometimes the school comes into the situation. The position of such a child is always in danger \& for this reason the pampered child feels inferior. We do not notice this feeling of inferiority among pampered children so long as they are in a favorable situation, but the moment an unfavorable situation arises we see these children either breaking down \& becoming depressed or developing a superiority complex.

The superiority complex \& inferiority complex agree on 1 point, namely, that they are always on the useless side. We can never find an arrogant, impertinent child, one with a superiority complex, on the useful side of life.

When these pampered children go to school, they are no longer in a favorable situation. From that moment on we see them adopting a hesitating attitude in life \& never finishing anything. So it was with the younger sister of whom we 1st spoke. She began to learn to sew, to play the piano, etc., but after a short time she stopped. At the same time she lost interest in society, did not like to go out any more \& felt depressed. She felt herself overshadowed by her sister with her more agreeable characteristics. Her hesitating attitude made her weaker \& caused a deterioration of her character.

Later in life she hesitated in the matter of occupations \& never finished anything. She also hesitated in love \& marriage, despite her desire to compete with her sister. When she reached 30 she looked around \& found a man who was suffering from tuberculosis. Of course we can readily see that this selection would be opposed by her parents. In this case it was not necessary for her to stop action, for her parents stopped the action, \& the marriage did not take place. A year later she married a man 35 years her senior. Now as such a man is not thought to be a man any more, this marriage which was not a marriage seemed useless. We often find an expression of an inferiority complex in the selection of a much older person for marriage or in the selection of a person who cannot be married; e.g., a married man or woman. There is always a suspicion of cowardice when there are hindrances. Because this girl did not justify her feeling of superiority in marriage, she found another way of acquiring a superiority complex.

She insisted that the most important thing in this world is duty. She had to wash herself all the time. If anybody or anything touched her, she had to wash again. In this way she became wholly isolated. As a matter of fact her hands were as dirty as they could be. The reason was obvious: because of her continual washing she acquired a very rough skin that collected dirt in great quantities.

Now all this looks like an inferiority complex, but she felt herself to be the only pure person in the world \& was continually criticizing \& accusing others because they did not have her washing mania. So she played her role as in a pantomime. She had always wanted to be superior \& now in a fictitious way she was. She was the purest person in the world. So we see that her inferiority complex had become a superiority complex, very distinctly expressed.

We see the same phenomenon in megalomaniacs who believe themselves to be Jesus Christ or an emperor. Such a person is on the useless side of life \& plays his role almost as if it were true. He is isolated in life, \& we shall find, if we go back to his past, that he felt inferior \& that, in a worthless way, he developed a superiority complex.

There is the case of a boy of 15 who entered an asylum for the insane because of his hallucinations. At that time, which was before the war, he fancied that the emperor of Austria was dead. This was not true, but he claimed that the emperor had appeared to him in a dream demanding that he lead the Austrian army against the enemy. \& he is a little undersized boy! He would not be convinced when he was shown the newspapers, which reported that the emperor was stopping at his castle or that he had been out driving in his car. He insisted that the emperor was dead \& had appeared to him in a dream.

At that time Individual Psychology was trying to find out the importance of positions in sleep in indicating a person's feeling of superiority or inferiority. One can see that such information might prove useful. Some persons lie in bed in a curved line like a hedgehog, covering their heads with the covers. This expresses an inferiority complex. Can we believe such a person to be courageous? Or if we see a person stretched out straight, can we believe him weak or bent in life? Both in a literal \& metaphorical way he will appear great, as he does in sleep. It has been observed that persons who sleep on their stomachs are stubborn \& pugnacious.

This boy was examined in an attempt to find correlations between his walking behavior \& his positions in sleep. It was found that he slept with arms crossed on his breast, like Napoleon. As we all know the pictures show Napoleon with his arms in such a position. Next day the boy was asked, ``Do you know somebody of whom this position reminds you?'' He answered, ``Yes, my teacher.'' The discovery was a little disturbing until it was suggested that the teacher might be like Napoleon. This proved to be the case. Moreover, the boy had loved this teacher \& wanted to be a teacher like him. But for lack of funds with which to assure him an education, his family had to put him to work in a restaurant where he patrons had all derided him because he was undersized. He could not bear this \& wanted to escape from this feeling of humiliation. But he escaped to the useless side of life.

We are able to understand what happened in the case of this boy. In the beginning he had an inferiority complex because he was undersized \& hence derided by the guests in the restaurant. But he was constantly striving for superiority. He wanted to be a teacher. But because he was blocked in attaining this occupation, he found another goal of superiority by making a detour to the useless side of life. He became superior in sleep \& dreams.

Thus we see that the goal of superiority may be on the useless or useful side of life. If a person is benevolent, e..g, it may mean either of 2 things -- it may mean that he is socially adjusted \& wants to help, or else it may mean simply that he wants to boast. The psychologist meets with many whose main goal is to boast. There is the case of a boy who was not very accomplished in school; in fact he was so bad that he became a truant \& stole things, but he was always boastful. He did these things because of his inferiority complex. He wanted to accomplish results in some line -- be it only the line of cheap vanity. Thus he stole money \& presented prostitutes with flowers \& other gifts. 1 day he drove a car far away to a little town \& there he demanded a carriage \& 6 horses. He rode all through the town in state until he was arrested. In all his behavior his great striving was to appear greater than others -- \& greater than he really was.

A similar tendency may be remarked in the behavior of criminals -- the tendency to claim easy success, which we have already discussed in another connection. The New York newspapers some time ago reported how a burglar broke into the home of some schoolteachers \& had a discussion with them. The burglar told the women they did not know how much trouble there was in ordinary honest occupations. It was much easier to be a burglar than to work. This man had escaped to the useless side of life. But by taking this road he had developed a certain superiority complex. He felt stronger than the women, particularly since he was armed \& they were not. But did he realize that he was a coward? We know he is because we see him as a person who had escaped his inferiority complex by going over to the useless side of life. He thought himself a hero, however, \& not a coward.

Some types turn to suicide \& desire in this way to throw off the whole world with its difficulties. They seem not to care for life \& so feel superior, although they are really cowards. We see that a superiority complex is a 2nd phase. It is a compensation for the inferiority complex. We must always try to find the organic connection -- the connection which may seem to be a contradiction but which is quite in the course of human nature, as we have already shown. Once this connection is found we are in a position to treat both the inferiority \& superiority complexes.

We should not conclude the general subject of inferiority \& superiority complexes without saying a few words as to the relation of these complexes to normal persons. Everyone, as we have said, has a feeling of inferiority. But the feeling of inferiority is not a disease, it is rather a stimulant to healthy normal striving \& development. It becomes a pathological condition only when the sense of inadequacy overwhelms the individual, \& so far from stimulating him to useful activity, makes him depressed \& incapable of development. Now the superiority complex is 1 of the ways which a person with an inferiority complex may use as a method of escape from his difficulties. He assumes that he is superior when he is not, \& this false success compensates him for the state of inferiority which he cannot bear. The normal person does not have a superiority complex, he does not even have a sense of superiority. He has the striving to be superior in the sense that we all have ambition to be successful, but so long as this striving is expressed in work it does not lead to false valuations, which is at the root of mental disease.'' -- \cite[pp. 78--97]{Adler2013}

%------------------------------------------------------------------------------%

\section{The Style of Life}
``If we look at a pine tree growing in the valley we will notice that it grows differently from one on top of a mountain. It is the same kind of a tree, a pine, but there are 2 distinct styles of life. Its style on top of the mountain is different from its style when growing in the valley. The style of life of a tree is the individuality of a tree expressing itself \& moulding itself in an environment. We recognize a style when we see it against a background of an environment different from what we expect, for then we realize that every tree has a life pattern \& is not merely a mechanical reaction to the environment.

It is much the same way with human beings. We see the style of life under certain conditions of environment \& it is our task to analyze its exact relation to the existing circumstances, inasmuch as mind changes with alteration of the environment. As long as a person is in a favorable situation we cannot see his style of life clearly. In new situations, however, where he is confronted with difficulties, the style of life appears clearly \& distinctly. A trained psychologist could perhaps understand a style of life of a human being even in a favorable situation, but it becomes apparent to everybody when the human subject is put into unfavorable or difficult situations.

Now life, being something more than a game, does not lack difficulties. There are always situations in which human beings find themselves confronted with difficulties. It is while the subject is confronted with these difficulties that we must study him \& find out his different movements \& characteristic distinguishing marks. As we have previously said, the style of life is a unity because it has grown out of the difficulties of early life \& out of the striving for a goal.

But we are interested not so much in the past as in the future. \& in order to understand a person's future we must understand his style of life. Even if we understand instincts, stimuli, drive, etc., we cannot predict what must happen. Some psychologists indeed try to reach conclusions by noting certain instincts, impressions or traumas, but on closer examination it will be found that all these elements presuppose a consistent style of life. Thus whatever stimulates, stimulates only to \textit{save} \& \textit{fix} a style of life.

How does the notion of the style of life tie up with what we have discussed in previous chapters? We have seen how human beings with weak organs, because they face difficulties \& feel insecure, suffer from a feeling or complex of inferiority. But as human beings cannot endure this for long, the inferiority feeling stimulates them, as we have seem, to movement \& action. This results in a person having a goal. Now Individual Psychology has long called the consistent movement toward this goal a plan of life. But because this name has sometimes led to mistakes among students, it is now called a style of life.

Because an individual has a style of life, it is possible to predict his future sometimes just on the basis of talking to him \& having him answer questions. It is like looking at the 5th act of a drama, where all the mysteries are solved. We can make predictions in this way because we know the phases, the difficulties \& the questions of life. Thus from experience \& knowledge of a few facts we can tell what will happen to children who always separate themselves from others, who are looking for support, who are pampered \& who hesitate in approaching situations. What happens in the case of a person whose goal it is to be supported by others? Hesitating, he stops or escapes the solution of the questions of life. We know how he can hesitate, stop, or escape, because we have seen the same thing happen a thousand times. We know that he does not want to proceed alone but wants to be pampered. He wants to stay far away from the great problems of life, \& he occupies himself with useless things rather than struggle with the useful ones. He lacks social interests, \& as a result he may develop into a problem child, a neurotic, a criminal or a suicide -- that final escape. All these things are now better understood than formerly.

We realize, e.g., that in looking for the style of a life of a human being we may use the normal style of life as a basis for measurement. We use the socially adjusted human being as a stand, \& we can measure the variations from the normal.

At this point perhaps it would be helpful to show how we determine the normal style of life \& how on the basis of it we understand mistakes \& peculiarities. But before we discuss this we ought to mention that we do not count types in such studies. We do not consider human beings types because every human being has an individual style of life. Just as one cannot find 2 leaves of a tree absolutely identical, so one cannot find 2 human beings absolutely alike. Nature is so rich \& the possibilities of stimuli, instincts \& mistakes are so numerous, that it is not possible for 2 persons to be exactly identical. If we speak of types, therefore, it is only as an intellectual device to make more understandable the similarities of individuals. We can judge better if we postulate an intellectual classification like a type \& study its special peculiarities. However, in doing so we do not commit ourselves to using the same classification at all times; we use the classification which is most useful for bringing out a particular similarity. People who take types \& classifications seriously, once they put a person in a pigeonhole, do not see how he can be put into any other classification.

An illustration will make our point clear. E.g., when we speak of a type of individual not socially adjusted, we refer to one who leads a barren life without any social interests. This is 1 way of classifying individuals, \& perhaps it is the most important way. But consider the individual, whose interest, however limited, is centered on visual things. Such a person differs entirely from one whose interests are largely concentrated on things oral, but both of them may be socially mal-adjusted \& find it difficult to establish contact with their fellow-men. The classification by types can thus be a source of confusion if we do not realize that types are merely convenient abstractions.

Let us return now to the normal man, who is our standard for measuring variations. The normal man is an individual who lives in society \& whose mode of life is so adapted that whether he wants it or not society derives a certain advantage from his work. Also from a psychological point of view he has enough energy \& courage to meet the problems \& difficulties as they come along. Both of these qualities are missing in the case of psychopathic persons: they are neither socially adjusted nor are they psychologically adjusted to the daily tasks of life. As an illustration we may take the care of a certain individual, a man of 30 who was always at the last moment escaping the solution of his problems. He had a friend but was very suspicious of him, \& as a result this friendship never prospered. Friendship cannot grow under such conditions because the other partner feels the tension in the relation. We can readily see how this man really had no friends despite the fact that he was on speaking terms with a large number of persons. He was not sufficiently interested nor adjusted socially to make friends. In fact he did not like society, \& was always silent in company. He explained this on the ground that in company he never had any ideas \& therefore he had nothing to say.

Moreover, the man was bashful. He had a pink skin which flushed from time to time when he talked. When he could overcome this bashfulness he would speak quite well. What he really needed was to be helped in this direction without criticism. Of course when he was in this state he did not present a nice picture \& was not very much liked by his neighbors. He felt this, \& as a result his dislike for speech increased. One might say that his style of life was such that if he approached other persons in society he called attention to himself.

Next to social life \& the art of getting along with friends, is the question of occupation. Now our patient always had the fear that he might fail in his occupation, \& so he studied day \& night. He overworked \& overstrained himself. \& because he overstrained himself he put himself out of commission for solving the question of occupation.

If we compare our patient's approach to the 1st \& 2nd questions in his life, we see that he was always in too great a tension. This is a sign that he had a great feeling of inferiority. He undervalued himself \& looked on others \& on new situations as things that were unfriendly to him. He acted as though he was in an enemy country.

We have now enough data to picture the style of life of this man. We can see that he wants to go on but at the same time he is blocked because he fears defeat. It is as if he stood before an abyss, straining \& always at a tension. He manages to go forward but only conditionally, \& he would prefer to stay at home \& not mingle with others.

The 3rd question with which this man was confronted -- \& it is a question on which most persons are not very well prepared -- is the question of love. He hesitated to approach the other sex. He found that he wanted to love \& to get married, but on account of his great feeling of inferiority he was too frightened to face the prospect. He could not accomplish what he wanted \&, so we see his whole behavior \& attitude summed up in the words, ``Yes $\ldots$ but!'' We see him in love with 1 girl \& then in love with another. This is of course a frequent occurrence with neurotic persons because in a sense 2 girls are less than one. This truth sometimes accounts for a tendency towards polygamy.

\& now let us take up the reasons for this style of life. Individual Psychology undertakes to analyze the causes for a style of life. This man established his style of life during the 1st 4 or 5 years. At that time some tragedy happened which moulded \& formed him, \& so we have to look for the tragedy. We can see that something made him lose his normal interest in others \& gave him the impression that life is simply 1 great difficulty \& that it is better not to go on at all than to be always confronting difficult situations. Therefore he became cautious, hesitant, \& a seeker of ways of escape.

We must mention the fact that he was a 1st child. We have already spoken about the great significance of this position. We have shown how the chief problem in the case of a 1st child arises from the fact that he is for years the center of attention, only to be displaced from his glory \& another preferred. In a great many cases where a person is bashful \& afraid to go on we find the reason to be that another person has been preferred. Hence in this case it is not difficult to find out where the trouble lies.

In many cases we need only ask a patient. Are you the 1st, 2nd, or 3rd child? Then we have all we need. We can also use an entirely different method: we can ask for old remembrances, which we shall discuss at some length in the next chapter. This method is worthwhile because these remembrances or 1st pictures are a part of the building up of the early style of life which we have called the prototype. One comes upon an actual part of the prototype when a person tells of his early remembrances. Looking back, everybody remembers certain important things, \& indeed what is fixed in memory is always important. There are schools of psychology which act on the opposite assumption. They believe that what a person has forgotten is the most important point, but there is really no great difference between the 2 ideas. Perhaps a person can tell us his conscious remembrances, but he does not know what they mean. He does not see their connection with his actions. Hence the result is the same, whether we emphasize the hidden or forgotten significance of conscious memories or the importance of forgotten memories.

Little descriptions of old remembrances are highly illuminating. Thus a man might tell you that when he was small, his mother took him \& his younger brother to market. That is enough. We can then discover his style of life. He pictures himself \& a younger brother. Therefore we see it must have been important to him to have had a younger brother. Lead him further \& you may find a situation similar to a certain one in which a man recalled that it began to rain that day. His mother took him in her arms, but when she saw the younger brother she put him down to carry the little one. Thus we can picture his style of life. He always hsa the expectation that another person will be preferred. \& so we can understand why he cannot speak in society for he is always looking around to see if another will not be preferred. The same is true with friendship. He is always thinking that another is more preferred by his friend, \& as a result he can never have a true friend. He is constantly suspicious, looking out for little things that disturb friendship.

We can also see how the tragedy he has experienced has hindered the development of his social interest. He recalls that his mother took the younger brother in her arms \& we see that he feels that this baby took more of his mother's attention than he did. He feels that the younger brother is preferred \& is looking constantly for confirmation of this idea. He is wholly convinced he is right, \& so he is always under strain -- always under the great difficulty of trying to accomplish things when some one else is preferred.

Now the only solution for such a suspicious person is complete isolation, so that he would not have to compete at all with others \& would be, so to speak, the only human being on this earth's crust. Sometimes indeed it appears in fancy to such a child that the whole world has broken down, that he is the only person left \& that hence no one else can be preferred. We see how he taps all the possibilities to save himself. But he does not go along the lines of logic, common sense, or truth -- rather along the lines of suspicion. He lives in a limited world, \& he has a private idea of escape. He has absolutely no connection with others \& no interest in others. But he is not to be blamed for we know that he is not really normal.

It is our task to give such a person the social interest demanded of a well-adjusted human being. How is this to be done? The great difficulty with persons trained in this way is that they are overstrained \& are always looking for a confirmation of their fixed ideas. It thus becomes impossible to change their ideas unless somehow we penetrate into their personality in a manner that will disarm their preconceptions. To accomplish this it is necessary to use a certain art \& a certain tact. \& it is best if the adviser is not closely related or interested in the patient. For if one is directly interested in the case, one will find that one is acting for one's own interest \& not for the interest of the patient. The patient will not fail to notice this \& will become suspicious.

The important thing is to decrease the patient's feeling of inferiority. It cannot be extirpated altogether, \& in fact we do not want to extirpate it because a feeling of inferiority can serve as a useful foundation on which to build. What we have to do is to change the goal. We have seen that his goal has been 1 of escape just because someone else is preferred, \& it is around this complex of ideas that we must work. We must decrease his feeling of inferiority by showing him that he really undervalues himself. We can show him the trouble with his movements \& explain to him his tendency to be over-tense, as if standing before a great abyss or as if living in an enemy country \& always in danger. We can indicate to him how his fear that others may be preferred, is standing in the way of his doing his best work \& making the best spontaneous impression.

If such a person could act as a host in society, making his friends have a good time \& being friendly with them \& thinking of their interests, he would improve tremendously. But in ordinary social life we see that he does not enjoy himself, does not have ideas \& as a result says: ``Stupid persons -- they cannot enjoy me, they cannot interest me.''

The trouble with such persons is that they do not understand the situation because of their private intelligence \& their lack of common sense. As we have said, it is as if they were always confronted by enemies \& were leading the life of a lone wolf. In the human situation such a life is a tragic abnormality.

Let us now look at another specific case -- the case of a man afflicted with melancholia. This is a very common illness, but it can be cured. Such persons are distinguishable very early in life. In fact we notice many children who in their approach to a new situation show signs of suffering from melancholia. This melancholy man of whom we are speaking had about 10 attacks, \& these always occurred when he took a new position. As long as he was in his old position he was nearly normal. But he did not want to go out into society \& he wanted to rule others. Consequently he had no friends \& at 50 he had not married.

Let us look at his childhood in order to study his style of life. He had been very sensitive \& quarrelsome, always ruling his older brothers \& sisters by emphasizing his pains \& weaknesses. When playing on a couch 1 day, he pushed them all off. When his aunt reproached him for this, he said, ``Now my whole life is ruined because you have blamed me!'' \& at that time he was only 4 or 5 years old.

Such was his style of life -- always trying to rule others, always complaining of his weakness \& of how he suffered. This trait led in his later life to melancholy, which in itself is simply an expression of weakness. Every patient with melancholia uses almost the same words: ``My whole life is ruined. I have lost everything.'' Frequently such a person has been pampered \& is so no longer, \& this influences his style of life.

Human beings in their reactions to situations are much like the different species of animals. A hare reacts differently to the same situation from a wolf or a tiger. So it is with human individuals. The experiment was once made of taking 3 different types of boys to a lion's cage in order to see how they would behave on seeing this terrible animal for the 1st time. The 1st boy turned \& said, ``Let's go home.'' The 2nd boy said, ``How nice!'' He wanted to appear courageous but he was trembling when he said it. He was a coward. The 3rd boy said, ``May I spit at him?'' Here then we see 3 different reactions, 3 different ways of experiencing the same situation. We see also that for the most part human beings have a tendency to be afraid.

This timidity, when expressed in a social situation, is 1 of the most frequent causes of maladjustment. There was a man of high-born family who never wanted to exert himself but always wished to be supported. He appeared weak, \& of course he could not find a position. Now when the situation at home changed for the worse, his brothers went after him, saying, ``You are so stupid that you cannot find a position. You do not understand anything.'' So this man began to drink. After some months he was a confirmed drunkard \& was put in an asylum for 2 years. It helped him but it did not benefit him permanently, for he was put back into society without preparation. He could find no work except as a laborer, although he was a scion of this well-known family. Soon he began to have hallucinations. He thought a man appeared to tease him so that he could not work. 1st he could not work because he was a drunkard \& then because he had hallucinations. \& so we see that it is not the right treatment merely to make a drunkard sober; we must find \& correct his style of life.

We discover on investigation that this man was a pampered child, always wanting to be helped. He was not prepared to work alone \& we see the results. We must make all children independent, \& this can be done only if we get them to understand the mistakes in their style of life. This child should have been trained to do something, \& then he would not have had to be ashamed in the presence of his brothers \& sisters.'' -- \cite[pp. 98--116]{Adler2013}

%------------------------------------------------------------------------------%

\section{Old Remembrances}
``Having analyzed the significance of an individual's style of life, we turn now to the topic of old remembrances, which are perhaps the most important means for getting at a style of life. By looking back through childhood memories we are able to uncover the prototype -- the core of the style of life -- better than by any other method.

If we want to find out the style of life of a person -- child or adult -- we should, after we have heard a little about his complaints, ask him for old remembrances \& then compare them with the other facts he has given. For the most part the style of life never changes. There is always the same person with the same personality, the same unity. A style of life, as we have shown, is built up through the striving for a particular goal of superiority, \& so we must except every word, act \& feeling to be an organic part of the whole ``action line.'' Now at some points this ``action line'' is more clearly expressed. This happens particularly in old remembrances.

We should not, however, distinguish too sharply between old \& new remembrances, for in new remembrances also the action line is involved. It is easier \& more illuminating to find the action line in the beginning, for then we discover the theme \& are able to understand how the style of life of a person does not really change. In the style of life formed at the age of 4 or 5 we find the connection between remembrances of the past \& actions of the present. \& so after many observations of this kind we can hold fast to the theory that in these old remembrances we can always find a real part of the patient's prototype.

When a patient looks back into his past we can be sure that anything his memory will turn up will be of emotional interest to him, \& thus we will find a clue to his personality. It is not to be denied that the forgotten experiences are also important for the style of life \& for the prototype, but many times it is more difficult to find out the forgotten remembrances, or, as they are called, the unconscious remembrances. Both conscious \& unconscious remembrances have the common quality of running towards the same goal of superiority. They are both a part of the complete prototype. It is well, therefore, to find both the conscious \& unconscious remembrances if possible. Both conscious \& unconscious remembrances are in the end about equally important, \& the individual himself generally understands neither. It is for the outsider to understand \& interpret both of them.

Let us begin with conscious remembrances. Some persons, when they are asked for old remembrances, answer, ``I do not know any.'' We must ask such persons to concentrate \& try to remember. After some effort we will find that they will recall something. But this hesitation may be considered as a sign that they do not want to look far back into their childhood \& we may then come to the conclusion that their childhood has not been pleasant. We have to lead such people. We must give them hints in order to find out what we want. They always remember something in the end.

Some persons claim that they can remember back to their 1st year. This is scarcely possible, \& the truth is probably that these are fancied memories, not true remembrances. But it does not matter whether they are fancies or true since they are parts of one's personality. Some persons insist they are not sure whether they remember a thing or whether their parents have told them about it. This, too, is not really important because even if their parents did tell them they have fixed it in their minds \& therefore it helps to tell us where their interest lies.

As we have explained in the last chapter it is convenient for certain purposes to classify individuals into types. Now old remembrances go according to types \& reveal what is to be expected of the behavior of a particular type. E.g., let us take the case of a person who remembers that he saw a marvelous Christmas tree, filled with lights, presents \& holiday cakes. What is the most interesting thing in this story? \textit{That he saw}. Why does he tell us that he has seen? Because he is always interested in visual things. He has struggled against some difficulties in sight, \&, having been trained, has always been interested \& attentive to seeing. Perhaps this is not the most important element of his style of life, but it is an interesting \& important part. It indicates that if we are to give him an occupation it should be one in which he will use his eyes.

In school the education of children too often disregards this principle of types. We may fnd a child interested in sight who will not listen because he always wants to be looking at something. In the case of such a child we ought to be patient in trying to educate him to hear. Many children at school are taught only in 1 way because they enjoy with 1 sense. They may be only good at listening or good at seeing. Some always like to be moving \& to be working. we cannot expect the same results for the 3 types of children, especially if the teacher prefers 1 method, as, e.g., the method for listening children. When such a method is used the lookers \& the doers will suffer \& will he hindered in their development.

Consider the case of a young man, 24 years old, who suffered from fainting spells. When asked for his remembrances, he recalled that when he was 4 years old he fainted when he heard an engine whistle. In other words, he was a man \textit{who had heard}, \& was therefore interested in hearing. It is not necessary to explain here how this young man later developed fainting spells, but it is sufficient to note that from his childhood he was very sensitive to sounds. He was very musical, for he could not bear noises, disharmonies or strident tones. We are not surprised, therefore, that he should have been so affected by the sound of a whistle. There are often things in which children or adults are interested because they have suffered through them. The reader will remember the case of the man with asthma mentioned in a previous chapter. He had been bound tightly about his lungs in childhood for some trouble, \& as a result had developed an extraordinary interest in ways to breathe.

One meets person whose whole interest seems to lie in things to eat. Their early remembrances have to do with eating. It seems the most important thing in the world for them -- how to eat, what to eat, \& what not to eat. We will often find that difficulties connected with eating in early life have enhanced the importance of eating for such an individual.

We turn now to a case of remembrance that has to do with movement \& walking. We have seen how many children cannot move very well at the beginning of life because they are weak or suffer from rickets. They become abnormally interested in movement \& always want  to hurry. The case is an illustration of this fact. A man of 50 came to a doctor complaining that whenever he accompanied a person across the street he suffered from a terrible fear that they would both be run over. When alone he was never bothered with this fear, \& in fact was very composed in crossing the street. It was only when another was with him that he wanted to save this person. He would then grasp his companion's arm, push him now right \& now left, \& generally annoy him. We meet with such persons occasionally, though not frequently. Let us analyze the reasons for his stupid actions.

Asked for his old remembrances, he explained that when he was 3 years old he could not move very well \& was suffering from rickets. He was twice run over when crossing a street. \& so, now that he was a man it was important for him to prove that he had overcome this weakness. He wanted to show, so to speak, that he was the only man who could cross a street. He was always looking for an opportunity to prove it whenever he was with a companion. Of course to be able to cross a street safety is not something that most people would take pride in or compete with others. But with such persons as our patient, the desire to move \& to show off about the ability to move can be quite lively.

We turn now to another case -- the case of a boy who was on the way to becoming a criminal. He stole, played ``hookey'' from school, etc. until his parents were in despair. His early remembrances were of how he had always wanted to move around \& to hurry. He was now working with his father \& was sitting still all day. From the nature of the case part of the treatment prescribed was that he be made a salesman -- a traveler for his father's business.

1 of the most significant types of old remembrances is the memory of a death during the period of childhood. When children see a person die suddenly \& abruptly, the effect on their minds is very marked. Sometimes such children become morbid. Sometimes, without becoming morbid, they devote their whole lives to the problem of death \& are always occupied in struggling against illness \& death in some form. We may find many of these children interested in medicine later in life, \& they may become physicians or chemists. Such a goal of course is on the useful side of life. They not only struggle against death but help others to do so. Sometimes, however, the prototype develops a very egotistical point of view. A child who was very much affected by the death of an older sister was asked what he wanted to be. The answer expected was that he would be a physician; instead he replied: ``A grave-digger.'' He was asked why he wanted to follow this occupation, \& he answered, ``Because I want to be the one to bury the others \& not the one buried.'' This goal, we see, is on the useless side of life, for the boy is interested only in himself.

Let us turn now to consider old remembrances of people who were pampered children. The old remembrances mirror the characteristics of this class very clearly. A child of this type often mentions his mother. Now perhaps this is natural but it is a sign that he has had to struggle for a favorable situation. Sometimes the old remembrances seem to be quite innocuous, but they repay analysis. E.g., a man tells you, ``I was sitting in my room \& my mother stood by the cabinet.'' This appears unimportant, but his mentioning his mother is a sign that this has been a matter of interest to him. Sometimes the mother is more hidden \& the study more complicated. We have to guess about the mother. Thus the man in question may tell you, ``I remember I made a trip.'' If you ask who accompanied him, you will discover it was his mother. Or, if children tell us, ``I remember I was in the country at a certain place 1 summer,'' we can presuppose that the father was in the city working \& the mother was with the children. We can ask, ``Who was with you?'' In this way we often see the hidden influence of the mother.

From a study of these remembrances we can see a struggle for preferment. We can see how a child in the course of his development begins to value the pampering his mother gives him. This is important for our understanding because if children or adults tell us about such remembrances, we may be sure that such persons always feels that they are in danger or that another will be preferred to them. We see the tension becoming increased \& more \& more obvious \& we see that their minds are sharply focused on this idea. Such a fact is important: it indicates that in later life such persons will be jealous.

Sometimes persons express interest on 1 point above all others. E.g., a child may say, ``I had to watch my little sister 1 day \& I wanted to protect her very well. I put her at the table but the cover caught \& my little sister fell down.'' This child was only 4 years old. It is of course an early age at which to permit an older child to watch a younger girl. We can see what a tragedy it is in the life of the older child who was doing everything possible to protect the younger one. This particular older girl grew up \& married a kind -- we might almost say, obedient -- husband. But she was always jealous \& critical, always afraid that her husband would prefer another. We can easily understand how the husband tired of her \& turned to the children.

Sometimes tension is more clearly expressed \& people remember that they actually wanted to hurt other members of their family, in fact to kill them. Such persons are people who are interested in their own affairs exclusively. They do not like other people. They feel a certain rivalry towards them. This feeling already exists in the prototype.

We have here the type of person who can never finish anything because he fears someone else will be preferred in friendship \& comradeship, or because he is suspicious of people always trying to surpass him. He can never really becomes a part of society because of the idea that another might outshine him \& be preferred. In every occupation he is extremely tense. This attitude appears specially in connection with love \& marriage.

Even if we cannot completely cure such persons, we can, with a certain art in the study of old remembrances, see that they improve.

1 of the subjects for our methods of treatment was the boy whom we described in another chapter as having gone to market with his mother \& younger brother 1 day. When it started to rain the mother took him up in her arms, but, on noticing the younger brother, she set him down \& took up the younger child. Hence he felt that the younger brother was preferred.

If we can obtain such old remembrances we can predict, as we have said, what will happen later in the life of our patients. However, it must be remembered that old remembrances are not reasons, they are hints. They are signs of what happened \& how development took place. They indicate the movement toward a goal \& what obstacles had to be overcome. They show how a person becomes more interested in 1 side of life than another. We see that he may have what we call a trauma, along the lines of sex, e.g.; i.e., he may be more interested in such matters than in others. We cannot be surprised if, when we ask for old remembrances, we hear some sex experiences. Some persons are interested in sex features more than in others at an early age. It is part of the usual human behavior to be interested in sex but, as I have said before, there are many varieties \& degrees of interest. We often find that in a case where a person tells us about sex remembrances, he later develops in this direction. The resulting life is not harmonious because his 1 side of human life is over-valued. There are persons who insist that everything has a sex bias. On the other hand, there are others who insist that the stomach is the most important organ \& we will find that old remembrances parallel later characteristics in such instances also.

There was a boy whose getting into high school was always a riddle. He wanted to be constantly moving, \& would never settle down to study. He was always thinking about something else, frequenting coffee houses \& visiting at friends' houses -- all when he should have been studying. It was therefore interesting to examine his old remembrances. He said, ``I can remember lying in my cradle \& looking at the wall. I noticed the paper on the wall, with all its flowers, figures, etc.'' This person was prepared only for lying in a cradle, not for taking examinations. He could not concentrate on his studies because he was always thinking of other things \& trying to go after 2 hares at once, which cannot be done. We can see that this man was a pampered child \& could not work alone.

We come now to the hated child. This type is rare \& represents extreme cases. If a child is really hated from the beginning of life, he cannot live. Such a child would perish. Usually children have parents or a nurse who \textit{pampers} them to some extent \& satisfies their desires. We find the hated children among illegitimate, criminal \& not wanted children, \& we often see these children becoming depressed. Frequently we find in their remembrances this feeling of being hated. E.g., there was the case of a man who said, ``I remember I was spanked; my mother scolded me, criticized me until I ran away.'' While running away he came very nearly being drowned.

This man came to a psychologist because he could not leave his home. We see from his old remembrances that he went out once \& met with great danger. This stuck in his memory \& he constantly looked for danger when he went out. He was a bright child but always feared that he might not make 1st place in examinations. So he hesitated \& could not go on. When he at last got to the university he feared that he could not compete in the prescribed way. We see how all this may be traced back to his old remembrances of danger.

Another case which may be taken as an illustration is that of an orphan whose parents died when he was only about a year old. He had rickets, \& being in an asylum, he was not cared for properly. Nobody looked after him, \& in later life it was very difficult for him to make friends or comrades. Looking back to his remembrances we see that he always felt that others were preferred. This feeling played an important part in his development. He always felt hated \& this hindered his approach to all problems. He was excluded from all questions \& situations of life, such as love, marriage, friendship, business -- all these situations which required contact with his fellows -- on account of his feeling of inferiority.

Another interesting case is that of a middle-aged man who was always complaining of sleeplessness. He was 46 or 48 years old, married, \& had children. He was very critical of everybody, \& was always trying to tyrannize, particularly over the members of his family. His actions made everyone feel miserable.

When asked for his old remembrances he explained that he had grown up in a home with quarrelsome parents, who were always fighting \& threatening each other, so that he was afraid of them both. He went to school dirty \& uncared for. 1 day his usual teacher was absent \& a substitute took her place. This substitute woman was interested in her task \& its possibilities. She saw that it was a good \& noble work. She saw possibilities in this ill-kept boy \& went out to encourage him. This was the 1st time in his life he had had any such treatment. From that time one he began to develop, but it was always as if he were pushed from behind. He did not really believe he was able to be superior, \& so he worked all day \& half the night. In this way he grew up trained to use half the night for his work or else not to sleep at all but to spend the time thinking of what he had to do. As a result he grew to think that it was necessary to be awake almost all night in order to accomplish results.

We see later his desire to be superior expressed in his attitude towards his family \& in his behavior towards others. His family being weaker than he, he could appear in the role of a conqueror before them. His wife \& children suffered through this type of behavior, as was inevitable.

Summing up the character of this man as a whole, we may say that he had a goal of superiority \& that it was the goal of a person with a great feeling of inferiority. This we often find among over-strained persons. Their tenseness is a sign of their doubt of their own success, \& their doubt in turn is covered up by a superiority complex which is really a superiority pose. A study of old remembrances reveals the situation in its true light.'' -- \cite[pp. 117--134]{Adler2013}

%------------------------------------------------------------------------------%

\section{Attitudes \& Movements}
``In the last chapter we endeavored to describe the manner in which old remembrances \& fancies may be used to illuminate the hidden style of life of an individual. Now the study of old remembrances is only 1 device of a whole class of devices for the study of personality. They all depend on the principle of using isolated parts for an interpretation of the whole. Besides old remembrances we can observe movements \& attitudes. The movements themselves are \textit{expressed} or imbedded in attitudes, \& the attitudes are an expression of that whole attitude to life which constitutes what we call the style of life.

Let us 1st speak about the movements of the body. Everybody knows that we judge a person by his manner of standing, walking, moving, expressing himself, etc. We do not always consciously judge, but there is always a feeling of sympathy or antipathy created by these impressions.

Let us consider attitudes in standing, e.g. We notice promptly whether a child for adult stands upright or whether he is crooked or bent. This is not very difficult. We have to watch specially for exaggerations. A person who stands too straight, in a stretched position, causes us to suspect that he is using too much power to assume this posture. We can suppose that this person feels much less great than he wants to appear. In this little point we can see how he mirrors what we have called the superiority complex. He wants to appear more courageous -- he wants to express himself more as he would be if he were not so tense.

On the other hand we see persons with just the opposite posture -- persons who appear bent \& who are always stooping. Such a posture implies to a certain extent that they are cowards. But it is a rule of our art \& science that we should always be cautious, looking for other points \& never judging solely by 1 consideration. Sometimes we feel that we are almost sure of being correct, but we still want to verify our judgment by other points. We ask, ``Are we right in insisting that persons who stoop are always cowards? What can we expcet of them in a difficult situation?''

To look at another point in this connection, we will notice how such a person always tries to rest upon something, to lean on a table or chair e.g. He does not trust his own power but wants to be supported. This reflects the same attitude of mind as when standing crooked, \& so when we find both types of action present our judgment is somewhat confirmed.

We will find that children who want always to be supported have not the same posture as independent children. We can tell the degree of independence by how a child stands, how he approaches other persons. In such cases we need not be in doubt, for we have many possibilities of confirming our conclusion. \& once we have confirmed our conclusion, we can take steps to remedy the situation \& put the child on the right path.

Thus we may experiment with such a child who wants to be supported. Sit his mother on a chair \& then let the child come into the room. We will find that he does not look at any other person but goes directly towards his mother \& leans on the chair or against his mother. This confirms what we expect -- that the child wants to be supported.

It is interesting also to note that child's approach, for it shows the degree of social interest \& adjustment. It expresses the confidence of the child in others. We will find that a person who does not want to approach others \& who always stands far away is also reserved in other respects. We will find that he does not speak enough \& is unusually silent.

We can see how all these things point the same way because every human being is a unity \& reacts as such towards the questions of life. As an illustration let us take the case of a woman who came to a doctor for treatment. The doctor expected that she would take a seat near him, but when she was offered a chair she looked around \& took a seat far away. It could only be concluded that this was a person who wanted to be connected with only 1 person. She said that she was married, \& from this the whole story could be guessed. It could be guessed that she wanted to be connected only with her husband. It could also be guessed that she wanted to be pampered, that she is the sort of person who would demand that her husband be very exact \& always on time in coming home. If she was alone she would suffer great anxiety, \& she would never want to go out of her house alone \& would not enjoy meeting other people. In short from her 1 physical movement we could guess the whole story. But we have also ways of confirming our theory.

She may tell us: ``I am suffering from anxiety.'' Now nobody would understand what this meant unless he knew that anxiety can be used as a weapon to rule another person. If a person or adult suffers from anxiety we can guess that there is another person who supports this child or adult.

There was once a couple who insisted that they were free thinkers. Such people believe that everybody can do what he wants in marriage, so long as each one tells the other what happens. The consequence was that the husband had some love affairs \& told all of them to his wife. She seemed perfectly content. But later on she began to suffer from anxiety. She would not go out alone. Her husband must always go with her. We can see then how this free thinking became modified by anxiety or phobia.

Some persons will always stay near a wall of a house \& lean on it. This is a sign that they are not courageous enough, not independent enough. Let us analyze the prototype of such a timid \& hesitating person. There was a boy who came to school appearing very shy. This is an important sign that he does not want to be connected with others. He had no friends \& was always waiting for school to close. He moved very slowly, \& would go down the stairs close to the wall, look down the street \& rush for his house. He was not a good pupil in school, \& in fact was very poor in his school work since he did not feel happy inside of school walls. He always wanted to go home to his mother, a widow who was weak \& pampered him very much.

In order to understand more about the case the doctor went to talk with his mother. He asked her, ``Does he want to go to bed?'' She said, ``Yes.'' ``Does he cry out at night?'' ``No.'' ``Does he wet the bed?'' ``No.''

The doctor thought that either he had made a mistake or that the boy had made a mistake. Then he concluded that the boy must sleep in bed with his mother. How was this conclusion arrived at? Well, to cry out at night is to demand attention of the mother. If he slept in her bed, this would not be necessary. Similarly to wet the bed is also to demand the mother's attention. The doctor's conclusion was verified: the boy slept in bed with his mother.

If we look carefully we will see that all the little things to which the psychologist pays attention form part of a consistent plan of life. Hence when we can see the goal -- in the child's case, to be always tied up with his mother -- we can conclude a great many things. We can conclude by this means whether a child is feeble-minded or not. A feeble-minded child would not be able to establish such an intelligent plan of life.

Now let us turn to the mental attitudes distinguishable in persons. Some persons are more or less pugnacious. Some on the other hand want to give up the ship. However, we never see a person who really gives up. It is not possible, for it is beyond human nature. The normal being cannot give up. If he seems to do so, it indicates even more of a struggle to carry on than otherwise.

There is a type of child who always wants to give up. He is usually the center of attention in a family. Everybody has to care for him, push him forward \& admonish him. He must be supported in life \& is always a burden to others. This is his goal of superiority -- he expresses his desire to dominate others in this fashion. Such a goal of superiority is of course the result of an inferiority complex, as we have already shown. If he had not been doubtful of his own powers, he would not take this easy way out for attaining success.

There was a boy of 17 who illustrated this trait. He was the oldest in the family. We have already seen how the oldest child usually experiences a tragedy when the coming of another child dethrones him from his place in the center of family affections. This was the case with this boy. He was very depressed \& peevish \& had no occupation. 1 day he tried to commit suicide. Soon after that he came to a doctor \& explained that he had had a dream before his attempt at suicide. He dreamt he had shot his father. We see how such a person -- depressed, lazy \& not moving -- has all the time the possibility of movement present in his mind. We also see how all these children who are indolent in school, \& all these indolent adults who seem incapable of doing anything may be on the brink of danger. Oftentimes this indolence is only on the surface. Then something happens, \& we have an attempt at suicide, or else a neurotic condition or insanity may appear. To ascertain the mental attitude of such persons is sometimes a difficult scientific task.

Shyness in a child is another thing that is full of danger. A shy child must be carefully treated. The shyness must be corrected or it will ruin his whole life. He will always have great difficulties unless his shyness is corrected, for in our culture things are so established that only courageous persons get good results \& the advantages of life. If a person is courageous \& suffers defeat he is not hurt so much, but a shy person makes his escape to the useless side of life as soon as he sees difficulties ahead. Such children will become neurotics or insane in later life.

We see such persons going about with a hangdog air, \& when they are with others they stammer \& will not speak or they will avoid people altogether.

The characteristics that we have been describing are mental attitudes. They are not inborn or inherited, but are simply reactions toward a situation. A given characteristic is the answer that my style of life gives to my apperception of a problem that confronts me. Of course it is not always the logical answer that the philosopher would expect. It is the answer that my childhood experiences \& mistakes have trained me to make.

We can see the functioning of these attitudes as well as the way in which they have been built up in children or in abnormal persons better than we can in the case of normal adults. The prototype stage of the style of life, as we have seen, is much clearer \& simpler than the later style. In fact one may compare the functioning of the prototype to an unripe fruit that will assimilate everything that comes along -- manure, water, food, air. All these things will be taken up in its development. The difference between a prototype \& the style of life is like the difference between an unripe \& a ripe fruit. The unripe fruit stage in human beings is much easier to open up \& examine, but what it reveals is to a large extent valid for the ripe fruit stage.

We can see, e.g., how a child who is a coward at the beginning of life expresses this cowardice in all his attitudes. A world of differences separate the cowardly child from the aggressive, fighting child. The fighting child always has a certain degree of courage which is the natural outgrowth of what we have called common sense. Sometimes, however, a very cowardly child may appear like a hero in a certain situation. This happens whenever he is deliberately trying to attain 1st place. This is clearly illustrated in the case of a boy who did not know how to swim. 1 day he went swimming with other boys who had asked him to join them. The water was very deep, \& the boy, who could not swim, nearly drowned. This of course is not real courage, \& is all on the useless side of life. The boy merely did what he did because he wanted to be admired. He ignored the danger he was in, \& hoped that the others would save him.

The question of courage \& timidity is psychologically closely related to the belief in predestination. The belief in predestination affects our capacity for useful action. There are persons who have such a feeling of superiority that they feel they can accomplish anything. They know everything \& do not want to learn anything. We all know the result of such ideas. Children who feel this way in school usually get poor marks. There are other people who always want to try the most dangerous things: they feel that nothing can happen to them, that they cannot suffer defeat. Very often the result is a bad one.

We find this feeling of predestination among people whenever something terrible has happened in their lives \& they have remained unhurt. E.g., they may have been present in a serious accident \& were not killed. As a result they feel that they are destined for higher purposes. There was once a man who had such a feeling but after going through an experience which resulted differently from his expectation he lost courage \& became depressed \& melancholy. His most important support had fallen away.

When asked for his early remembrances he related a very significant experience. He said he was once about to go to a theater in Vienna, but had to attend to something 1st. When he finally arrived at the theater it had burned down. Everything was over, but he was saved. One can well understand how such a person felt himself destined for higher things. All went well until he suffered defeat in his relations with his wife. Then he broke down.

Much could be said \& written about the significance of the belief in fatalism. It affects whole peoples \& civilizations as well as individuals, but for our part we desire to point out only its connection with the sprints of psychological activity \& the style of life. The belief in predestination is in many ways a cowardly escape from the task of striving \& building up activity along the useful line. For that reason it will prove a false support.

1 of the basic attitudes of mind that affects our relations with our fellow-men is the attitude of envy. Now to be envious is a sign of inferiority. True, we all have a certain amount of envy in our make-up. A small amount does no harm \& is quite common. We must, however, demand that envy be useful. It must result in work, in a going in, \& in a facing of problems. In such cases it is not useless. For that reason we should pardon the bit of envy which is found in all of us.

On the other hand jealousy is a much more difficult \& dangerous mental attitude, because it cannot be made useful. There is no single way in which a jealous person can be useful.

Moreover, we see in jealousy the result of a great \& deep feeling of inferiority. A jealous person is afraid of his inability to hold his or her partner. \& so at the very moment when he wants to influence his partner in some manner, he betrays his weakness by his expressions of jealousy. If we look in the prototype of such a person we shall see a sense of curtailment. In fact whenever we meet with jealous persons it is well to look back into their past \& see whether we have not to do with a dethroned person who expects that he will be dethroned again.

From the general problem of envy \& jealousy we may pass to the consideration of a very peculiar type of envy -- the envy on the part of the female sex of the superior social position of the male sex. We find many women \& girls who want to be boys. This attitude is quite understandable, for if we look at things impartially we can see that in our culture the men are always in the lead; they are always more appreciated, valued \& esteemed than women. Morally this is not right \& ought to be corrected. Now girls see that in the family the men \& boys are much more comfortable \& do not have to bother with little things. They see that they are freer in many ways, \& this superior freedom of the male sex makes them dissatisfied with their own role. They therefore try to act like boys. This imitation of boys may appear in various ways. We see them, e.g., trying to dress like boys, \& in this they are sometimes supported by their parents since boys' clothes are admittedly more comfortable. Now a number of these acts are useful \& need not be discouraged. But there are some useless attitudes, as when a girl wants to be called by a boy's name \& not by the name of a girl. Such girls get very angry if others do not call them by the boy's name which they have chosen. This attitude is very dangerous if it reflects something below the surface \& is not a mere prank. In such a case it may appear later in life as a dissatisfaction with the sex role \& a distaste for marriage -- or, when married, a distaste for the sex role of woman.

One should not find fault with women for wearing short clothes, because it is an advantage. It is also fitting for them to develop like men in many ways, \& to have a job like men. But it is dangerous for them to be dissatisfied with their feminine role \& try to adopt the vices of men.

This dangerous tendency makes its appearance in the adolescent period, for it is then that the prototype becomes poisoned. The immature minds of the girls become jealous of the privileges of the boys. It reacts in the desire to imitate boys. Now this is a superiority complex -- it is an escape from proper development.

As we have said, this can lead to a great disinclination for love \& marriage. This is not to say that girls who have this disinclination do not want to be married, for in our culture not to be married is taken as a sign of defeat. Even the girls who are not interested in marriage want to get married.

One who believes in regulating the basis of the relations of the sexes on the principle of equality should not encourage this ``masculine protest'' of women. The equality of the sexes must be fitted into the natural scheme of things, while the masculine protest is a blind revolt against reality \& is thus a superiority complex. As a matter of fact through this masculine protest all the sex functions can be disturbed \& affected. Many serious symptoms can be produced, \& if we trace them back we shall see that the conditions started in childhood.

Not so frequently as in the case of girls who want to be boys, we also meet with the boy who wants to be like a girl. He wants to imitate not the ordinary girl, but the type of girl who flirts in an exaggerated manner. Such boys use face powder, they wear flowers, \& try to act in the manner of a frivolous girl. This is also a form of superiority complex.

We find in fact that in many such cases the boy had grown up in an environment in which a woman was at the head. Thus the boy grew up to imitate the traits of the mother, not of the father.

There was a boy who came for consultation because o certain sex troubles. He related how he was always with his mother. The father was almost a nonentity in the home. Now his mother had been a dressmaker before she was married \& continued something of her occupation after her marriage. The boy being always near her got to be interested in the things she made. He began to sew \& draw pictures of dresses for women, etc. One can judge how interested he was in his mother from the fact  that at 4 years he had learned to tell time because his mother always went out at 4 \& came back at 5 o'clock. Impelled by his pleasure on seeing her return, he learned to read the clock.

Later in life, when he went to school, he acted like a girl. He took no part in sports or games. The boys made fun of him, \& at times they even kissed him, as they frequently do in such cases. 1 day they had to give a theatrical play, \& as we can imagine this boy had the part of a girl. He acted it so well that many in the audience actually thought he was a girl. 1 man in the audience even fell in love with him. In this way this boy got to see that even if he could not be much appreciated as a man he could be greatly appreciated as a woman. This was the genesis of his later sexual troubles.'' -- \cite[pp. 135--153]{Adler2013}

%------------------------------------------------------------------------------%

\section{Dreams \& Their Interpretation}
``For Individual Psychology consciousness \& unconsciousness form a single unity, as we have already explained in a number of contexts. In the last 2 chapters we have been interpreting conscious parts -- remembrances, attitudes, movements -- in terms of the individual whole. We shall now apply the same method of interpretation to our unconscious or semiconscious life -- the life of our dreams. The justification for this method is that our dream lief is just as much a part of the whole, as our waking life -- no more \& no less. Followers of other schools of psychology are constantly trying to find new views concerning dreams, but our understanding of dreams has been developed along the same line as our understanding of all the integral parts manifested in the expressions \& movements of the psyche.

Now just as our waking life, we have seen, is determined by the goal of superiority, so we may see that dreams are determined by the individual goal of superiority. A dream is always a part of the style of life \& we always find the prototype involved in it. In fact it is only when you see how the prototype is bound up to a particular dream that you can be sure that you have really understood the dream. Also, if you know a person well, you can pretty nearly guess the character of his dreams.

Take, e.g., our knowledge that mankind as a whole is really cowardly. From this general fact we can presuppose that the largest number of dreams will be dreams of fear, danger, or anxiety. \& so if we know a person \& see that his goal is to escape the solution of life's problems, we can guess that he often dreams that he falls down. Such a dream is like a warning to him: ``Do not go on -- you will be defeated.'' He expresses his view of the future in this way -- by falling. The large majority of men have these dreams of falling.

A specific case is a student on the eve of an examination -- a student whom we know to be a quitter. We can guess what will happen with him. He is worried the whole day, cannot concentrate, \& finally says to himself, ``The time is too short.'' He wants to postpone the examination. His dream will be 1 of falling down. \& this expresses his style of life, for to attain his goal, he must dream in such a way.

Take another student who makes progress in his studies is courageous \& not afraid, \& never uses subterfuges. We can also guess his dreams. Before an examination he will dream that he climbs a high mountain, is enchanted with the view from the mountain top, \& in this way awakes. This is an expression of his current of life, \& we can see how it reflects his goal of accomplishment.

Then there is the person who is limited -- the person who can proceed only up to a certain point. Such a person dreams about limits, \& about being unable to escape persons \& difficulties. He often has dreams of being chased \& hunted.

Before we go on to the next type of dream it may be well to remark that the psychologist is never discouraged if somebody says to him, ``I will not tell you any dreams for I cannot remember them. But I will make up some dreams.'' The psychologist knows that his fancy cannot create anything other than that which his style of life commands. His made-up dreams are just as good as his genuinely remembered dreams, for his imagination \& fancy will also be an expression of his style of life.

Fancy need not literally copy a man's real movements in order to be an expression of his style of life. We find, e.g., the type of person who lives more in fancies than in reality. He is the type that is very cowardly in the daytime but quite courageous in dreams. But we will always find some manifestations which indicate that he does not want to finish his work. Such manifestations will be quite evident even in his courageous dreams.

It is always the purpose of a dream to pave the way towards the goal of superiority -- that is to say, the individual's private goal of superiority. All the symptoms, movements \& dreams of a person are a form of training to enable one to find this dominating goal -- to be goal 1 of being the center of attention, of domineering, or of escape.

The purpose of a dream is neither logically or truthfully expressed. It exists in order to create a certain feeling, mood or emotion, \& it is impossible fully to unravel its obscurities. But in this it differs from waking life \& the movements of waking life only in degree, not in kind. We have seen that the answers of the psyche to life's problems are relative to the individual scheme of life: they do not fit into a pre-established frame of logic, although it is our aim, for purposes of social intercourse, to make them do so more \& more. Now once we give up the absolute point of view for waking life, dream life loses its mystery. It becomes a further expression of the same relativity \& the same mixture of fact \& emotion that we find in waking life.

Historically dreams have always appeared very mysterious to primitive peoples, \& they have generally resorted to the prophetic interpretation. Dreams were regarded as prophecies of events to come. In this there was a half-truth. It is true that a dream is a bridge that connects the problem which confronts the dreamer with his goal of attainment. In this way a dream will often come true, because the dreamer will be training his part during the dream \& will be thus preparing for it to come true.

Another way of saying the same thing is that there is the same interconnectedness revealed in dreams as in our waking life. If a person is keen \& intelligent he can foresee the future whether he analyzes his waking life or his dream life. What he does is to diagnose. E.g. if somebody dreams that an acquaintance has died \& the person does die, this might be no more than what a physician or a close relative could foresee. What a dreamer does is to think in his sleep rather than in waking life.

The prophetic view of dreams, precisely because it contains a certain half-truth, is a superstition. It is generally clung to by persons who believe in other superstitions. Or else it is championed by men who seek importance by giving the impression that they are prophets.

To dispel the prophetic superstition \& the mystery that surrounds dreams we have to explain of course why most people do not understand of their own dreams. The explanation is to be found in the fact that few people know themselves even in waking life. Few persons have the power of reflective self-analysis which permits them to see whither they are headed, \& the analysis of dreams is, as we have said, a more complicated \& obscure affair than the analysis of waking behavior. It is thus no wonder that the analysis of dreams should be beyond the scope of most persons -- \& it is also no wonder that in their ignorance of what is involved they should turn to charlatans.

It will help us to understand the logic of dreams if we compare it, not directly with the movements of normal waking life, but with the type of phenomena which we have described in previous chapters as a manifestation of private intelligence. The reader will remember how we described the attitudes of criminals, problem children \& neurotics -- how they create a certain feeling, temper or mood in order to convince themselves of a given fact. Thus the murderer justifies himself by saying, ``Life has no place for this man; therefore I must kill him.'' By emphasizing in his own mind the view that there is not sufficient place on earth he creates a certain feeling which prepares him for the murder.

Such a person may also reason that so-\&-so has nice trousers \& he has not. He puts such value on this circumstance that he becomes envious. His goal of superiority becomes to have nice trousers, \& so we may find him dreaming a dream which creates a certain emotion which will lead to the accomplishment of that goal. We see this illustrated, in fact, in well-known dreams. There are, e.g., the dreams of Joseph in the Bible. He dreamt that all the others bent before him. Now we can see how this dream fitted in with the whole episode of the coat of many colors -- \& with his banishment by his brothers.

Another well-known dream is that of the Greek poet Simonides, who was invited to go to Asia Minor to lecture. He hesitated \& continually postponed the trip in spite of the fact that the ship was in the harbor waiting for him. His friends tried to make him go, but to no avail. Then he had a dream. He dreamt that a dead man whom he had once found in a forest appeared to him \& said, ``Because you were sopious \& cared for me in the forest, I now warn you not to go to Asia Minor.'' Simonides arose \& said, ``I will not go.'' But he had already been inclined not to go before he ever had the dream. He had simply created a certain feeling or emotion to back up a conclusion that he had already reached, although he did not understand his own dream.

If one understands it is clear that one creates a certain fantasy for purposes of self-deception, which results in a desired feeling or emotion. Frequently this is all that is remembered of the dream.

In considering this dream of Simonides we come to another point. What should be the procedure in interpreting dreams. 1stly, we must bear in mind that a dream is part of a person's creative power. Simonides, dreaming, used his fancy \& built up a sequence. He selected the incident of the dead man. Why should this poet pick the experience of the dead man from out of all his experiences? Obviously because he was very much concerned with ideas of death, due to the fact that he was terrified at the thought of sailing on a ship. In those days a sea voyage presented real danger, \& so he hesitated. It is a sign that he was probably not only afraid of seasickness but also that he feared the ship might sink. As a result of this preoccupation with the thought of death, his dream selected the episode of the dead man.

If we consider dreams in this manner, the task of interpretation does not become too difficult. We should remember that the selection of pictures, remembrances \& fancies is an indication of the direction in which the mind is moving. It shows you the dreamer's tendency, \& eventually we can see the goal at which he wants to arrive.

Let us consider, e.g., the dream of a certain married man. He was not content with his family life. He had 2 children, but was always worried, thinking that his wife did not take care of them \& was too much interested in other things. He was always criticizing his wife about these things \& tried to reform her. 1 night he dreamt that he had a 3rd child. This child got lost \& was not to be found. He reproached his wife because she had not taken care of him.

Here we see his tendency: he had in mind the thought that 1 of his 2 children might get lost, but he was not courageous enough to make it 1 of them in his dream. \& so he invented a 3rd child \& made him get lost.

Another point to be observed is that he liked his children \& did not want them to get lost. Also that he felt that his wife was overburdened with 2 children \& could not care for 3. This 3rd child would perish. Hence we find another aspect of the dream, which, when interpreted, reads: ``Should I have a 3rd child or not?''

The real result of the dream was that he had created an emotion against his wife. No child really got lost, but he got up in the morning criticizing \& feeling antagonistic towards her. Thus people frequently get up in the morning -- argumentative \& critical as a result of an emotion created by the night's dream. It is like a state of intoxication \& not unlike what one finds in melancholia, where the patient intoxicates himself with ideas of defeat, of death \& of all being lost.

We may also see that this man selected things in which he was sure to be superior, as, e.g., the feeling, ``I am careful of the children, but my wife is not \& therefore one got lost.'' Thus his tendency to dominate is revealed in his dream.

The modern interpretation of dreams is about 25 years old. Dreams were 1st regarded by Freud as the fulfillment of infantile sex desires. We cannot agree with this, inasmuch as if dreams are such a fulfillment then everything can be expressed in terms of a fulfillment. Every idea behaves in this way -- going from the depths of the subconscious up into consciousness. The formula of sex-fulfillment thus explains nothing in particular.

Later Freud suggested that the desire for death was involved. But it is certain that this last dream could not be explained very well in this way, for we cannot say that the father wanted the child to get lost \& die.

The truth is that there is no specific formula which will explain dreams, except the general postulates which we have discussed about the unity of psychical life \& about the special affective character of dream life. This affective character, \& its accompaniment of self-deception is a theme with many variations. Thus it is expressed in the preoccupation with comparisons \& metaphors. The use of comparisons is 1 of the best means of deceiving oneself \& others. For we may be sure that if a person uses comparisons he does not feel sure that he can convince you with reality \& logic. He always wants to influence you by means of useless \& far-fetched comparisons.

Even poets deceive, but pleasantly, \& we enjoy being entertained by their metaphors \& poetic comparisons. We may be sure, however, that they are meant to influence us more than we would be influenced by usual words. If Homer, e.g., speaks of an army of Greek soldiers overrunning a field like lions, the metaphor will not deceive us when we think sharply but it will certainly intoxicate us when we are in a poetic mood. The author makes us believe he has marvelous power. He could not do this if he were merely to describe the clothes the soldiers wore \& the arms they carried, etc.

We see the same thing in the case of a person who is in difficulty about explaining things: if he sees he cannot convince you, he will use comparisons. This use of comparisons, as we have said, is self-deceptive, \& this is the reason it is so prominently manifested in dreams in the selection of pictures, images, etc. This is an artistic way of intoxicating oneself.

The fact the dreams are emotionally intoxicating offers, curiously enough, a method for preventing dreams. If a person understands what he has been dreaming about \& realizes that he has been intoxicating himself, he will stop dreaming. To dream will have no more purpose for him. At least this is the case with the present writer, who stopped dreaming as soon as he realized what dreaming meant.

Incidentally it may be said that this realization, to be effective, must have the aspects of a thorough-going emotional conversion. This was brought about, in the case of the writer, by his last dream. The dream occurred during war time. In connection with his duties he was making a great effort to keep a certain man from being sent to the front in a place of danger. In the dream the idea came to him that he had murdered someone, but he did not know whom. He got himself into a bad state wondering, ``Whom have I murdered?'' The fact is he was simply intoxicated with the idea of making the greatest possible effort to put the soldier in the most favorable position for avoiding death. The dream emotion was meant to be conducive to this idea, but when he understood the subterfuge of the dream, he gave up dreaming altogether, since he did not need to deceive himself in order to do the things that for reasons of logic he might want either to do or to leave undone.

What we have said may be taken as an answer to the question that is frequently asked, ``Why do some persons never dream?'' These are persons who do not want to deceive themselves. They are too much tied up with movement \& logic, \& want to face problems. Persons of this sort, if they dream, often forget their dreams very soon. They forget so quickly that they believe they have not dreamed.

This brings up the theory that we always dream \& that we forget most of our dreams. If we accepted such a theory it would put a different construction on the fact that some persons never dream: they would then become persons who dream but who always forget their dreams. The present writer does not accept this theory. He rather believes that there are persons who never dream \& that there are also dreamers who sometimes forget their dreams. In the nature of the case such a theory is hard to refute, but perhaps the burden of proof should be put on the propounders of the theory.

Why do we have the same dream repeatedly? This is a curious fact for which no definite explanation can be given. However, in such repeated dreams we are able to find the style of life expressed with much more clarity. Such a repeated dream gives us a definite \& unmistakable indication where the individual goal of superiority lies.

In the case of long \& extended dreams we must believe that the dreamer is not fully ready. He is looking for the bridge from the problem to the attainment of the goal. For this reason the dreams which can be best understood are short dreams. Sometimes a dream consists of only 1 picture, a few words, \& it shows how the dreamer is really trying to find a short way to deceive himself.

We may close our discussion with the question of sleep. A great many persons put to themselves needless questions about sleep. They imagine that sleep is the contradiction of being awake, \& that it is the ``brother of death.'' But such views are erroneous. Sleep is not a contradiction of being awake, but is rather a degree of being awake. We are not separated from life in sleep. On the contrary we are thinking, \& hearing in sleep. The same tendencies are generally expressed in sleep as in waking life. Thus there are mothers who cannot be awakened by any of the street noises, but if the children move in the least bit they immediately jump up. We see how their interest is really awake. Also from the fact that we do not fall out of bed we can see that we realize limits in sleep.

The whole personality is expressed by night \& by day. This explains the phenomena of hypnotism. What superstition has made to appear as a magic power is for the most nothing more than a variety of sleep. But it is a variety in which 1 person wants to obey another \& knows that the 2nd person wants to make him sleep. A simple form of the same thing is when parents say, ``It is enough -- now sleep!'' \& the children obey. In hypnotism, too, the results take place because the person is obedient. \& in proportion to his obedience is the ease with which he may become hypnotized.

In hypnotism we have an opportunity of making a person create pictures, ideas, remembrances which he would not do with his waking inhibitions. The only requirement is obedience. By this method we can find some solutions -- some old remembrances -- which may have been forgotten before.

As a method of treatment \& cure, hypnotism has its dangers, however. The present writer does not like hypnotism \& uses it only when a patient trusts no other method. One will find that hypnotized persons are rather revengeful. In the beginning they overcome their difficulties, but they do not really change their style of life. It is like a drug or a mechanical means: the person's true nature has not been touched. What we have to do is to give a person courage, self-confidence \& better understanding of his mistakes, if we are really to help him. Hypnotism does not do this, \& should not be used except in rare cases.'' -- \cite[pp. 154--172]{Adler2013}

%------------------------------------------------------------------------------%

\section{Problem Children \& Their Education}

%------------------------------------------------------------------------------%

\section{Social Problems \& Social Adjustment}

%------------------------------------------------------------------------------%

\section{Social Feeling, Common Sense \& the Inferiority Complex}

%------------------------------------------------------------------------------%

\section{Love \& Marriage}

%------------------------------------------------------------------------------%

\section{Sexuality \& Sex Problems}

%------------------------------------------------------------------------------%

\section{Conclusion}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
	
\end{document}