\documentclass{article}
\usepackage[backend=biber,natbib=true,style=alphabetic,maxbibnames=50]{biblatex}
\addbibresource{/home/nqbh/reference/bib.bib}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=1cm,right=1cm,top=5mm,bottom=5mm,footskip=4mm]{geometry}
\def\labelitemii{$\circ$}

\title{Character Disturbance: The Phenomenon of Our Age}
\author{George K. Simon Jr.}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\begin{quotation}
	``\textit{For Sherry}, whose heart is truer than any I know, \& who inspires me daily to be a better person. \& for all those silent but committed souls of noble character, upon whom the very survival of freedom depends.'' -- \cite{Simon2011}
\end{quotation}

%------------------------------------------------------------------------------%

\section*{Preface}
``Imagine you recently read a newspaper article about a young girl who suddenly \& inexplicably lost her eyesight. Her frantic parents took her from doctor to doctor, specialist to specialist, \& clinic to clinic, yet no one could find the reason for her blindness. In desperation, 1 day they decided to take her to a psychologist. After months of traditional psychoanalysis, the therapist revealed that the child's blindness resulted from severe emotional trauma. It seems that several months before, while riding on the school bus, this young lady just happened to glance at a boy seated with some friends across the aisle. She thought to herself: ``This guy is really cute.'' Before long, she also began thinking things like: ``I wonder what it would be like to kiss him.'' But almost immediately after having these thoughts, she started to feel badly. She began fretting about what kind of horrible person she must be to entertain such ``impure'' thoughts. She worried that they could only lead to other impure urges \& temptations, \& even perhaps some impure action on her part. Eventually, she became consumed with guilt \& shame. She remembered times in the past when she had looked at boys, \& how hard it was to resist impure thoughts \& urges. Surely worse would follow, she feared, if she didn't keep herself in check. Shortly after this incident, she lost her vision.

When she 1st saw the psychologist, this troubled little girl didn't even remember the bus incident. She certainly didn't remember what she was thinking or feeling at the time. She'd even forgotten how deeply the incident unnerved her, \& how she dealt with her anxiety over the situation. The psychologist helped her see that she had ``repeated'' her memoires as a way of easing the intensity of her emotional pain. Her lengthy analysis eventually helped her not only recover her memory of that fateful day's events, but also enabled her to reconnect with her conflicted emotions. She came to realize this: She was so deeply distressed by what she thought were her unforgivable, impure desires that she actually believed it was better not to see at all than risk having such thoughts about boys again.

Once she had confessed her sins to the doctor \& he did not condemn her, the young girl slowly began to feel better. She took heart in the notion that he appeared to accept her just as she was. She slowly began to feel that she wasn't such a horrible person after all merely for the kinds of thoughts she sometimes had about boys. In time, she came to believe that the level of her fear, guilt, \& shame was excessive \& unwarranted, given the nature of the situation. No longer believing that simply having thoughts about kissing boys was as evil as she once did, she allowed herself to see again.

Now, I would pose to you, the reader, the same question I ask of professionals \& non-professionals alike at every 1 of the hundreds of workshops I've given over the past 25 years. How many scenarios similar to the one I described have you read or heard about in the last year? How about the past 5 years? How about the past, 10--20--30 years? It should come as no surprise that the answer I get is always the same: \textbf{\textit{zero}}. What should shock you, however (\& always gets the attention of my audience), is this: \textit{Almost all of the principles of classical-psychology paradigms stemmed from various theorists' attempts to explain this \& similar phenomena} (sometimes referred to by adherents to these theories as ``hysterical'' blindness).

You see, in Sigmund Freud's day, \textit{some} individuals actually suffered from such strange maladies. It's important to note, however, that these extreme psychological illnesses were \textit{never} widespread phenomena. But in the intensely socially repressive Victorian era, cases appeared of persons experiencing extraordinary levels of guilt or shame merely for being tempted to act on a primal human instinct, \& displaying pathological symptoms as a result. If there were a motto or saying that might best describe the ``zeitgeist'' (i.e. social or cultural milieu) of that time it would be: ``Don't even \textit{think} about it!'' So, some individuals were quite unnecessarily consumed with excessive guilt \& shame about their most basic human urges. Freud treated some of these individuals, most of whom were women, typically subjected to more oppression than men. He eventually coined the term ``neurosis'' to describe the internal struggle he believed went on between a person's instinctual urges (i.e. \textit{id}) \& their conscience (i.e. \textit{superego}), the excessive \textit{anxiety} or nervous tension that often accompanies this internal war, \& the unusual psychological \textit{symptoms} a person can develop when attempting to mitigate (i.e. through the use of marginally effective \textit{defense mechanisms}) the intense emotional pain associated with these inner conflicts. He then developed a set of theories \& constructs that appeared to adequately explain how his patients developed their bizarre maladies. In the process, however, \textit{he also came to believe that he had discovered universal, fundamental principles that explained personality formation \& the entire spectrum of human psychological functioning}.

Many of the constructs \& terms 1st articulated by Freud, \& several other ``psychodynamic'' theorists who followed him, found their way into common parlance over the years. For a significant period of time, these tenets also gained widespread acceptance by mental health professionals. \& many of the principles of what I refer to in this book as ``traditional psychology'' still enjoy a fair degree of acceptance, not only among professionals but also among lay persons. This is true despite the fact that, in recent years, several of the most important assumptions \& doctrines have been proven completely false or significantly flawed.

Not only are some of traditional psychology's central tenets of questionable validity, but also times have changed dramatically since the days of Sigmund Freud. If a motto or saying best befits our modern era's zeitgeist, it would be much like the once popular commercial that urged: ``Just \textit{do} it!'' As a result, highly pathological levels of neurosis (as exemplified by the girl who went blind simply because she ``lusted'' after a boy she found attractive) have all but completely disappeared, especially in industrialized free societies. Instead of being dominated by individuals overly riddled with unfounded guilt \& shame (i.e., ``hung-up'' as children of the `60s used to say), modern western culture has produced increasing numbers of individuals who aren't ``hung-up'' enough about the things they let themselves do.

So, today we are facing a near epidemic of what some theorists refer to as \textit{character disturbance}. Neurosis is still with us, but for the most part at \textit{functional} as opposed to pathological levels. I.e., most people today experience just enough apprehension \& internal turmoil when it comes to simply acting on their primal urges, that they don't in fact ``just do it.'' Instead, they experience sufficient anticipatory guilt or shame to restrain their impulses \& conform their conduct to more socially acceptable standards. So, one can easily say that their neurosis is \textit{functional}. It's largely what makes society work.

Freud used to say that civilization is the cause of neurosis, \& given the climate of his time, it's easy to see how he came to that conclusion. In a sense, he had a point; but his observation was more than a bit narrow-sighted. True, the prohibitions any society imposes on the unrestrained expression of primal urges can give rise to a fair degree of anxiety in some of us. \&, a brutally oppressive culture can breed excessive degrees of neurosis. But in large measure, \textit{it's most people's \textbf{capacity} to become unnerved when contemplating acting like an animal (i.e. their capacity to be ``neurotic'' to some degree) that makes civilization itself possible}. So, it's precisely because significant numbers of people still get a little ``hung up'' when they contemplate punching someone's lights out, or have some apprehension \& qualms of conscience when they blink about taking something that doesn't belong to them, that there's any degree of civilization left at all.

Today many types of professionals span a wide variety of disciplines that deal with mental health issues \& personal problems of 1 variety or another. Most of these professionals have \textit{never} encountered -- let alone treated -- a case of ``hysterical blindness,'' pseudo-paralysis, or any similar phenomenon. In fact, it's becoming increasingly rare for professionals to encounter a case of neurosis at a highly pathological level of intensity. Therapists rarely deal with problems that stem from a conscience so overactive or oppressive that it causes a person to develop bizarre or severely debilitating psychosomatic or other pathological symptoms. Instead, mental health clinicians in all disciplines increasingly find themselves intervening with individuals whose problems are related to their dysfunctional \textit{attitudes} \& \textit{thinking patterns}, their shallow, self-centered relationships, their moral immaturity \& social irresponsibility, \& their habitual, \textit{dysfunctional behavior patterns}. All of these stem from an \textit{underdeveloped conscience} \& reflect significant \textbf{\textit{deficiencies or disturbances of character}}.

This wouldn't be so much of a problem if it weren't for the fact that many mental health professionals not only are trained primarily in classical theories of human behavior, but also cling to beliefs about human nature \& the underpinnings of psycho-social dysfunction that originally emanated from these theories. For this reason, they often attempt to use the tenets \& the principles that flow from more traditional paradigms to guide them in their efforts to solve today's very different kinds of psychological problems. In short, they attempt to understand \& treat character disturbance with approaches \& methods originally designed to treat extreme levels of neurosis.

Retaining outdated notions about why people do the things they do can put anyone -- lay persons \& mental health professionals alike -- at a great disadvantage when it comes to understanding \& dealing with the disturbed character. Even some of our more modern frameworks for understanding human behavior are inadequate to address the phenomenon. The problem is compounded by this fact: It's not ``politically correct'' to consider people's emotional \& behavioral problems as stemming from or reflecting deficiencies in their character. So, sometimes problems actually rooted in character pathology might be framed as almost anything else (e.g., an ``addiction,'' ADHD, Bipolar Disorder, a ``chemical imbalance,'' etc.). Their symptoms can be somewhat managed with medication or other forms of treatment. Then, if at least some degree of change is observed, that mere fact validates the perspective that disease of some sort -- as opposed to character -- caused the problem. Sometimes, professionals actually do recognize personality or character disturbances, but regard them as unchangeable or untreatable. So, they target issues other than character concerns in therapy. At other times a professional's over-immersion in traditional paradigms will prompt them to view everyone -- even the most severely character-disturbed individual -- as neurotic, at least to some degree, \& then attempt to treat their supported neurosis. But the reality is this: \textbf{\textit{character disturbance is 1 of the most pressing psychological realities of our age; it's becoming increasingly prevalent; \& it's an entirely different phenomenon from neurosis, requiring a different perspective to adequately understand \& treat}.}

Understanding character disturbance requires viewing human beings, \& the reasons they do some of the things they do, in a very different light. Further, character disturbance simply can't be dealt with effectively using traditional approaches. The tools \& techniques that have proven effective in treating character disturbance are radically different from those originally developed to treat neurosis.

Over 14 years ago, I wrote \textit{In Sheep's Clothing: Understanding \& Dealing with Manipulative People} to help both therapists \& average folks understand what certain people are really like, \& how they manage to manipulate \& control others. The book has grown increasingly popular every year since its 1st printing, has been translated into several foreign languages, \& consistently draws highly laudatory reviews from professionals as well as lay persons. The comments are generally of a similar nature. Readers report that once they cast off old notions about the nature \& behavior of the manipulative persons in their lives, ``got it'' w.r.t. what really makes such individuals ``tick,'' \& adopted not only a new perspective for understanding such persons but also a new set of rules \& principles to guide their relationships with them, their circumstances dramatically improved. This kind of feedback has been more than edifying for me, \& inspired me to write this book. \textit{In Sheep's Clothing} was about 1 type of disturbed character. This book is about all of the various disturbed character types you are likely to encounter in your life. \textit{In Sheep's Clothing} briefly delved into the general topic of character disturbance. This book takes a much deeper look at this significant \& disturbing phenomenon, its growing prevalence, \& some of the socio-cultural features of the current era responsible for promoting it.

My primary purpose in writing this book is to help you: (1) cast off faulty assumptions that can place you at a distinct disadvantage, \& (2) understand the real reasons disturbed characters behave the way they do. It's important to recognize that disturbed characters differ dramatically from neurotics on almost every imaginable dimension of interpersonal functioning. They don't hold the same values, believe the same things, harbor the same attitudes, think the same way, or behave in the same manner as neurotics. This book will help you understand why you might have had such a difficult time dealing with the disturbed characters in your life, what makes these individuals so different, \& why aid you might have sought, hoping to deal with problems involving them, proved inadequate.

As with \textit{In Sheep's Clothing}, I've written this book in a concise manner, easy to read \& understand. It should be equally helpful to the layperson as well as the professional who is primarily versed in or aligned with traditional perspectives. I have deliberately not included mounds of difficult-to-understand scientific research data \& have attempted to translate sophisticated \& highly technical material into simpler, yet reasonably accurate language. My intention was not to craft an authoritative \& comprehensive textbook-style treatise on human nature, personality, or psychopathology. Rather, I present principles \& perspectives derived from years of experience working with disturbed characters \& their victims, illustrate those principles with real-life examples, \& support the most important contentions \& perspectives with relevant research findings when appropriate or necessary.

My secondary purpose in writing this book is to expose character disturbance for the significant social problem that it is, \& address some of the socio-cultural influences responsible for it. I am not the 1st to sound the alarm on the issue, \& I hope I will not be the last. In their book on character lengths \& virtues, the eminent psychology pioneers \& researchers Christopher Peterson \& Martin E.P. Seligman address the issue directly:

``After a detour through the hedonism of the 1960's, the narcissism of the 1970's, the materialism of the 1980's \& the apathy of the 1990's, most everyone today seems to believe that character is important after all \& that the United States is facing a character crisis on many fronts, from the playground to the classroom to the sports arena to the Hollywood screen to business corporations to politics.''\footnote{Peterson, C., \& Seligman, M., \textit{Character Strengths \& Values: A Handbook \& Classification}, (American Psychological Association \& Oxford University Press, 2004), p. 5.}

But no problem can be resolved until it's fully acknowledged \& adequately defined. That's the 1st step. The 2nd step is to take a serious look at the characteristics of our social milieu that contribute to the problem. This book will do both.

I have been working with disturbed characters for over 25 years now. Yet I know that much of what I assert in this book is likely to be controversial. When I 1st started doing workshops on manipulators \& other problem characters, especially with professionals, several attendees were uncomfortable with my perspective. Some even walked out! That's because a lot of what I had to say challenged some longstanding notions \& deeply-held beliefs about why people experience psychological problems in the 1st place, \& how professionals must assist them in achieving mental \& emotional health. But time \& recent research has validated much about the perspective I introduced those many years ago. These days, it's very common for me to see many heads in the audience frequently nod in approval as I articulate the principles I've adopted to understand \& deal with disturbed characters. It's also common for individuals who have already been to 1 of my workshops to come back again several times for ``refreshers'' on the tenets I will outline in this book. 1 of the more frequent comments I get from professionals involves how much more satisfying working with disturbed characters has become after they adopted some of these principles.

But old notions \& biases aren't easily shed, so I still expect some controversy \& debate. The same was initially true for \textit{In Sheep's Clothing}. But I am confident that -- if you approach the concepts I will introduce with an open mind, allow my disturbed-character descriptions to \textit{validate your experience}, accept the reality that our very different times have spawned problems the major helping professionals have only recently begun to face by developing sound paradigms, \& at least try out some of the principles \& tools I will outline -- you'll be better prepared to understand \& deal with the individuals in your life who display significant character pathology.

There are 2 principal reasons I felt it imperative to write this book at this time. 1st, I have come to a deep realization about how serious the problem of character disturbance has become, \& the degree to which it undermines the very foundations of our free society. For the past 15 years, I have been working with some of the most severely disturbed characters, most of whom have been incarcerated at some point in their lives, \& several of whom have been incarcerated multiple times. You should know that the United States has a higher percentage of its population incarcerated than any industrialized country, \& that percentage is increasing. You should also know that contrary to popular belief, it generally takes some very serious \&\texttt{/}or chronic misbehavior to end up incarcerated. Yet, prisons \& jails are overcrowded, \& the judicial system is overburdened. Although very few may be incarcerated who don't necessarily need to be, many others lead lives of wanton major social norm violations for which they rightfully could have been sanctioned many times, but weren't because of the scarcity of space \& resources. I've also worked with many individuals who, although they engaged in no major societal transgressions \& do not have criminal records, nonetheless harbor significant deficiencies of character. I've increasingly counseled individuals who didn't begin functioning in a truly responsible manner until they were in their late 40s, 50s, or even 60s. (I recently attended a discussion group in which a group member discussed his problematic relationship with a brother who, even at age 67, continues to exercise little control over his impulses \& engages in deeply irresponsible behavior!)

The greatness of American society has deteriorated dramatically in recent years because of our culturally-spawned character crisis. There is a saying often attributed to Alexis De Tocqueville: ``America is great because she is good; \& if America eve ceases to be good, she will cease to be great.'' De Tocqueville never actually said this, yet the saying contains a powerful truth. The overall character of a country can be no better than the collective character of its constituents. I grew up in an era of unprecedented American greatness, in significant measure spawned by the distinctive character of what Tom Brokaw rightly called ``the greatest generation.'' I have now lived long enough to witness what I regard as the tarnishing of the best American ideals; \& although I am no longer shocked by it, I am nonetheless outraged. I'm mostly outraged because I still see our unique brands of free expression \& free enterprise as the last, best hope for improving the human condition. \textbf{\textit{freedom \& character are inextricably interdependent}}. John Adams adamantly asserted that ``no government'' deliberately designed to be limited in power \& scope would be ``capable of contending with human passions unbridled by morality \& religion.'' He further noted that the American ``constitution was made \textit{only} for a religious \& moral people,'' \& that ``it is \textit{wholly inadequate} for the government'' of people ``of any other'' character.\footnote{Adams, J., Address to the Military, Oct 1798.}

In the absence of sound individual conscience, society is naturally tempted to legislate its way out of the problems that ensue, imposing rules \& restrictions that inevitably increase the role, size \& scope of government, \& limit individual freedoms. While it may seem like a necessary evil, such attempts to legislate social responsibility \textit{never} really fix the problem; they only set the stage for both economic \& cultural deterioration. So, it bothers me greatly that the populace's complacency \& denial, with a fair amount of aiding \& abetting by some mental health professionals, in recent times have all allowed this character crisis to surreptitiously \& steadily spin out of control, to fray our culture's once distinctly noble fabric, to erode the integrity of our national character, \& to hasten our country's fall from greatness.

Sometimes I wonder just how many fast-talking politicians, greedy executives, \& crafty religious ``leaders'' -- finally caught in their lies \& indiscretions, failed relationships, or major corporate scandals -- it will take for us to realize how much character \& integrity really matter. For a long time, it seems like we had resigned ourselves to the notion that people simply can't be any better or do any better than we've come to expect. But the 2nd major reason I felt compelled to write this book now is this: I know from years of experience that the problem of character disturbance is not at all hopeless. There are some straightforward things we can do to address the cultural contributors to the character crisis. There are interventions we can initiate at the individual level to encourage a person to develop the character they need to function responsibly in society, \& to form healthy, happy, intimate relationships. From a therapist's perspective, the interventions necessary are not of the same sort that one would use to deal with neurosis. Nor are they busy, given the challenging nature of the problem. But \textit{there is hope} for dealing effectively with the disturbed character, especially once you know \& accept the approach that's necessary to deal with the problem.

When I was developing \textit{In Sheep's Clothing}, the cognitive-behavioral therapy (CBT) revolution was in its infancy. Today, it has become more widely accepted. Nonetheless, many individuals who outwardly assert that they understand \& accept the tenets of cognitive-behavioral therapy still have a hard divorcing themselves from outdated \& ineffective paradigms. Furthermore, although they might readily focus on a person's distorted cognitions, some clinicians have an aversion to aligning themselves with the most important aspect of the cognitive-behavioral model: focusing on \& modifying \textit{behavior}. As a result, they often don't fully \& faithfully implement the principles of CBT (doing largely C instead of C{\bf B}T). This is a genuine shame because CBT has demonstrated its superiority to other forms of intervention when it comes to treating persons of disturbed character. The principles advanced in this book are very much in harmony with the principles of CBT.

We are all in this vast human experiment together. Like it or not, ours is an extremely interconnected \& interdependent world. My personal mission for the last several years has been to call attention to a significant social problem, \& to inspire people to address \& overcome it. This book is the culmination of my most recent efforts toward that end.

At a fundamental level, we are all savages seeking the survival of the fittest. But we also have within us the power \& capacity to elevate ourselves \& inspire others to a much higher plane of functioning. Doing so, however, requires the development of \textit{character}.'' -- \cite[pp. 8--19]{Simon2011}

%------------------------------------------------------------------------------%

\section*{Introduction}

\subsection*{Personality \& Character}
``This is a book about character, \& the disturbances of character it seeks to examine differ in significant ways from the various personality disturbances we'll discuss. Therefore, it's important that we clearly distinguish \& define these 2 terms. It's common to hear \textit{personality} \& \textit{character} used in a manner suggesting that they are virtually synonymous. Even professionals frequently employ the terms interchangeably. Although the 2 words are related, I think it not only important but also helpful to draw a firm distinction between these 2 concepts. A close look at the terms' origins \& the evolutions of their meanings will help clarify the working definitions we'll employ throughout this book.'' -- \cite[p. 20]{Simon2011}

\subsection*{Traditional Conceptualizations of Personality}
``The word ``personality'' derives from the Latin word \textit{persona}, which means ``mask.'' In the ancient theater, males played all the roles, including those of female characters. Also, the art of dramatizing situations \& conveying emotion was not as evolved as it is today. So actors used masks of various types not only to denote gender, but also to depict \& emphasize various emotional states. Classical psychology theories borrowed the term ``persona'' because they generally conceptualized \textit{personality} as the social ``mask'' a person wears to conceal the more authentic or ``true'' self.

Classical theories of personality also conceptualize all individuals as fundamentally struggling with \textit{fears} of various kinds, especially fears of social rejection, condemnation, or abandonment. These theories regard basic human needs, desires \& emotions as universal, \& potential environmental ``threats'' to them as ever-present. So, the theories postulate that (1) to a greater or lesser extent, everyone struggles with fears that those basic wants \& needs will be thwarted in some way; \& (2) people fear that they might encounter disappointment \& disapproval if they ever reveal their true wishes \& desires. Classical theories also propose that people engage in certain stylized but unconscious ways of ``defending'' themselves against a potentially hostile \& rejecting world, \& in the process end up estranged from their more authentic selves. The classical model, then, sees personality as an \textit{unconsciously constructed fa\c{c}ade} -- a mask that hides the genuine person behind a veritable wall of defenses. This conceptualization still has some value when it comes to understanding personalities who are best thought of as \textit{neurotic} (more about this later).

Traditional theories of personality also postulate that a person wouldn't need to hide his or her authentic self behind the fa\c{c}ade of personality if our environments weren't so hostile, cold, or rejecting. In other words, everyone would be ``authentic,'' healthy, \& without defensive armor if the world didn't force so much emotional trauma into our lives. From traditional perspectives, it's the slings \& arrows of life that have the greatest influence in shaping personality. Traditional theories view everyone as \textit{essentially the same} underneath their structure of defenses. Furthermore, the types of defenses a person most likely will employ are seen as logical growth -- responses to the kind \& severity of traumatic experience to which they've been subjected, especially during their formative years. Within traditional frameworks, the various personality types are defined primarily by the cluster of defenses they have learned to employ.'' -- \cite[pp. 20--21]{Simon2011}

\subsection*{A More Contemporary Conceptualization of Personality}
``More recent conceptualizations, most eloquently described by the eminent theorist \& researcher Theodore Millon, generally define personality as an individual's distinctive \& relatively engrained ``style'' of interrelating.\footnote{Millon, T., \textit{Disorders of Personality}, (Wiley Interscience, 1981), p. 4.} Such conceptualizations of personality often incorporate a multidimensional perspective: They recognize a person's constitutional predispositions (i.e. hereditary, hormonal, temperamental, \& other biological factors), environmental factors (e.g., early learning, significant life events, peer \& role-model influences, degree of nurturance available, etc.), \& a \textit{dynamic interaction} between what the world teaches a person \& how they are predisposed to respond to events. Multidimensional personality theorists believe all the aforementioned factors contribute to a person's distinctive manner of perceiving, relating to, \& interacting with others \& the world at large.

The multidimensional perspective of personality better explains how some individuals can develop very dysfunctional personalities, despite being reared in relatively benign, supportive environments; whereas others can develop remarkable \& admirable character despite experiencing the most egregious circumstances. If you recall, traditional theories assume this: Underneath, we're all the same, \& we'd be perfectly healthy \& authentic individuals if it weren't for the fact that we grew up in a world full of pain, rejection, \& emotional trauma. Within the multidimensional framework, both the environment \& an individual's innate predispositions play roles. The weight various factors might carry in shaping personality varies considerably from person to person. The choices a person makes about how to best cope with life's challenges also play a role. In short, the multidimensional model allows for the increasing scientific evidence that we are not all the same. Each of us has a unique collection of traits. Some developed as a result of learning, some we were simply endowed with by nature, \& some developed as a result of the dynamic interaction between our innate predispositions, the environment in which we were raised, \& the choices we've made. \& the model also allows for the fact that, once our ``preferred'' ways of thinking \& behaving congeal \& become ingrained, they fairly much define who we are as individuals. Personality, therefore, is seen -- not so much as a false-face or a pretense -- but rather as a stable set of traits, preferred thinking \& behavior patterns that define our unique \textit{style} of interaction over a wide variety of situations, \& for most of our lifetime.

So, the multidimensional conceptualizations of personality are more comprehensive than the classical definitions. They also appear more accurate \& useful when we try to understand the thinking \& behavior patterns of individuals best described as \textit{character-disturbed} rather than \textit{neurotic}. This will become increasingly evident to you when we take a more in-depth look at the nature of character disturbance \& its vast differences from neurosis.'' -- \cite[pp. 21--23]{Simon2011}

\subsection*{Character}
``The word ``character'' derives from both Old French \& Greek words meaning to engrave or furrow a \textit{distinctive mark}. The word has been used to denote the most distinguishing traits of overall personality that uniquely define or ``mark'' an individual as a social being. Most especially, the term commonly reflects an individual's positive personality aspects -- those \textit{socially desirable qualities} \& virtues such as self-control, ethics, loyalty, \& fortitude.

As mentioned before, it's not uncommon for professionals as well as lay persons to use the terms ``character'' \& ``personality'' interchangeably. It's also not uncommon for folks to speak of character as if it were synonymous with \textit{strength} of character. Many also erroneously equate the terms \textit{personality disorder} \& \textit{character disorder} (I'll have more to say about this later). But, again, this is a book about character \& the social \& psychological consequences of significant disturbances or deficiencies of character. So it's of paramount importance that we highlight the key differences between someone who possesses all of the various traits \& quirks of a certain \textit{personality} \& someone whose dominant personality features reflect significant deficiencies or defects in their character. Therefore, in this book, the term ``character'' will refer to those \textit{distinct aspects of personality} that reflect the presence \& strength of a person's virtues, personal ethics, social conscientiousness, \& depth of commitment to respect-worthy \& meritorious social conduct.'' -- \cite[p. 23]{Simon2011}

%------------------------------------------------------------------------------%

\section{Neurosis \& Character Disturbance}
``Before we can even begin a meaningful discussion about the differences between neurosis \& character disturbance, it's important to understand the strengths \& limitations of any scientific or philosophical metaphor. We simply can't have a discourse about the important matters of life without theories, definitions, \& constructs. But in the end, absolute truth is illusive. Some might say it's unknowable. We have to settle for ``metaphors'' that, indicated by evidence, approximate the incomprehensible truth as closely as possible.

Some metaphors appear more ideally suited to understanding \& dealing with various aspects of reality. E.g., Sir Isaac Newton tried to explain the phenomenon we know as gravity, as well as planets' motion \& orbits, with the ``law'' that every material object attracts every other in proportion to their mass, \& in inverse proportion to their distance from one another. In a competing ``metaphor'' for explaining the same phenomenon, Einstein proposed that what appears as 1 body attracting another is really the distortion or warping of ``space-time'' (a concept so complex it's far beyond the scope of this book); \& that all clusters of matter (including particles of light) simply follow the resulting ``curvature'' of distorted space-time. A landmark test of Einstein's theory during a solar eclipse demonstrated convincingly that his metaphor provides a more accurate description of gravitation. Newton's metaphor, however, still works very well when you're trying to plot a course for a rocket to the moon, or determine an orbital path around the earth. So, there are still situations in which the Newtonian metaphor is useful, despite its limitations. \& neither Einstein's metaphor nor Newton's gives us the whole truth about gravity. We still don't really know exactly what it is or how it works.

Social science metaphors are no less subject to limitation or appropriate application than any other scientific metaphor. \& any metaphor can become problematic when we over-generalize it, or stretch it to fit or explain phenomena it was never really meant to explain. Most branches of science have come to grips with this. All encompassing explanations of reality such as string theory, the theory of everything, or the unified field theory are still illusive. So, for the most part, scientists apply various theories in the areas they appear to suitably explain. But in the social sciences, we've been slow to recognize \& deal with limits to some of our traditional explanatory models.

Traditional psychodynamic metaphors still apply \& have value when you're trying to understand or deal with the phenomenon we call ``neurosis.'' But the evidence is mounting that such metaphors are inadequate at best, \& potentially counterproductive or even harmful at worst, when we're trying to understand or deal with character disturbance. The constructs of neurosis \& character disturbance are themselves only metaphors. Each attempts to describe a psychological reality we cannot adequately or completely define; but we have to attempt to give it some structure if we're to understand it at all. The terms themselves are not reality; they're metaphorical ways we attempt to describe some aspects of reality. \& each way has its strengths \& limitations, as well as areas of applicability \& areas of poor fit.

I find it useful to conceptualize neurosis \& character disturbance as constructs lying at different ends of a continuum. A representation of this concept is provided in the following diagram: Neurosis $\longleftrightarrow$ Self-Actualization Altruism, Character Disturbance.

At 1st glance, the representation depicted above might appear a bit confusing. It's important to realize, however, that very few individuals have ever walked our planet who could rightfully be considered so socially \& morally evolved (e.g., Jesus, the Buddha) that they're considered truly altruistic. By altruistic, I mean that their commitment to selfless, humanistic behavior is not rooted in pangs of guilt for doing otherwise, or in an unacknowledged selfish desire for admiration or immortal reputation. Altruists, by definition, are individuals who \textit{freely} \& completely commit themselves to advancing the greater good. They are not neurotic because they have no driving desire to avoid guilt or shame for doing otherwise. Also, they're not out for personal glory or to be revered by society. By definition, altruists simply, freely, \& nobly choose to subordinate their own selfish desires for the good of all.

Fully \textit{self-actualized} (altruistic) individuals are extremely rare indeed; some assert they don't really exist. That's because most people restrain their baser instincts out of a sense of guilt \&\texttt{/}or shame for doing otherwise. It's theoretically possible to be without neurosis \& not be character disordered, but the absence of neurosis usually results in some degree of character disturbance. That's because without potential pangs of guilt or shame influencing our decisions, most of us would act in socially irresponsible ways. So, character disturbance \& neurosis are at opposite poles of a continuum, with \textit{each individual personality falling somewhere along the continuum}. \&, as you might expect, personalities rightfully classified as more neurotic will differ significantly from those rightfully classified as more character-disturbed on several dimensions.

The framework advanced here ultimately views character disturbance as a crucial \textit{dimension} of personality. It varies in degree \& results from a number of shaping influences that affect an individual's personality development. E.g, there are people who started out in life with some degree of neurosis; but extraordinary pain \& trauma ``hardened'' their hearts, \textit{solidifying} their preferred styles of coping to such a degree that their neurosis eventually became relegated to their souls' deepest recesses. In such individuals, any neurosis they harbor can usually only re-emerge when they are truly ``broken down'' by sufficiently humiliating or spirit-fracturing life events, abject failures of their coping style, or sufficiently intensive \& confrontational therapeutic techniques. But some individuals were never very neurotic to start with. Either they didn't experience the usual neurosis-fostering \& civilizing experiences like most of us, or their innate predispositions prevented those typical civilizing influences from impacting them as such influences impact most of us.'' -- \cite[pp. 24--27]{Simon2011}

\subsection{Disturbance vs. Disorder}
``Throughout this book you will see the following terms: ``personality disturbance,'' ``personality disorder,'' ``character disturbance,'' \& ``character disorder.'' It's important to recognize that not all problematic aspects of personality or character rise to the level of a true disorder. For a disturbance of personality \&\texttt{/}or character to be considered a disorder, it must be of such intensity, inflexibility, \& intractability that it impairs adaptive functioning in a wide variety of situations. I've included a more in-depth discussion on this in the next chapter.'' -- \cite[p. 27]{Simon2011}

\subsection{Key Neurotic vs. Disturbed Character Differences}
``Character-disturbed personalities differ from neurotic personalities on almost any dimension of interpersonal functioning imaginable. The differences greatly affect how they function in interpersonal relationships, as well as how they respond to therapeutic interventions. It's virtually impossible to list all of the ways disturbed characters \& neurotics differ, but I think it's helpful to examine some of the more significant ones (summarized in the chart on p. 60):

\subsubsection{Anxiety}
Disturbed characters are significantly different from neurotics w.r.t. to their levels \& quality of anxiety. Anxiety is the primal emotion (i.e. fear response) we experience when we feel threatened in some way. It has always been thought to play the most important role in both creating \& maintaining neurosis. When our fear is rooted in a specific, identifiable circumstances -- such as being in a room filled with a lot of people, having to take a test, or coming face to face with a snake -- we call it a \textit{phobia}. When our apprehension does not appear connected to a specific thing or circumstance, is unidentifiable, unknown, or unconscious, we call it \textit{anxiety}.
	
Psychodynamic theorists of all persuasions give pre-eminence to anxiety's role in all forms of neurosis. Individuals who are overly anxious, excessively apprehensive, inordinately fretful, or too easily unnerved can suffer a host of maladies either directly caused or exacerbated by their anxiety. Further, therapists have traditionally thought neurotic ``symptoms'' -- whether stress-related ulcers, tension headaches, avoidance of crowds or open places (i.e. agoraphobia), obsessive worry, desperate actions to prevent abandonment, etc. -- to be rooted in anxiety.

It's also noteworthy that, whether a fear is conscious \& specific (i.e. a phobia) or unconscious or unidentified, research indicates that there is a common essential factor in getting rid of it: exposure. When phobic clients undergo treatment with behavioral or cognitive-behavioral methods, they are encouraged to systematically come into increasing contact with the very situations they fear, thus gaining an increased sense of personal empowerment, \& reducing their apprehension levels. In traditional psychotherapy, clients gradually come into conscious contact with, \& therefore ``face,'' their previously unconscious fears in the safe, supportive, accepting atmosphere promoted by the therapist; \& their fears are eventually reduced.

Anxiety is minimally present or plays a negligible role in the disturbed character's problems. In some cases, it's absent altogether. Character-disordered individuals are notoriously nonchalant about the things that upset most other people. Some, especially the aggressive personalities (more about them later) appear to \textit{lack adaptive levels} of fearfulness.\footnote{Millon, T., \textit{Disorders of Personality}, (Wiley Interscience, 1981), p. 198.} They don't get apprehensive enough about their circumstances or their conduct. They're not unnerved enough at the prospect of conflict, \& they readily leap into risky situations when others would hesitate. For the most part, disordered characters don't do dysfunctional things because some past trauma has them too ``hung-up'' to do otherwise. Instead, they do them because, unlike neurotics, they lack the capacity to get hung-up enough to think twice about their behavior, inhibit their impulses, or restrain their conduct. A little of the neurotic's typical apprehension would go a long way toward helping the disturbed character be more cautious or hesitant when it comes to frequently doing the  things that cause problems.

For several reasons that I have never fully understood, traditionally-minded therapists, as well as relatively neurotic individuals, appear determined to ascribe fears \& insecurities to disordered characters that simply don't exist. They will frequently misinterpret the behavior \& motivations of character-disordered individuals \& frame their behaviors inaccurately. E.g., some disordered characters have a \textit{passion} for novelty \& a \textit{craving} for excitement. So they constantly seek shallow, intense, short-lived, \& high-risk sexual involvements \& other interpersonal entanglements. But this characteristic thrill-seeking behavior is often framed as a ``fear'' of intimacy or commitment. This mistake is made because it's difficult for neurotic individuals (or traditionally-minded therapists for that matter) to imagine why a person wouldn't necessarily prefer a stable, intimate relationship over multiple risky encounters, unless they were in some way apprehensive about the prospects of engaging in something more substantive. This kind of thinking also reflects a long-held, but never proven, tenet of classical psychology: Everyone will naturally gravitate toward the healthiest life choices unless they are hung-up by unconscious fears that stem from past emotional trauma. It's also possible that some therapists are so married to their traditional metaphors about the nature of human behavior that, even when they encounter a ``square peg,'' they still try to fit it into the proverbial round hole.

\subsubsection{Conscience}
The neurotic individual is basically a person with a very well-developed conscience or superego. Sometimes that conscience can be overly active to the point of being oppressive. Neurotics have a huge sense of right \& wrong, \& always want to do what they think is the most correct. They can sometimes set impossibly high standards, engendering a significant amount of stress. Neurotics are also prone to judge themselves overly harshly when they fail to meet their own expectations. They take on inordinate burdens, carrying the world's weight on their shoulders. When something goes wrong, they quickly ask themselves what more  \textit{they} can do to make the situation better.
	

The disordered character's conscience is remarkably under-developed \& impaired. Disturbed characters don't hear that little voice that urges most of us to do right, or admonishes us when we're contemplating doing wrong. Or if they do hear it, they can easily ignore it, silence it, or put it in a ``lock box'' (i.e. \textit{compartmentalize it}). As opposed to persons with a sound conscience, they don't push themselves to take on unattractive burdens \& responsibilities; \& they don't hold themselves back when they want something they really shouldn't have.

In the most sense disturbances of character, conscience is not simply weak, underdeveloped, or flawed, but \textit{absent altogether}. The little voice most of us have simply isn't there, \& the capacity to even form a conscience is grossly deficient. Robert Hare aptly named his book about the most severely disordered character, the psychopath, \textit{Without Conscience}.\footnote{Hare, R., Without Conscience: \textit{The Disturbing World of the Psychopaths Among Us}, (Guilford Press, 1991).}

It's hard to imagine individuals with no conscience at all. \& as mentioned earlier, our traditional psychology metaphors have conditioned us to believe that underneath it all, we're all the same. Most folks find it unimaginable that some people are simply not ``normal.'' That's why psychopaths are able to prey on the unsuspecting so effectively. Their victims dupe themselves with their inability to accept that the predators they've been dealing with are heartless \& devoid of normal human empathy. It's very important to realize \& accept that not everyone is the same. Not everyone has an active, mature conscience. All the disordered characters have deficiencies of conscience. They vary from those whose conscience is markedly immature to those with no conscience at all. Such individuals can do others great harm with absolutely no compunction. Those who fail to recognize this are extremely vulnerable.

Lacking a mature conscience, possessing a diminished capacity to experience distress in the face of injury to others, \& deficient in empathy, many disturbed characters don't experience genuine remorse for their hurtful acts, whether acts of commission or omission. They might have some after-the-fact \& practical regret for a behavior, especially if it results in some clear \& appreciable cost to them. But because they don't have a healthy conscience, there's nothing to keep them from acting on their destructive impulses in the 1st place. So any practical regret for their own behavior is usually too little \& comes far too late.

\subsubsection{Shame \& Guilt}
Because they are persons of conscience, neurotics can experience high (\& sometimes toxic) levels of both shame \& guilt. Shame is the emotional state we experience when we feel badly about \textit{who we are}. Guilt is the emotional state we experience when we feel badly about \textit{what we've done}. Neurotics tend to judge themselves harshly, so they're quick to feel ashamed when they fail to measure up to their own high standards \& the self-image they try to maintain. They're also quick to feel guilty when they think they've fallen short, done something poorly, or caused injury to someone else.
	
Most neurotics' levels of shame \& guilt are adaptive to some degree. I have encountered a few neurotics who experienced such toxic levels of shame \& guilt growing up that it led them to develop truly pathological symptoms. But even these individuals rarely experienced the extreme levels of guilt or shame necessary to produce the kinds of bizarre psychological phenomena Freud used to treat. Most neurotics are not sick from unreasonable or extreme levels of guilt or shame. But they are hypersensitive to these feelings. They are quick to feel badly about themselves when they've done something that reflects negatively on their character; \& they're too quick to beat themselves up emotionally when they think they've committed unpardonable sins.

Disturbed characters lack sufficient pangs of guilt or shame when they do things that are harmful or hurtful. Such feelings can only emanate from a well-developed conscience, which, as we discussed earlier, they lack. so, shamelessness \& guiltlessness are 2 of the disturbed character's most distinguishing features. They don't feel badly enough about the kind of person they are when they repeated do things that negatively affect or injure others. They also don't feel badly enough about the harmful things they do, at least not badly enough to keep from doing them over \& over again. If they were able to experience enough guilt or shame, they might refrain from doing socially harmful things in the 1st place, or from doing the same wrongful acts repeatedly.

The plethora of books dealing with shame \& guilt that dominated the self-help \& ``recovery'' market of the `60s, `70s, \& even `80s, was largely written \textbf{\textit{by, for, \& about}} neurotics. Shame \& blame were the names of their game, \& most of those books blamed toxic levels of guilt \&\texttt{/}or shame for a wide variety of psychological problems that damaged a person's self-esteem. These books largely made us believe there was no such thing as good shame or healthy guilt. Some authors \& theorists later relented regarding guilt, acknowledging that at least some measure of guilt is necessary to keep us civilized. But even today, the dominant opinion about shame (\& supported by empirical research), casts it as a bad thing, period. The general consensus seems to be this: While it's a relatively good thing to feel badly about something harmful you've \textit{done}, feeling badly about oneself -- about whom one literally is -- is \textit{never} a good thing.

But after working for many years with disturbed characters, I quickly came to question the validity of this premise.

Most of us experience some genuine self-disgust with  the \textit{kind of person} we might find ourselves becoming when we engage in behaviors disordered characters display. This is precisely what prompts us to change our ways, \& restore a self-image we can live with. I've known many individuals who made significant changes in their characters. But when they did so, it was not only because they regretted their irresponsible behaviors, but also because they became unsettled enough with the person they had allowed themselves to become (i.e. become too ashamed of themselves) that they decided to change course. So, it appears that one must have the capacity to experience both shame \& guilt in order to forge a sound character. As is usually the case, however, it's a matter of degree. When individuals experience toxic \& unwarranted levels of either guilt or shame, there can indeed be a negative impact on psychological health.

Some professionals (\& non-professionals) take issue with positions I outlined above. They insist that even disordered characters actually do feel guilt \& shame, but that they effectively utilize -- or perhaps over-utilize -- certain ``defense mechanisms'' such as ``denial'' \& ``projection'' to assuage any emotional pain. This is largely because they still adhere to the tenets of classical psychology (i.e. that \textit{everyone} is to a greater or lesser degree neurotic, if they are not in fact psychotic, \& that all individuals are fundamentally the same at a deeper psychological level). So, it's impossible for them to imagine how anyone could behave so shamelessly unless they were, in fact, defending themselves against real pain \textit{underneath} it all. I'll address the pitfalls of the all-too-common tendency to over-extend \& over-generalize classical psychology's metaphors in more depth later in the book. But it's important to remember that -- although a disordered character's failure to acknowledge or ``come clean'' about wrongdoing can be viewed as a ``state of denial'' prompted by feelings of shame \& guilt -- such a perspective is often inaccurate. \& it's potentially dangerous when trying to deduce the true nature of their problems \& degree of character pathology.

Being embarrassed at being uncovered or found out is not the same as feeling genuine shame. Shame is 1 of those mechanisms that \textit{make a person think twice} about doing something wrong in the 1st place. Moreover, a person who truly feels ashamed is certainly not likely to do the same things over \& over again with no compunction. Character-disturbed people will sometimes claim they didn't come clean with themselves or others, or didn't seek help for their problematic conduct, because they were too ashamed to do so. This is often simply a lie they tell. They know a neurotic person is likely to find such an explanation more acceptable than the truth about their lack of motivation to change.

Individuals overly invested in human behavior's classical explanations also tend to misinterpret the careless, reckless, impulsive propensities of criminal behavior. They might assume criminal individuals had a subconscious desire (arising out of pangs of conscience) to be stopped when their seemingly thoughtless actions lead them to be easily apprehended. There has never been any empirical support for this notion, but that has not kept some from adhering to it. As I have stated several times already: Using metaphors that once appropriately described some neurotic behaviors to explain a disordered character's behavior almost always puts a person at a disadvantage when trying to understand or intervene in troublesome situations. A key thing to remember: Whereas neurotics have a propensity to experience guilt \& shame too readily \& with too great an intensity, the exact opposite is true for the disturbed character.

\subsubsection{Level of awareness}
The distress neurotic individuals experience generally stems from emotional conflicts  that are mostly \textit{unconscious}. A woman experiences an unexplained ``funk.'' She doesn't know it's related to her suppressed feelings of grief \& loss re-surfacing near the anniversary of her mother's death. If she \textit{did} know it, she might not even need to see a therapist to help her sort out why she was suddenly feeling so blue. A man with an ulcer is unaware that his obsessive worry over losing his job, in turn fueled by his deep-seated mistrust of authority figures, arises out of his experience with his abusive father. If he \textit{were} aware of it, he might never have needed to knock on the counselor's door. Neurotics are often in considerable emotional turmoil; but the deeper roots of their distress, \& sometimes even the very nature of their emotions, are often unknown to them.
	
The problems the disordered character experiences might be so ingrained that they fairly ``automatically'' occur. However, it's important to remember that the disordered character is \textit{fully conscious} of his problem behaviors. He not only knows exactly what he's doing, but also is fully aware of his \textit{motivations} for doing it.

Lying is 1 of the more common \& problematic behaviors of the disordered character. Sometimes this lying is done so ``automatically'' that the disturbed character finds himself lying without thinking much about it, \& even when the truth would have done just fine. That doesn't mean he doesn't know he's lying. He knows. He just does it so often \& so readily that he doesn't give it a second thought.

In my classical training I was taught never to ask clients ``why'' they did something because it would likely ``throw them on the \textit{defensive}'' \& they would be ``afraid'' to disclose. So, in my early work with disturbed characters, even though I was very interested in the motivations for their behaviors, I did not ask them directly why they did what they did. A fair amount of the time, when I did broach motivation issues, they would reply with something like: ``To tell you the truth, doctor, I really don't know,'' or ``That's what I'm in therapy to find out.'' This, of course, would reinforce the old notion that 1 of the major tasks for therapy would be to ``uncover'' the unconscious underpinnings of the conduct, so they could eventually ``see'' \& understand the reasons for their behavior \& ``work through'' their ``issues.''

What a surprise it was for me to learn that disordered characters are most often keenly aware, not only of their actions, but also of their motivations for them. This has proven to be true despite the fact that they might use the manipulation \& impression-management tactics of ``playing dumb,'' ``feigning ignorance,'' or ``feigning innocence'' to skirt responsibility. I eventually learned (as have other colleagues \& researchers in the area of character disturbance) that most of the time ``I don't know'' doesn't really mean the disturbed character is oblivious about his actions. It almost always means something else. It can mean:
\begin{itemize}
	\item ``I never really think about it that much.''
	\item ``I don't like to think about it.''
	\item ``I don't want to talk to you about it.''
	\item ``I know very well why I did it, but certainly don't want you to know. That would put you in a position of equal advantage with me, or possibly even give you an advantage over me -- having my number, so to speak -- \& I won't be able to manipulate you as easily or manage your impression of me.''
	\item ``I hope you'll buy the notion that I'm basically a good person whose intentions were benign. That I simply made un unwitting mistake, oblivious about the harm I caused; \& that I am willing to increase my awareness with your guidance.'' (This kind of implied message is the epitome of effective \textit{manipulation \& impression management}, \& it's amazing how many times it is successfully ``sold.'')
\end{itemize}
So, ``I don't know'' can mean any of the above \& a whole host of other things. But in the case of the disturbed character, it \textbf{\textit{rarely}}, truly means ``I don't know.'' So I no longer accept it for an answer. \& once I politely but firmly stopped taking it for an acceptable response, I immediately started getting explanations that made much more sense. Most importantly, the games of manipulation \& impression management my character-disordered clients tried to engage me in diminished dramatically.

Realizing how hoodwinked I had been by accepting their initial non-explanations for behavior helped me become dramatically aware of how expert disordered characters generally are on the subject of neurosis, as well as the mindsets of many mental health professionals, especially those steeped in traditional paradigms. They know very well how neurotics tend to think. They know the attitudes neurotics hold, \& the naiveties that make them vulnerable to tactics of manipulation \& impression management. They often know the neurotics in their lives better than those neurotics know themselves. They're also frequently quite ``couch broken'' (i.e. have made the rounds to many professionals \& become very familiar with psychological concepts, terms, \& paradigms); \& they know what to say that might be easily believed, especially if it's not scrutinized carefully. So, they can manipulate even a seasoned professional.

These days, I focus very little attention on the underpinnings of behavior, even though it's sometimes very helpful to know what all the nefarious motivations might be. I primarily assume that the disordered character is very aware -- not only of his behavior but also the reasons for it. So I direct attention mostly to the behavior itself with an emphasis on changing it.

\subsubsection{The role of feelings}
When a neurotic person seeks counseling for 1 reason to another, you can safely assume some emotional issues need to be attended to or resolved. Perhaps those feelings have been long repressed. Perhaps those feelings are very mixed \& conflicted. In any case, helping persons establish deeper contact with their feelings \& sort through their troubled emotions represents the hallmark of traditional psychotherapy. Regardless of the neurotic individual's problems, traditional psychotherapy almost always devotes considerable focus to the person's \textit{feelings}.
	
The disturbed character's problems with functioning well in a social context are not so much a consequence of the way he feels, but the way he \textit{thinks}. It's his ill-gotten \textit{attitudes}, erroneous \& distorted thinking patterns, \& dysfunctional core beliefs that lie at the root of problems. Awareness of this fact has in recent years spurred a revolution in therapeutic approaches.

The term \textit{cognitive-behavioral therapy} refers to an orientation founded on the principle of an inextricable relationship between a person's core beliefs, attitudes, \& thinking patterns \& that person's behavior. E.g., say a man's core beliefs include the view that any woman is naturally inferior to him, is designed by nature to be submissive, is a rightful personal possession if involved in a serious relationship with him, \& has value only as a sexual object or toy. One would not be particularly surprised to learn that this man had a history of abusive conduct with his wife or girlfriend. \textit{How we think in large measure determines how we will act}. When dealing with disordered characters, their kinds of problematic thinking become the bigger issue to address, as opposed to how they are feeling.

1 of the ways that folks become embroiled in abusive or exploitive relationships is by falling prey to concerns about the way their character-disordered partner is feeling. E.g., they might focus all too much on why their partner seems angry all the time, wondering what they might have done to engender such ire. They almost never consider that the brandishing of anger is sometimes a tactic that character-disturbed individuals use to manipulate \& control others, as opposed to a genuine feeling. What's more, they don't consider that it's their abuser's attitudes toward them, their value system, \& their distorted manner of thinking about things that constitutes the main problem, \& might be the precipitant of any unwarranted anger in the 1st place. If they manage to lure their partner into counseling, \& the counselor is of the traditional mindset \& primarily focuses on ``feelings,'' things are not likely to get much better. They \textit{may possibly get worse} as the non-character-disordered party bares his or her soul in the counseling process \& exposes even more of his or her vulnerability to the abusive party.

When I 1st began using more appropriate methods to counsel character-disordered persons, a frequent comment I heard from clients was, ``I don't think you really care.'' This kind of comment really put me on the defensive (as it would any good neurotic) initially. After all, I was a therapist, \& I was supposed to be a warm, empathic person dedicated to helping alleviate human misery. It took me awhile to realize that a comment like that was often a counter-therapeutic \& manipulative tactic on the disturbed characters' part. They wanted to get me to align with their point of view on things, to justify their conduct, \& most especially to see if they could convince me that they deserved just as much sympathy as those they had brought great pain. Of course I \textit{cared}. But I also came to know that if I were to \textit{care properly} \& \textit{therapeutically} for the disordered characters I was working with, I would have to confront the real impediments to their inner psychological health \& healthy interpersonal functioning. That meant calling them on their tactics \& letting them know that the genuineness of my ``caring'' would become all too evident in time. That would be through my unwavering dedication to challenging \& helping them correct the distorted thinking patterns that had made a mess of their lives \& the lives of others for too many years already. The door was always open to more traditional feelings-based counseling after character issues were resolved. Many did indeed walk through that door when the time came. That made me a believer in the notion that, when it comes to dealing with significant disturbances of character, focusing on feelings is not an exercise in ``caring,'' but rather a perfect example of \textit{enabling}.

In a later chapter, I will outline several of the major thinking errors disordered characters have in common, \& the dysfunctional social attitudes those erroneous ways of thinking breed. Knowing disturbed characters' frequent dysfunctional thinking patterns is crucial to understanding their problematic behavior, \& learning how to confront \& deal with them effectively.

\subsubsection{The role of defense mechanisms}
Neurotics are thought to use a variety of intra-psychic \textit{mechanisms} to \textit{defend} themselves against the experience of emotional pain, \& especially to alleviate anxiety associated with conflicts between their primal urges \& their consciences. Almost everyone has heard of these classic ``defense mechanisms.'' Such mechanisms, by definition, operate \textit{unconsciously}. I.e., the person doesn't deliberately engage in an action, but rather the unconscious mind employs the mechanism so the conscious mind never has to experience the pain in the 1st place. Here's the reason neurotic individuals develop problematic symptoms: These unconscious tools of anxiety mitigation, though powerful, are neither adequate, nor are they always fully adaptive ways to mitigate emotional pain. Many times, the symptoms the neurotic individual brings to the therapist's attention result from residual anxiety -- or the emotional pain left over after ineffective use of 1 or more of the typical defenses. At other times, people might seek help because their defenses have become increasingly inadequate or have begun to break down, letting their emotional pain underneath rise to the surface. In those cases, it's the emotional pain that brings someone into treatment. All successful therapies for neurosis depend upon building a trusting rapport with the therapist, so that an atmosphere of safety \& encouragement allows clients to relax their defenses \& reveal their inner pain or emotional conflicts.

Disordered characters engage in certain behaviors so ``automatic'' that it's tempting to think they do them unconsciously. On the surface, these behaviors often so resemble \textit{defense mechanisms} that they can be easily misinterpreted as such, especially by individuals overly immersed in traditional paradigms. However, on closer inspection, many of these behaviors are more accurately regarded as \textit{tactics} of \textit{manipulation, impression-management}, \& \textit{responsibility-resistance} (much more about this later).

In workshops, I always illustrate the contrast between a true ``defense'' mechanism \& a tactic of manipulation \& responsibility-avoidance using the concept of ``denial.'' 1 of the 10 most commonly misused terms in mental health (more about this later), denial can indeed be an unconscious defense mechanism. Let's take the example of a woman who has been married to the same man for 40 years. She has just rushed him to the hospital because, while they were out in the yard working, he began having trouble speaking \& looked in some distress. The doctors then tell her that he has suffered a stroke, is now virtually brain-dead, \& will not recover. Yet, every day she is by his bedside, holding his hand \& talking to him. The nurses tell her that he cannot hear, but she talks to him anyway. The doctors tell her he will not recover, but she only replies, ``I know he'll pull through, he's such a strong man.'' This woman is in a unique \textit{psychological state} -- the state of \textit{denial}. She can hardly believe what has happened. Not long ago she was in the yard with her darling, enjoying 1 of their favorite activities. The day before, they were at a friend's home for a get-together. He seemed the picture of happiness \& health. He didn't even seem that sick when she brought him to the hospital. Now -- in a blink of an eye -- they're telling her he's gone. This is far more emotional pain than she can bear just yet. She's not ready to accept that her partner of 40 years won't be coming home with her. She's not quite ready to face a life without him. So, her unconscious mind has provided her with an effective (albeit most likely temporary) \textit{defense} against the pain. Eventually, as she becomes better able to accept the distressing reality, her denial will break down. When it does, the pain it served to contain will gush forth \& she will grieve.

Now, let's take another example of so-called ``denial.'' Joe, the class bully, strolls up to 1 of his unsuspecting classmates \& engages in 1 of his favorite mischievous pastimes -- pushing the books out of her arms \& spilling them on the floor. It just so happens that the hall monitor catches the event \& sternly hollers: ``Joe!'' to which Joe, spreading his arms wide open \& with a look of great shock, surprise, \& innocence on his face retorts: ``Whaaaat?'' Does Joe really not understand the reality of what was happened? Does he actually think he didn't do what the hall monitor saw him do? Is he in some kind of altered psychological state? Is his possible altered state brought about by more emotional pain than he could possibly stand to bear? Is he so consumed with shame \&\texttt{/}or guilt for what he's done that he simply can't allow himself to believe he actually did such a horrible thing? More than likely, none of the aforementioned possibilities is correct. Joe is probably more concerned that he has another detention hall coming, which means another note to his parents, \& possibly even a suspension. So, he's got 1 long-shot \textit{tactic} to try. He'll do his best to make the hall monitor believe she didn't really see what she thought she saw. The hallway was crowded. Maybe it was someone else. Maybe it was just an ``accident.'' If he \textit{acts} surprised, innocent, \& righteously indignant enough, maybe, just maybe, she'll begin to doubt herself. He hopes that, unlike him, she might be just \textit{neurotic} enough (i.e. has an overactive conscience \& excessive sense of guilt or shame) to think she might have misjudged the situation. Maybe she'll even berate herself for jumping to conclusions or for causing a possibly innocent person unwarranted emotional pain. This tactic might have worked before. Maybe it will work again.

This preceding example is based on a real case. It is noteworthy that when ``Joe'' realized that he simply couldn't manipulate the hall monitor, he reluctantly stopped ``denying,'' saying: ``Well, maybe I did do it, but she had it coming because she's always bad-mouthing me to her friends.'' Now, we could engage in some discussion about the other tactics Joe is using to continue the game of manipulation \& impression management, but the most important thing to recognize is this: Unlike what happens in the case of real psychological denial as the defense mechanism, in Joe's case we don't see an outpouring of anguish \& grief when the denial ends. The reason is simple. Joe was never ``in denial'' (the psychological state) per se in the 1st place. He was simply \textit{lying}. He eventually stopped lying because it wasn't getting him anywhere. He moved on from the tactic of lying to excuse-making \& playing the victim, also effective tactics of manipulation \& impression-management.

I can't stress this enough: The ``denial'' of the unfortunate elderly woman mentioned above is nothing like the ``denial'' of Joe the school bully. One is truly an ego \textit{defense mechanism}; the other a \textit{manipulation \& responsibility-resistance tactic}. One is an unconscious mechanism of protection from deep emotional pain; the other is a deliberate, calculated lie. Yet many professionals (as well as lay persons) use the same term to describe these very different behaviors. \& because they presume there is only 1 type of denial, whenever their clients display behavior like Joe's, they regard it as a sign of shame-based, unconscious denial. I can't count the number of times clinicians have spoken to me of clients who are still ``in denial'' about 1 problem behavior or another, when what they were really describing was a client still ``lying \& manipulating'' as part of the game of impression management \& responsibility-resistance. As a result, they end up wasting precious time \& missing the mark in multiple sessions designed to help their clients ``come out of their denial.''

It's important for both lay persons \& clinicians not only to know the difference between a true defense mechanism \& a tactic of responsibility-avoidance \& interpersonal manipulation, but also to know how to appropriately respond to these very different types of behaviors. I presented the principal tactics manipulators use \& how to respond to them in \textit{In Sheep's Clothing}.\footnote{Simon, G., \textit{In Sheep's Clothing: Understanding and Dealing with Manipulative People}, (A.J. Christopher \& Co., 1996).} In a later chapter, we'll take a more in-depth look at those tactics, as well as other responsibility-avoidance \& impression-management maneuvers used by the various disordered characters.

\subsubsection{The genuineness of \textit{style}}
Some of the more prominent traditionally-oriented theorists have conceptualized the neurotic personality as essentially a fraud. The neurotic's true self was thought to be hidden behind a social fa\c{c}ade. So, a particularly gregarious person might be perceived in reality as quite shy \& interpersonally anxious ``underneath'' the social face he or she presents, \& to ``compensate'' for this social insecurity with feigned sociability. Similarly, bullies have been perceived as cowards underneath their brash exterior, \& haughty individuals viewed as compensating for low-self esteem. In short, their outward presentation is an unconsciously constructed ``front'' to mask their inner insecurities. Classical theorists also believed they could essentially define the principal personality types by the defense mechanisms they typically used to protect their true selves. Even relatively recently, this kind of conceptualization was articulated in David Shapiro's landmark work \textit{Neurotic Styles}.\footnote{Shapiro, D., \textit{Neurotic Styles}, (Basic Books, 1999).} He eloquently describes the various personality ``styles,'' but still views them as essentially an expression of a person's neurosis.

Have you ever noticed the consistency in these traditional notions about a very different kind of reality lying underneath the fa\c{c}ade? They always involve an outward appearance that's not very appealing, \& a more pitiable reality underneath. In other words, traditional notions about personality tend to view egomaniacs as really having low self-esteem underneath, bullies being scared little kids underneath, \& abusers being traumatized victims underneath it all, etc. But you've probably never heard a devotee of classical perspectives claim that a shy person is really a ravenous animal underneath it all, wanting to jump the bones of everyone they meet. Or that a particularly sensitive person is really a vicious monster with a heart of stone underneath. I think we're too quick to align with psychology metaphors that have outlined their usefulness for this reason: Most of us still don't like to face the unpleasant things in life, \& want to explain them away with a perspective that makes the unnerving more palatable. After all, genuine \textit{denial} is 1 of the things neurotics do best!

While it might be true to some extent that a neurotic is quite different under the exterior, with the disturbed character, \textbf{\textit{what you see is what you get}}, unless a deliberate con game is being played. W.r.t. primary personality traits, there is no pretense. Disordered characters are who they are, as unfortunate as that may be, \& often to the core.

I once worked in a residential treatment program that specialized in young persons already displaying significant disturbances of character \& conduct. 1 day, a young man was admitted who, within minutes of arrival, began listing on a notepad program improvements he thought the staff needed to make. He wanted to present this list to the facility administrator \& demanded an audience to discuss matters. Steeped in traditional psychological theories, the head nurse -- when this young man's treatment plan was fashioned -- recalled his haughtiness, \& proposed a 1st treatment goal of increasing his sense of self-esteem. She assumed, as is common to neurotics \& devotees of traditional perspectives, that his pompous attitude simply \textit{must} have been a \textit{compensation} for underlying feelings of inferiority. But in time, it became quite apparent that this young man in fact had no feelings of inferiority. Rather, he possessed only a deeply-rooted sense of superiority \& \textit{entitlement} common to individuals over-indulged \& over-valued all their lives -- people who end up as deeply disturbed characters.

\subsubsection{Self-esteem}
The discussion above alludes to another very key difference between neurotics \& disordered characters. Neurotics have significantly damaged self-images, typically arising out of a deflated sense of self-worth or self-esteem. The neurotic often feels defective in some way, \& therefore not truly lovable. Sometimes this impaired self-esteem can lead to a profound sense of inadequacy \& serious weaknesses of character. Sometimes it prompts the neurotic to try hard to please others, \& to do their very best to ``earn'' approval. If not excessive, this tendency can provide a critical incentive to behave in socially acceptable or responsible ways. In other words, when people think their self-worth largely depends on how clearly they demonstrate their value to society, they can be motivated to conduct themselves in a manner that increases the likelihood of benefiting the greater good.

Disordered characters have an inflated sense of self-worth. They see themselves as superior to others. As a result, they often feel \textit{entitled} to use \& exploit others as they see fit. As noted earlier, their ego-inflation is not rooted in underlying feelings of insecurity or inferiority. It's not a pretense. They really do think they're something special \& above the common throng. \&, their sense of superiority fuels their attitudes of entitlement. Later in this book, we'll take a deeper look at 1 type of character disorder, the psychopath, in which this disturbing trait is present to its most pathological extreme.

In a later chapter, I'll also be introducing a critical distinction between the concepts of self-esteem \& self-respect. This distinction will hopefully not only give some additional clarity to self-esteem's definition, but also will help explain why there's often such confusion \& misunderstanding about how \& why some people's self-image gets bent out of shape.

\subsubsection{Response to adverse consequence}
Neurotics try very hard to effect positive outcomes, \& they easily become anxious \& upset when things go badly. They are \textit{hypersensitive to adverse consequence}. A co-occurring trait accompanying this hypersensitivity is the tendency to make \textit{internal attributions} about the reasons for problems. When a neurotic worker doesn't get the ``good job'' comment she craves from her boss, she might well beat herself up with self-criticism. When the neurotic therapist doesn't see positive change in her therapy group, she might well berate herself as a sub-standard counselor. Neurotics always want things to be right with the world, \& they get internally unnerved when things go wrong. They want the world to be at peace, for everyone to love them, \& for everyone's dreams to come true. When things go badly, they take it hard. They're unnerved by the cruelty \& unfairness of life, \& feeling inordinately responsible, they take it upon themselves to make things better. When things go poorly, they chide themselves, question themselves, conduct an internal debate about how they might do thing differently, \& try all the harder to make things right. Even when it's fairly impossible to blame themselves for a painful happenstance (e.g., a natural disaster or catastrophe), they still question \& prod themselves about what they can or should do to make the situation better. \& they're deeply affected by any unintended adverse consequences of their own actions, \& strive to make amends.

Disordered characters are largely unaffected \& \textit{undeterred by adverse consequence}. They have a characteristic imperturbability when it comes to dealing with the negative fallout of their behavior. They are typically not unnerved by situations that would make the neurotic upset. More importantly, they remain undeterred about their basic way of handling things. They might even be strengthened in their resolve to keep doing just as they have been doing, despite the objective fact that their way is clearly not working.

A co-occurring character trait is the disordered character's tendency to make \textit{external attributions} whenever anything bad does happen. They see others \& circumstances outside themselves as the source of problems. So, even if they've lost another job, had another marriage fall apart, run into legal difficulties, or even lost their freedom (if they have become incarcerated), they take it in stride. They never blame themselves, \& keep on behaving the same way they've always behaved, despite where it's gotten them. They even pride themselves in the notion that they will remain who they are, doing things as they've always done, despite the hardships in their lives, including those often resulting from their own behavior.

\subsubsection{Level of internal discomfort}
Neurotics are also different from character disorders on another very key dimension. In large measure, they experience the signs \& symptoms of their neuroses as unpleasant \& unwanted. Perhaps an individual has been worrying to the point that he's developed an ulcer, \& is now in frequent gastric distress. Perhaps this circumstance has slowed him down at work, making him less productive. He still drives hard like always, but his frequent bouts of pain have stymied his usual effectiveness. It's very likely that he does not like the person he has become as a result. Clinicians say that it is \textit{ego-dystonic} (unpalatable to his image of himself) for him not to be performing at his best. What's more, he doesn't like his symptoms either (i.e. the pain of his ulcer). As a result, he will likely take the initiative to get some help. He wants the problem gone, to be out of pain, to be at his best again; \& he'll probably do what he has to do to make things better again. If he seeks help from a therapist, he'll even be open to changing some things about himself so he'll be a better person \& have fewer disturbing symptoms. He'll likely be motivated to at least try what the therapist suggests, to regard the therapist's guidance as valuable as he begins to gain relief, \& to remain in treatment until he has overcome his problem \& doesn't need therapy services anymore.

Disturbed characters also display telltale signs \& symptoms of their disorder. Lying, conning, manipulating, defaulting on social obligations, etc., are several of the disordered character's defining features. The negative attitudes they hold, the distorted way they tend to think, \& the irresponsible ways they tend to behave are likely to be greatly upsetting to others. But these things are what clinicians call \textit{ego-syntonic} to the disordered character. I.e., the disordered character doesn't see anything wrong or disturbing about them. Moreover, he is not upset by the kind of person he has become as a result of these characteristics. He likes who he is \& how he operates. Others may complain that he tends to use \& exploit people. His answer might be that gullible or weaker people deserve to be taken advantage of. As he sees it, if others have a problem with him, it's because \textit{they} are all screwed up. When someone points out 1 of his most disturbing characteristics, he might retort: ``You got a problem with that?'' Disturbed characters also don't seek help in the manner neurotics do. Rather, they're more dragged into counseling by distraught neurotics who are in some kind of relationship with them. Disturbed characters are rarely in the kind of inner distress that prompts most people to seek \& appreciate guidance or counseling on their own.

\subsubsection{Needs in treatment}
Another important difference between neurotics \& disordered characters is what the need most from any kind of therapeutic experience. This is, perhaps, 1 of the most important points I need to make for the benefit of therapists. Because neurotics are struggling with inadequacies, insecurities, \& emotional conflicts, they both crave \& need positive regard as well as supportive guidance. In short, they need \textit{help}. Also, because many of their unresolved issues are mostly rooted in the unconscious, neurotics both need \& benefit from \textit{insight}. They literally don't know what they're doing to perpetuate their difficulties. So they benefit greatly from listening to their counselors \textit{interpret} the ``dynamics'' of their circumstances, thus shedding ``new light'' on their situation. Because they were largely unable to come to such insights on their own, they not only seek help but also appreciate it when they get it in the forms of new insights, emotional support, \& guidance.

In contrast to insight-deficient neurotics, disordered characters are already keenly aware of the ways their thinking \& behavior cause problems. There isn't 1 thing anyone can say or bring to their attention that they haven't heard a thousand times before from a variety of sources, or experienced in a variety of circumstances. In many of my workshops, I introduce the trite little saying: ``They already \textit{see}; they simply \textit{disagree}.'' They're just not disturbed enough by their way of doing things; or they may have been successful enough getting their way by doing those things, so they're resolved not to change that modus operandi. Because they do the things they do so automatically \& habitually, it can seem like they're unaware.

'' -- \cite[p. 27--]{Simon2011}

%------------------------------------------------------------------------------%

\section{Major Disturbances of Personality \& Character}

%------------------------------------------------------------------------------%

\section{The Aggressive Pattern}

%------------------------------------------------------------------------------%

\section{The Process of Character Development}

%------------------------------------------------------------------------------%

\section{Thinking Patterns \& Attitudes Predisposing Character Disturbance}

%------------------------------------------------------------------------------%

\section{Habitual Behavior Patterns Fostering \& Perpetuating Character Disturbance}

%------------------------------------------------------------------------------%

\section{Engaging Effectively \& Intervening Therapeutically with Disturbed Characters}

%------------------------------------------------------------------------------%

\section{Epilogue: Neurosis \& Character Disturbance}

%------------------------------------------------------------------------------%

\section*{Endnotes}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
	
\end{document}