\documentclass{article}
\usepackage[backend=biber,natbib=true,style=alphabetic]{biblatex}
\addbibresource{/home/nqbh/reference/bib.bib}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=1cm,right=1cm,top=5mm,bottom=5mm,footskip=4mm]{geometry}
\def\labelitemii{$\circ$}

\title{On Writing: A Memoir of the Craft}
\author{Stephen King}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\vspace{5mm}
\begin{quotation}
	\textit{``Honesty's the best policy.''} -- Miguel de Cervantes
	
	\textit{``Liars prosper.''} -- Anonymous
\end{quotation}

%------------------------------------------------------------------------------%

\section*{1st Foreword}
``In the early 90s (it might have been 1992, but it's hard to remember when you're having a good time) I joined a rock-\&-roll band composed mostly of writers. The Rock Bottom Remainders were the brainchild of Kathi Kamen Goldmark, a book publicist \& musician from San Francisco. The group included Dave Barry on lead guitar, Ridley Pearson on bass, Barbara Kingsolver on keyboards, Robert Fulghum on mandolin, \& me on rhythm guitar. There was also a trio of ``chick singer,'' \textit{\`a la} the Dixie cups, made up (usually) of Kathi, Tad, Bartimus, \& Amy Tan.

The group was intended as a 1-shot deal -- we would play 2 shows at the American Booksellers Convention, get a few laughs, recapture our misspent youth for 3 or 4 hours, then go our separate ways.

It didn't happen that way, because the group never quite broke up. We found that we liked playing together too much to quit, \& with a couple of ``ringer'' musicians on sax \& drums (plus, in the early days, our musical guru, Al Kooper, at the heart of the group), we sounded pretty good. You'd pay to hear us. Not a lot, not U2 or E Street Bans prices, but maybe what the oldtimers call ``roadhouse money.'' We took the group on tour, wrote a book about it (my wife took the photos \& danced whenever the spirit took her, which was quite often), \& continue to play now \& then, sometimes as The Remainders, sometimes as Raymond Burr's Legs. The personnel comes \& goes -- columnist Mitch Albom has replaced Barbara on keyboards, \& Al doesn't play with the group anymore `cause he \& Kathi don't get along -- but the core has remained Kathi, Amy, Ridley, Dave, Mitch Albom, \& me $\ldots$ plus Josh Kelly on drums \& Erasmo Paolo on sax.

We do it for the music, but we also do it for the companionship. We like each other, \& we like having a chance to talk sometimes about the real job, the day job people are always telling us not to quit. We are writers, \& we never ask one another where we get our ideas; we know we don't know.

1 night while we were eating Chinese before a big in Miami Beach, I asked Amy if there was any 1 question she was \textit{never} asked during the Q-\&-A that follows almost every writer's talk -- that question you never get to answer when you're standing in front of a group of author-struck fans \& pretending you don't put your pants on 1 leg at a time like everyone else. Amy paused, thinking it over very carefully, \& then said: ``No one ever asks about the language.''

I owe an immense debt of gratitude to her for saying that. I had been playing with the idea of writing a little book about writing for a year or more at that time, but had held back because I didn't trust my own motivations -- \textit{why} did I want to write about writing? What made me think I had anything worth saying?

The easy answer is that someone who has sold as many books of fiction as I have must have \textit{something} worthwhile to say about writing it, but the easy answer isn't always the truth. Colonel Sanders sold a hell of a lot of fried chicken, but I'm not sure anyone wants to know how he made it. If I was going to be presumptuous enough to tell people how to write, I felt there had to be a better reason than my popular success. Put another way, I didn't want to write a book, even a short one like this, that would leave me feeling like either a literary gasbag or a transcendental asshole. There are enough of those books -- \& those writers -- on the market already, thanks.

But Amy was right: nobody ever asks about the language. They ask the DeLillos \& the Updikes \& the Styrons, but they don't ask popular novelists. Yet many of us proles also care about the language, in our humble way, \& care passionately about the art \& craft of telling stories on paper. What follows is an attempt to put down, briefly \& simply, how I came to the craft, what I know about it now, \& how it's done. It's about the day job; it's about the language.

This book is dedicated to Amy Tan, who told me in a very simple \& direct way that it was okay to write it.'' -- \cite[p. 9]{King2010}

%------------------------------------------------------------------------------%

\section*{2nd Foreword}
``This is a short book because most books about writing are filled with bullshit. Fiction writers, present company included, don't understand very much about what they do -- not why it works when it's good, not why it doesn't when it's bad. I figured the shorter the book, the less the bullshit.

1 notable exception to the bullshit rule is \textit{The Elements of Style}, by William Strunk Jr. \& E. B. White. There is little or no detectable bullshit in that book. (Of course it's short; at 85 pages it's much shorter than this one.) I'll tell you right now that every aspiring writer should read \textit{The Elements of Style}. Rule 17 in the chapter titled Principles of Composition is ``Omit needless words.'' I will try to do that here.'' -- \cite[p. 10]{King2010}

%------------------------------------------------------------------------------%

\section*{3rd Foreword}
``1 rule of the road not directly stated elsewhere in this book: ``The editor is always right.'' The corollary is that no writer will take all of his or her editor's advice; for all have sinned \& fallen short of editorial perfection. Put another way, to write is human, to edit is divine. Chuck Verrill edited this book, as he has so many of my novels. And as usual, Chuck, you were divine. -- Steve'' -- \cite[p. 11]{King2010}

%------------------------------------------------------------------------------%

\section{C.V.}
``I was stunned by Mary Karr's memoir, \textit{The Liar's Club}. Not just by its ferocity, its beauty, \& by her delightful grasp of the vernacular, but by its \textit{totality} -- she is a woman who remembers \textit{everything} about her early years.

I'm not that way. I lived an odd, herky-jerky childhood, raised by a single parent who moved around a lot in my earliest years \& who -- I am not completely sure of this -- may have farmed my brother \& me out to 1 of her sisters for awhile because she was economically or emotionally unable to cope with us for a time. Perhaps she was only chasing our father, who piled up all sorts of bills \& then did a runout when I was 2 \& my brother David was 4. If so, she never succeeded in finding him. My mom, Nellie Ruth Pillsbury King, was 1 of America's early liberated women, but not by choice.

Mary Karr presents her childhood in an almost unbroken panorama. Mine is a fogged-out landscape from which occasional memories appear like isolated trees $\ldots$ the kind that look as if they might like to grab \& eat you.

What follows are some of those memories, plus assorted snapshots from the somewhat more coherent days of my adolescence \& young manhood. This is not an autobiography. It is, rather, a kind of \textit{curriculum vitae} -- my attempt to show ow 1 writer was formed. Not how 1 writer was \textit{made}; I don't believe writers \textit{can} be made, either by circumstances or by self-will (although I did believe those things once). The equipment comes with the original package. Yet it is by no means unusual equipment; I believe large numbers of people have at least some talent as writers \& storytellers, \& that those talents can be strengthened \& sharpened. If I didn't believe that, writing a book like this would be a waste of time.

This is how it was for me, that's all -- a disjointed growth process in which ambition, desire, luck, \& a little talent all played a part. Don't bother trying to read between the lines, \& don't look for a through-line. There are \textit{no} lines -- only snapshots, most out of focus.

\fbox{\bf1} My earliest memory is of imaging I was someone else -- imagining that I was, in fact, the Ringing Brothers Circus Strongboy. This was at my Aunt Ethelyn \& Uncle Oren's house in Durham, Maine. My aunt remembers this quite clearly, \& says I was 2 \& a half or maybe 3 years old.

I had found a cement cinderblock in a corner of the garage \& had managed to pick it up. I carried it slowly across the garage's smooth cement floor, except in my mind I was dressed in an animal skin singlet (probably a leopard skin) \& carrying the cinderblock across the center ring. The vast crowd was silent. A brilliant bluewhite spotlight marked my remarkable progress. Their wondering faces told the story: never had they seen such an incredibly strong kid. ``\& he's only \textit{2}!'' someone muttered in disbelief.

Unknown to me, wasps had constructed a small nest in the lower half of the cinderblock. 1 of them, perhaps pissed off at being relocated, flew out \& stung me on the ear. The pain was brilliant, like a poisonous inspiration. It was the worst pain I had ever suffered in my short life, but it only held the top spot for a few seconds. When I dropped the cinderblock on 1 bare foot, mashing all 5 toes, I forgot all about the wasp. I can't remember if I was taken to the doctor, \& neither can my Aunt Ethelyn (Uncle Oren, to whom the Evil Cinderblock surely belonged, is almost 20 years dead), but she remembers the sting, the mashed toes, \& my reaction. ``How you howled, Stephen!'' she said. ``You were certainly in fine voice that day.''

\fbox{\bf2} A year or so later, my mother, my brother, \& I were in West De Pere, Wisconsin. I don't know why. Another of my mother's sisters, Cal (a WAAC beauty queen during World War II), lived in Wisconsin with her convivial beer-drinking husband, \& maybe Mom had moved to be near them. If so, I don't remember seeing much of the Weimers. \textit{Any} of them, actually. My mother was working, but I can't remember what her job was, either. I want to say it was a bakery she worked in, but I think that came later, when we moved to Connecticut to live near her sister Lois \& \textit{her} husband (no beer for Fred, \& not much in the way of conviviality, either; he was a crewcut daddy who was proud of driving his convertible with the top \textit{up}, God knows why).

There was a stream of babysitters during our Wisconsin period. I don't know if they left because David \& I were a handful, or because they found better-paying jobs, or because my mother insisted on higher standards than they were willing to rise to; all I know is that there were a lot of them. The only one I remember with any clarity is Eula, or maybe she was Beulah. She was a teenager, she was as big as a house, \& she laughed a lot. Eula-Beulah had a wonderful sense of humor, even at 4 I could recognize that, but it was a \textit{dangerous} sense of humor -- there seemed to be a potential thunderclap hidden inside each hand-patting, butt-rocking, head-tossing outburst of glee. When I see those hidden-camera sequences where real-life babysitters \& nannies just all of a sudden wind up \& clout the kids, it's my days with Eula-Beulah I always think of.

Was she as hard on my brother David as she was on me? I don't know. He's not in any of these pictures. Besides, he would have been less at risk from Hurricane Eula-Beulah's dangerous winds; at 6, he would have been in the 1st grade \& off the gunnery range for most of the day.

Eula-Beulah would be on the phone, laughing with someone, \& beckon me over. She would hug me, tickle me, get me laughing, \& then, still laughing, go upside my head enough to knock me down. Then she would tickle me with her bare feet until we were both laughing again.

Eula-Beulah was prone to farts -- the kind that are both loud \& smelly. Sometimes when she was so afflicted, she would throw me on the couch, drop her wool-skirted butt on my face, \& let loose. ``Pow!'' she'd cry in high glee. It was like being buried in marshgas fireworks. I remember the dark, the sense that I was suffocating, \& I remember laughing. Because, while what was happening was sort of horrible, it was also sort of funny. In many ways, Eula-Beulah prepared me for literary criticism. After having a 200-pound babysitter fart on your face \& yell \textit{Pow!, The Village Voice} holds few terrors.

I don't know what happened to the other sitters, but Eula-Beulah was fired. It was because of the eggs. 1 morning Eula-Beulah fried me an egg for breakfast. I ate it \& asked for another one. Eula-Beulah fried me a 2nd egg, then asked if I wanted another one. \& another one. \& so on. I stopped after 7, I think -- 7 is the number that sticks in my mind, \& quite clearly. Maybe we ran out of eggs. Maybe I cried off. Or maybe Eula-Beulah got scared. I don't know, but probably it was good that the game ended at 7. 7 eggs is quite a few for a 4-year-old.

I felt all right for awhile, \& then I yarked all over the floor. Eula-Beulah laughed, then went upside my head, then shoved me into the closet \& locked the door. Pow. If she'd locked me in the bathroom, she might saved her job, but she didn't. As for me, I didn't really mind being in the closet. It was dark, but it smelled of my mother's Coty perfume, \& there was a comforting line of light under the door.

I crawled to the back of the closet, Mom's coats \& dresses brushing along my back. I began to belch -- long loud belches that burned like fire. I don't remember being sick to my stomach but I must have been, because when I opened my mouth to let out another burning belch, I yarked again instead. All over my mother's shoes. That was the end for Eula-Beulah. When my mother came home from work that day, the babysitter was fast asleep on the couch \& little Stevie was locked in the closet, fast asleep with half-digested fried eggs drying in his hair.

\fbox{\bf3} Our stay in West De Pere was neither long nor successful. We were evicted from our 3rd-floor apartment when a neighbor spotted my 6-year-old brother crawling around on the roof \& called the police. I don't know where my mother was when this happened. I don't know where the babysitter of the week was, either. I only know that I was in the bathroom, standing with my bare feet on the heater, watching to see if my brother would fall off the roof or make it back into the bathroom okay. He made it back. He is now 55 \& living in New Hampshire.

\fbox{\bf4} When I was 5 or 6, I asked my mother if she had ever seen anyone die. Yes, she said, she had been 1 person die \& had heard another one. I asked how you could hear a person die \& she told me that it was a girl who had drowned off Prout's Neck in the 1920s. She said the girl swam out past the rip, couldn't get back in, \& began screaming for help. Several men tried to reach her, but that day's rip had developed a vicious undertow, \& they were all forced back. In the end they could only stand around, tourists \& townies, the teenager who became my mother among them, waiting for a rescue boat that never came \& listening to that girl scream until her strength gave out \& she went under. Her body washed up in New Hampshire, my mother said. I asked how old the girl was. Mom said she was 14, then read me a comic book \& packed me off to bed. On some other day she told me about the one she saw -- a sailor who jumped off the roof of the Graymore Hotel in Portland, Maine, \& landed in the street.

``He splattered,'' my mother said in her most matter-of-fact tone. She paused, \& added, ``The stuff that came out of him was green. I have never forgotten it.''

That makes 2 of us, Mom.

\fbox{\bf5} Most of the 9 months I should have spent in the 1st grade I spent in bed. My problems started with the measles -- a perfectly ordinary case -- \& then got steadily worse. I had bout after bout of what I mistakenly thought was called ``stripe throat''; I lay in bed drinking cold water \& imagining my throat in alternating stripes of red \& white (this was probably not so far wrong).

At some point my ears become involved, \& 1 day my mother called a taxi (she did not drive) \& took me to a doctor too important to make house calls -- an ear specialist. (For some reason I got the idea that this sort of doctor was called an otiologist.) I didn't care whether he specialized in ears or assholes. I had a fever of 104 degrees, \& each time I swallowed, pain lit up the sides of my face like a jukebox.

The doctor looked in my ears, spending most of his time (I think) on the left one. Then he laid me down on his examining table. ``Lift up a minute, Stevie,'' his nurse said, \& put a large absorbent cloth -- it might have been a diaper -- under my head, so that my cheek rested on it when I lay back down. I should have guessed that something was rotten in Denmark. Who knows, maybe I did.

There was a sharp smell of alcohol. A clank as the ear doctor opened his sterilizer. I saw the needle in his hand -- it looked as long as the ruler in my school pencil-box -- \& tensed. The ear doctor smiled reassuringly \& spoke the lie for which doctors should be immediately jailed (time of incarceration to be doubled when the lie is told to a child): ``Relax, Stevie, this won't hurt.'' I believed him.

He slid the needle into my ear \& punctured my eardrum with it. The pain was beyond anything I have ever felt since -- the only thing close was the 1st month of recovery after being struck by a van in the summer 1999. That pain was longer in duration but not so intense. The puncturing of my eardrum was pain beyond the world. I screamed. There was a sound inside my head -- a loud kissing sound. Hot fluid ran out of my ear -- it was as if I had started to cry out of the wrong hole. God knows I was crying enough out of the right ones by then. I raised my streaming face \& looked unbelieving at the ear doctor \& the ear doctor's nurse. Then I looked at the cloth the nurse had spread over the top 3rd of the exam table. It had a big wet patch on it. There were fine tendrils of yellow pus on it as well.

``There,'' the ear doctor said, patting my shoulder. ``You were very brave, Stevie, \& it's all over.''

The next week my mother called another taxi, we went back to the ear doctor's, \& I found myself once more lying on my side with the absorbent square of cloth under my head. The ear doctor once again produced the smell of alcohol -- a smell I still associate, as I suppose many people do, with pain \& sickness \& terror -- \& with it, the long needle. He once more assured me that it wouldn't hurt, \& I once more believed him. Not completely, but enough to be quiet while the needle slid into my ear.

It \textit{did} hurt. Almost as much as the 1st time, in fact. The smoothing sound in my head was louder, too; this time it was giants kissing (``suckin' face \& rotatin' tongues,'' as we used to say). ``There,'' the ear doctor's nurse said when it was over \& I lay there crying in a puddle of watery pus. ``It only hurts a little, \& you don't want to be deaf, do you? Besides, it's all over.''

I believed that for about 5 days, \& then another taxi came. We went back to the ear doctor's. I remember the cab driver telling my mother that he was going to pull over \& let us out if she couldn't shut that kid up.

Once again it was me on the exam table with the diaper under my head \& my mom out in the waiting room with a magazine she was probably incapable of reading (or so I like to imagine), Once again the pungent smell of alcohol \& the doctor turning to me with a needle that looked as long as my school ruler. Once more the smile, the approach, the assurance that \textit{this} time it wouldn't hurt.

Since the repeated eardrum-lancings when I was 6, 1 of my life's firmest principles has been this: Fool me once, shame on you. Fool me twice, shame on me. Fool me 3 times, shame on both of us. The 3rd time on the ear doctor's table I struggled \& screamed \& thrashed \& fought. Each time the needle came near the side of my face, I knocked it away. Finally the nurse called my mother in from the waiting room, \& the 2 of them managed to hold me long enough for the doctor to get his needle in. I screamed so long \& so loud that I can still hear it. In fact, I think that in some deep valley of my head that last scream is still echoing.

\fbox{\bf6} In a dull cold month not too long after that -- it would have been Jan or Feb of 1954, if I've got the sequence right -- the taxi came again. This time the specialist wasn't the ear doctor but a throat doctor. Once again my mother sat in the waiting room, once again I sat on the examining table with a nurse hovering nearby, \& once again there was that sharp smell of alcohol, an aroma that still has the power to double my heartbeat in the space of 5 seconds.

All that appeared this time, however, was some sort of throat swab. It stung, \& it tasted awful, but after the ear doctor's long needle it was a walk in the park. The throat doctor donned an interesting gadget that went around his head on a strap. It had a mirror in the middle, \& a bright fierce light that shone out of it like a 3rd eye. He looked down my gullet for a long time, urging me to open wider until my jaws creaked, but he did not put needles into me \& so I loved him. After awhile he allowed me to close my mouth \& summoned my mother.

``The problem is his tonsils,'' the doctor said. ``They look like a cat clawed them. They'll have to come out.''

At some point after that, I remember being wheeled under bright lights. A man in a white mask bent over me. He was standing at the head of the table I was lying on (1953 \& 1954 were my years for lying on tables), \& to me he looked upside down.

``Stephen,'' he said. ``Can you hear me?''

I said I could.

``I want you to breathe deep,'' he said. ``When you wake up, you can have all the ice cream you want.''

He lowered a gadget over my face. In the eye of my memory, it looks like an outboard motor. I took a deep breath, \& everything went black. When I woke up I was indeed allowed all the ice cream I wanted, which was a fine joke on me because I didn't want any. My throat felt swollen \& fat. But it was better than the old needle-in-the-ear trick. Oh yes. \textit{Anything} would have been better than the old needle-in-the-ear trick. Take my tonsils if you have to, put a steel birdcage on my leg if you must, but God save me from the otiologist.

\fbox{\bf7} That year my brother David jumped ahead to the 4th grade \& I was pulled out of school entirely. I had missed too much of the 1st grade, my mother \& the school agreed; I could start it fresh in the fall of the year, if my health was good.

Most of that year I spent either in bed or housebound. I read my way through approximately 6 tons of comic books, progressed to Tom Swift \& Dave Dawson (a heroic World War II pilot whose various planes were always ``prop-clawing for altitude''), then moved on to Jack London's bloodcurdling animal tales. At some point I began to write my own stories. Imitation preceded creation; I would copy \textit{Combat Casey} comics word for word in my Blue Horse tablet, sometimes adding my own descriptions where they seemed appropriate. ``They were camped in a big dratty farmhouse room,'' I might write; it was another year or 2 before I discovered that \textit{drat} \& \textit{draft} were different words. During that same period I remember believing that \textit{details} were \textit{dentals} \& that a bitch was an extremely tall woman. A son of a bitch was apt to be a basketball player. When you're 6, most of your Bingo balls are still floating around in the draw-tank.

Eventually I showed 1 of these copycat hybrids to my mother, \& she was charmed -- I remember he slightly amazed smile, as if she was unable to believe a kid of hers could be so smart -- practically a damned prodigy, for God's sake. I had never seen that look on her face before -- not on my account, anyway -- \& I absolutely loved it.

She asked me if I had made the story up myself, \& I was forced to admit that I had copied most of it out of a funnybook. She seemed disappointed, \& that drained away much of my pleasure. At last she handed back my tablet. ``Write 1 of your own, Stevie,'' she said. ``Those \textit{Combat Casey} funny-books are just junk -- he's always knocking someone's teeth out. I bet you could do better. Write 1 of your own.''

I remember an immense feeling of \textit{possibility} at the idea, as if I had been ushered into a vast building filled with closed doors \& had been given leave to open any I liked. There were more doors than 1 person could ever open in a lifetime, I thought (\& still think).

I eventually wrote a story about 4 magic animals who rode around in an old car, helping out little kids. Their leader was a large white bunny named Mr. Rabbit Trick. He got to drive the car. The story was 4 pages long, laboriously printed in pencil. No one in it, so far as I can remember, jumped from the roof of the Graymore Hotel. When I finished, I gave it to my mother, who sat down in the living room, put her pocketbook on the floor beside her, \& read it all at once. I could tell she liked it -- she laughed in all the right places -- but I couldn't tell if that was because she liked me \& wanted me to feel good or because it really \textit{was} good.

``You didn't copy this one?'' she asked when she had finished. I said no, I hadn't. She said it was good enough to be in a book. Nothing anyone has said to me since has made me feel any happier. I wrote 4 more stories about Mr. Rabbit Trick \& his friends. She gave me a quarter apiece for them \& sent them around to her 4 sisters, who pitied her a little, I think. \textit{They} were all still married, after all; their men had stuck. It was true that Uncle Fred didn't have much sense of humor \& was stubborn about keeping the top of his convertible up, it was also true that Uncle Oren drank quite a bit \& had dark theories about how the Jews were running the world, but they were \textit{there}. Ruth, on the other hand, had been left holding the baby when Don ran out. She wanted them to see that he was a talented baby, at least.

4 stories. A quarter apiece. That was the 1st buck I made in this business.

\fbox{\bf9} We moved to Stratford, Connecticut. By then I was in the 2nd grade \& stone in love with the pretty teenage girl who lived next door. She never looked twice at me in the daytime, but at night, as I lay in bed \& drifted toward sleep, we ran away from the cruel world of reality again \& again. My new teacher was Mrs. Taylor, a kind lady with gray Elsa Lanchester -- \textit{Bride of Frankenstein} hair \& protruding eyes. ``When we're talking I always want to cup my hands under Mrs. Taylor's peepers in case they fall out,'' my mom said.

Our new 3rd-floor apartment was on West Broad Street. A block down the hill, not far from Teddy's Market \& across from Burrets Building Materials, was a huge tangled wilderness area with a junkyard on the far side \& a train track running through the middle. This is 1 of the places I keep returning to in my imagination; it turns up in my books \& stories again \& again, under a variety of names. The kids in \textit{It} called it the Barrens; we called it the jungle. Dave \& I explored it for the 1st time not long after we had moved into our new place. It was summer. It was hot. It was great. We were deep into the green mysteries of this cool new playground when I was struck by an urgent need to move my bowels.

``Dave,'' I said. ``Take me home! I have to push!'' (This was the word we were given for this particular function.)

David didn't want to hear it. ``Go do it in the woods,'' he said. It would take at least half an hour to walk me home, \& he had no intention of giving up such a shining stretch of time just because his little brother had to take a dump.

``I can't!'' I said, shocked by the idea. ``I won't be able to wipe!''

``Sure you will,'' Dave said. ``Wipe yourself with some leaves. That's how the cowboys \& Indians did it.''

By then it was probably too late to get home, anyway; I have an idea I was out of options. Besides, I was enchanted by the idea of shitting like a cowboy. I pretended I was Hopalong Cassidy, squatting in the underbrush with my gun drawn, not to be caught unawares even at such a personal moment. I did my business, \& took care of the cleanup as my older brother had suggested, carefully wiping my ass with big handfuls of shiny green leaves. These turned out to be poison ivy.

2 days later I was bright red from the backs of my knees to my shoulderblades. My penis was spared, but my testicles turned into stoplights. My ass itched all the way up to my ribcage, it seemed. Yet most of all was the hand I had wiped with; it swelled to the size of Mickey Mouse's after Donald Duck had bopped it with a hammer, \& gigantic blisters formed at the places where the fingers rubbed together. When they burst they left deep divots of raw pink flesh. For 6 weeks I sat in lukewarm starch baths, feeling miserable \& humiliated \& stupid, listening through the open door as my mother \& brother laughed \& listened to Peter Tripp's countdown on the radio \& played Crazy 8s.

\fbox{\bf 10} Dave was a great brother, but too smart for a 10-year-old. His brains were always getting him in trouble, \& he learned at some point (probably after I had wiped my ass with poison ivy) that it was usually possible to get Brother Stevie to join him in the point position when trouble was in the wind. Dave never asked me to shoulder \textit{all} the blame for his often brilliant fuck-ups -- he was neither a sneak nor a coward -- but on several occasions I was asked to share it. Which was, I think, why we both got in trouble when Dave dammed up the stream running through the jungle \& flooded much of lower West Broad Street. Sharing the blame was also the reason we both ran the risk of getting killed while implementing his potentially lethal school science project.

This was probably 1958. I was at Center Grammar School; Dave was at Stratford Junior High. Mom was working at the Stratford Laundry, where she was the only white lady on the mangle crew. That's what she was doing -- feeling sheets into the mangle -- while Dave constructed his Science Fair project. My big brother wasn't the sort of boy to content himself drawing frog-diagrams on construction paper or making The House of the Future out of plastic Tyco bricks \& pained toilet-tissue rolls; Dave aimed for the stars. His project that year was Dave's Super Duper Electromagnet. My brother had great affection for things which were super duper \& things which began with his own name; this latter habit culminated with \textit{Dave's Rag}, which we will come to shortly.

His 1st stab at the Super Duper Electromagnet wasn't very super duper; in fact, it may not have worked at all -- I don't remember for sure. It \textit{did} come out of an actual book, rather than Dave's head, however. The idea was this: you magnetized a spike nail by rubbing it against a regular magnet. The magnetic charge imparted to the spike would be weak, the book said, but enough to pick up a few iron fillings. After trying this, you were supposed to wrap a length of copper wire around the barrel of the spike, \& attach the ends of the wire to the terminals of a dry-cell battery. According to the book, the electricity would strengthen the magnetism, \& you could pick up a lot more iron filings.

Dave didn't just want to pick up a stupid pile of metal flakes, though; Dave wanted to pick up Buicks, railroad boxcars, possibly Army transport planes. Dave wanted to turn on the juice \& move the world in its orbit.

Pow! Super!

We each had our part to play in creating the Super Duper Electromagnet. Dave's part was to build it. My part would be to test it. Little Stevie King, Stratford's answer to Chuck Yeager.

Dave's new version of the experiment bypassed the pokey old dry cell (which was probably flat anyway when we bought it at the hardware store, he reasoned) in favor of actual wall-current. Dave cut the electrical cord off an old lamp someone had put out on the curb with the trash, stripped the coating all the way down to the plug, then wrapped his magnetized spike in spirals of bare wire. Then, sitting on the floor in the kitchen of our West Broad Street apartment, he offered me the Super Duper Electromagnet \& bade me do my part \& plug it in.

I hesitated -- give me at least that much credit -- but in the end, Dave's manic enthusiasm was too much to withstand. I plugged it in. There was no noticeable magnetism, but the gadget \textit{did} blow out every light \& electrical appliance in our apartment, every light \& electrical appliance in the building, \& every light \& electrical appliance in the building next door (where my dream-girl lived in the ground-floor apartment). Something popped in the electrical transformer out front, \& some cops came. Dave \& I spent a horrible hour watching from out mother's bedroom window, the only one that looked out on the street (all the others had a good view of the grassless, turd-studded yard behind us, where the only living thing was a mangy canine named Roop-Roop). When the cops left, a power truck arrived. A man in spiked shoes climbed the pole between the 2 apartment houses to examine the transformer. Under other circumstances, this would have absorbed us completely, but not that day. That day we could only wonder if our mother would come \& see us in reform school. Eventually, the lights came back on \& the power truck went away. We were not caught \& lived to fight another day. Dave decided he might build a Super Duper Glider instead of a Super Duper Electromagnet for his science project. I, he told me, would get to take the 1st ride. Wouldn't that be great?

\fbox{\bf 11} I was born in 1947 \& we didn't get our 1st television until 1958. The 1st thing I remember watching on it was \textit{Robot Monster}, a film in which a guy dressed in an ape-suit with a goldfish bowl on his head -- Ro-Man, he was called -- ran around trying to kill the last survivors of a nuclear war. I felt this was art of quite a high nature.

I also watched \textit{Highway Patrol} with Broderick Crawford as the fearless Dan Matthews, \& \textit{1 Step Beyond}, hosted by John Newland, the man with the world's spookiest eyes. There was \textit{Cheyenne} \& \textit{Sea Hunt, Your Hit Parade} \& \textit{Annie Oakley}; there was Tommy Rettig as the 1st of Lassie's many friends, Jock Mahoney as \textit{The Range Rider}, \& Andy Devine yowling, ``Hey, Wild Bill, wait for me!'' in his odd, high voice. There was a whole world of vicarious adventure which came packaged in black-\&-white, 14 inches across \& sponsored by brand names which still sound like poetry to me. I loved it all.

But TV came relatively late to the King household, \& I'm glad. I am, when you stop to think of it, a member of a fairly select group: the final handful of American novelists who learned to read \& write before they learned to eat a daily helping of video bullshit. This might not be important. On the other hand, if you're just starting out as a writer, you could do worse than strip your television's electric plug-wire, wrap a spike around it, \& then stick it back into the wall. See what blows, \& how far.

Just an idea.

\fbox{\bf12} In the late 1950s, a literary agent \& compulsive science fiction memorabilia collector named Forrest J. Ackerman changed the lives of thousands of kids -- I was one -- when he began editing a magazine called \textit{Famous Monsters of Filmland}. Ask anyone who has been associated with the fantasy-horror-science fiction genres in the last 30 years about this magazine, \& you'll get a laugh, a flash of the eyes, \& a stream of bright memories -- I practically guarantee it.

Around 1960, Forry (who sometimes referred to himself as ``the Ackermonster'') spun off the short-lived by interesting \textit{Spacemen}, a magazine which covered science fiction films. In 1960, I sent a story to \textit{Spacemen}. It was, as well as I can remember, the 1st story I ever submitted for publication. I don't recall the title, but I was still in the Ro-Man phase of my development, \& this particular tale undoubtedly owed a great deal to the killer ape with the goldfish bowl on his head.

My story was rejected, but Forry kept it. (Forry keeps \textit{everything}, which anyone who has ever toured his house -- the Ackermansion -- will tell you.) About 20 years later, while I was signing autographs at a Los Angeles bookstore, Forry turned up in line $\ldots$ with my story, single-spaced \& typed with the long-vanished Royal typewriter my mom gave me for Christmas the year I was 11. He wanted me to sign it to him, \& I guess I did, although the whole encounter was so surreal I can't be completely sure. Talk about your ghosts. Man oh man.

\fbox{\bf13} The 1st story I did actually publish was in a horror fanzine issued by Mike Garrett of Birmingham, Alabama (Mike is still around, \& still in the biz). He published this novella under the title ``In a Half-World of Terror,'' but I still like my title much better. Mine was ``I Was a Teen-Age Grave-robber.'' Super Duper! Pow!

\fbox{\bf14} My 1st really original story idea -- you always know the 1st one, I think -- came near the end of Ike's 8-year reign of benignity. I was sitting at the kitchen table of our house in Durham, Maine, \& watching my mother stick sheets of S\&H Green Stamps into a book. (For more colorful stories about Green Stamps, see \textit{The Liar's Club}.) Our little family troika had moved back to Maine so our mom could take care of her parents in their declining years. Mama was about 80 at that time, obese \& hypertensive \& mostly blind; Daddy Guy was 82, scrawny, morose, \& prone to the occasional Donald Duck outburst which only my mother could understand. Mom called Daddy Guy ``Fazza.''

My mother's sisters had gotten my mom this job, perhaps thinking they could kill 2 birds with 1 stone -- the aged Ps would be taken care of in a homey environment by a loving daughter, \& The Nagging Problem of Ruth would be solved. She would no longer be adrift, trying to take care of 2 boys while she floated almost aimlessly from Indiana to Wisconsin to Connecticut, baking cookies at 5 in the morning or pressing sheets in a laundry where the temperatures often soared to 110 in the summer \& the foreman gave out salt pills at 1 \& 3 every afternoon from July to the end of September.

She hated her new job, I think -- in their effort to take care of her, her sisters turned our self-sufficient, funny, slightly nutty mother into a sharecropper living a largely cashless existence. The money the sisters sent her each month covered the groceries but little else. They sent boxes of clothes for us. Toward the end of each summer, Uncle Clayt \& Aunt Ella (who were not, I think, real relatives at all) would bring cartons of canned vegetables \& preserves. The house we lived in belonged to Aunt Ethelyn \& Uncle Oren. \& once she was there, Mom was caught. She got another actual job after the old folks died, but she lived in that house until the cancer got her. When she left Durham for the last time -- David \& his wife Linda cared for her during the final weeks of her final illness -- I have an idea she was probably more than ready to go.

\fbox{\bf15} Let's get 1 thing clear right now, shall we? There is no idea Dump, no Story Central, no Island of the Buried Bestsellers; good story ideas seem to come quite literally from nowhere, sailing at you right out of the empty sky: 2 previously unrelated ideas come together \& making something new under the sun. Your job isn't to find these ideas but to recognize them when they show up.

On the day this particular idea -- the 1st really good one -- came sailing at me, my mother remarked that she needed 6 more books of stamps to get a lamp she wanted to give her sister Molly for Christmas, \& she didn't think she would make it in time. ``I guess it will have to be for her birthday, instead,'' she said. ``These cussed things always look like a lot until you stick them in a book.'' Then she crossed her eyes \& ran her tongue out at me. When she did, I saw her tongue was S\&H green. I thought how nice it would be if you could make those damned stamps in your basement, \& in that instant a story called ``Happy Stamps'' was born. The concept of counterfeiting Green Stamps \& the sight of my mother's green tongue created it in an instant.

The hero of my story was your classic Poor Schmuck, a guy named Roger who had done jail time twice for counterfeiting money -- 1 more burst could make him a 3-time loser. Instead of money, he began to counterfeit Happy Stamps $\ldots$ except, he discovered, the design of Happy Stamps was so moronically simple that he wasn't really counterfeiting at all; he was creating reams of the actual article. In a funny scene -- probably the 1st really competent scene I ever wrote -- Roger sits in the living room with his old mom, the 2 of them mooning over the Happy Stamps catalogue while the printing press runs downstairs, ejecting bale after bale of those same trading stamps.

``Great Scott!'' Mom says. ``According to the fine print, you can get \textit{anything} with Happy Stamps, Roger -- you tell them what you want, \& they figure out how many books you need to get it. Why, for 6 or 7 million books, we could probably get a Happy Stamps house in the suburbs!''

Roger discovers, however, that although the \textit{stamps} are perfect, the \textit{glue} is defective. If you lap the stamps \& stick them in the book they're fine, but if you send them through a mechanical licker, the pink Happy Stamps turn blue. At the end of the story, Roger is in the basement, standing in front of a mirror. Behind him, on the table, are roughly 90 books of Happy Stamps, each book filled with individually licked sheets of stamps. Our hero's lips are pink. He runs out his tongue; that's even pinker. Even his teeth are turning pink. Mom calls cheerily down the stairs, saying she has just gotten off the phone with the Happy Stamps National Redemption Center in Terre Haute, \& the lady said they could probably get a nice Tudor home in Weston for only 11 600 000 books of Happy Stamps.

``That's nice, Mom,'' Roger says. He looks at himself a moment longer in the mirror, lips pink \& eyes bleak, then slowly returns to the table. Behind him, billions of Happy Stamps are stuffed into basement storage bins. Slowly, our hero opens a fresh stamp-book, then begins to lick sheets \& stick them in. Only 11 590 000 books to go, he thinks as the story ends, \& Mom can have her Tudor.

There were things wrong with this story (the biggest hole was probably Roger's failure simply to start over with a different glue), but it was cute, it was fairly original, \& I knew I had done some pretty good writing. After a long time spent studying the markets in my beat-up \textit{Writer's Digest}, I sent ``Happy Stamps'' off to \textit{Alfred Hitchcock's Mystery Magazine}. It came back 3 weeks later with a form rejection slip attached. This slip bore Alfred Hitchcock's unmistakable profile in red rink \& wished me good luck with my story. At the bottom was an unsigned jotted message, the only personal response I got from \textit{AHMM} over 8 years of periodic submissions. ``Don't staple manuscripts,'' the postscript read. ``Loose pages plus paperclip equal correct way to submit copy.'' This was pretty cold advice, I thought, but useful in its way. I have never stapled a manuscript since.

\fbox{\bf16} My room in our Durham house was upstairs, under the eaves. At night I could lie in bed beneath 1 of these eaves -- if I sat up suddenly, I was apt to whack my head a good one -- \& read by the light of a gooseneck lamp that put an amusing boa constrictor of shadow on the ceiling. Sometimes the house was quiet except for the whoosh of the furnace \& the patter of rats in the attic; sometimes my grandmother would spend an hour or so around midnight yelling for someone to check Dick -- she was afraid he hadn't been fed. Dick, a horse she'd had in her days as a schoolteacher, was at least 40 years dead. I had a desk beneath the room's other eave, my old Royal typewriter, \& 100 or so paperback books, mostly science fiction, which I lined up along the baseboard. On my bureau was a Bible won for memorizing verses in Methodist Youth Fellowship \& a Webcor phonograph with an automatic changer \& a turntable covered in soft green velvet. On it I played my records, mostly 45s by Elvis, Chuck Berry, Freddy Cannon, \& Fats Domino. I liked Fats; he knew how to rock, \& you could tell he was having fun.

When I got the rejection slip from \textit{AHMM}, I pounded a nail into the wall above the Webcor, wrote ``Happy Stamps'' on the rejection slip, \& poked it onto the nail. Then I sat on my bed \& listened to Fats using ``I'm Ready.'' I felt pretty good, actually. When you're still too young to shave, optimism is a perfectly legitimate response to failure.

By the time I was 14 (\& shaving twice a week whether I needed to or not) the nail in my wall would no longer support the weight of the rejection slips impaled upon it. I replaced the nail with a spike \& went on writing. By the time I was 16 I'd begun to get rejection slips with handwritten notes a little more encouraging than the advice to stop using staples \& using paperclips. The 1st of these hopeful notes was from Algis Budrys, then the editor of \textit{Fantasy \& Science Fiction}, who read a story of mine called ``The Night of the Tiger'' (the inspiration was, I think, an episode of \textit{The Fugitive} in which Dr. Richard Kimble worked as an attendant cleaning out cages in a zoo or a circus) \& wrote: ``This is good. Not for us, but good. You have talent. Submit again.''

Those 4 brief sentences, scribbled by a fountain pen that left big ragged blotches in its wake, brightened the dismal winter of my 16th year. 10 years or so later, after I'd sold a couple of novels, I discovered ``The Night of the Tiger'' in a box of old manuscripts \& thought it was still a perfectly respectable tale, albeit one obviously written by a guy who had only begun to learn his chops. I rewrote it \& on a whim resubmitted it to \textit{F\&SF}. This time they bought it. 1 thing I've noticed is that when you've had a little success, magazines are a lot less apt to use that phrase, ``Not for us.''

\fbox{\bf17} Although he was a year younger than his classmates, my big brother was bored with high school. Some of this had to do with his intellect -- Dave's IQ tested in the 150s or 160s -- but I think it was mostly his restless nature. For Dave, high school just wasn't super duper enough -- there was no pow, no wham, no \textit{fun}. He solved the problem, at least temporarily, by creating a newspaper which he called \textit{Dave's Rag}.

The \textit{Rag}'s office was a table located in the dirt-floored, rock-walled, spider-infested confines of our basement, somewhere north of the furnace \& east of the root-cellar, where Clayt \& Ella's endless cartons of preserves \& canned vegetables were kept. The \textit{Rag} was an odd combination of family newsletter \& small-town biweekly. Sometimes it was a monthly, if Dave got sidetracked by other interests (maple-sugaring, cider-making, rocket-building, \& car-customizing, just to name a few), \& then there would be jokes I didn't understand about how Dave's \textit{Rag} was a little late this month or how we shouldn't bother Dave, because he was down in the basement, on the \textit{Rag}.

Jokes or no jokes, circulation rose slowly from about 5 copies per issue (sold to nearby family members) to something like 50 or 60, with our relatives \& the relatives of neighbors in our small town (Durham's population in 1962 was about 900) eagerly awaiting each new edition. A typical number would let people know how Charley Harrington's broken leg was mending, what guest speakers might be coming to the West Durham Methodist Church, how much water the King boys were hauling from the town pump to keep from draining the well behind the house (of course it went dry every fucking summer no matter how much water we hauled), who was visiting the Browns or the Halls on the other side of Methodist Corners, \& whose relatives were due to hit town each summer. Dave also included sports, word-games, weather reports (``It's been pretty dry, but local farmer Harold Davis says if we don't have at least 1 good rain in August he will smile \& kiss a pig''), recipes, a continuing story (I wrote that), \& Dave's Jokes \& Humor, which included nuggets like these:

Stand: ``What did the beaver say to the oak tree?''

Jan: ``It was nice gnawing you!''

1st Beatnik: ``How do you get to Carnegie Hall?''

2nd Beatnik: ``Practice man practice!''

During the \textit{Rag}'s 1st year, the print was purple -- those issues were produced on a flat plate of jelly called a hectograph. My brother quickly decided the hectograph was a pain in the butt. It was just too slow for him. Even as a kid in short pants, Dave hated to be halted. Whenever Milt, our mom's boyfriend (``Sweeter than smart,'' Mom said to me 1 day a few months after she dropped him), got stuck in traffic or at a stoplight, Dave would lean over from the back seat of Milt's Buick \& yell, ``Drive over em, Uncle Milt! Drive over em!''

As a teenager, waiting for the hectograph to ``freshen'' between pages printed (while ``freshening,'' the print would melt into a vague purple membrane which hung in the jelly like a manatee's shadow) drove David all but insane with impatience. Also, he badly wanted to add photographs to the newspaper. He took good ones, \& by age 16 he was developing them, as well. He rigged a darkroom in a closet \& from its tiny, chemical-stinking confines produced pictures which were often startling in their clarity \& composition (the photo on the back of \textit{The Regulators}, showing me with a copy of the magazine containing my 1st published story, was taken by Dave with an old Kodak \& developed in his closet darkroom).

In addition to these frustrations, the flats of hectograph jelly had a tendency to incubate \& support colonies of strange, sporelike growths in the unsavory atmosphere of our basement, no matter how meticulous we were about covering the damned old slowcoach thing once the day's printing chores were done. What looked fairly ordinary on Monday sometimes looked like something out of an H.P. Lovecraft horror tale by the weekend.

In Brunswick, where he went to high school, Dave found a shop with a small drum printing press for sale. It worked -- barely. You typed up your copy on stencils which could be purchased in a local office-supply store for 19 cents apiece -- my brother called this chore ``cutting stencil,'' \& it was usually my job, as I was less prone to make typing errors. The stencils were attached to the drum of the press, lathered up with the world's stinkiest, oogiest ink, \& then you were off to the races -- crank `til your arm falls off, son. We were able to put together in 2 nights what had previously taken a week with the hectograph, \& while the drum-press was messy, it did not look inflected with a potentially fatal disease. \textit{Dave's Rag} entered its brief golden age.

\fbox{\bf18} I wasn't much interested in the printing process, \& I wasn't interested at all in the arcana of 1st developing \& then reproducing photographs. I didn't care about putting Hearst shifters in cars, making cider, or seeing if a certain formula would send a plastic rocket into the stratosphere (usually they didn't even make it over the house). What I cared about most between 1958 \& 1966 was movies.

As the 50s gave way to the 60s, there was only 2 movie theaters in the area, both in Lewiston. The Empire was the 1st-run house, showing Disney pictures, Bible epics, \& musicals in which widescreen ensembles of well-scrubbed folks danced \& sang. I went to these if I had a ride -- a movie was a movie, after all -- but I didn't like them very much. They were boringly wholesome. They were predictable. During \textit{The Parent Trap}, I kept hoping Hayley Mills would run into Vic Morrow from \textit{The Blackboard Jungle}. That would have livened things up a little, by God. I felt that 1 look at Vic's switchblade knife \& gimlet gaze would have put Hayley's piddling domestic problems in some kind of reasonable perspective. \& when I lay in bed at night under my eave, listening to the wind in the trees or the rats in the attic, it was not Debbie Reynolds as Tammy or Sandra Dee as Gidget that I dreamed of, but Yvette Vickers from \textit{Attack of the Giant Leeches} or Luana Anders from \textit{Dementia 13}. Never mind sweet; never mind uplifting; never mind Snow White \& the 7 Goddam Dwarfs. At 13 I wanted monsters that ate whole cities, radioactive corpses that came out of the ocean \& ate surfers, \& girls in black bras who looked like trailer trash.

Horror movies, science fiction movies, movies about teenage gangs on the prowl, movies about losers on motorcycles -- this was the stuff that turned my dials up to 10. The place to get all of this was not at the Empire, on the upper end of Lisbon Street, but at the Ritz, down at the lower end, amid the pawnshops \& not far from Louie's Clothing, where in 1964 I bought my 1st pair of Beatle boots. The distance from my house to the Ritz was 14 miles, \& I hitchhiked there almost every weekend during the 8 years between 1958 \& 1966, when I finally got my driver's license. Sometimes I went with my friend Chris Chesley, sometimes I went alone, but unless I was sick or something, I always went. It was at the Ritz that I saw \textit{I Married a Monster from Outer Space}. with Tom Tryon; \textit{The Haunting}, with Claire Bloom \& Julie Harris; \textit{The Wild Angels}, with Peter Fonda \& Nancy Sinatra. I saw Olivia de Havilland put out James Caan's eyes with makeshift knives in \textit{Lady in a Cage}, saw Joseph Cotton come back from the dead in \textit{Hush $\ldots$ Hush, Sweet Charlotte}, \& watched with held breath (\& not a little prurient interest) to see if Allison Hayes would grow all the way out of her clothes in \textit{Attack of the 50 Ft. Woman}. At the Ritz, all the finer things in life were available $\ldots$ or \textit{might be} available, if you only sat in the 3rd row, paid close attention, \& did not blink at the wrong moment.

Chris \& I liked just about any horror movie, but our faves were the string of American-International films, most directed by Roger Corman, with titles cribbed from Edgar Allan Poe. I wouldn't say \textit{based upon} the works of Edgar Allan Poe, because there is little in any of them which has anything to do with Poe's actual stories \& poems (\textit{The Raven} was filmed as a comedy -- no kidding). \& yet the best of them -- \textit{The Haunted Palace, The Conqueror Worm, The Masque of the Red Death} -- achieved a hallucinatory eeriness that made the special. Chris \& I had our own name for these films, one that made them into a separate genre. There were westerns, there were love stories, there were war stories $\ldots$ \& there were Poepictures.

``Wanna hitch to the show Saturday afternoon?'' Chris would ask. ``Go to the Ritz?''

``What's on?'' I'd ask.

``A motorcycle picture \& a Poepicture,'' he'd say. I, of course, was on that combo like white on rice. Bruce Dern going batshit on a Harley \& Vincent Price going batshit in a haunted castle overlooking a restless ocean: who could ask for me? You might even get Hazel Court wandering around in a lacy low-cut nightgown, if you were lucky.

Of all the Poepictures, the one that affected Chris \& me the most deeply was \textit{The Pit \& the Pendulum}. Written by Richard Matheson \& filmed in both widescreen \& Technicolor (color horror pictures were still a rarity in 1961, when this one came out), \textit{Pit} took a bunch of standard gothic ingredients \& turned them into something special. It might have been the last day really great studio horror picture before George Romero's ferocious indie \textit{The Night of the Living Dead} came along \& changed everything forever (in some few cases for the better, in most for the worse). The best scene -- the one which froze Chris \& me into our seats -- depicted John Kerr digging into a castle wall \& discovering the corpse of his sister, who was obviously buried alive. I have never forgotten the corpse's close-up, shot through a red filter \& a distorting lens which elongated the face into a huge silent scream.

On the long hitch home that night (if rides were slow in coming, you might end up walking 4 or 5 miles \& not get home until well after dark) I had a wonderful idea: I would turn \textit{The Pit \& the Pendulum} into a book! Would novelize it, as Monarch Books had novelized such undying film classics as \textit{Jack the Ripper, Gorgo}, \& Konga. But I wouldn't just write this masterpiece; I would also print it, using the drum-press in our basement, \& sell copies at school! Zap! Ka-pow!

As it was conceived, so was it done. Working with the care \& deliberation for which I would later be critically acclaimed, I turned out my ``novel version'' of \textit{The Pit \& the Pendulum} in 2 days, composing directly onto the stencils from which I'd print. Although no copies of that particular masterpiece survive (at least to my knowledge), I believe it was 8 pages long, each page single-spaced \& paragraph breaks kept to an absolute minimum (each stencil cost 19 cents, remember). I printed sheets on both sides, just as in a standard book, \& added a title page on which I drew a rudimentary pendulum dripping small black blotches which I hoped would look like blood. At the last moment I realized I had forgotten to identify the publishing house. After a half-hour or so of pleasant mulling, I typed the words \textbf{A V.I.B. BOOK} in the upper right corner of my title page. V.I.B. stood for Very Important Book.

I ran off about 40 copies of \textit{The Pit \& the Pendulum}, blissfully unaware that I was in violation of every plagiarism \& copyright statue in the history of the world; my thoughts were focused almost entirely on how much money I might make if my story was a hit at school. The stencils had cost me \$1.71 (having to use up 1 whole stencil for the title page seemed a hideous waste of money, but you had to look good, I'd reluctantly decided; you had to go out there with a bit of the old attitude), the paper cost another 2 bits or so, the staples were free, cribbed from my brother (you might have to paperclip stories you were sending out to magazines, but this was a \textit{book}, this was the bigtime). After some further thought, I priced V.I.B. \#1, \textit{The Pit \& the Pendulum} by Steve King, at a quarter a copy. I thought I might be able to sell 10 (my mother would buy 1 to get me started; she could always be counted on), \& that would add up to \$2.50. I'd make about 40 cents, which would be enough to finance another educational trip to the Ritz. If I sold 2 more, I could get a big sack of popcorn \& a Coke, as well.

\textit{The Pit \& the Pendulum} turned out to be my 1st best-seller. I took the entire print-run to school in my book-bag (in 1961 I would have been an 8th-grader at Durham's newly built 4-room elementary school), \& by noon that day I had sold 2 dozen. By the end of lunch hour, when word had gotten around about the lady buried in the wall (``They stared with horror at the bones sticking out from the ends of her fingers, realizing she had died scratcheing madley for escape''), I had sold 3 dozen. I had \$9 in change weighing down the bottom of my book-bag (upon which Durham's answer to Daddy Cool had carefully printed most of the lyrics to ``The Lion Sleeps Tonight'') \& was walking around in a kind of dream, unable to believe my sudden ascension to previously unsuspected realms of wealth. It all seemed too good to be true.

It was. When the school day ended at 2 o'clock, I was summoned to the principle's office, where I was told I couldn't turn the school into a marketplace, especially not, Miss Hisler said, to sell such trash as \textit{The Pit \& the Pendulum}. Her attitude didn't much surprise me. Miss Hisler had been the teacher at my previous school, the 1-roomer at Methodist Corners, where I went to the 5th \& 6th grades. During that time she had spied me reading a rather sensational ``teenage rumble'' novel (\textit{The Amboy Dukes}, by Irving Shulman), \& had taken it away. This was just more of the same, \& I was disgusted with myself for not seeing the outcome in advance. In those days we called someone who did an idiotic thing a dubber (pronounced \textit{dubba} if you were from Maine). I had just dubbed up bigtime.

``What I don't understand, Stevie,'' she said, ``is why you'd write junk like this in the 1st place. You're talented. Why do you want to waste your abilities?'' She had rolled up a copy of V.I.B. \#1 \& was brandishing it at me the way a person might brandish a rolled-up newspaper at a dog that has piddled on the rug. She waited for me to answer -- to her credit, the question was not entirely rhetorical -- but I had no answer to give. I was ashamed. I have spent a good many years since -- too many, I think -- being ashamed about what I write. I think I was 40 before I realized that almost every writer of fiction \& poetry who has ever published a line has been accused by someone of wasting his or her God-given talent. If you write (or paint or dance or sculpt or sing, I suppose), someone will try to make you feel lousy about it, that's all. I'm not editorializing, just trying to give you the facts as I see them.

Miss Hisler told me I would have to give everyone's money back. I did so with no argument, even to those kids (\& there were a quite a few, I'm happy to say) who insisted on keeping their copies of V.I.B. \#1. I ended up losing money on the deal after all, but when summer vacation came I printed 4 dozen copies of a new story, an original called \textit{The Invasion of the Star-Creatures}, \& sold all but 4 or 5. I guess that means I won in the end, at least in a financial sense. But in my heart I stayed ashamed. I kept hearing Miss Hisler asking why I wanted to waste my talent, why I wanted to waste my time, why I wanted to write junk.

\fbox{\bf19} Doing a serial story for \textit{Dave's Rag} was fun, but my other journalistic duties bored me. Still, I had worked for a newspaper of sorts, word got around, \& during my sophomore year at Lisbon High I became editor of our school newspaper, \textit{The Drum}. I don't recall being given any choice in this matter; I think I was simply appointed. My 2nd-in-command, Danny Emond, had even less interest in the paper than I did. Danny just liked the idea that Room 4, where we did our work, was near the girls' bathroom. ``Someday I'll just go crazy \& hack my way in there, Steve,;; he told me on more than 1 occasion. ``Hack, hack, hack.'' Once he added, perhaps in an effort to justify himself: ``The prettiest girls in school pull up their skirts in there.'' This struck me as so fundamentally stupid it might actually be wise, like a Zen koan or an early story by John Updike.

\textit{The Drum} did not prosper under my editorship. Then as now, I tend to go through periods of idleness followed by periods of workaholic frenzy. In the schoolyear 1963--1964. \textit{The Drum} published just 1 issue, but that one was a monster thicker than the Lisbon Falls telephone book. 1 night -- sick to death of Class Reports, Cheerleading Updates, \& some lamebrain's efforts to write a school poem -- I created a satiric high school newspaper of my own when I should have been captioning photographs for \textit{The Drum}. What resulted was a 4-sheet which I called \textit{The Village Vomit}. The boxed motto in the upper lefthand corner was not ``All the News That's Fit to Print'' but ``All the Shit That Will Stick.'' That piece of dimwit humor got me into the only real trouble of my high school career. It also led me to the most useful writing lesson I ever got.

In typical \textit{Mad} magazine style (``What, me worry?''), I filled the \textit{Vomit} with fictional tidbits about the LHS faculty, using teacher nicknames the student body would immediately recognize. Thus Miss Raypach, the study-hall monitor, became Miss Rat Pack; Mr. Ricker, the college-track English teacher (\& the school's most urbane faculty member -- he looked quite a bit like Craig Stevens in \textit{Peter Gunn}), became Cow Man because his family owned Ricker Dairy; Mr. Diehl, the earth-science teacher, became Old Raw Diehl.

As all sophomoric humorists must be, I was totally blown away by my own wit. What a funny fellow I as! A regular mill-town H. L. Mencken! I simply must take the \textit{Vomit} to school \& show all my friends! They would bust a collective gut!

As a matter of fact, they \textit{did} bust a collective gut; I had some good ideas about what tickled the funnybones of high school kids, \& most of them were showcased in \textit{The Village Vomit}. In 1 article, Cow Man's prize Jersey won a livestock farting contest at Topsham Fair; in another, Old Raw Diehl was fired for sticking the eyeballs of specimen fetal pigs up his nostrils. Humor in the grand Swiftian manner, you see. Pretty sophisticated, eh?

During period 4, 3 of my friends were laughing so hard in the back of study-hall that Miss Raypach (Rat Pack to you, chum) crept up on them to see what was so funny. She confiscated \textit{The Village Vomit}, on which I had, either out of overweening pride or almost unbelievable naivet\'e, put my name as Editor in Chief \& Grand High Poobah, \& at the close of school I was for the 2nd time in my student career summoned to the office on account of something I had written.

This time the trouble was a good deal more serious. Most of the teachers were inclined to be good sports about my teasing -- even Old Raw Diehl was willing to let bygones by bygones concerning the pigs' eyeballs -- but one was not. This was Miss Margitan, who taught shorthand \& typing to the girls in the business courses. She commanded both respect \& fear; in the tradition of teachers from an earlier era, Miss Margitan did not want to be your pal, your psychologist, or your inspiration. She was there to teach business skills, \& she wanted all learning to be done by the rules. \textit{Her} rules. Girls in Miss Margitan's classes were sometimes asked to kneel on the floor, \& if the hems of their skirts didn't touch the linoleum, they were sent home to change. No amount of tearful begging could soften her, no reasoning could modify her view of the world. Her detention lists were the longest of any teacher in the school, but her girls were routinely selected as valedictorians or salutatorians \& usually went on to good jobs. Many came to love her. Others loathed her then \& likely still do now, all these years later. These later girls called her ``Maggot'' Margitan, as their mothers had no doubt before them. \& in \textit{The Village Vomit} I had an item which began, ``Miss Margitan, known affectionately to Lisbonians everywhere as Maggot $\ldots$''

Mr. Higgins, our bald principal (breezily referred to in the \textit{Vomit} as Old Cue-Ball), told me that Miss Margitan had been very hurt \& very upset by what I had written. She was apparently not too hurt to remember that old scriptural admonition which goes ``Vengeance is mine, saith the shorthand teacher,'' however; Mr. Higgins said she wanted me suspended from school.

In my character, a kind of wildness \& a deep conservatism are wound together like hair in a braid. It was the crazy part of me that had 1st written \textit{The Village Vomit} \& then carried it to school; now that troublesome Mr. Hyde had dubbed up \& slunk out the back door. Dr. Jekyll was left to consider how my mom would look at me if she found out I had been suspended -- her hurt eyes. I had to put thoughts of her out of my mind, \& fast. I was a sophomore, I was a year older than most others in my class, \& at 6.2 feet I was 1 of the bigger boys in school. I desperately didn't want to cry in Mr. Higgin's office -- not with kids surging through the halls \& looking curiously in the window at us: Mr. Higgins behind his desk, me in the Bad Boy Seat.

In the end, Miss Margitan settled for a formal apology \& 2 weeks of detention for the bad boy who had dared call her Maggot in print. It was bad, but what in high school is not? At the time we're stuck in it, like hostages locked in a Turkish bath, high school seems the most serious business in the world to just about all of us. It's not until the 2nd or 3rd class reunion that we start realizing how absurd the whole thing was.

A day or 2 later I was ushered into Mr. Higgins's office \& made to stand in front of her. Miss Margitan sat ramrod-straight with her arthritic hands folded in her lap \& here gray eyes fixed unflinchingly on my face, \& I realized that something about her was different from any other adult I had ever met. I didn't pinpoint that difference at once, but I knew that there would be no charming this lady, no winning her over. Later, while I was flying paper planes with the other bad boys \& bad girls in detention hall (detention turned out to be not so bad), I decided that it was pretty simple: Miss Margitan didn't like boys. She was the 1st woman I ever met in my life who didn't like boys, not even 1 little bit.

If it makes any difference, my apology was heartfelt. Miss Margitan really had been hurt by what I wrote, \& that much I could understand. I doubt that she hated me -- she was probably too busy -- but she was the National Honor Society advisor at LHS, \& when my name showed up on the candidate list 2 years later, she vetoed me. The Honor Society did not need boys ``of his type,'' she said. I have come to believe she was right. A boy who once wiped his ass with poison ivy probably doesn't belong in a smart people's club.

I haven't trucked much with satire since then.

\fbox{\bf20} Hardly a week after being sprung from detention hall, I was once more invited to step down to the principal's office. I went with a sinking heart, wondering what new shit I'd stepped in.

It wasn't Mr. Higgins who wanted to see me, at least; this time the school guidance counselor had issued the summons. There had been discussions about me, he said, \& how to turn my ``restless pen'' into more constructive channels. He had enquired of John Gould, editor of Lisbon's weekly newspaper, \& had discovered Gould had an opening for a sports reporter. While the school couldn't \textit{insist} that I take this job, everyone in the front office felt it would be a good idea. \textit{Do it or die}, the G.C.'s eyes suggested. Maybe that was just paranoia, but even now, almost 40 years later, I don't think so.

I groaned inside. I was shut of \textit{Dave's Rag}, almost shut of \textit{The Drum}, \& now here was the Lisbon \textit{Weekly Enterprise}. Instead of being haunted by waters, like Norman Maclean in \textit{A River Runs Through It}, I was as a teenager haunted by newspapers. Still, what could I do? I rechecked the look in the guidance counselor's eyes \& said I would be delighted to interview for the job.

Gould -- not the well-known New England humorist or the novelist who wrote \textit{The Greenleaf Fires} but a relation of both, I think -- greeted me warily but with some interest. We would try each other out, he said, if that suited me.

Now that I was away from the administrative offices of Lisbon High, I felt able to muster a little honesty. I told Mr. Gould that I didn't know much about sports. Gould said, ``These are games people understand when they're watching them drunk in bars. You'll learn if you try.''

He gave me a huge roll of yellow paper on which to type me copy -- I think I still have it somewhere -- \& promised me a wage of half a cent a word. It was the 1st time someone had promised me wages for writing.

The 1st 2 pieces I turned in had to do with a basketball game in which an LHS player broke the school scoring record. One was a straight piece of reporting. The other was a sidebar about Robert Ransom's record-breaking performance. I brought both on Gould the day after the game so he'd have them for Friday, which was when the paper came out. He read the game piece, made 2 minor corrections, \& spiked it. Then he started in on the feature piece with a large black pen.

I took my fair share of English Lit classes in my 2 remaining years at Lisbon, \& my fair share of composition, fiction, \& poetry classes in college, but John Gould taught me more than any of them, \& in no more than 10 minutes. I wish I still had the piece -- it deserves to be framed, editorial corrections \& all -- but I can remember pretty well how it went \& how it looked after Gould had combed through it with that black pen of his. Here's an example: {\sf[excerpt]}

Gould stopped at ``the years of Korea'' \& looked up at me. ``What year was the last record made?'' he asked.

Luckily, I had my notes. ``1953,'' I said. Gould grunted \& went back to work. When he finished marking my copy in the manner indicated above, he looked up \& saw something on my face. I think he must have mistaken it for horror. It wasn't; it was pure revelation. Why, I wondered, didn't English teachers ever do this? It was like the Visible Man Old Raw Diehl had on his desk in the biology room.

``I only took out the bad parts, you know,'' Gould said. ``Most of it's pretty good.''

``I know,'' I said, meaning both things: yes, most of it was good -- okay anyway, serviceable -- \& yes, he had only taken out the bad parts. ``I won't do it again.''

He laughed. ``If that's true, you'll never have to work for a living. You can do \textit{this} instead. Do I have to explain any of these marks?''

``No,'' I said.

``When you write a story, you're telling yourself the story,'' he said. ``When you rewrite, your main job is taking out all the things that are \textit{not} the story.''

Gould said something else that was interesting on the day I turned in my 1st 2 pieces: write with the door closed, rewrite with the door open. Your stuff starts out being just for you, in other words, but then it goes out. Once you know what the story is \& get it right -- as right as you can, anyway -- it belongs to anyone who wants to read it. Or criticize it. If you're very lucky (this is my idea, not John Gould's, but I believe he would have subscribed to the notion), more will want to do the former than the latter.

\fbox{\bf21} Just after the senior class trip to Washington, D.C., I got a job at Worumbo Mills \& Weaving, in Lisbon Falls. I didn't want it -- the work was hard \& boring, the mill itself a dingy fuckhole overhanging the polluted Androscoggin River like a workhouse in a Charles Dickens novel -- but I needed the paycheck. My mother was marking lousy wages as a housekeeper at a facility for the mentally ill in New Gloucester, but she was determined I was going to college like my brother David (University of Maine, class of `66, \textit{cum laude}). In her mind, the education had become almost secondary. Durham \& Lisbon Falls \& the University of Maine at Orono were part of a small world where folks neighbored \& still minded each other's business on the 4- \& 6-party lines which then served the Sticksville townships. In the big world, boys who didn't go to college were being sent overseas to fight in Mr. Johnson's undeclared war, \& many of them were coming home in boxes. My mother like Lyndon's War on Poverty (``That's the war \textit{I'm} in,'' she sometimes said), but not what he was up to in Southeast Asia. Once I told her that enlisting \& going over there might be good for me -- surely there would be a book in it, I said.

``Don't be an idiot, Stephen,'' she said. ``With your eyes, you'd be the 1st one to get shot. You can't write if you're dead.''

She meant it; her head was set \& so was her heart. Consequently, I applied for scholarships, I applied for loans, \& I went to work in the mill. I certainly wouldn't get far on the \$5 \& \$6 a week I could make writing about bowling tournaments \& Soap Box Derby races for the \textit{Enterprise}.

During my final weeks at Lisbon High, my schedule looked like this: up at 7, off to school at 7:30, last bell at 2 o'clock, punch in on the 3rd floor of Worumbo at 2:58, bag loose fabric for 8 hours, punch out at 11:02, get home around 12:15, eat a bowl of cereal, fall into bed, get up the next morning, do it all again. On a few occasions I worked double shifts, slept in my `60 Ford Galaxie (Dave's old car) for an hour or so before school, then slept through periods 5 \& 6 in the nurse's cubicle after lunch.

Once summer vacation came, things got easier. I was moved down to the dyehouse in the basement, for 1 thing, where it was 30 degrees cooler. My job was dyeing swatches of melton cloth purple or navy blue. I imagine there are still folks in New England with jackets in their closets dyed by yours truly. It wasn't the best summer I ever spent, but I managed to avoid being sucked into the machinery or stitching my fingers together with 1 of the heavy-duty sewing machines we used to belt the undyed cloth.

During 4th of July week, the mill closed. Employees with 5 years or more at Worumbo got the week off with pay. Those with fewer than 5 years were offered work on a crew that was going to clean the mill from top to bottom, including the basement, which hadn't been touched in 40 or 50 years. I probably would have agreed to work on this crew -- it was time \& a half -- but all the positions were filled long before the foreman got down to the high school kids, who'd be gone in September. When I got back to work the following week, 1 of the dyehouse guys told me I should have been there, it was wild. ``The rats down in that basement were big as cats,'' he said. ``Some of them, goddam if they weren't as big as \textit{dogs}.''

Rats as big as dogs! Yow!

1 day late in my final semester at college, finals over \& at loose ends, I recalled the dyehouse guy's story about the rats under the mill -- big as cats, some as big as \textit{dogs} -- \& started writing a story called ``Graveyard Shift.'' I was only passing the time on a late spring afternoon, but 2 months later \textit{Cavalier} magazine bought the story for \$200. I had sold 2 other stories previous to this, but they bad brought in a total of just \$65. This was 3 times that, \& at a single stroke. It took my breath away, it did. I was rich.

\fbox{\bf22} During the summer of 1969 I got a work-study job in the University of Maine library. That was a season both fair \& foul. In Vietnam, Nixon was executing his plan to end the war, which seemed to consist of bombing most of Southeast Asia into Kibbles `n Bits. ``Meet the new boss,'' The Who sang, ``same as the old boss.'' Eugene McCarthy was concentrating on his poetry, \& happy hippies wore bell-bottom pants \& tee-shirts that said things like \textsc{killing for peace is like fucking for chastity}. I had a great set of muttonchop sideburns. Creedence Clearwater Revival was singing ``Green River'' -- barefoot girls, dancing in the moonlight -- \& Kenny Rogers was still with The 1st Edition. Martin Luther King \& Robert Kennedy were dead, but Janis Joplin, Jim Morrison, Bob ``The Bear'' Hite, Jimi Hendrix, Cass Elliot, John Lennon, \& Elvis Presley were still alive \& making music. I was staying just off campus in Ed Price's Rooms (7 bucks a week, 1 change of sheets included). Men had landed on the moon, \& I had landed on the Dean's List. Miracles \& wonders abounded.

1 day in late June of that summer, a bunch of us library guys had lunch on the grass behind the university bookstore. Sitting between Paolo Silva \& Eddie Marsh was a trim girl with a raucous laugh, red-tinted hair, \& the prettiest legs I had ever seen, well-displayed beneath a short yellow skirt. She was carrying a copy of \textit{Soul on Ice}, by Eldridge Cleaver. I hadn't run across her in the library, \& I didn't believe a college student could utter such a wonderful, unafraid laugh. Also, heavy reading or no heavy reading, she swore like a millworker instead of a coed. (Having been a millworker, I was qualified to judge.) Her name was Tabitha Spruce. We got married a year \& a half later. We're still married, \& she has never let me forget that the 1st time I met her I thought she was Eddie Marsh's townie girlfriend. Maybe a book-reading waitress from the local pizza joint on her afternoon off.

\fbox{\bf23} It's worked. Our marriage has outlasted all of the world's leaders except for Castro, \& if we keep talking, arguing, making love, \& dancing to the Ramones -- gabba-gabba-hey -- it'll probably keep working. We came from different religions, but as a feminist Tabby has never been crazy about the Cathelics, where the man make the rules (including the God-given directive to always go in bareback) \& the women wash the underwear. \& while I believe in God I have no use for organized religion. We came from similar working-class backgrounds, we both ate meat, we were both political Democrats with typical Yankee suspicions of life outside New England. We were sexually compatible \& monogamous by nature. Yet what ties us most strongly are the words, the language, \& the work of our lives.

We met when we were working in a library, \& I fell in love with her during a poetry workshop in the fall of 1969, when I was a senior \& Tabby was a junior. I fell in love with her partly because I understood what she was doing with her work. I fell because \textit{she} understood what she was doing with it. I also fell because she was wearing a sexy black dress \& silk stockings, the kind that hook with garters.

I don't want to speak too disparagingly of my generation (actually I do, we had a chance ot change the world \& opted for the Home Shopping Network instead), but there was a view among the student writers I knew at that time that good writing came spontaneously, in an uprush of feeling that had to be caught at once; when you were building that all-important stairway to heaven, you couldn't just stand around with your hammer in your hand. \textit{Ars poetica} in 1969 was perhaps best expressed by a Donovan Leitch song that went, ``1st there is a mountain {\tt/} Then there is no mountain {\tt/} Then there is.'' Would-be poets were living in a dewy Tolkien-tinged world, catching poems out of the ether. It was pretty much unanimous: serious art came from $\ldots$ \textit{out there}! Writers were blessed stenographers taking divine dictation. I don't want to embarrass any of my old mates from that period, so here is a fictionalized version of what I'm talking about, created from bits of many actual poems:
\begin{center}
	i close my eyes\\in th dark i see\\Rodan Rimbaud\\in th dark\\ i swallow th cloth\\ of loneliness\\crow i am here\\raven i am here	
\end{center}
If you were to ask the poet what this poem \textit{meant}, you'd likely get a look of contempt. A slightly uncomfortable silence was apt to emanate from the rest. Certainly the fact that the poet would likely have been unable to tell you anything about the mechanics of creation would not have been considered important. If pressed, he or she might have said that there \textit{were} no mechanics, only that seminal spurt of feeling: 1st there is a mountain, then there is no mountain, then there is. \& if the resulting poem is sloppy, based on the assumption that such general words as ``loneliness'' mean the same thing to all of us -- hey man, so what, let go of that outdated bullshit \& just dig the heaviness. I didn't cop too much of this attitude (although I didn't dare say so out loud, at lest not in so many words), \& was overjoyed to find that the pretty girl in the black dress \& the silk stockings didn't cop too much of it, either. She didn't come right out \& say so, but she didn't need to. Her work spoke for her.

The workshop group met once or twice a week in the living room of instructor Jim Bishop's house, perhaps a dozen undergrads \& 3 or 4 faculty members working in a marvelous atmosphere of equality. Poems were typed up \& mimeographed in the English Department office on the day of each workshop. Poets read while the rest of us followed along on our copies. Here is 1 of Tabby's poems from that fall:
\begin{center}
	\textsc{A gradual canticle for Augustine}
	
	The thinnest bear is awakened in the winter\\by the sleep-laughter of locusts,\\by the dream-blustering of bees,\\by the honeyed scent of desert sands\\that the wind carries in her womb\\into the distant hills, into the houses of Cedar.\\The bear has heard a sure promise.\\Certain words are edible; they nourish\\more than snow heaped upon silver plates\\or ice overflowing golden bowls. Chips of ice\\from the mouth of a lover are not always better,\\Nor a desert dreaming always a mirage.\\The rising bear sings a gradual canticle\\woven of sand that conquers cities\\by a slow cycle. His praise seduces\\a passing wind, traveling to the sea\\wherein a fish, caught in a careful net,\\hears a bear's song in the cool-scented snow.
\end{center}
There was silence when Tabby finished reading. No one knew exactly how to react. Cables seemed to run through the poem, tightening the lines until they almost hummed. I found the combination of crafty diction \& delirious imagery exciting \& illuminating. Her poem also made me feel that I wasn't alone in my belief that good writing can be simultaneously intoxicating \& idea-driven. If stone-sober people can fuck like they're out of their minds -- can actually be out of their minds while caught in that throe -- why shouldn't writers be able to go bonkers \& still stay sane?

There was also a work-ethic in the poem that I liked, something that suggested writing poems (or stories, or essays) had as much in common with sweeping the floor as with mythy moments of revelation. There's a place in \textit{A Raisin in the Sun} where a character cries out: ``I want to fly! I want to touch the sun!'' to which his wife replies, ``1st eat your eggs.''

In the discussion that followed Tab's reading, it became clear to me that she understood her own poem. She knew exactly what she had meant to say, \& had said most of it. Saint Augustine (A.D. 354--430) she knew both as a Catholic \& as a history major. Augustine's mother (a saint herself) was a Christian, his father a pagan. Before his conversion, Augustine pursued both money \& women. Following it he continued to struggle with his sexual impulses, \& is known for the Libertine's Prayer, which goes: ``O Lord, make me chaste $\ldots$ but not yet.'' In his writing he focused on man's struggle to give up belief in self in favor of belief in God. \& he sometimes likened himself to a bear. Tabby has a way of tilting her chin down when she smiles -- it makes her look both wise \& severely cute. She did that then, I remember, \& said, ``Besides, I like bears.''

The canticle is gradual perhaps because the bear's awakening is gradual. The bear is powerful \& sensual, although thin because he is out of his time. In a way, Tabby said when called upon to explicate, the bear can be seen as a symbol of mankind's troubling \& wonderful habit of dreaming the right dreams at the wrong time. Such dreams are difficult because they're inappropriate, but also wonderful in their promise. The poem also suggests that dreams are powerful -- the bear's is strong enough to seduce the wind into bringing his song to a fish caught in a net.

I won't try to argue that ``A Gradual Canticle'' is a great poem (although I think it's a pretty good one). The point is that it was a reasonable poem in a hysterical time, one sprung from a writing ethic that resonated all through my heart \& soul.

Tabby was in 1 of Jim Bishop's rocking chairs that night. I was sitting on the floor beside her. I put my hand on her calf as she spoke, cupping the curve of warm flesh through her stocking. She smiled at me. I smiled back. Sometimes these things are not accidents. I'm almost sure of it.

\fbox{\bf24} We had 2 kids by the time we'd been married 3 years. They were neither planned nor unplanned; they came when they came, \& we were glad to have them. Naomi was prone to ear infections. Joe was healthy enough but never seemed to sleep. When Tabby went into labor with him, i was at a drive-in movie in Brewer with a friend -- it was a Memorial Day triple feature, 3 horror films. We were on the 3rd movie (\textit{The Corpse Grinders}) \& the 2nd sixpack when the guy in the office broke in with an announcement. There were still pole-speakers in those days; when you parked your car you lifted one off \& hung it over your window. The manager's announcement thus rang across the entire parking lot: ``\textsc{Steve King, please go home! Your wife is in labor! Steve King, please go home! Your wife is going to have the baby!}''

As I drove our old Plymouth toward the exit, a couple of hundred horns blared a satiric salute. Many people flicked their headlights on \& off, bathing me in a stuttery glow. My friend Jimmy Smith laughed so hard he slid into the footwell on the passenger side of the front seat. There he remained for most of the tripe back to Bangor, chortling among the beercans. When I got home, Tabby was calm \& packed. She gave birth to Joe less than 3 hours later. He entered the world easily. For the next 5 years or so, nothing else about Joe was easy. But he was a treat. Both of them were, really. Even when Naomi was tearing off the wallpaper above her crib (maybe she thought she was housekeeping) \& Joe was shitting in the wicker seat of the rocker we kept on the porch of our apartment on Sanford Street, they were a treat.

\fbox{\bf25} My mother knew I wanted to be a writer (with all those rejection slips hanging from the spike on my bedroom wall, how could she not?), but she encouraged me to get a teacher's credential ``so you'll have something to fall back on.''

``You may want to get married, Stephen, \& a garret by the Seine is only romantic if you're a bachelor,'' she'd said once. ``It's no place to raise a family.''

I did as she suggested, entering the College of Education at UMO \& emerging 4 years later with a teacher's certificate $\ldots$ sort of like a golden retriever emerging from a pond with a dead duck in its jaws. It was dead, all right. I couldn't find a teaching job \& so went to work at New Franklin Laundry for wages not much higher than those I had been making at Worumbo Mills \& Weaving 4 years before. I was keeping my family in a series of garrets which overlooked not the Seine but some of Bangor's less appetizing streets, the ones where the police cruisers always seemed to show up at 2 o'clock on Saturday morning.

I never saw personal laundry at New Franklin unless it was a ``fire order'' being paid for by an insurance company (most fire orders consisted of clothes that \textit{looked} okay but smelled like barbecued monkeymeat). The greater part of what I loaded \& pulled were motel sheets from Maine's coastal towns \& table linen from Maine's coastal restaurants. The table linen was desperately nasty. When tourists go out to dinner in Maine, they usually want clams \& lobster. Mostly lobster. By the time the tablecloths upon which these delicacies had been served reached me, they stank to high heaven \& were often boiling with maggots. The maggots would try to crawl up your arms as you loaded the washers; it was as if the little fuckers knew you were planning to cook them. I thought I'd get used to them in time but I never did. The maggots were bad; the smell of decomposing clams \& lobster-meat was even worse. \textit{Why are people such slobs?} I would wonder, loading feverish linens from Testa's of Bar Harbor into my machines. \textit{Why are people such fucking slobs?}

Hospital sheets \& linens were even worse. These also crawled with maggots in the summertime, but it was blood they were feeding on instead of lobster-meat \& clam-jelly. Clothes, sheets, \& pillowslips deemed to be infected were stuffed inside what we called ``plague-bags'' which dissolved when the hot water hit them, but blood was not, in those times, considered to be especially dangerous. There were often little extras in the hospital laundry; those loads were like nasty boxes of Cracker Jacks with weird prizes in them. I found a steel bedpan in 1 load \& a pair of surgical shears in another (the bedpan was of no practical use, but the shears were a damned handy kitchen implement). Ernest ``Rocky'' Rockwell, the guy I worked with, found \$20 in a load from Eastern Maine Medical Center \& punched out at noon to start drinking. (Rocky referred to quitting time as ``Slitz o'clock.'')

On 1 occasion I heard a strange clicking from inside 1 of the Washex 3-pockets which were my responsibility. I hit the Emergency Stop button, thinking the goddam thing was stripping its gears or something. I opened the doors \& hauled out a huge wad of dripping surgical tunics \& green caps, soaking myself in the process. Below them, lying scattered across the colander-like inner sleeve of the middle pocket, was what looked like a complete set of human teeth. It crossed my mind that they would make an interesting necklace, then I scooped them out \& tossed them in the trash. My wife has put up with a lot from me over the years, but her sense of humor stretches only so far.

\fbox{\bf26} From a financial point of view, 2 kids were probably 2 too many for college grads working in a laundry \& the 2nd shift at Dunkin' Donuts. The only edge we had came courtesy of magazines like \textit{Dude, Cavalier, Adam}, \& \textit{Swank} -- what my Uncle Oren used to call ``the titty books.'' By 1972 they were showing quite a lot more than bare breasts \& fiction was on its way out, but I was lucky enough to ride the last wave. I wrote after work; when we lived on Grove Street, which was close to the New Franklin, I would sometimes write a little on my lunch hour, too. I suppose that sounds almost impossibly Abe Lincoln, but it was no big deal -- I was having fun. Those stories, grim as some of them were, served as brief escapes from the boss, Mr. Brooks, \& Harry the floor-man.

Harry had hooks instead of hands as a result of a tumble into the sheet-mangler during World War II (he was dusting the beams above the machine \& felt off). A comedian at heart, he would sometimes duck into the bathroom \& run water from the cold tap over 1 hook \& water from the hot tap over the other. Then he'd sneak up behind you while you were loading laundry \& lay the steel hooks on the back of your neck. Rocky \& I spent a fair amount of time speculating on how Harry accomplished certain bathroom cleanup activities. ``Well,'' Rocky said 1 day while we were drinking our lunch in his car, ``at least he don't need to wash his hands.''

There were times -- especially in summer, while swallowing my afternoon salt-pill -- when it occurred to me that I was simply repeating my mother's life. Usually this thought struck me as funny. But if I happened to be tired, or if there were extra bills to pay \& no money to pay them with, it seemed awful. I'd think \textit{This isn't the way our lives are supposed to be going}. Then I'd think \textit{Half the world has the same idea}.

The stories I sold to the men's magazines between Aug of 1970, when I got my 200-dollar check for ``Graveyard Shift,'' \& the winter of 1973--1974 were just enough to create a rough sliding margin between us \& the welfare office (my mother, a Republician all her life, had communicated her deep horror of ``going on the county'' to me; Tabby had some of that same horror).

My clearest memory of those days is of our coming back to the Grove Street apartment 1 Sunday afternoon after spending the weekend at my mother's house in Durham -- this would have been right around the time the symptoms of the cancer which killed her started to show themselves. I have a picture from that day -- Mom, looking both tired \& amused, is sitting in a chair in her dooryard, holding Joe in her lap while Naomi stands sturdily beside her. Naomi wasn't so sturdy by Sunday afternoon, however; she had come down with an ear infection, \& was burning with fever.

Trudging from the car to our apartment building on that summer afternoon was a low point. I was carrying Naomi \& a tote-bag full of baby survival equipment (bottles, lotions, diapers, sleep suits, undershirts, socks) while Tabby carried Joe, who had spit up on her. She was dragging a sack of dirty diapers behind her. We both knew Naomi needed \textsc{the pink stuff}, which was what we called liquid amoxicillin. \textsc{The pink stuff} was expensive, \& we were broke. I mean stony.

I managed to get the downstairs door open without dropping my daughter \& was easing her inside (she was so feverish she glowed against my chest like a banked coal) when I saw there was an envelope sticking out of our mailbox -- a rare Saturday delivery. Young marrieds don't get much mail; everyone but the gas \& electric companies seems to forget they are alive. I snagged it, praying it wouldn't turn out to be another bill. It wasn't. My friends at the Dugent Publishing Corporation, purveyors of \textit{Cavalier} \& many other fine adult publications, had sent me a check for ``Sometimes They Come Back,'' a long story I hadn't believed would sell anywhere. The check was for \$500, easily the largest sum I'd ever received. Suddenly we were able to afford not only a doctor's visit \& a bottle of \textsc{the pink stuff}, but also a nice Sunday-night meal. \& I imagine that once the kids were asleep, Tabby \& I got friendly.

I think we had a lot of happiness in those days, but we were scared a lot, too. We weren't much more than kids ourselves (as the saying goes), \& being friendly helped keep the mean reds away. We took care of ourselves \& the kids \& each other as best we could. Tabby wore her pink uniform out to Dunkin's Donuts \& called the cops when the drunks who came in for coffee got obstreperous. I washed motel sheets \& kept writing 1-reel horror movies.

\fbox{\bf27} By the time I started \textit{Carrie}, I had landed a job teaching English in the nearby town of Hampden. I would be paid \$6400{\tt/}year, which seemed an unthinkable sum after earning a dollar-60 a hour at the laundry. If I'd done the math, being careful to add in all the time spent in after-school conferences \& correcting papers at home, I might have seen it was a very thinkable sum indeed, \& that our situation was worse than ever. By the late winter of 1973 we were living in a doublewide trailer in Hermon, a little town west of Bangor. (Much later, when asked to do the \textit{Playboy} Interview, I called Hermon ``The asshole of the world.'' Hermonites were infuriated by that, \& I hereby apologize. Hermon is really no more than the armpit of the world.) I was driving a Buick with transmission problems we couldn't afford to fix, Tabby was still working at Dunkin' Donuts, \& we had no telephone. We simply couldn't afford the monthly charge. Tabby tried her hand at confession stories during that period (``Too Pretty to Be a Virgin'' -- stuff like that), \& got personal responses of the this-isn't-quite-right-for-us-but-try-again type immediately. She would have broken through if given an extra hour or 2 in every day, but she was stuck with the usual 24. Besides, any amusement value the confession-mag formula (it's called the 3 R's -- Rebellion, Ruin, \& Redemption) might have had for her at the start wore off in a hurry.

I wasn't having much success with my own writing, either. Horror, science fiction, \& crime stories in the men's magazines were being replaced by increasingly graphic tales of sex. That was part of the trouble, but not all of it. The bigger deal was that, for the 1st time in my life, writing was \textit{hard}. The problem was the teaching. I liked my coworkers \& loved the kids -- even the Beavis \& Butt-Head types in Living with English could be interesting -- but by most Friday afternoons I felt as if I'd spent the week with jumper cables clamped to my brain. If I ever came close to despairing about my future as a writer, it was then. I could see myself 30 years on, wearing the same shabby tweed coats with patches on the elbows, potbelly rolling over my Gap khakis from too much beer. I'd have a cigarette cough from too many packs of Pall Malls, thicker glasses, more dandruff, \& in my desk drawer, 6 or 7 unfinished manuscripts which I would take out \& tinker with from time to time, usually when drunk. If asked what I did in my spare time, I'd tell people I was writing a book -- what else does \textit{any} self-respecting creative-writing teacher do with his or her spare time? \& of course I'd lie to myself, telling myself there was still time, it wasn't too late, there were novelists who didn't get started until they were 50, hell, even 60. Probably plenty of them.

My wife made a crucial difference during those 2 years I spent teaching at Hampden (\& washing sheets at New Franklin Laundry during the summer vacation). If she had suggested that the time I spent writing stories on the front porch of our rented house on Pond Street or in the laundry room of of our rented trailer on Klatt Road in Hermon was wasted time, I think a lot of the heart would have gone out of me. Tabby never voiced a single doubt, however. Her support was a constant, 1 of the few good things I could take as a given. \& whenever I see a 1st novel dedicated to a wife (or a husband), I smile \& think, \textit{There's someone who knows}. Writing is a lonely job. Having someone who believes in your makes a lot of difference. They don't have to make speeches. Just believing is usually enough.

\fbox{\bf28} While he was going to college my brother Dave worked summers as a janitor at Brunswick High, his old alma mater. For part of 1 summer I worked there, too. I can't remember which year, only that it was before I met Tabby but after I started to smoke. That would have made me 19 or 20, I suppose. I got paired with a guy named Harry, who wore green fatigues, a big keychain, \& walked with a limp. (He \textit{did} have hands instead of hooks, however.) 1 lunch hour Harry told me what it had been like to face a Japanese \textit{banzai} charge on the island of Tarawa, all the Japanese officers waving swords made out of Maxwell House coffee cans, all the screaming enlisted men behind them stoned out of their gourds \& smelling of burned poppies. Quite a raconteur was my pal Harry.

1 day he \& I were supposed to scrub the rust-strains off the walls in the girls' shower. I looked around the locker room with the interest of a Muslim youth who for some reason finds himself deep within the women's quarters. It was the same as the boys' locker room, \& yet completely different. There were no urinals, of course, \& there were 2 extra metal boxes on the tile walls -- unmarked, \& the wrong size for paper towels. I asked what was in them. ``Pussy-plugs,'' Harry said. ``For them certain days of the month.''

I also noticed that the showers, unlike those in the boys' locker room, had chrome U-rings with pink plastic curtains attached. You could actually shower in privacy. I mentioned this to Harry, \& he shrugged. ``I guess young girls are a bit more shy about being undressed.''

This memory came back to me 1 day while I was working at the laundry, \& I started seeing the opening scene of a story: girls showering in a locker room where there were no U-rings, pink plastic curtains, or privacy. \& this 1 girl starts to have her period. Only she doesn't know what it is, \& the other girls -- grossed out, horrified, amused -- start pelting her with sanitary napkins. Or with tampons, which Harry had called pussy-plugs. The girl begins to scream. All that blood! She thinks she's dying, that the other girls are making fun of her even while she's bleeding to death $\ldots$ she reacts $\ldots$ fights back $\ldots$ but how?

I'd read an article in \textit{Life} magazine some years before, suggesting that at least some reported poltergeist activity might actually be telekinetic phenomena -- telekinesis being the ability to move objects just by thinking about them. There was some evidence to suggest that young people might have such powers, the article said, especially girls in early adolescence, right around the time of their 1st --

Pow! 2 unrelated ideas, adolescent cruelty \& telekinesis, came together, \& I had an idea. I didn't leave my post at Washex \#2, didn't go running around the laundry waving my arms \& shouting ``Eureka!,'' however. I'd had many other ideas as good \& some that were better. Still, I thought I might have the basis for a good \textit{Cavalier} yarn, with the possibility of \textit{Playboy} lurking in the back of my mind. \textit{Playboy} paid up to \$2000 for short fiction. 2000 bucks would buy a new transmission for the Buick with plenty left over for groceries. The story remained on the back burner for awhile, simmering away in that place that's not quite the conscious but not quite the subconscious, either. I had started my teaching career before I sat down 1 night to give it a shot. I did 3 single-spaced pages of a 1st draft, then crumpled them up in disgust \& threw them away.

The next night, when I came home from school, Tabby had the pages. She'd spied them while emptying my wastebasket, had shaken the cigarette ashes off the crumpled balls of paper, smoothed them out, \& sat down to read them. She wanted me to go on with it, she said. She wanted to know the rest of the story. I told her I didn't know jack-shit about high school girls. She said she'd help me with that part. She had her chin tilted down \& was smiling in that severely cute way of hers. ``You've got something here,'' she said. ``I really think you do.''

\fbox{\bf29} I never got to like Carrie White \& I never trusted Sue Snell's motives in sending her boyfriend to the prom with her, but I \textit{did} have something there. Like a whole career. Tabby somehow knew it, \& by the time I had piled up 50 single-spaced pages, I knew it, too. For 1 thing, I didn't think any of the characters who went to Carrie White's prom would ever forget it. Those few who lived through it, that was.

I had written 3 other novels before \textit{Carrie -- Rage, The Long Walk}, \& \textit{The Running Man} were later published. \textit{Rage} is the most troubling of them. \textit{The Long Walk} may be the best of them. But none of them taught me the things I learned from Carrie White. The most important is that the writer's original perception of a character or characters may be as erroneous as the reader's. Running a close second was the realization that stopping a piece of work just because it's hard, either emotionally or imaginatively, is a bad idea. Sometimes you have to go on when you don't feel like it, \& sometimes you're doing good work when it feels like all you're managing is to shovel shit from a sitting position.

Tabby helped me, beginning with the information that the sanitary-napkin dispensers in high schools were usually not coin-op -- faculty \& administration didn't like the idea of girls' walking around with blood all over their skirts just because they happened to come to school short a quarter, my wife said. \& I also helped myself, digging back to my memories of high school (my job teaching English didn't help; I was 26 by then, \& on the wrong side of the desk), remembering what I knew about the 2 loneliest, most reviled girls in my class -- how they looked, how they acted, how they were treated. Very rarely in my career have I explored more distasteful territory.

I'll call 1 of these girls Sondra. She \& her mother lived in a trailer home not too far from me, with their dog, Cheddar Cheese. Sondra had a burbly, uneven voice, as if she were always speaking through a throatful of tightly packed phlegm. She wasn't fat, but her flesh had a loose, pale look, like the undersides of some mushrooms. Her hair clung to her pimply cheeks in tight Little Orphan Annie curls. She had no friends (except for Cheddar Cheese, I guess). 1 day her mother hired me to move some furniture. Dominating the trailer's living room was a nearly life-sized crucified Jesus, eyes turned up, mouth turned down, blood dribbling from beneath the crown of thorns on his head. He was naked except for a rag twisted around his hips \& loins. Above this bit of breechclout were the hollowed belly \& the jutting ribs of a concentration-camp inmate. It occurred to me that Sondra had grown up beneath the agonal gaze of this dying god, \& doing so had undoubtedly played a part in making her what she was when I knew her: a timid \& homely outcast who went scuttling through the halls of Lisbon High like a frightened mouse.

``That's Jesus Christ, my Lord \& Savior,'' Sondra's mother said, following my gaze. ``Have \textit{you} been saved, Steve?''

I hastened to tell her I was saved as saved could be, although I didn't think you could ever be good enough to have \textit{that} version of Jesus intervene on your behalf. The pain had driven him out of his mind. You could see it on his face. If \textit{that} guy came back, he probably wouldn't be in a saving mood.

The other girl I'll call Dodie Franklin, only the other girls called her Dodo or Doodoo. Her parents were interested in only 1 thing, \& that was entering contests. They were good at them, too; they had won all sorts of odd stuff, including a year's supply of 3 Diamonds Brand Fancy Tuna \& Jack Benny's Maxwell automobile. The Maxwell sat off to the left of their house in that part of Durham known as Southwest Bend, gradually sinking into the landscape. Every year or 2, 1 of the local papers -- the Portland \textit{Press-Herald}, the Lewiston \textit{Sun}, the Lisbon \textit{Weekly Enterprise} -- would do a piece on all the weird shit Dodie's folks had won in raffles \& sweepstakes \& giant prize drawings. Usually there would a photo of the Maxwell, or Jack Benny with his violin, or both.

Whatever the Franklins might have won, a supply of clothes for growing teenagers wasn't part of the haul. Dodie \& her brother Bill wore the same stuff every day for the 1st year \& a half of high school: black pants \& a short-sleeved checked sport shirt for him, a long black skirt, gray knee-socks, \& a sleeveless white blouse for her. Some of my readers may not believe I am being literal when I say \textit{every day}, but those who grew up in country towns during the 50s \& 60s will know that I am. In the Durham of my childhood, life wore little or any makeup. I went to school with kids who wore the same neck dirt for months, kids whose skin festered with sores \& rashes, kids with the eerie dried-apple-doll faces that result from untreated burns, kids who were sent to school with stones in their dinnerbuckets \& nothing but air in their Thermoses. It wasn't Arcadia; for the most part it was Dogpatch with no sense of humor.

Dodie \& Bill Franklin got on all right at Durham Elementary, but high school meant a much bigger town, \& for children like Dodie \& Bill, Lisbon Falls meant ridicule \& ruin. We watched in amusement \& horror as Bill's sport shirt faded \& began to unravel from the short sleeves up. He replaced a missing button with a paperclip. Tape, carefully colored black with a crayon to match his pants, appeared over a rip behind 1 knee. Dodie's sleeveless white blouse began to grow yellow with wear, age, \& accumulated sweat-stains. As it grew thinner, the straps of her bra showed through more \& more clearly. The other girls made fun of her, at 1st behind her back \& then to her face. Teasing became taunting. The boys weren't a part of it; we had Bill to take care of (yes, I helped -- not a whole lot, but I was there). Dodie had it worse, I think. The girls didn't just laugh at Dodie; they hated her, too. Dodie was everything they were afraid of.

After Christmas vacation of our sophomore year, Dodie came back to school resplendent. The dowdy old black skirt had been replaced by a cranberry-colored one that stopped at her knees instead of halfway down her shins. The tatty knee-socks had been replaced by nylon stockings, which looked pretty good because she had finally shaved the luxuriant mat of black hair off her legs. The ancient sleeveless blouse had given way to a soft wool sweater. She'd even had a permanent. Dodie was a girl transformed, \& you could see by her face that she knew it. I have no idea if she saved for those new clothes, if they were given to her for Christmas by her parents, or if she went through a hell of begging that finally bore dividends. It doesn't matter, because mere clothes changed nothing. The teasing that day was worse than ever. Her peers had no intention of letting her out of the box they'd put her in; she was punished for even trying to break free. I had several classes with her, \& was able to observe Dodie's ruination at 1st hand. I saw her smile fade, saw the light in her eyes 1st dim \& then go out. By the end of the day she was the girl she'd been before Christmas vacation -- a dough-faced \& freckle-cheeked wraith, scurrying through the halls with her eyes down \& her books clasped to her chest.

She wore the new skirt \& sweater the next day. \& the next. \& the next. When the school year ended she was still wearing them, although by then the weather was much too hot for wool \& there were always beads of sweat at her temples \& on her upper lip. The home permanent wasn't repeated \& the new clothes took on a matted, dispirited look, but the teasing had dropped back to its pre-Christmas levels \& the taunting stopped entirely. Someone made a break for the fence \& had to be knocked down, that was all. Once the escape was foiled \& the entire company of prisoners was once more accounted for, life could go back to normal.

Both Sondra \& Dodie was dead by the time I started writing \textit{Carrie}. Sondra moved out of the trailer in Durham, out from beneath the agonal gaze of the dying savior, \& into an apartment in Lisbon Falls. She must have worked somewhere close by, probably in 1 of the mills or shoe factories. She was epileptic \& died during a seizure. She lived alone, so there was no one to help her when she went down with her head bent the wrong way. Dodie married a TV weatherman who gained something of a reputation in New England for his drawling downeast delivery. Following the birth of a child -- I think it was their 2nd -- Dodie went into the cellar \& put a .22 bullet in her abdomen. It was a lucky shot (or unlucky, depending on your point of view, I guess), hitting the portal vein \& killing her. In town they said it was postpartum depression, how sad. Myself, I suspected high school hangover might have had something to do with it.

I never liked Carrie, that female version of Eric Harris \& Dylan Klebold, but through Sondra \& Dodie I came at last to understand her a little. I pitied her \& I pitied her classmates as well, because I had been 1 of them once upon a time.

\fbox{\bf30} The manuscript of \textit{Carrie} went off to Doubleday, where I had made a friend named William Thompson. I pretty much forgot about it \& moved on with my life, which at that time consisted of teaching school, raising kids, loving my wife, getting drunk on Friday afternoons, \& writing stories.

My free period that semester was 5, right after lunch. I usually spent it in the teacher's room, grading papers \& wishing I could stretch out on the couch \& take a nap -- in the early afternoon I have all the energy of a boa constrictor that's just swallowed a goat. The intercom came on \& Colleen Sites in the office asked if I was there. I said I was, \& she asked me to come to the office. I had a phone call. My wife.

The walk from the teachers' room in the lower wing to the main office seemed long even with classes in session \& the halls mostly empty. I hurried, not quite running, my heart beating hard. Tabby would have had to dress the kids in their boots \& jackets to use the neighbors' phone, \& I could think of only 2 reasons she might have done so. Either Joe or Naomi had fallen off the stoop \& broken a leg, or I had sold \textit{Carrie}.

My wife, sounding out of breath but deliriously happy, read me a telegram. Bill Thompson (who would later go on to discover a Mississippi scribbler named John Grisham) had sent it after trying to call \& discovering the Kings no longer had a phone. \textsc{Congratulations}, it read. \textsc{Carrie officially a doubleday book. is \$2500 advance okay? The future lies head. Love, Bill}.

\$2500 was a very small advance, even for the early 70s, but I didn't know that \& had no literary agent to know it for me. Before it occurred to me that I might actually need an agent, I had generated well over \$3000000' worth of income, a good deal of it for the publisher. (The standard Doubleday contract in those days was better than indentured servitude, but not much.) \& my little high school horror novel marched toward publication with excruciating slowness. Although it was accepted in late Mar or early Apr of 1973, publication wasn't slated until the spring of 1974. This wasn't unusual. In those days Double day was an enormous fiction-mill churning out mysteries, romances, science fiction yarns, \& double D westerns at a rate of 50 or more a month, all of this in addition to a robust frontlist including books by heavy hitters like Leon Uris \& Allen Drury. I was only 1 small fish in a very busy river.

Tabby asked if I could quit teaching. I told her no, not based on a \$2500 advance \& only nebulous possibilities beyond that. If I'd been on my own, maybe (hell, \textit{probably}). But with a wife \& 2 kids? Not happening. I remember the 2 of us lying in bed that night, eating toast \& talking until the small hours of the morning. Tabby asked me how much we'd make if Doubleday was able to sell paperback reprint rights to \textit{Carrie}, \& I said I didn't know. I'd read that Mario Puzo had just scored a huge advance for paperback rights to \textit{The Godfather} -- \$400000 according to the newspaper -- but I didn't believe \textit{Carrie} would fetch anything near that, assuming it sold to paperback at all.

Tabby asked -- rather timidly for my normally outspoken wife -- if \textit{I} thought the book would find a paperback publisher. I told her I thought the chances were pretty good, maybe 7 or 8 in 10. She asked how much it might bring. I said my best guess would be somewhere between \$10000 \& \$60000.

``\$60000?'' She sounded almost shocked. ``Is that much even possible?''

I said it was -- not \textit{likely}, perhaps, but \textit{possible}. I also reminded her that my contract specified a 50-50 paperback split, which meant that if Ballantine or Dell \textit{did} pay 60 grand, we'd only get 30. Tabby didn't dignify this with a reply -- she didn't have to. \$30000 was what I could expect to make in 4 years of teaching, even with annual salary increases thrown in. It was a lot of money. Probably just pie in the sky, but it was a night for dreaming.

\fbox{\bf31} \textit{Carrie} inched along toward publication. We spent the advance on a new car (a standard shift which Tabby hated \& reviled in her most colorful millworker's language) \& I signed a teaching contract for the 1973--1974 academic year. I was writing a new novel, a peculiar combination of \textit{Peyton Place} \& \textit{Dracula} which I called \textit{Second Coming}. We had moved to a ground-floor apartment back in Bangor, a real pit, but we were in town again, we had a car covered by an actual warranty, \& we had a telephone.

To tell you the truth, \textit{Carrie} had fallen off my radar screen almost completely. The kids were a handful, both the ones at school \& the ones at home, \& I had begun to worry about my mother. She was 61, still working at Pineland Training Center \& as funny as ever, but Dave said she didn't feel very well a lot of the time. Her bedside table was covered with prescription painkillers, \& he was afraid there might be something seriously wrong with her. ``She's always smoked like a chimney, you know,'' Dave said. He was a great one to talk, since he smoked like a chimney himself (so did I, \& how my wife hated the expense \& the constant ashy dirt of it), but I knew what he meant. \& although I didn't live as close to her as Dave \& didn't see her as often, the last time I \textit{had} seen her I could tell she had lost weight.

``What can we do?'' I asked. Behind the question was all we knew of our mother, who ``kept herself to herself,'' as she liked to say. The result of that philosophy was a vast gray space where other families have histories; Dave \& I knew almost nothing about our father or his family, \& little enough about or own mother's past, which included an incredible (to me, at least) 8 dead brothers \& sisters \& her own failed ambition to become a concert pianist (she did play the organ on some of the NBC radio soaps \& Sunday church shows during the war, she claimed).

``We can't do anything,'' Dave replied, ``until she asks.''

1 Sunday not long after that call, I got another one from Bill Thompson at Doubleday. I was alone in the apartment; Tabby had packed the kids off to her mother's for a visit, \& I was working on the new book, which I thought of as \textit{Vampires in} Our Town.

``Are you sitting down?'' Bill asked.

``No,'' I said. Our phone hung on the kitchen wall, \& I was standing in the doorway between the kitchen \& the living room. ``Do I need to?''

``You might,'' he said. ``The paperback rights to \textit{Carrie} went to Signet Books for \$400000.''

When I was a little kid, Daddy Guy had once said to my mother: ``Why don't you shut that kid up, Ruth? When Stephen opens his mouth, all his guts fall out.'' It was true then, has been true all my life, but on that Mother's Day in May of 1973 I was completely speechless. I stood there in the doorway, casting the same shadow as always, but I couldn't talk. Bill asked if I was still there, kind of laughing as he said it. He knew I was.

I hadn't heard him right. Couldn't have. The idea allowed me to find my voice again, at least. ``Did you say it went for \$40000?''

``\$\textit{400000},'' he said. ``Under the rules of the road'' -- meaning the contract I'd signed --``200 K of it's yours. Congratulations, Steve.''

I was still standing in the doorway, looking across the living room toward our bedroom \& the crib where Joe slept. Our place on Sanford Street rented for \$90 a month \& this man I'd only met once face-to-face was telling me I'd just won the lottery. The strength ran out of my legs. I didn't fall, exactly, but I kind of whooshed down to a sitting position there in the doorway.

``Are you sure?'' I asked Bill.

He said he was. I asked him to say the number again, very slowly \& very clearly, so I could be sure I hadn't misunderstood. He said the number was a 4 followed by 5 0s. ``After that a decimal point \& 2 more 0s,'' he added.

We talked for another half an hour, but I don't remember a single word of what we said. When the conversation was over, I tried to call Tabby at her mother's. Her youngest sister, Marcella, said Tab had already left. I walked back \& forth through the apartment in my stocking feet, exploding with good news \& without an ear to hear it. I was shaking all over. At last I pulled on my shoes \& walked downtown. The only store that was open on Bangor's Main Street was LaVerdiere's Drug. I suddenly felt that I had to buy Tabby a Mother's Day present, something wild \& extravagant. I tried, but here's 1 of life's true facts: there's nothing really wild \& extravagant for sale at LaVerdiere's. I did the best I could. I got her a hair-dryer.

When I got back home she was in the kitchen, unpacking the baby bags \& singing along with the radio. I gave her the hair-dryer. She looked at it as if she'd never seen one before. ``What's this for?'' she asked.

I took her by the shoulders. I told her about the paperback sale. She didn't appear to understand. I told her again. Tabby looked over my shoulder at our shitty little 4-room apartment, just as I had, \& began to cry.

\fbox{\bf32} I got drunk for the 1st time in 1966. This was on the senior class trip to Washington. We went on a bus, about 40 kids \& 3 chaperones (1 of them was old Cue-Ball, as a matter of fact), \& spent the 1st night in New York, where the drinking age was then 18. Thanks to my bad ears \& shitty tonsils, I was almost 19. Room to spare.

A bunch of us more adventurous boys found a package store around the corner from the hotel. I cast an eye over the shelves, aware that my spending money was far from a fortune. There was too much -- too many bottles, too many brands, too many prices over \$10. Finally I gave up \& asked the guy behind the counter (the same bald, bored-looking, gray-coated guy who has, I'm convinced, sold alcohol virgins their 1st bottle since the dawn of commerce) what was cheap. Without a world, he put a pint of Old Log Cabin whiskey down on the Winston mat beside the cash register. The sticker on the label said \$1.95. The price was right.

I have a memory of being led onto the elevator later that night -- or maybe it was early the next morning -- by Peter Higgins (Old Cue-Ball's son), Butch Michaud, Lenny Partridge, \& John Chizmar. This memory is more like a scene from a TV show than a real memory. I seem to be outside of myself, watching the whole thing. There's just enough of me left inside to know that I am globally, perhaps even galactically, fucked up.

The camera watches as we go up to the girls' floor. The camera watches as I am propelled up \& down the hall, a kind of rolling exhibit. An amusing one, it seems. The girls are in nighties, robes, curlers, cold cream. They are all laughing at me, but their laughter seems good-natured enough. The sound is muted, as if I am hearing them through cotton. I am trying to tell Carole Lemke that I love the way she wears her hair, \& that she has the most beautiful blue eyes in the world. What comes out is something like ``Uggin-wugin-blue eyes, wuggin-ruggin-whole world.'' Carole laughs \& nods as if she understands completely. I am very happy. The world is seeing an asshole, no doubt, but he is a \textit{happy} asshole, \& everyone loves him. I spend several minutes trying to tell Gloria Moore that I've discovered The Secrete Life of Dean Martin.

At some point after that I am in my bed. The bed holds still but the room starts to spin around it, faster \& faster. It occurs to me that it's spinning like the turntable of my Webcor phonograph, on which I used to play Fats Domino \& now play Dylan \& the Dave Clark 5. The room is the turntable, I am the spindle, \& pretty soon the spindle is going to start tossing its platters.

I go away for a little bit. When I wake up, I'm on my knees in the bathroom of the double room I'm sharing with my friend Louis Purington. I have no idea how I got in there, but it's good that I did because the toilet is full of bright yellow puke. \textit{Looks like Niblets}, I think, \& that's all it takes to get me going again. Nothing comes up but whiskey-flavored strings of spit, bu my head feels like it's going to explode. I can't walk. I crawl back to bed with my sweaty hair hanging in my eyes. \textit{I'll feel better tomorrow}, I think, \& then I go away again.

In the morning my stomach has settled a little but my diaphragm is sore from vomiting \& my head is throbbing like a mouthful of infected teeth. My eyes have turned into magnifying glasses; the hideously bright morning light coming in through the hotel windows is being concentrated by them \& will soon set my brains on fire.

Participating in that day's schedule activities -- a walk to Times Square, a boat ride to the Statute of Liberty, a climb to the top of the Empire State Building -- is out of the question. Walking? Urk. Boats? Double urk. Elevators? Urk to the 4th power. Christ, I can hardly move. I make some soft of feeble excuse \& spend most of the day in bed. By late afternoon I'm feeling a little better. I dress, creep down the hall to the elevator, \& descend to the 1st floor. Eating is still impossible, but I believe I'm ready for a ginger ale, a cigarette, \& a magazine. \& who should I see in the lobby, sitting in a chair \& reading a newspaper, but Mr. Earl Higgins, alias Old Cue-Ball. I pass him as silently as I can, but it's no good. When I come back from the gift shop he's sitting with his newspaper in his lap, looking at me. I feel my stomach drop. Here is more trouble with the principal, probably even worse than the trouble I got into over \textit{The Village Vomit}. He calls me over \& I discover something interesting: Mr. Higgins is actually an okay guy. He bounced me pretty hard over my joke newspaper, but perhaps Miss Margitan had insisted on that. \& I'd just been 16, after all. On the day of my 1st hangover I'm going on 19, I've been accepted at the state university, \& I have a mill job waiting for me when the class trip is over.

``I understand you were too sick to tour New York with the rest of the boys \& girls,'' Old Cue-Ball says. He eyes me up \& down.

I say that's right, I'd been sick.

``A shame for you to miss the fun,'' Old Cue-Ball says. ``Feeling better now?''

Yes, I was feeling better. Probably stomach flu, 1 of those 24-hour bugs.

``I hope you won't get that bug again,'' he says. ``At least not on this trip.'' He looks at me for a moment longer, his eyes asking if we understand each other.

``I'm sure I won't,'' I say, meaning it. I know what drunk is like, now -- a vague sense of roaring goodwill, a clearer sense that most of your consciousness is out of your body, hovering like a camera in a science fiction movie \& filming everything, \& then the sickness, the puking, the aching head. No, I won't get that bug again, I tell myself, not on this trip, not ever. Once is enough, just to find out what it's like. Only an idiot would make a 2nd experiment, \& only a lunatic -- a \textit{masochistic} lunatic -- would make booze a regular part of his life.

The next day we go on to Washington, making 1 stop in Amish country on the way. There's a liquor store near where the bus parks. I go in \& look around. Although the drinking age in Pennsylvania is 21, I must look easily that in my 1 good suit \& Fazza's old black overcoat -- in fact, I probably look like a freshly released young convict, tall \& hungry \& very likely not bolted together right. The clerk sells me a 5th of 4 Roses without asking to see any ID, \& by the time we stop for the night I'm drunk again.

10 years or so later I'm in an Irish saloon with Bill Thompson. We have lots to celebrate, not the least of which is the completion of my 3rd book, \textit{The Shining}. That's the on which just happens to be about an alcoholic writer \& ex-school teacher. It's July, the night of the All-Star baseball game. Our plan is to eat a good old-fashioned meal from the dishes set out on the steam table, then get shitfaced. We begin with a couple at the bar, \& I start reading all the signs. \textsc{Have a Manhattan in Manhattan}, says one. \textsc{Tuesdays are twofors}, says another. \textsc{Work is the cure of the drinking class}, says a 3rd. \& there, right in front of me, is one which reads: \textsc{Early bird special! Screwdrivers a buck Monday--Friday 8--10 a.m.}

I motion to the bartender. He comes over. He's bald, he's wearing a gray jacket, he could be the guy who sold me my 1st pint back in 1966. Probably he is. I point to the sign \& ask, ``Who comes in at 8:15 in the morning \& orders a screwdriver?''

I'm smiling but he doesn't smile back. ``College boys,'' he replies. ``Just like you.''

\fbox{\bf33} In 1971 or '72, Mom's sister Carolyn Weimer died of breast cancer. My mother \& my Aunt Ethelyn (Carolyn's twin) flew out to Aunt Cal's funeral in Minnesota. It was the 1st time my mother had flown in 20 years. On the plane trip back, she began to bleed profusely from what she would have called ``her privates.'' Although long past her change of life by that point, she told herself it was simply 1 final menstrual period. Locked in the tiny bathroom of a bouncing TWA jet, she stanched the bleeding with tampons (\textit{plug it up, plug it up}, as Sue Snell \& her friends might have cried), then returned to her seat. She said nothing to Ethelyn \& nothing to David \& me. She didn't go to see Joe Mendes in Lisbon Falls, her physician since time out of mind. Instead of any of those things, she did what she always did in times of trouble: kept herself to herself. For awhile, things seemed to be all right. She enjoyed her job, she enjoyed her friends, \& she enjoyed her 4 grandchildren, 2 from Dave's family \& 2 from mine. Then things stopped being all right. In Aug of 1973, during a checkup following an operation to ``strip'' some of her outrageously varicose veins, my mother was diagnosed with uterine cancer. I think Nellie Ruth Pillsbury King, who once dumped a bowl of Jell-O on the floor \& then danced in it while her 2 boys lay collapsed in the corner, screaming with laughter, actually died of embarrassment.

The end came in Feb of 1974. By then a little of the money from \textit{Carrie} had begun to flow \& I was able to help with some of the medical expenses -- there was that much to be glad about. \& I was there fore the last of it, staying in the back bedroom of Dave \& Linda's place. I'd been drunk the night before but was only moderately hungover, which was good. One wouldn't want to be too hungover at the deathbed of one's mother.

Dave woke me at 6:15 in the morning, calling softly through the door that he thought she was going. When I got into the master bedroom he was sitting beside her on the bed \& holding a Kool for her to smoke. This she did between harsh gasps for breath. She was only semiconscious, her eyes going from Dave to me \& then back to Dave again. I sat next to Dave, took the cigarette, \& held it to her mouth. Her lips stretched out to clamp on the filter. Beside her bed, reflected over \& over again in a cluster of glasses, was an early bound galley of \textit{Carrie}. Aunt Ethelyn had read it to her aloud a month or so before she died.

Mom's eyes went from Dave to me, Dave to me, Dave to me. She had gone from 160 pounds to about 90. Her skin was yellow \& so tightly stretched that she looked like 1 of those mummies they parade through the streets of Mexico on the Day of the Dead. We took turns holding the cigarette for her, \& when it was down to the filter, I put it out.

``My boys,'' she said, then lapsed into what might have been sleep or unconsciousness. My head ached. I took a couple of aspirin from 1 of the many bottles of medicine on her table. Dave held 1 of her hands \& I held the other. Under the sheet was not the body of our mother but that of a starved \& deformed child. Dave \& I smoked \& talked a little. I don't remember what we said. It had rained the night before, then the temperature had dropped \& the morning streets were filled with ice. We could hear the pause after each rasping breath she drew growing longer \& longer. Finally there were no more breaths \& it was all pause.

\fbox{\bf34} My mother was buried out of the Congregational Church at Southwest Bend; the church she'd attended in Methodist Corners, where my brother \& I grew up, was closed because of the cold. I gave the eulogy. I think I did a pretty good job, considering how drunk I was.

\fbox{\bf35} Alcoholics build defenses like the Dutch build dikes. I spent the 1st 12 years or so of my married life assuring myself that I ``just liked to drink.'' I also employed the world-famous Hemingway Defense. Although never clearly articulated (it would not be manly to do so), the Hemingway Defense goes something like this: as a writer, I am a very sensitive fellow, but I am also a man, \& a real men don't give in to their sensitivities. Only \textit{sissy}-men do that. Therefore I drink. How else can I face the existential horror of it all \& continue to work? Besides, come on, I can handle it. A real man always can.

Then, in the early 80s, Maine's legislature enacted a returnable-bottle \& -can law. Instead of going into the trash, my 16-ounce cans of Miller Lite started going into a plastic container in the garage. 1 Thursday night I went out there to toss in a few dead soldiers \& saw that this container, which had been empty on Monday night, was now almost full. \& since I was the only one in the house who drank Miller Lite --

\textit{Holy shit, I'm an alcoholic}, I thought, \& there was no dissenting opinion from inside my head -- I was, after all, the guy who had written \textit{The Shining} without even realizing (at least until that night) that I was writing about myself. My reaction to this idea wasn't denial or disagreement; it was what I'd call frightened determination. \textit{You have to be careful, then}, I clearly remember thinking. \textit{Because if you fuck up --}

If I fucked up, rolled my car over on a back road some night or blew an interview on live TV, someone would tell me I ought to get control of my drinking, \& telling an alcoholic to control his drinking is like telling a guy suffering the world's most cataclysmic case of diarrhea to control his shitting. A friend of mine who has been through this tells an amusing story about his 1st tentative effort to get a grip on his increasingly slippery life. He went to a counselor \& said to his wife was worried that he was drinking too much.

``How much do you drink?'' the counselor asked.

My friend looked at the counselor with disbelief. ``All of it,'' he said, as if that should have been self-evident.

I know how he felt. It's been almost 12 years since I took a drink, \& I'm still struck by disbelief when I see someone in a restaurant with a half-finished glass of wine near at hand. I want to get up, go over, \& yell ``Finish that! Why don't you finish that?'' into his or her face. I found the idea of social drinking ludicrous -- if you didn't want to get drunk, why not just have a Coke?

My nights during the last 5 years of my drinking always ended with the same ritual: I'd pour any beers left in the refrigerator down the sink. If I didn't, they'd talk to me as I lay in bed until I got up \& had another. \& another. \& 1 more.

\fbox{\bf36} By 1985 I had added drug addiction to my alcohol problem, yet I continued to function, as a good many substance abusers do, on a marginally competent level. I was terrified not to; by then I had no idea of how to live any other life. I hid the drugs I was taking as well as I could, both out of terror -- what would happen to me without dope? I had forgotten the trick of being straight -- \& out of shame. I was wiping my ass with poison ivy again, this time on a daily basis, but I couldn't ask for help. That's not the way you did things in my family. In my family what you did was smoke your cigarettes \& dance in the Jell-O \& keep yourself to yourself.

Yet the part of me that writes the stories, the deep part that knew I was an alcoholic as early as 1975, when I wrote \textit{The Shining}, wouldn't accept that. Silence isn't what that part is about. It began to scream for help in the only way it knew how, through my fiction \& through my monsters. In late 1985 \& early 1986 I wrote \textit{Misery} (the title quite aptly described my state of mind), in which a writer is held prisoner \& tortured by a psychotic nurse. In the spring \& summer of 1986 I wrote \textit{The Tommyknockers}, often working until midnight with my heart running at 130 beats a minute \& cotton swabs stuck up my nose to stem the coke-induced bleeding.

\textit{Tommyknockers} is a 40s-style science fiction tale in which the writer-heroine discovers an alien spacecraft buried in the ground. The crew is still on board, not dead but only hibernating. These alien creatures got into your head \& just started $\ldots$ well, tommyknocking around in there. What you got was energy \& a kind of superficial intelligence (the writer, Bobbi Anderson, creates a telepathic typewriter \& an atomic hot-water heater, among other things). What you gave up in exchange was your soul. It was the best metaphor for drugs \& alcohol my tired, overstressed mind could come up with.

Not long after that my wife, finally convinced that I wasn't going to pull out of this ugly downward spiral on my own, stepped in. It couldn't have been easy -- by then I was no longer within shouting distance of my right mind -- but she did it. She organized an intervention group formed of family \& friends, \& I was treated to a kind of \textit{This Is Your Life} in hell. Tabby began by dumping a trashbag full of stuff from my office out on the rug: beercans, cigarette butts, cocaine in gram bottles \& cocaine in plastic Baggies, coke spoons caked with snot \& blood, Valium, Xanax, bottles of Robitussin could syrup \& NyQuil cold medicine, even bottles of mouthwash. A year or so before, observing the rapidity with which huge bottles of Listerine were disappearing from the bathroom, Tabby asked me if I drank the stuff. I responded with self-righteous hauteur that I most certainly did not. Nor did I. I drank the Scope instead. It was tastier, had that hint of mint.

The point of this intervention, which was certainly as unpleasant for my wife \& kids \& friends as it was for me, was that I was dying in front of them. Tabby said I had my choice: I could get help at a rehab or I could get the hell out of the house. She said that she \& the kids loved me, \& for that very reason none of them wanted to witness my suicide.

I bargained, because that's what addicts do. I was charming, because that's what addicts are. In the end I got 2 weeks to think about it. In retrospect, this seems to summarize all the insanity of that time. Guy is standing on top of a burning building. Helicopter arrives, hovers, drops a rope ladder. \textit{Climb up!} the man leaning out of the helicopter's door shouts. Guy on top of the burning building responds, \textit{Give me 2 weeks to think about it}.

I did think, though -- as well as I could in my addled state -- \& what finally decided me was Annie Wilkes, the psycho nurse in \textit{Misery}. Annie was coke, Annie was booze, \& I decided I was tired of being Annie's pet writer. I was afraid that I wouldn't be able to work anymore if I quit drinking \& drugging, but I decided (again, so far as I was able to decide anything in my distraught \& depressed state of mind) that I would trade writing for staying married \& watching the kids grow up. If it came to that.

It didn't, of course. The idea that creative endeavor \& mind-altering substances are entwined is 1 of the great pop-intellectual myths of our time. The 4 20th-century writers whose work is most responsible for it are probably Hemingway, Fitzgerald, Sherwood Anderson, \& the poet Dylan Thomas. They are the writers who largely formed our vision of an existential English-speaking wasteland where people have been cut off from one another \& live in an atmosphere of emotional strangulation \& despair. These concepts are very familiar to most alcoholics; the common reaction to them is amusement. Substance-abusing writers are just substance abusers -- common garden-variety drunks \& druggies, in other words. Any claims that the drugs \& alcohol are necessary to a dull a finer sensibility are just the usual self-serving bullshit. I've heard alcoholic snowplow drivers make the same claim, that they drink to still the demons. It doesn't matter if you're James Jones, John Cheever, or a stewbum snoozing in Penn Station; for an addict, the right to the drink or drug of choice must be preserved at all costs. Hemingway \& Fitzgerald didn't drink because they were creative, alienated, or morally weak. They drank because it's what alkies are wired up to do. Creative people probably \textit{do} run a greater risk of alcoholism \& addiction than those in some other jobs, but so what? We all look pretty much the same when we're puking in the gutter.

\fbox{\bf37} At the end of my adventures I was drinking a case of 16-ounce tallboys a night, \& there's 1 novel, \textit{Cujo}, that I barely remember writing at all. I don't say that with pride or shame, only with a vague sense of sorrow \& loss. I like that book. I wish I could remember enjoying the good parts as I put them down on the page.

At the worst of it I no longer wanted to drink \& no longer wanted to be sober, either. I felt evicted from life. At the start of the road back I just tried to believe the people who said that things would get better if I gave them time to do  so. \& I never stopped writing. Some of the stuff that came out was tentative \& flat, but at least it was there. I buried those unhappy, lackluster pages in the bottom drawer of my desk \& got on to the next project. Little by little I found the beat again, \& after that I found the joy again. I came back to my family with gratitude, \& back to my work with relief -- I came back to it the way folks came back to a summer cottage after a long winter, checking 1st to make sure nothing has been stolen or broken during the cold season. Nothing had been. It was still all there, still all whole. Once the pipes were thawed out \& the electricity was turned back on, everything worked fine.

\fbox{\bf38} The last thing I want to tell you in this part is about my desk. For years I dreamed of having the sort of massive oak slab that would dominate a room -- no more child's desk in a trailer laundry-closet, no more cramped knee-hole in a rented house. In 1981 I got the one I wanted \& placed it in the middle of a spacious, skylighted study (it's a converted stable loft at the rear of the house). For 6 years I sat behind that desk either drunk or wrecked out of my mind, like a ship's captain in charge of a voyage to nowhere.

A year or 2 after I sobered up, I got rid of that monstrosity \& put in a living-room suite where it had been, picking out the pieces \& a nice Turkish rug with my wife's help. In the early 90s, before they moved on to their own lives, my kids sometimes came up in the evening to watch a basketball game or a movie \& eat pizza. They usually left a boxful of crusts behind when they moved on, but I didn't care. They came, they seemed to enjoy being with me, \& I know I enjoyed being with them. I got another desk -- it's handmade, beautiful, \& half the size of the \textit{T.rex} desk. I put it at the far west end of the office, in a corner under the eave. That eave is very like the one I slept under in Durham, but there are not rats in the walls \& no senile grandmother downstairs yelling for someone to feed Dick the horse. I'm sitting under it now, a 53-year-old man with bad eyes, a gimp leg, \& no hangover. I'm doing what I know how to do, \& as well as I know how to do it. I came through all the stuff I told you about (\& plenty more that I didn't), \& now I'm going to tell you as much as I can about the job. As promised, it won't take long.

It starts with this: put your desk in the corner, \& every time you sit down there to write, remind yourself why it isn't in the middle of the room. Life isn't a support-system for art. It's the other way around.'' -- \cite[pp. 13--81]{King2010}

%------------------------------------------------------------------------------%

\section{What Writing Is}
``Telepathy, of course. It's amusing when you stop to think about it -- for years people have argued about whether or not such a thing exists, folks like J. B. Rhine have busted their brains trying to create a valid testing process to isolate it, \& all the time it's been right there, lying out in the open like Mr. Poe's Purloined Letter. All the arts depend upon telepathy to some degree, but I believe that writing offers the purest distillation. Perhaps I'm prejudiced, but even if I am we may as well stick with writing, since it's what we came here to think \& talk about.

My name is Stephen King. I'm writing the 1st draft of this part at my desk (the one under the eave) on a snowy morning in December of 1997. There are things on my mind. Some are worries (bad eyes, Christmas shopping not even started, wife under the weather with a virus), some are good things (our younger son made a surprise visit home from college, I got to play Vince Taylor's ``Brand New Cadillac'' with The Wallflowers at a concert), but right now all that stuff is up top. I'm in another place, a basement place where there are lots of bright lights \& clear images. This is a place I've built for myself over the years. It's far-seeing place. I know it's a little strange, a little bit of a contradiction, that a far-seeing place should also be a basement place, but that's how it is with me. If you construct your own far-seeing place, you might put it in a treetop or on the roof of the World Trade Center or on the edge of the Grand Canyon. That's your little red wagon, as Robert McCammon says in 1 of his novels.

This book is scheduled to be published in the late summer or early fall of 2000. If that's how things work out, then you are somewhere downstream on the timeline from me $\ldots$ but you're quite likely in your own far-seeing place, the one where you go to receive telepathic messages. Not that you \textit{have} to be there; books are a uniquely portable magic. I usually listen to one in the car (always unabridged; I think abridged audiobooks are the pits), \& carry another wherever I go. You just never know when you'll want an escape hatch: mile-long lines at tollbooth plazas, the 15 minutes you have to spend in the hall of some boring college building waiting for your advisor (who's got some yank-off in there threatening to commit suicide because he\texttt{/}she is flunking Custom Kurmfurling 101) to come out so you can get his signature on a drop-card, airport boarding lounges, laundromats on rainy afternoons, \& the absolute worst, which is the doctor's office when the guy is running late \& you have to wait half an hour in order to have something sensitive mauled. At such times I find a book vital. If I have to spend time in purgatory before going to 1 place or the other, I guess I'll be all right as long as there's a lending library (if there is it's probably stocked with nothing but novels by Danielle Steel \& \textit{Chicken Soup} books, ha-ha, joke's on you, Steve).

So I read where I can, but I have a favorite place \& probably you do, too -- a place where the light is good \& the vibe is usually strong. For me it's the blue chair in my study. For you it might be the couch on the sun-porch, the rocker in the kitchen, or maybe it's propped up in your bed -- reading in bed can be heaven, assuming you can get just the right amount of light on the page \& aren't prone to spilling your coffee or cognac on the sheets.

So let's assume that you're in your favorite receiving place just as I am in the place where I do my best transmitting. We'll have to perform our mentalist routine not just over distance but over time as well, yet that presents no real problem; if we can still read Dickens, Shakespeare, \& (with the help of a footnote or 2) Herodotus, I think we can manage the gap between 1997 \& 2000. \& here we go -- actually telepathy in action. You'll notice I have nothing up my sleeves \& that my lips never move. Neither, most likely, do yours.

Look -- here's a table covered with a red cloth. On it is a cage the size of a small fish aquarium. In the cage is a white rabbit with a pink nose \& pink-rimmed eyes. In its front paws is a carrot-stub upon which it is contentedly munching. On its back, clearly marked in blue ink, is the numeral 8.

Do we see the same thing? We'd have to get together \& compare notes to make absolutely sure, but I think we do. There will be necessary variations, of course: some receivers will see a cloth which is turkey red, some will see one that's scarlet, while others may see still other shades. (To color-blind receivers, the red tablecloth is the dark gray of cigar ashes.) Some may see scalloped edges, some may see straight ones. Decorative souls may add a little lace, \& welcome -- my tablecloth is your tablecloth, knock yourself out.

Likewise, the matter of the cage leaves quite a lot of room for individual interpretation. For 1 thing, it is described in terms of \textit{rough comparison}, which is useful only if you \& I see the world \& measure the things in it with similar eyes. It's easy to become careless when making rough comparisons, but the alternative is a prissy attention to detail that takes all the fun out of writing. What am I going to say, ``on the table is a cage 3 feet, 6 inches in length, 2 feet in width, \& 14 inches high''? That's not prose, that's an instruction manual. The paragraph also doesn't tell us what sort of material the cage is made of -- wire mesh? steel rods? glass? -- but does it really matter? We all understand the cage is a see-through medium; beyond that, we don't care. The most interesting thing here isn't even the carrot-munching rabbit in the cage, but the number on its back. Not a 6, not a 4, not 19.5. It's an 8. This is what we're looking at, \& we all see it. I didn't tell you. You didn't ask me. I never opened my mouth \& you never opened yours. We're not even in the same \textit{year} together, let alone the same room $\ldots$ except we \textit{are} together. We're close.

We're having a meeting of the minds.

I sent you a table with a red cloth on it, a cage, a rabbit, \& the number 8 in blue ink. You got them all, especially that blue 8. We've engaged in an act of telepathy. No mythy-mountain shit; real telepathy. I'm not going to belabor the point, but before we go any further you have to understand that I'm not trying to be cute; there \textit{is} a point to be made.

You can approach the act of writing with nervousness, excitement, hopefulness, or even despair -- the sense that you can never completely put on the page what's in your mind \& heart. You can come to the act with your fists clenched \& your eyes narrowed, ready to kick ass \& take down names. You can come to it because you want a girl to marry you or because you want to change the world. Come to it any way but lightly. Let me say it again: \textit{you must not come lightly to the blank page}.

I'm not asking you to come reverently or unquestioningly; I'm not asking you to be politically correct or cast aside your sense of humor (please God you have one). This isn't a popularity contest, it's not the moral Olympics, \& it's not church. But it's \textit{writing}, damn it, not washing the car or putting an eyeliner. If you can take it seriously, we can do business. If you can't or won't, it's time for you to close the book \& do something else.

Wash the car, maybe.'' -- \cite[pp. 83--85]{King2010}

%------------------------------------------------------------------------------%

\section{TOOLBOX}

\begin{quotation}
	``Grandpa was a carpenter,\\he built houses, stores \& banks,\\he chain-smoked Camel cigarettes\\\& hammered nails in planks.\\He was level-on-the-level,\\shaved even every door,\\\& voted for Eisenhower\\`cause Lincoln won the war.
\end{quotation}
That's 1 of my favorite John Prine lyrics, probably because my grandpa was also a carpenter. I don't know about stores \& banks, but Guy Pillsbury built his share of houses \& spent a good many years making sure the Atlantic Ocean \& the harsh seacoast winters didn't wash away the Winslow Homer estate in Prout's Neck. Fazza smoked cigars, though, not Camels. It was my Uncle Oren, also a carpenter, who smoked the Camels. \& when Fazza retired, it was Uncle Oren who inherited the old fellow's toolbox. I don't remember its being there in the garage on the day I dropped the cinderblock on my foot, but it probably was sitting in its accustomed place just outside the nook where my cousin Donald kept his hockey sticks, ice skates, \& baseball glove.

The toolbox was what we called a big `un. It had 3 levels, the top 3 removable, all 3 containing little drawers as cunning as Chinese boxes. It was handmade, of course. Dark wooden slats were bound together by tiny nails \& strips of brass. The lid was held down by big latches; to my child's eye they looked like the latches on a giant's lunchbox. Inside the top was a silk lining, rather odd in such a context \& made more striking still by the pattern, which was pinkish-red cabbage roses fading into a smog of grease \& dirt. On the sides were great big grabhandles. You never saw a toolbox like this one for sale at Wal-Mart or Western Auto, believe me. When my uncle 1st got it, he found a brass etching of a famous Homer painting -- I believe it was \textit{The Undertow} -- lying in the bottom. Some years later Uncle Oren had it authenticated by a Homer expert in New York, \& a few years after that I believe he sold it for a good piece of money. Exactly how or why Fazza came by the engraving in the 1st place is a mystery, but there was no mystery about the origins of the toolbox -- he made it himself.

1 summer day I helped Uncle Oren replace a broken screen on the far side of the house. I might have been 8 or 9 at the time. I remember following him with the replacement screen balanced on my head, like a native bearer in a Tarzan movie. He had the toolbox by the grabhandles, horsing it along at thigh level. As always, Uncle Oren was wearing khaki pants \& a clean white tee-shirt. Sweat gleamed in this graying Army crewcut. A Camel hung from his lower lip. (When I came in years later with a pack of Chesterfields in my breast pocket, Uncle Oren sneered at them \& called them ``stockade cigarettes.'')

We finally reached the window with the broken screen \& he set the toobox down with audible sign of relief. When Dave \& I tried to lift it from its place on the garage floor, each of us holding 1 of the handles, we could barely budge it. Of course we were just little kids back then, but even so I'd guess that Fazza's fully loaded toolbox weighed between 80 \& 120 pounds.

Uncle Oren let me undo the big latches. The common tools were all on the top layer of the box. There was a hammer, a saw, the pliers, a couple of sized wrenches \& an adjustable; there was a level with that mystic yellow window in the middle, a drill (the various bits were neatly drawered farther down in the depths), \& 2 screwdrivers. Uncle Oren asked me for a screwdriver.

``Which one?'' I asked.

``Either-or,'' he replied.

The broken screen was held on by loophead screws, \& it really didn't matter whether he used a regular screwdriver or the Phillips on them; with loopheads you just stuck the screwdriver's barrel through the hole at the top of the screw \& then spun it the way you spin a tire iron once you've got the lugnuts loose.

Uncle Oren took the screws out -- there were 8, which he handed to me for safekeeping -- \& then removed the old screen. He set it against the house \& held up the new one. The holes in the screen's frame mated up heatly with the holes in the window-frame. Uncle Oren grunted with approval when he saw this. He took the loophead screws back to me, one after the other, got them started with his fingers, then tightened them down just as he'd loosened them, by inserting the screwdriver's barrel through the loops \& turning them.

When the screen was secure, Uncle Oren gave me the screwdriver \& told me to put it back in the toolbox \& ``latch her up.'' I did, but I was puzzled. I asked him why he'd lugged Fazza's toolbox all the way around the house, if all he'd needed was that 1 screwdriver. He could have carried a screwdriver in the back pocket of his khakis.

``Yeah, but Stevie,'' he said, bending to grasp the handles, ``I didn't know what else I might find to do once I got out here, did I? It's best to have your tools with you. If you don't, you're apt to find something you didn't expect \& get discouraged.''

I want to suggest that to write to your best abilities, it behooves you to construct your own toolbox \& then build up enough muscle so you can carry it with you. Then, instead of looking at a hard job \& getting discouraged, you will perhaps seize the correct tool \& get immediately to work.

Fazza's toolbox had 3 levels. I think that yours should have at least 4. You could have 5 or 6, I suppose, but there comes a point where a toolbox becomes too large to be portable \& thus loses its chief virtue. You'all also want all those little drawers for your screws \& nuts \& bolts, but where you put those drawers \& what you put in them $\ldots$ well, that's your little red wagon, isn't it? You'll find you have most of the tools you need already, but I advise you to look at each one again as you load it into your box. Try to see each one new, remind yourself of its function, \& if some are rusty (as they may be if you haven't done this seriously in awhile), clean them off.

Common tools go on top. The commonest of all, the bread of writing, is vocabulary. In this case, you can happily pack what you have without the slightest bit of guilt \& inferiority. As the whore said to the bashful sailor, ``It ain't how much you've got, honey, it's how you use it.''

Some writers have enormous vocabularies; these are folks who'd know if there really \textit{is} such a thing as an insalubrious dithyramb or a cozening raconteur, people who haven't missed a multiple-choice answer in Wilfred Funk's \textit{It Pays to Increase Your Word Power} in oh, 30 years or so. E.g.:
\begin{quotation}
	The leathery, undeteriorative, \& almost indestructible quality was an inherent attibute of the thing's form of organization, \& pertained to some paleogean cycle of invertebrate evolution utterly beyond our powers of speculation. -- H.P. Lovecraft, \textit{At the Mountains of Madness}
\end{quotation}
Like it? Here's another:
\begin{quotation}
	In some [of the cups] there was no evidence whatever that anything had been planted; in others, wilted brown stalks gave testimony to some inscrutable depredation. -- T. Coraghessan Boyle, \textit{Budding Prospects}
\end{quotation}
\& yet a 3rd -- this is a good one, you'll like it:
\begin{quotation}
	Someone snatched the old woman's blindfold from her \& she \& the juggler were clouted away \& when the company turned in to sleep \& the low fire was roaring in the blast like a thing alive these 4 yet crouched at the edge of the firelight among their strange chattels \& watched how the ragged flames fled down the wind as if sucked by some maelstrom out there in the void, some vortex in that waste apposite to which man's transit \& his reckonings alike lay abrogate. -- Cormac McCarthy, \textit{Blood Meridian}
\end{quotation}
Other writers use smaller, simpler vocabularies. Examples of this hardly seem necessary, but I'll offer a couple of my favorites, just the same:
\begin{quotation}
	He came to the river. The river was there. -- Ernest Hemingway, ``Big 2-hearted river''
	
	They caught the kid doing something nasty under the bleachers. -- Theodore Sturgeon, \textit{Some of Your Blood}
	
	This is what happened. -- Douglas Fairbairn, \textit{Shoot}
	
	Some of the owner men were kind because they hated what they had to do, \& some of them were angry because they hated to be cruel, \& some of them were cold because they had long ago found that one could not be an owner unless one were cold. -- John Steinbeck, \textit{The Grapes of Wrath}
\end{quotation}
The Steinbeck sentence is especially interesting. It's 50 words long. Of those 50 words, 39 have but 1 syllable. That leaves 11, but even that number is deceptive; Steinbeck uses \textit{because} 3 times, \textit{owner} twice, \& \textit{hated} twice. There is no word longer than 2 syllables in the entire sentence. The structure is complex; the vocabulary is no far removed from the old Dick \& Jane primers. \textit{The Grapes of Wrath} is, of course, a fine novel. I believe that \textit{Blood Meridian} is another, although there are great whacks of it that I don't fully understand. What of that? I can't decipher the words to many of the popular songs I love, either.

There's also stuff you'll never find in the dictionary, but it's still vocabulary. Check out the following:
\begin{quotation}
	``Egggh, whaddaya? Whaddaya want from me?''
	
	``Here come Hymie!''
	
	``Unnh! Unnnh! Unnnhh!''
	
	``Chew my willie, Yo' Honor.''
	
	``Yeggghh, fuck you, too, man!'' -- Tom Wolfe, \textit{Bonfire of the Vanities}
\end{quotation}
This last is phonetically rendered street vocabulary. Few writers have Wolfe's ability to translate such stuff to the page. (Elmore Leonard is another writer who can do it.) Some street-rap gets into the dictionary eventually, but not until it's safely dead. \& I don't think you'll ever find \textit{Yeggghhh} in Webster's Unabridged.

Put your vocabulary on the top shelf of your toolbox, \& don't make any conscious effort to improve it. (You'll be doing that as you read, of course $\ldots$ but that comes later.) 1 of the really bad things you can do to your writing is to dress up the vocabulary, looking for long words because you're maybe a little bit ashamed of your short ones. This is like dressing up a household pet in evening clothes. The pet is embarrassed \& the person who committed this act of premeditated cuteness should be even more embarrassed. Make yourself a solemn promise right now that you'll never use ``emolument'' when you mean ``tip'' \& you'll never say \textit{John stopped long enough to perform an act of excretion} when you mean \textit{John stopped long enough to take a shit}. If you believe ``take a shit'' would be considered offensive or inappropriate by your audience, feel free to say \textit{John stopped long enough to move his bowels} (or perhaps \textit{John stopped long enough to ``push''}). I'm not trying to get you to talk dirty, only plain \& direct. Remember that the basic rule of vocabulary is \textit{use the 1st word that comes to your mind, if it is appropriate \& colorful}. If you hesitate \& cogitate, you will come up with another word -- of course you will, there's always another word -- but it probably won't be as good as your 1st one, or as close to what you really mean.

This business of meaning is a very big deal. If you doubt it, think of all the times you've heard someone say ``I just can't describe it'' or ``That isn't what I mean.'' Think of all the times you've said those things yourself, usually in tone of mild or serious frustration. The word is only a representation of the meaning; even at its best, writing almost always falls short of full meaning. Given that, why in God's name would you want to make things worse by choosing a word which is only cousin to the one you really wanted to use?

\& \textit{do} feel free to take appropriateness into account; as George Carlin once observed, in some company it's perfectly all right to prick your finger, but very bad form to finger your prick.

\fbox{\bf2} You'll also want grammar on the top shelf of your toolbox, \& don't annoy me with your moans of exasperation or your cries that you \textit{don't understand grammar}, you \textit{never did understand} grammar, you flunked that \textit{whole semester} in Sophomore English, writing is fun but grammar sucks the big one.

Relax. Chill. We don't spend much time here because we don't need to. One either absorbs the grammatical principles of one's native language in conversation \& in reading or one does not. What Sophomore English does (or tries to do) is little more than the naming of parts.

\& this isn't high school. Now that you're not worried that (a) your skirt is too short or too long \& the other kids will laugh at you, (b) you're not going to make the varsity swimming team, (c) you're still going to be a pimple-studded virgin when you graduate (probably when you die, for that matter), (d) the physics teacher won't grade the final on a curve, or (e) nobody really likes you anyway \textsc{\& they never did} $\ldots$ now that all that  extraneous shit is out of the way, you can study certain academic matters with a degree of concentration you could never manage while attending the local textbook loonybin. \& once you start, you'll find you know almost all of the stuff anyway -- it is, as I said, mostly a matter of cleaning the rust off the drillbits \& sharpening the blade of your saw.

Plus $\ldots$ oh, to hell with it. If you can remember all the accessories that go with your best outfit, the contents of your purse, the starting lineup of the New York Yankees or the Houston Oilers, or what label ``Hang On Sloopy'' by The McCoys was on, you are capable of remembering the difference between a gerund (verb form used as a noun) \& a participle (verb form used as an adjective).

I thought long \& hard about whether or not to include a detailed section on grammar in this little book. Part of me would actually like to; I taught it successfully at high school (where it hid under the name Business English), \& I enjoyed it as a student. American grammar doesn't have the sturdiness of British grammar (a British advertising man with a proper education can make magazine copy for ribbed condoms sound like the Magma goddam Carta), but it has its own scruffy charm.

In the end I decided against it, probably for the same reason William Strunk decided not to recap the basics when he wrote the 1st edition of \textit{The Elements of Style}: if you don't know, it's too late. \& those really incapable of grasping grammar -- as I am incapable of playing certain guitar riffs \& progressions -- will have little or no use for a book like this, anyway. In that sense I am preaching to the converted. Yet allow me to go on just a little bit further -- will you indulge me?

Vocabulary used in speech or writing organizes itself in 7 parts of speech (8, if you count interjections such as \textit{Oh!} \& \textit{Gosh!} \& \textit{Fuhgeddaboudit!}). Communication composed of these parts of speech must be organized by rules of grammar upon which we agree. When these rules break down, confusion \& misunderstanding result. Bad grammar produced bad sentences. My favorite example from Strunk \& White \cite{Strunk_White2019} is this one: \textit{``As a mother of 5, with another one on the way, my ironing board is always up.''}

Nouns \& verbs are the 2 indispensable parts of writing. Without 1 of each, no group of words can be a sentence, since a sentence is, by definition, a group of words containing a subject (noun) \& a predicate (verb); these strings of words begin with a capital letter, end with a period, \& combine to make a complete thought which starts in the writer's head \& then leaps to the reader's.

\textit{Must} you write complete sentences each time, every time? Perish the thought. If your work consists only of fragments \& floating clauses, the Grammar Police aren't going to come \& take you away. Even William Strunk, that Mussolini of rhetoric, recognized the delicious pliability of language. ``It is an old observation,'' he writes, ``that the best writers sometimes disregard the rules of rhetoric.'' yet he goes on to add this thought, which I urge you to consider: ``Unless he is certain of doing well, [the writer] will probably do best to follow the rules.''

The telling clause here is \textit{Unless he is certain of doing well}. If you don't have a rudimentary grasp of how the parts of speech translate into coherent sentences, how can you be certain that you \textit{are} doing well? How will you know if you're doing ill, for that matter? The answer, of course, is that you can't, you won't. One who does grasp the rudiments of grammar finds a comforting simplicity at its heart, where there need be only nouns, the words that name, \& verbs, the words that act.

Take any noun, put it with any verb, \& you have a sentence. It never fails. \textit{Rocks explode. Jane transmits. Mountains float}. These are all perfect sentences. Many such thoughts make little rational sense, but even the strange ones (\textit{Plums deify!}) have a kind of poetic weight that's nice. The simplicity of noun-verb construction is useful -- at the very least it can provide a safety net for your writing. Strunk \& White caution against too many simple sentences in a row, but simple sentences provide a path you can follow when you fear getting lost in the tangles of rhetoric -- all those restrictive \& nonrestrictive clauses, those modifying phrases, those appositives \& compound-complex sentences. If you start to freak out at the sight of such unmapped territory (unmapped by you, at least), just remind yourself that rocks explode, Jane transmits, mountains float, \& plums deify. Grammar is not just a pain in the ass; it's the pole you grab to get your thoughts up to their feet \& walking. Besides, all those simple sentences worked for Hemingway, didn't they? Even when he was drunk on his ass, he was a fucking genius.

If you want to refurbish your grammar, go to your local used-book store \& find a copy of \textit{Warriner's English Grammar \& Composition} -- the same book most of us took home \& dutifully covered with brown paper shopping-bags when we were sophomores \& juniors in high school. You'll be relieved \& delighted, I think, to find that almost all you need is summarized on the front \& back endpapers of the book.

\fbox{\bf3} Despite the brevity of his style manual, William Strunk found room to discuss his own dislikes in matters of grammar \& usage. He hated the phrase ``student body,'' e.g., insisting that ``studentry''s was both clearer \& without the gloulish connotations he saw in the former term. He thought ``personalize'' a pretentious word. (Strunk suggests ``Get up a letterhead'' to replace ``Personalize your stationery.'') He hated phrases such as ``the fact that'' \& ``along these lines.''

I have my own dislikes -- I believe that anyone using the phrase ``That's so cool'' should have to stand in the corner \& that those using the far more odious phrases ``at this point in time'' \& ``at the end of the day'' should be sent to bed without supper (or writing-paper, for that matter). 2 of my other pet peeves have to do with this most basic level of writing, \& I want to get them off my chest before we move along.

Verbs come in 2 types, active \& passive. With an active verb, the subject of the sentence is doing something. With a passive verb, something is being done \textit{to} the subject of the sentence. The subject is just letting it happen. \textit{You should avoid the passive tense}. I'm not the only one who says so; you can find the same advice in \textit{The Elements of Style}.

Messrs. Strunk \& White don't speculate as to why so many writers are attracted to passive verbs, but I'm willing to; I think timid writers like them for the same reason timid lovers like passive partners. The passive voice is safe. There is no troublesome action to contend with; the subject just has to close its eyes \& think of England, to paraphrase Queen Victoria. I think unsure writers also feel the passive voice somehow lends their work authority, perhaps even a quality of majesty. If you find instruction manuals \& lawyers' torts majestic, I guess it does.

The timid fellow writes \textit{The meeting will be held at 7 o'clock} because that somehow says to him, ``Put it this way \& people will believe \textit{you really know}.'' Purge this quisling thought! Don't be a muggle! Throw back your shoulders, stick out your chin, \& put that meeting in charge! Write \textit{The meeting's at 7}. There, by God! Don't you feel better?

I won't say there's no place for the passive tense. Suppose, e.g., a fellow dies in the kitchen but ends up somewhere else. \textit{The body was carried from the kitchen \& placed on the parlor sofa} is a fair way to put this, although ``was carried'' \& ``was placed'' still irk the shit out of me. I accept them but I don't embrace them. What I would embrace is \textit{Freddy \& Myra carried the body out of the kitchen \& laid it on the parlor sofa}. Why does the body have to be the subject of the sentence, anyway? It's dead, for Christ's sake! Fuhgeddaboudit!

2 pages of the passive voice -- just about any business document ever written, in other words, not to mention reams of bad fiction -- make me want to scream. It's weak, it's circuitous, \& it's frequently tortuous, as well. How about this: \textit{My 1st kiss will always be recalled by me as how my romance with Shayna was begun}. Oh, man -- who farted, right? A simpler way to express this idea -- sweeter \& more forceful, as well -- might be this: \textit{My romance with Shayna began with our 1st kiss. I'll never forget it}. I'm not in love with this because it uses \textit{with} twice in 4 words, but at least we're out of that awful passive voice.

You might also notice how much simpler the thought is to understand when it's broken up into \textit{2} thoughts. This makes matters easier for the reader, \& the reader must always be your main concern; without Constant Reader, you are just a voice quacking in the void. \& it's no walk in the park being the guy on the receiving end. ``[Will Strunk] felt the reader was in serious trouble most of the time,'' E. B. White writes in his introduction to \textit{The Elements of Style}, ``a man floundering in a swamp, \& that it was the duty of anyone trying to write English to drain this swamp quickly \& get his man up on dry ground, or at least throw him a rope.'' \& remember: \textit{The writer threw the rope}, not \textit{The rope was thrown by the writer}. Please oh please.

The other piece of advice I want to give you before moving on to the next level of the toolbox is this: \textit{The adverb is not your friend}.

Adverbs, you will remember from your own version of Business English, are words that modify verbs, adjectives, or other adverbs. They're the ones that usually end in \textit{-ly}. Adverbs, like the passive voice, seem to have been created with the timid writer in mind. With the passive voice, the writer usually expresses fear of not being taken seriously; it is the voice of little boys wearing shoepolish mustaches \& little girls clumping around in Mommy's high heels. With adverbs, the writer usually tells us he or she is afraid he{\tt/}she isn't expressing himself{\tt/}herself clearly, that he or she is not getting the point or the picture across.

Consider the sentence \textit{He closed the door firmly}. It's by no means a terrible sentence (at least it's got an active verb going for it), but ask yourself if \textit{firmly} really has to be there. You can argue that it expresses a degree of difference between \textit{He closed the door} \& \textit{He slammed the door}, \& you'll get no argument from me $\ldots$ but what about context? What about all the enlightening (not to say emotionally moving) prose which came \textit{before `He closed the door firmly?'} Shouldn't this tell us how he closed the door? \& if the foregoing prose \textit{does} tell us, isn't \textit{firmly} an extra word? Isn't it redundant?

Someone out there is now accusing me of being tiresome \& anal-retentive. I deny it. I believe the road to hell is paved with adverbs, \& I will shout it from the rooftops. To put it another way, they're like dandelions. If you have one on your lawn, it looks pretty \& unique. If you fail to root it out, however, you find 5 the next day $\ldots$ 50 the day after that $\ldots$ \& then, my brothers \& sisters, your lawn is \textit{totally, completely}, \& \textit{profligately} covered with dandelions. By then you see them for the weeds they really are, but by then it's -- \textsc{gasp}!! -- too late.

I can be a good sport about adverbs, though. Yes I can. With 1 exception: dialogue attribution. I insist that you use the adverb in dialogue attribution only in the rarest \& most special of occasions $\ldots$ \& not even then, if you can avoid it. Just to make sure we all know that we're talking about, examine these 3 sentences:
\begin{quotation}
	``Put it down!'' she shouted.
	
	``Give it back,'' he pleaded, ``it's mine.''
	
	``Don't be such a fool, Jekyll,'' Utterson said.
\end{quotation}
In these sentences, \textit{shouted, pleaded}, \& \textit{said} are verbs of dialogue attribution. Now look at these dubious revisions:
\begin{quotation}
	``Put it down!'' she shouted menacingly.
	
	``Give it back,'' he pleaded abjectly, ``it's mine.''
	
	``Don't be such a fool, Jekyll,'' Utterson said contemptuously.
\end{quotation}
The 3 latter sentences are all weaker than the 3 former ones, \& most readers will see why immediately. \textit{``Don't be such a fool, Jekyll,'' Utterson said contemptuously} is the best of the lot; it is only a clich\'e, while the other 2 are actively ludicrous. Such dialogue attributions are sometimes known as ``Swifties,'' after Tom Swift, the brave inventor-hero in a series of boys' adventure novels written by Victor Appleton II. Appleton was fond of such sentences as \textit{``Do your worst!'' Tom cried bravely} \& \textit{``My father helped with the equations,'' Tom said modestly}. When I was a teenager there was a party-game based on one's ability to create witty (or half-witty) Swifties. \textit{``You got a nice butt, lady,'' he said cheekily} is one I remember; another is \textit{``I'm the plumber,'' he said, with a flush}. (In this case the modifier is an adverbial phrase.) When debating whether or not to make some pernicious dandelion of an adverb part of your dialogue attribution, I suggest you ask yourself if you really want to write the sort of prose that might wind up in a party-game.

Some writers try to evade the no-adverb rule by shooting the attribution verb full of steroids. The result is familiar to any reader of pulp fiction or paperback originals:
\begin{quotation}
	``Put down the gun, Utterson!'' Jekyll grated.
	
	``Never stop kissing me!'' Shayna gasped.
	
	``You damned tease!'' Bill jerked out.
\end{quotation}
Don't do these things. Please oh please.

The best form of dialogue attribution is \textit{said}, as in \textit{he said, she said, Bill said, Monica said}. If you want to see this put stringently into practice, I urge you to read or reread a novel by Larry McMurtry, the Shane of dialogue attribution. That looks damned snide on the page, but I'm speaking with completely sincerity. McMurtry has allowed few adverbial dandelions to grow on his lawn. He believes in he-said{\tt/}she-said even in moments of emotional crisis (\& in Larry McMurtry novels there are a lot of those). Go \& do thou likewise.

Is this a case of ``Do as I say, not as I do?'' The reader has a perfect right to ask the question, \& I have a duty to provide an honest answer. Yes. It is. You need only look back through some of my own fiction to know that I'm just another ordinary sinner. I've been pretty good about avoiding the passive tense, but I've spilled out my share of adverbs in my time, including some (it shames me to say it) in dialogue attribution. (I have never fallen so low as ``he grated'' or ``Bill jerked out,'' though.) When I do it, it's usually for the same reason any writer does it: because I am afraid the reader won't understand me if I don't.

I'm convinced that fear is at the root of most bad writing. If one is writing for one's own pleasure, that fear may be mild -- \textit{timidity} is the word I've used here. If, however, one is working under deadline -- a school paper, a newspaper article, the SAT writing sample -- that fear may be intense. Dumbo got airborne with the help of a magic feather; you may feel the urge to grasp a passive verb or 1 of those nasty adverbs for the same reason. Just remember before you do that Dumbo didn't need the feather; the magic was in him.

You probably \textit{do} know what you're talking about, \& can safely energize your prose with active verbs. \& you probably \textit{have} told your story well enough to believe that when you use \textit{he said}, the reader will know how he said it -- fast or slowly, happily or sadly. Your man may be floundering in a swamp, \& by all means throw him a rope if he is $\ldots$ but there's no need to knock him unconscious with 90 feet of steel cable.

Good writing is often about letting go of fear \& affectation. Affectation itself, beginning with the need to define some sorts of writing as ``good'' \& other sorts as ``bad,'' is fearful behavior. Good writing is also about making good choices when it comes to picking the tools you plan to work with.

No writer is entirely without sin in these matters. Although William Strunk got E. B. White in his clutches when White was but a naive undergraduate at Cornell (give them to me when they're young \& they're mine forever, heh-heh-heh), \& although White both understood \& shared Strunk's prejudice against loose writing \& the loose thinking which prompts it, he admits, ``I suppose I have written \textit{the fact that} a thousand times in the heat of composition, revised it out maybe 500 times in the cool aftermath. To be batting only. 500 this late in the season, to fail half the time to connect with this fat pitch, saddens me $\ldots$'' Yet E. B. White went on to write for a good many years following his initial revisions of Strunk's ``little book'' in 1957. I will go on writing in spite of such stupid lapses as \textit{``You can't be serious,'' Bill said unbelievingly}. I expect you to do the same thing. There is a core simplicity to the English language \& its American variant, but it's slippery core. All I ask is that you do as well as you can, \& remember that, while to write adverbs is human, to write \textit{he said} or \textit{she said} is divine.

\fbox{\bf4} Lift out the top layer of your toolbox -- your vocabulary \& all the grammar stuff. On the layer beneath go those elements of style upon which I've already touched. Strunk \& White offer the best tools (\& the best rules) you could hope for, describing them simply \& clearly. (They are offered with a refreshing strictness, beginning with the rule on how to form possessives: you always add \textit{'s}, even when the word you're modifying ends in \textit{s} -- always write \textit{Thomas's bike} \& never \textit{Thomas' bike} -- \& ending with ideas about where it's best to place the most important parts of a sentence. They say at the end, \& everyone's entitled to his{\tt/}her opinion, but I don't believe \textit{With a hammer he killed Frank} will ever place \textit{He killed Frank with a hammer}.)

Before leaving the basic elements of form \& style, we ought to think for a moment about the paragraph, the form of organization which comes after the sentence. That that end, grab a novel -- preferably one you haven't yet read -- down your your shelf (the stuff I'm telling you applies to most prose, but since I'm a fiction writer, it's fiction I usually think about when I think about writing). Open the book in the middle \& look at any 2 pages. Observe the pattern -- the lines of type, the margins, \& most particularly the block of white space where paragraphs begin or leave off.

You can tell \textit{without even reading} if the book you've chosen is apt to be easy or hard, right? Easy books contain lots of short paragraphs -- including dialogue paragraphs which may only be a word or 2 long -- \& lots of white space. They're as airy as Dairy Queen ice cream cones. Hard books, ones full of ideas, narration, or description, have a stouter look. A \textit{packed} look. Paragraphs are almost as important for how they look as for what they say; they are maps of intent.

In expository prose, paragraphs can (\& should) be neat \& utilitarian. The ideal expository graf contains a topic sentence followed by others which explain or amplify the 1st. Here are 2 paragraphs from the ever-popular ``informal essay'' which illustrate this simple but powerful form of writing:
\begin{quotation}
	When I was 10, I feared my sister Megan. It was impossible for her to come into my room without breaking at least 1 of my favorite toys, usually the favorite of favorites. Her gaze had some magical tape-destroying quality; any poster she looked at seemed to fall off the wall only seconds later. Well-loved articles of clothing disappeared from the closet. She didn't take them (at least I don't think so), only made them vanish. I'd usually find that treasured tee-shirt or my favorite Nikes deep under the bed months later, looking sad \& abandoned among the dust kitties. When Megan was in my room, stereo speakers below, window-shades flew up with a bang, \& the lamp on my desk usually went dead.
	
	She could be consciously cruel, too. On 1 occasion, Megan poured orange juice into my cereal. On another, she squirted toothpaste into the toes of my socks while I was taking a shower. \& although she never admitted it, I am positive that whenever I fell asleep on the couch during halftime of the Sunday afternoon pro football games on TV, she rubbed boogers in my hair.
\end{quotation}
Informal essays are, by \& large, silly \& insubstantial things; unless you get a job as a columnist at your local newspaper, writing such fluffery is a skill you'll never use in the actual mall-\&-filling-station world. Teachers assign them when they can't think of any other way to waste your time. The most notorious subject, of course, is ``How I Spent My Summer Vacation.'' I taught writing for a year at the University of Maine in Orono \& had 1 class loaded with athletes \& cheerleaders. They liked informal essays, greeting them like the old high school friends they were. I spent 1 whole semester fighting the urge to ask them to write 2 pages of well-turned prose on the subject of ``If Jesus Were My Teammate.'' What held me back was the sure \& terrible knowledge that most of them would take to the task with enthusiasm. Some might actually weep while in the throes of composition.

Even in the informal essay, however, it's possible to see how strong the basic paragraph form can be. Topic-sentence-followed-by-support-\&-description insists that the writer organize his{\tt/}her thoughts, \& it also provides good insurance against wandering away from the topic. Wandering isn't a big deal in an informal essay, is practically \textit{de rigueur}, as a matter of fact -- but it's a very bad habit to get into when working on more serious subjects in a more formal manner. Writing is refined thinking. If your master's thesis is no more organized than a high school essay titled ``Why Shania Twain Turns Me On,'' you're in big trouble.

In fiction, the paragraph is less structured -- it's the beat instead of the actual melody. The more fiction you read \& write, the more you'll find your paragraphs forming on their own. \& that's what you want. When composing it's best not to think too much about where paragraphs begin \& end; the trick is to let nature take its course. If you don't like it later on, fix it then. That's what rewrite is all about. Now check out the following:
\begin{quotation}
	Big Tony's room wasn't what Dale had expected. The light had an odd yellowish cast that reminded him of cheap motels he'd stayed in, the ones where he always seemed to end up with a scenic view of the parking lot. The only picture was Miss May hanging askew on a push-pin. 1 shiny black shoe stuck out from under the bed.
	
	``I dunno why you keep askin me about O'Leary,'' Big Tony said. ``You think my story's gonna change?''
	
	``Is it?'' Dale asked.
	
	``When your story's true it don't change. The truth is always the same boring shit, day in \& day out.''
	
	Big Tony sat down, lit a cigarette, ran a hand through his hair.
	
	``I ain't seen that fuckin mick since last summer. I let him hang around because he made me laugh, once showed me this thing he wrote about what it woulda been like if Jesus was on his high school football team, had a picture of Christ in a helmet \& kneepads \& everythin, but what a troublesome little fuck he turned out to be! I wish I'd never seen him!''
\end{quotation}
We could have a 50-minute writing class on just this brief passage. It would encompass dialogue attribution (not necessary if we know who's speaking; Rule 17, omit needless words, in action), phonetically rendered language (\textit{dunno, gonna}), the use of the comma (there is none in the line \textit{When your story's true it don't change} because I want you to hear it coming out all in 1 breath, without a pause), the decision not to use the apostrophe where the speaker has dropped a \textit{g} $\ldots$ \& that stuff is just from the top level of the toolbox.

Let's stick with the paragraphs, though. Notice how easily they flow, with the turns \& rhythms of the story dictating where each one begins \& ends. The opening graf is of the classic type, beginning with a topic sentence that is supported by the sentences which follow. Others, however, exist solely to differentiate between Dale's dialogue \& Big Tony's.

The most interesting paragraph is the 5th one: \textit{Big Tony sat down, lit a cigarette, ran a hand through his hair}. It's only a single sentence long, \& expository paragraphs almost never consist of a single sentence. It's not even a very \textit{good} sentence, technically speaking; to make it perfect in the \textit{Warriner's} sense, there should be a conjunction (\textit{\&}). Also, what exactly is the purpose of this paragraph?

1st, the sentence may be flawed in a technical sense, but it's a good one in terms of the entire passage. Its brevity \& telegraphic style vary the pace \& keep the writing fresh. Suspense novelist Jonathan Kellerman uses this technique very successfully. In \textit{Survival of the Fittest}, he writes:
\begin{quotation}
	The boat was 30 feet of sleek white fiberglass with gray trim. Tall masts, the sails tied. \textit{Satori} painted on the hull in black script edged with gold.
\end{quotation}
It is possible to overuse the well-turned fragment (\& Kellerman sometimes does), but frags can also work beautifully to streamline narration, create clear images, \& create tension as well as to vary the prose-line. A series of grammatically proper sentences can stiffen that line, make it less pliable. Purists hate to hear that \& will deny it to their dying breath, but it's true. Language does not always have to wear a tie \& lace-up shoes. The object of fiction isn't grammatical correctness but to make the reader welcome \& then tell a story $\ldots$ to make him{\tt/}her forget, whenever possible, that he{\tt/}she is reading a story at all. The single-sentence paragraph more closely resembles talk than writing, \& that's good. Writing is seduction. Good talk is part of seduction. If not so, why do so many couples who start the evening at dinner wind up in bed?

The other uses of this paragraph include stage direction, minor but useful enhancement of character \& setting, \& a vital moment of transition. From protesting that his story is true, Big Tony moves on to his memories of O'Leary. Since the source of dialogue doesn't change, Tony's sitting down \& lighting up could take place in the same paragraph, with the dialogue picking up again afterward, but the writer doesn't elect to do it that way. Because Big Tony takes a new tack, the writer breaks the dialogue into 2 paragraphs. It's a decision made instantaneously in the course of writing, one based entirely on the beat the writer hears in his{\tt/}her own head. That beat is part of the genetic hardwiring (Kellerman writes a lot of frags because he \textit{hears} a lot of frags), but it's also the result of the thousands of hours that writer has spent composing, \& the \textit{10s} of thousands of hours he{\tt/}she may have spent reading the compositions of others.

I would argue that the paragraph, not the sentence, is the basic unit of writing -- the place where coherence begins \& words stand a chance of becoming more than mere words. If the moment of quickening is to come, it comes at the level of the paragraph. It is a marvelous \& flexible instrument that can be a single word long or run on for pages (1 paragraph in Don Robertson's historical novel \textit{Paradise Falls} is 16 pages long; there are paragraphs in Ross Lockridge's \textit{Raintree County} which are nearly that). You must learn to use it well if you are to write well. What this means is lots of practice; you have to learn the beat.

\fbox{\bf5} Grab that book you were looking at off the shelf again, would you? The weight of it in your hands tells you other stuff that you can take in without reading a single word. The book's length, naturally, but more: the commitment the writer shouldered in order to create the work, the commitment Constant Reader must take to digest it. Not that length \& weight alone indicate excellence; many epic tales are pretty much epic crap -- just ask my critics, who will moan about entire Canadian forests massacred in order to print my drivel. Conversely, short doesn't always mean sweet. In some cases (e.g., \textit{The Bridges of Madison County}), short means far \textit{too} sweet. But there is that matter of commitment, whether a book is good or bad, a failure or a success. Words have weight. Ask anyone who works in the shipping department of a book company warehouse, or in the storage room of a large bookstore.

Words create sentences; sentences create paragraphs; sometimes paragraphs quicken \& begin to breathe. Imagine, if you like, Frankenstein's monster on its slab. Here comes lightning, not from the sky but from a humble paragraph of English words. Maybe it's the 1st really good paragraph you ever wrote, something so fragile \& yet full of possibility that you are frightened. You feel as Victor Frankenstein must have when the dead conglomeration of sewn-together spare parts suddenly opened its watery yellow eyes. \textit{Oh my God, it's breathing}, you realize. \textit{Maybe it's even thinking. What the hell's name do I do next?}

You go to the 3rd level, of course, \& begin to write real fiction. Why shouldn't you? Why should you fear? Carpenters don't build monsters, after all; they build houses, stores, \& banks. They build some of wood a plank at a time \& some of brick a brick at a time. You will build a paragraph at a time, constructing these of your vocabulary \& your knowledge of grammar \& basic style. As long as you stay level-on-the-level \& shave even every door, you can build whatever you like -- whole mansions, if you have the energy.

Is there any rationale for building entire mansions of words? I think there is, \& that the readers of Margaret Michell's \textit{Gone with the Wind}\& Charles Dickens's \textit{Bleak House} understand it: sometimes even a monster is no monster. Sometimes it's beautiful \& we fall in love with all that story, more than any film or TV program could ever hope to provide. Even after a thousand pages we don't want to leave the world the writer has made for us, or the make-believe people who live there. You wouldn't leave after 2000 pages, if there were 2000. The \textit{Rings} trilogy of J. R. R. Tolkien is a perfect example of this. 1000 pages of hobbits hasn't been enough for 3 generations of post-World War II fantasy fans; even when you add in that clumsy, galumphing dirigible of an epilogue, \textit{The Silmarillion}, it hasn't been enough. Hence Terry Brooks, Piers Anthony, Robert Jordan, the questing rabbits of \textit{Watership Down}, \& half 100 others. The writers of these books are creating the hobbits they still love \& pine for; they are trying to bring Frodo \& Sam back from the Grey Havens because Tolkien is no longer around to do it for them.

At its most basic we are only discussing a learned skill, but do we not agree that sometimes the most basic skills can create things far beyond our expectations? We are talking about tools \& carpentry, about words \& style $\ldots$ but as we move along, you'd do well to remember that we are also talking about magic.'' -- \cite[pp. 86--107]{King2010}

%------------------------------------------------------------------------------%

\section{ON WRITING}
``There are no bad dogs, according to the title of a popular training manual, but don't tell that to the parent of a child mauled by a pit bull or a rottweiler; he or she is apt to bust your beak for you. \& no matter how much I want to encourage the man or woman trying for the 1st time to write seriously, I can't lie \& say there are no bad writers. Sorry, but there are \textit{lots} of bad writers. Some are on-staff at your local newspaper, usually reviewing little-theater productions or pontificating about the local sports teams. Some have scribbled their way to homes in the Caribbean, leaving a trail of pulsing adverbs, wooden characters, \& vile passive-voice constructions behind them. Others hold forth at open-mike poetry slams, wearing black turtlenecks \& wrinkled khaki pants; they spout doggerel about ``my angry lesbian breasts'' \& ``the titled alley where I cried my mother's name.''

Writers form themselves into the pyramid we see in all areas of human talent \& human creativity. At the bottom are the bad ones. Above them is a group which is slightly smaller but still large \& welcoming; these are the competent writers. They may also be found on the staff of your local newspaper, on the racks at your local bookstore, \& at poetry readings on Open Mike Night. These are folks who somehow understand that although a lesbian may be angry, her breasts will remain breasts.

The next level is much small. These are the really good writers. Above them -- above most all of us -- are the Shakespeares, the Faulkners, the Yeatses, Shaws, \& Eudora Weltys. They are genius, divine accidents, gifted in a way which is beyond our ability to understand, let alone attain. Shit, most geniuses aren't able to understand themselves, \& many of them lead miserable lives, realizing (at least on some level) that they are nothing but fortunate freaks, the intellectual version of runway models who just happen to be born with the right cheekbones \& with breasts which fit the image of an age.

I am approaching the heart of this book with 2 theses, both simple. The 1st is that good writing consists of mastering the fundamentals (vocabulary, grammar, the elements of style) \& then filling the 3rd level of your toolbox with the right instruments. The 2nd is that while it is impossible to make a competent writer out of a bad writer, \& while it is equally impossible to make a great writer our of a good one, it \textit{is} possible, with lots of hard work, dedication, \& timely help, to make a good writer out of a merely competent one.

I'm afraid this idea is rejected by lots of critics \& plenty of writing teachers, as well. Many of these are liberals in their politics but crustaceans in their chosen fields. Men \& women who would take to the streets to protest the exclusion of African-Americans or Native Americans (I can imagine what Mr. Strunk would have made of these politically correct but clunky terms) from the local country club are often the same men \& women who tell their classes that writing ability is fixed \& immutable; once a hack, always a hack. Even if a writer rises in the estimation of an influential critic or 2, he{\tt/}she always carries his{\tt/}her early reputation along, like a respectable married woman who was a wild child as a teenager. Some people never forget, that's all, \& a good deal of literary criticism serves only to reinforce a caste system which is as old as the intellectual snobbery which nurtured it. Raymond Chandler may be recognized now as an important figure in 20th-century American literature, an early voice describing the anomie of urban life in the years after World War II, but there are plenty of critics who will reject such a judgment out of hand. He's a hack! they cry indignantly. A hack with pretensions! The worst kind! The kind who thinks he can pass for 1 of \textit{us}!

Critics who try to rise above his intellectual hardening of the arteries usually meet with limited success. Their colleagues may accept Chandler into the company of the great, but are apt to seat him at the foot of the table. \& there are always those whispers: \textit{Came out of the pulp tradition, you know $\ldots$ carries himself well for 1 of those, doesn't he? $\ldots$ did you know he wrote for} Black Mash \textit{in the 30s $\ldots$ yes, regrettable} $\ldots$

Even Charles Dickens, the Shakespeare of the novel, has faced a constant critical attack as a result of his often sensational subject matter, his cheerful fecundity (when he wasn't creating novels, he \& his wife were creating children), \&, of course, his success with the book-reading groundlings of his time \& ours. Critics \& scholars have always been suspicious of popular success. Often their suspicions are justified. In other cases, these suspicions are used as an excuse not to think. No one can be as intellectually slothful as a really smart person; give smart people half a chance \& they will ship their oars \& drift $\ldots$ dozing to Byzantium, you might say.

So yes -- I expect to be accused by some of promoting a brainless \& happy Horatio Alger philosophy, defending my own less-than-spotless reputation while I'm at it, \& of encouraging people who are ``just not our sort, old chap'' to apply for membership at the country club. I guess I can live with t hat. But before we go on, let me repeat my basic premise: if you're a bad writer, no one can help you become a good one, or even a competent one. If you're good \& want to be great $\ldots$ fuhgeddaboudit.

What follows is everything I know about how to write good fiction. I'll be as brief as possible, because your time is valuable \& so is mine, \& we both understand that the hours we spend talking about writing is time we don't spend actually \textit{doing} it. I'll be as encouraging as possible, because it's my nature \& because I love this job. I want you to love it, too. But if you don't want to work your ass off, you have no business trying to write well -- settle back into competency \& be grateful you have even that much to fall back on. There is a muse,\footnote{Traditionally, the muses were women, but mine's a guy; I'm afraid we'll just have to live with that.} but he's not going to come fluttering down into your writing room \& scatter creative fairy-dust all over your typewriter or computer station. He lives in the ground. He's a basement guy. You have to descent to his level, \& once you get down there you have to furnish an apartment for him to live in. You have to do all the grunt labor, in other words, while the muse sits \& smokes cigars \& admires his bowling trophies \& pretends to ignore you. Do you think this is fair? \textit{I} think it's fair. He may not be much to look at, that muse-guy, \& he may not be much of a conversationalist (what I get out of mine is mostly surly grunts, unless he's on duty), but he's got the inspiration. It's right that you should do all the work \& burn all the midnight oil, because the guy with the cigar \& the little wings has got a bag of magic. There's stuff in there that can change your life.

Believe me, I know.

\fbox{\bf1} If you want to be a writer, you must do 2 things above all others: read a lot \& write a lot. There's no way around these 2 things that I'm aware of, no shortcut.

I'm a slow reader, but I usually get through 70 or 80 books a year, mostly fiction. I don't read in order to study the craft; I read because I like to read. It's what I do at night, kicked back in my blue chair. Similarly, I don't read fiction to study the art of fiction, but simply because I like stories. Yet there is a learning process going on. Every book you pick up has its own lesson or lessons, \& quite often the bad books have more to teach than the good ones.

When I was in the 8th grade, I happened upon a paperback novel by Murray Leinster, a science fiction pulp writer who did most of his work during the 40s \& 50s, when magazines like \textit{Amazing Stories} paid a penny a word. I had read other books by Mr. Leinster, enough to know that the quality of his writing was uneven. This particular tale, which was about mining in the asteroid belt, was 1 of his less successful efforts. Only that's too kind. It was terrible, actually, a story populated by paper-thin characters \& driven by outlandish plot developments. Worst of all (or so it seemed to met at the time), Leinster had fallen in love with the word \textit{zestful}. Characters watched the approach of ore-bearing asteroids with \textit{zestful smiles}. Characters sat down to supper aboard their mining ship with \textit{zestful anticipation}. Near the end of the book, the hero swept the large-breasted, blonde heroine into a \textit{zestful embrace}. For me, it was the literary equivalent of a smallpox vaccination: I have never, so far as I know, used the word \textit{zestful} in a novel or a story. God willing, I never will.

\textit{Asteroid Miners} (which wasn't the title, but that's close enough) was an important book in my life as a reader. Almost everyone can remember losing his or her virginity, \& most writers can remember the 1st book he{\tt/}she put down thinking: \textit{I can do better than this. Hell, I} am \textit{doing better than this!} What could be more encouraging to the struggling writer than to realize his{\tt/}her work is unquestionably better than that of someone who actually got paid for his{\tt/}her stuff?

One learns most clearly what not to do by reading bad prose -- 1 novel like \textit{Asteroid Miners} (or \textit{Valley of the Dolls, Flowers in the Attic}, \& \textit{The Bridges of Madison County}, to name just a few) is worth a semester at a good writing school, even with the superstar guest lecturers thrown in.

Good writing, on the other hand, teaches the learning writer about style, graceful narration, plot development, the creation of believable characters, \& truth-telling. A novel like \textit{The Grapes of Wrath} may fill a new writer with feelings of despair \& good old-fashioned jealousy -- ``I'll never be able to write anything that good, not if I live to be a thousand'' -- but such feelings can also serve as a spur, goading the writer to work harder \& aim higher. Being swept away by a combination of great story \& great writing -- of being flattened, in fact -- is part of every writer's necessary formation. You cannot hope to sweep someone else away by the force of your writing until it has been done to you.

So we read to experience the mediocre \& the outright rotten; such experience helps us to recognize those things when they begin to creep into our own work, \& to steer clear of them. We also read in order to measure ourselves against the good \& the great, to get a sense of all that can be done. \& we read in order to experience different styles.

You may find yourself adopting a style you find particularly exciting, \& there's nothing wrong with that. When I read Ray Bradbury as a kid, I wrote like Ray Bradburry -- everything green \& wondrous \& seen through a lens smeared with the grease of nostalgia. When I read James M. Cain, everything I wrote came out clipped \& stripped \& hard-boiled. When I read Lovecraft, my prose became luxurious \& Byzantine. I wrote stories in my teenage years where all these styles merged, creating a kind of hilarious stew. This sort of stylistic blending is a necessary part of developing one's own style, but it doesn't occur in a vacuum. You have to read widely, constantly refining (\& redefining) your own work as you do so. It's hard for me to believe that people who read very little (or not at all in some cases) should presume to write \& expect people to like what they have written, but I know it's true. If I had a nickel for every person who ever told me he{\tt/}she wanted to become a writer but ``didn't have time to read,'' I could buy myself a pretty good steak dinner. Can I be blunt on this subject? If you don't have time to read, you don't have the time (or the tools) to write. Simple as that.

Reading is the creative center of a writer's life. I take a book with me everywhere I go, \& find there are all sorts of opportunities to dip in. The trick is to teach yourself to read in small sips as well as in long swallows. Waiting rooms were made for books -- of course! But so are theater lobbies before the show, long \& boring checkout lines, \& everyone's favorite, the john. You can even read while you're driving, thanks to the audiobook revolution. Of the books I read each year, anywhere from 6 to a dozen are on tape. As for all the wonderful radio you will be missing, come on -- how many times can you listen to Deep Purple sing ``Highway Star''?

Reading at meals is considered rude in polite society, but if you expect to succeed as a writer, rudeness should be the 2nd-to-least of your concerns. The least of all should be polite society \& what it expects. If you intend to write as truthfully as you can, your days as a member of polite society are numbered, anyway.

Where else can you read? There's always the treadmill, or whatever you use down at the local health club to get aerobic. I try to spend an hour doing that every day, \& I think I'd go mad without a good novel to keep me company. Most exercise facilities (at home as well as outside it) are now equipped with TVs, but TV -- while working out or anywhere else -- really is about the last thing an aspiring writer needs. If you feel you must have the news analyst blowhards on CNN while you exercise, or the stock market blowhards on MSNBC, or the sports blowhards on ESPN, it's time for you to question how serious you really are about becoming a writer. You must be prepared to do some serious turning inward toward the life of the imagination, \& that means, I'm afraid, that Geraldo, Keith Obermann, \& Jay Leno must go. Reading takes time, \& the glass teat takes too much of it.

Once weaned from the ephemeral craving for TV, most people will find they enjoy the time they spend reading. I'd like to suggest that turning off that endlessly quacking box is apt to improve the quality of your life as well as the quality of your writing. \& how much of a sacrifice are we talking about here? How many \textit{Frasier} \& \textit{ER} reruns does it take to make 1 American life complete? How many Richard Simmons infomercials? How many whiteboy{\tt/}fatboy Beltway insiders on CNN? Oh man, don't get me started. Jerry-Springer-Dr.-Dre-Judge-Judy-Jerry-Falwell-Donny-and-Marie, I rest my case.

When my son Owen was 7 or so, he fell in love with Bruce Springsteen's E Street Band, particularly with Clarence Clemons, the band's burly sax player. Owen decided he wanted to learn to play like Clarence. My wife \& I were amused \& delighted by this ambition. We were also hopeful, as any parent would be, that our kid would turn out to be talented, perhaps even some sort of prodigy. We got Owen a tenor saxophone for Christmas \& lessons with Gordon Bowie, 1 of the local music men. Then we crossed our fingers \& hoped for the best.

7 months later I suggested to my wife that it was time to discontinue the sax lessons, if Owen concurred. Owen did, \& with palpable relief -- he hadn't wanted to say it himself, especially not after asking for the sax in the 1st place, but 7 months had been long enough for him to realize that, while he might love Clarence Clemons's big sound, the saxophone was simply not for him -- God had not given him that particular talent.

I knew, not because Owen stopped practicing, but because he was practicing only during the periods Mr. Bowie had set for him: half an hour after school 4 days a week, plus an hour on the weekends. Owen mastered the scales \& the notes -- nothing wrong with his memory, his lungs, or his eye-hand coordination -- but we never heard him taking off, surprising himself with something new, blissing himself out. \& as soon as his practice time was over, it was back into the case with the horn, \& there it stayed until the next lesson or practice-time. What this suggested to me was that when it came to the sax \& my son, there was never going to be any real play-time; it was all going to be rehearsal. That's no good. If there's no joy in it, it's just no good. It's best to go on to some other area, where the deposits of talent may be richer \& the fun quotient higher.

Talent renders the whole idea of rehearsal meaningless; when you find something at which you are talented, you do it (whatever \textit{it} is) until your fingers bleed or your eyes are ready to fall out of your head. Even when no one is listening (or reading, or watching), every outing is a bravura performance, because you as the creator are happy. Perhaps even ecstatic. That goes for reading \& writing as well as for playing a musical instrument, hitting a baseball, or running the 4-40. The sort of strenuous reading \& writing program I advocate -- 4--6 hours a day, every day -- will not seem strenuous if you really enjoy doing these things \& have an aptitude for them; in fact, you may be following such a program already. If you feel you need permission to do all the reading \& writing your little heart desires, however, consider it hereby granted by yours truly.

The real importance of reading is that it creates an ease \& intimacy with the process of writing; one comes to the country of the writer with one's papers \& identification pretty much in order. Constant reading will pull you into a place (a mindset, if you like the phrase) where you can write eagerly \& without self-consciousness. It also offers you a constantly growing knowledge of what has been done \& what hasn't, what is trite \& what is fresh, what works \& what just lies there dying (or dead) on the page. The more you read, the less apt you are to make a fool of yourself with your pen or word processor.

\fbox{\bf2} If ``read a lot, write a lot'' is the Great Commandment -- \& I assure you that it is -- how much writing constitutes a lot? That varies, of course, from writer to writer. 1 of my favorite stories on the subject - probably more myth than truth -- concerns James Joyce.\footnote{There are some great stories about Joyce. My absolute favorite is that, as his vision failed, he took to wearing a milkman's uniform while writing. Supposedly he believed it caught the sunlight \& reflected it down on his page.} According to the story, a friend came to visit him 1 day \& found the great man sprawled across his writing desk in a posture of utter despair.

``James, what's wrong?'' the friend asked. ``Is it the work?''

Joyce indicated assent without even raising his head to look at the friend. Of course it was the work; isn't it always?

``How many words did you get today?'' the friend pursued.

Joyce (still in despair, still sprawled facedown on his desk): ``7.''

``7? But James $\ldots$ that's \textit{good}, at least for you!''

``Yes,'' Joyce said, finally looking up. ``I suppose it is $\ldots$ but I don't know what \textit{order} they go in!''

At the other end of the spectrum, there are writers like Anthony Trollope. He wrote humongous novels (\textit{Can You Forgive Her?} is a fair enough example; for modern audiences it might be retitled \textit{Can You Possibly Finish It?}), \& he pumped them out with amazing regularity. His day job was as a clerk in the British Postal Department (the red public mailboxes all over Britain were Anthony Trollope's invention); he wrote for 2.5 hours each morning before leaving for work. This schedule was ironclad. If he was in mid-sentence when the 2.5 hours expired, he left that sentence unfinished until the next morning. \& if he happened to finish 1 of his 600-page heavyweights with 15 minutes of the session remaining, he wrote \textbf{The End}, set the manuscript aside, \& began work on the next book.

John Creasey, a British mystery novelist, wrote 500 (yes, you read it correctly) novels under 10 different names. I've written 35 or so -- some of Trollopian length -- \& am considered prolific, but I look positively blocked next to Creasey. Several other contemporary novelists (they include Ruth Rendell{\tt/}Barbara Vine, Evan Hunter{\tt/}Ed McBain, Dean Koontz, \& Joyce Carol Oates) have written easily as much as I have; some have written a good deal more.

On the other hand -- the James Joyce hand -- there is Harper Lee, who wrote only 1 book (the brilliant \textit{To Kill a Mockingbird}). Any number of others, including James Agee, Malcolm Lowry, \& Thomas Harris (so far), wrote under 5. Which is okay, but I always wonder 2 things about these folks: how long did it take them to write the books they \textit{did} write, \& what did they do the rest of their time? Knit afghans? Organize church bazaars? Deify plums? I'm probably being snotty here, but I am also, believe me, honestly curious. If God gives you something you can do, why in God's name wouldn't you do it?

My own schedule is pretty clear-cut. Mornings belong to whatever is new -- the current composition. Afternoons are for naps \& letters. Evenings are for reading, family, Red Sox games on TV, \& any revisions that just cannot wait. Basically, mornings are my prime writing time.

Once I start work on a project, I don't stop \& I don't slow down unless I absolutely have to. If I don't write every day, the characters begin to stale off in my mind -- they begin to \textit{seem} like characters instead of real people. The tale's narrative cutting edge starts to rust \& I begin to lose my hold on the story's plot \& peace. Worst of all, the excitement of spinning something new begins to fade. The work starts to \textit{feel} like work, \& for most writers that is the smooch of death. Writing is at its best -- always, always, always -- when it is a kind of inspired play for the writer. I can write in cold blood if I have to, but I like it best when it's fresh \& almost too hot to handle.

I used to tell interviewers that I wrote every day except for Christmas, the 4th of July, \& my birthday. That was a lie. I told them that because if you agree to an interview you have to say \textit{something}, \& it plays better if it's something at least half-clever. Also, I didn't want to sound like a workaholic dweeb (just a workaholic, I guess). The truth is that when I'm writing, I write every day, workaholic dweeb or not. That \textit{includes} Christmas, the 4th, \& my birthday (at my age you try to ignore your goddam birthday anyway). \& when I'm not working, I'm not working at all, although during those periods of full stop I usually feel at loose ends with myself \& have trouble sleeping. For me, not working is the real work. When I'm writing, it's all the playground, \& the worst 3 hours I ever spent there were still pretty damned good.

I used to be faster than I am now; 1 of my books (\textit{The Running Man}) was written in a single week, an accomplishment John Creasey would perhaps have appreciated (although I have read that Creasey wrote several of his mysteries in \textit{2 days}). I think it was quitting smoking that slowed me down; nicotine is a great synapse enhancer. The problem, of course, is that it's killing you at the same time it's helping you compose. Still, I believe the 1st draft of a book -- even a long one -- should take no more than 3 months, the lengths of a season. Any longer \& -- for me, at least -- the story begins to take on an odd foreign feel, like a dispatch from the Romanian Department of Public Affairs, or something broadcast on high-band shortwave during a period of severe sunspot activity.

I like to get 10 pages a day, which amounts to 2000 words. That's 180000 words over a 3-month span, a goodish length for a book -- something in which the reader can get happily lost, if the table is done well \& stays fresh. On some days those 10 pages come easily; I'm up \& out \& doing errands by 11:30 in the morning, perky as a rat in liverwurst. More frequently, as I grow older, I find myself eating lunch at my desk \& finishing the day's work around 1:30 in the afternoon. Sometimes, when the words come hard, I'm still fiddling around at teatime. Either way is fine with me, but only under dire circumstances do I allow myself to shut down before I get my 2000 words.

The biggest aid to regular (Trollopian?) production is working in a serene atmosphere. It's difficult for even the most naturally productive writer to work in an environment where alarms \& excursions are the rule rather than the exception. When I'm asked for ``the secret of my success'' (an absurd idea, that, but impossible to get away from), I sometimes say there are 2: I stayed physically healthy (at least until a van knocked me down by the side of the road in the summer of 1999), \& I stayed married. It's a good answer because it makes the question go away, \& because there is an element of truth in it. The combination of a healthy body \& a stable relationship with a self-reliant woman who takes zero shit from me or anyone else has made the continuity of my working life possible. \& I believe the converse is also true: that my writing \& the pleasure I take in it has contributed to the stability of my health \& my home life.

\fbox{\bf3} You can read anywhere, almost, but when it comes to writing, library carrels, park benches, \& rented flats should be courts of last resort -- Truman Capote said he did his best work in motel rooms, but he is an exception; most of us do our best in a place of our own. Until you get one, you'll find your new resolution to write a lot hard to take seriously.

Your writing room doesn't have to sport a Playboy Philosophy decor, \& you don't need an Early American rolltop desk in which to house your writing implements. I wrote my 1st 2 published novels, \textit{Carrie} \& \textit{`Salem's Lot}, in the laundry room of a doublewide trailer, pounding away on my wife's portable Olivetti typewriter \& balancing a child's desk on my thighs; John Cheever reputedly wrote in the basement of his Park Avenue apartment building, near the furnace. The space can be humble (probably \textit{should} be, as I think I have already suggested), \& it really needs only 1 thing: a door which you are willing to shut. The closed door is your way of telling the world \& yourself that you mean business; you have made a serious commitment to write \& intend to walk the walk as well as talk the talk.

By the time you step into your new writing space \& close the door, you should have settled on a daily writing goal. As with physical exercise, it would be best to set this goal low at 1st, to avoid discouragement. I suggest 1000 words a day, \& because I'm feeling magnanimous, I'll also suggest that you can take 1 day a week off, at least to begin with. No more; you'll lose the urgency \& immediacy of your story if you do. With that goal set, resolve to yourself that the door stays closed until that goal is met. Get busy putting those thousand words on paper or on a floppy disk. In an early interview (this was to promote \textit{Carrie}, I think), a radio talk-show host asked me how I wrote. My reply -- ``1 word at a time'' -- seemingly left him without a reply. I think he was trying to decide whether or not I was joking. I wasn't. In the end, it's always that simple. Whether it's a vignette of a single page or an epic trilogy like \textit{The Lord of the Rings}, the work is always accomplished 1 word at a time. The door closes the rest of the world out; it also serves to close you in \& keep you focused on the job at hand.

If possible, there should be no telephone in your writing room, certainly no TV or videogames for you to fool around with. If there's a window, draw the curtains or pull down the shades unless it looks out at a blank wall. For any writer, but for the beginning writer in particular, it's wise to eliminate every possible distraction. If you continue to write, you will begin to filter out these distractions naturally, but at the start it's best to try \& take care of them before you write. I work to loud music -- hard-rock stuff like AC{\tt/}DC, Guns `n Roses, \& Metallica have always been particular favorites -- but for me the music is just another way of shutting the door. It surrounds me, keeps the mundane world out. When you write, you want to get rid of the world, do you not? Of course you do. When you're writing, you're creating your own worlds.

I think we're actually talking about creative sleep. Like your bedroom, your writing room should be a private, a place where you go to dream. Your schedule -- in at about the same time every day, out when your thousand words are on paper or disk -- exists in order to habituate yourself, to make yourself ready to dream just as you make yourself ready to sleep by going to bed at roughly the same time each night \& following the same rituals as you go. In both writing \& sleeping, we learn to be physically still at the same time we are encouraging our minds to unlock from the humdrum rational thinking of our daytime lives. \& as your mind \& body grow accustomed to a certain amount of sleep each night -- 6 hours, 7, maybe the recommended 8 -- so can you train your waking mind to sleep creatively \& work out the vividly imagined waking dreams which are successful works of fiction.

But you need the room, you need the door, \& you need the determination to shut the door. You need a concrete goal, as well. The longer you keep to these basics, the easier the act of writing will become. Don't wait for the muse. As I've said, he's a hardheaded guy who's not susceptible to a lot of creative fluttering. This isn't the Ouija board or the spirit-world we're talking about here, but just another job like laying pipe or driving long-haul trucks. Your job is to make sure the muse knows where you're going to be every day from 9 `til noon or 7 `til 3. If he does know, I assure you that sooner or later he'll start showing up, chomping his cigar \& making his magic.

\fbox{\bf4} So okay -- there you are in your room with the shade down \& the door shut \& the plug pulled out of the base of the telephone. You've blown up your TV \& committed yourself to 1000 words a day, come hell or high water. Now comes the big question: What are you going to write about? \& the equally big answer: Anything you damn well want. Anything at all $\ldots$ \textit{as long as you tell the truth}.

The dictum in writing classes used to ``write what you know.'' Which sounds good, but what if you want to write about starships exploring other planets or a man who murders his wife \& then tries to dispose of her body with a wood-chipper? How does the writer square either of these, or 1000 other fanciful ideas, with the ``write-what-you-know'' directive?

I think you begin by interpreting ``write what you know'' as broadly \& inclusively as possible. If you're a plumber, you know plumbing, but that is far from the extent of your knowledge; the heart also knows things, \& so does the imagination. Thank God. If not for heart \& imagination, the world of fiction would be a pretty seedy place. It might not even exist at all.

In terms of genre, it's probably fair to assume that you will begin by writing what you love to read -- certainly I have recounted my early love affair with the EC horror comics until the tale has gone stale. But I \textit{did} love them, ditto horror movies like \textit{I Married a Monster from Outer Space}, \& the result was stories like ``I Was a Teenage Graverobber.'' Even today I'm not above writing slightly more sophisticated versions of that tale; I was built with a love of the night \& the unquiet coffin, that's all. If you disapprove, I can only shrug my shoulders. It's what I have.

If you happen to be a science fiction fan, it's natural that you should want to write science fiction (\& the more sf you've read, the less likely it is that you'll simply revisit the field's well-mined conventions, such as space opera \& dystopian satire). If you're a mystery fan, you'll want to write mysteries, \& if you enjoy romances, it's natural for you to want to write romances of your own. There's nothing wrong with writing any of these things. What could be very wrong, I think, is to turn away from what you know \& like (or love, the way I loved those old ECs \& black-\&-white horror flicks) in favor of things you believe will impress your friends, relatives, \& writing-circle colleagues. What's equally wrong is the deliberate turning toward some genre or type of fiction in order to make money. It's morally wonky, for 1 thing -- the job of fiction is to find the truth inside the story's web of lies, not to commit intellectual dishonesty in the hunt for the buck. Also, brothers \& sisters, it doesn't work.

When I'm asked why I decided to write the sort of thing I do write, I always think the question is more revealing than any answer I could possibly give. Wrapped within it, like the chewy stuff in the center of a Tootsie Pop, is the assumption that the writer controls the material instead of the other way around.\footnote{irby McCauley, my 1st real agent, used to quote science fiction writer Alfred Bester (\textit{The Stars My Destination, The Demolished Man}) on this subject. ``The book is the boss,'' Alfie used to say in tones indicating that that closed the subject.} The writer who is serious \& committed is incapable of sizing up story material the way an investor might size up various stock offerings, picking out the ones which seem likely to provide a good return. If it could indeed be done that way, every novel published would be a best-seller \& the huge advances paid to a dozen or so ``big-name writers'' would not exist (publishers would like that).

Grisham, Clancy, Crichton, \& myself -- among others -- are paid these large sums of money because we are selling uncommonly large numbers of books to uncommonly large audiences. A critical assumption is sometimes made that we have access to some mystical vulgate that other (\& often better) writers either cannot find or will not deign to use. I doubt if this is true. Nor do I believe the contention of some popular novelists (although she was not the only one, I am thinking of the late Jacqueline Susann) that their success is based on literary merit -- that the public understands true greatness in ways the tight-assed, consumed-by-jealousy literary establishment cannot. This idea is ridiculous, a product of vanity \& insecurity.

Book-buyers aren't attracted, by \& large, by the literary merits of a novel; book-buyers want a good story to take with them on the airplane, something that will 1st fascinate them, then pull them in \& keep them turning the pages. This happens, I think, when readers recognize the people in a book, their behaviors, their surroundings, \& their talk. When the reader hears strong echoes of his or her own life \& beliefs, he or she is apt to become more invested in the story. I'd argue that it's impossible to make this sort of connection in a premeditated way, gauging the market like a racetrack tout with a hot top.

Stylistic imitation is 1 thing, a perfectly honorable way to get started as a writer (\& impossible to avoid, really; some sort of imitation marks each new stage of a writer's development), but one cannot imitate a writer's approach to a particular genre, no matter how simple what that writer is doing may seem. You can't aim a book like a cruise missile, in other words. People who decide to make a fortune writing like John Grisham or Tom Clancy produce nothing but pale imitations, by \& large, because vocabulary is not the same thing as feeling \& plot is light-years from the truth as it is understood by the mind \& the heart. When you see a novel with \textit{``In the tradition of} (John Grisham{\tt/}Patricia Cornwell{\tt/}Mary Higgins Clark{\tt/}Dean Koontz)'' on the cover, you know you are looking at 1 of these overcalculated (\& likely boring) imitations.

Write what you like, them imbue it with life \& make it unique by blending in your own personal knowledge of life, friendship, relationships, sex, \& work. Especially work. People love to read about work. God knows why, but they do. If you're a plumber who enjoys science fiction, you might well consider a novel about a plumber aboard a starship or on an alien planet. Sound ludicrous? The late Clifford D. Simak wrote a novel called \textit{Cosmic Engineers} which is close to just that. \& it's a terrific read. What you need to remember is that there's a difference between lecturing about what you know \& using it to enrich the story. The latter is good. The former is not.

Consider John Grisham's breakout novel, \textit{The Firm}. In this story, a young lawyer discovers that his 1st job, which seemed too good to be true, really is -- he's working for the Mafia. Suspenseful, involving, \& paced at breakneck speed, \textit{The Firm} sold roughly 9 gazillion copies. What seemed to fascinate its audience was the moral dilemma in which the young lawyer finds himself: working for the mob is bad, no argument here, but the frocking pay is \textit{great!} You can drive a beemer, \& that's just for openers!

Audiences also enjoyed the lawyers's resourceful efforts to extricate himself from his dilemma. It might not be the way most people would behave, \& the \textit{deus ex machina} clanks pretty steadily in the last 50 pages, but it \textit{is} the way most of us would \textit{like} to behave. \& wouldn't we also like to have a \textit{deus ex machina} in our lives?

Although I don't know for sure, I'd bet my dog \& lot that John Grisham never worked for the mob. All of that is total fabrication (\& total fabrication is the fiction-writer's purest delight). He \textit{was} once a young lawyer, though, \& he has a clearly forgotten none of the struggle. Nor has he forgotten the location of the various financial pitfalls \& honeytraps that make the field of corporate law so difficult. Using plainspun humor as a brilliant counterpoint \& never substituting cant for story, he sketches a world of Darwinian struggle where all the savages wear 3-piece suits. \& -- here's the good part -- \textit{this is a world impossible not to believe}. Grisham has been there, spied out the land \& the enemy positions, \& brought back a full report. He told the truth of what he knew, \& for that if nothing else, he deserves every buck \textit{The Firm} made.

Critics who dismissed \textit{The Firm} \& Grisham's later books as poorly written \& who profess themselves to be mystified by his success are either missing the point because it's so big \& obvious or because they are being deliberately obtuse. Grisham's make-believe tale is solidly based in a reality he knows, has personally experienced, \& which he wrote about with total (almost naive) honesty. The result is a book which is -- cardboard characters or no, we could argue about that -- both brave \& uniquely satisfying. You as a beginning writer would do well not to imitate the lawyers-in-trouble genre Grisham seems to have created but to emulate Grisham's openness \& inability to do anything other than get right to the point.

John Grisham, of course, knows lawyers. What \textit{you} know makes you unique in some other way. Be brave. Map the enemy's positions, come back, tell us all you know. \& remember that plumbers in space is not such a bad setup for a story.

\fbox{\bf5} In my view, stories \& novels consist of 3 parts: narration, which moves the story from point A to point B \& finally to point Z; description, which creates a sensory reality for the reader; \& dialogue, which brings characters to life through their speech.

You may wonder where plot is in all this. The answer -- my answer, anyway -- is nowhere. I won't try to convince you that I've never plotted any more than I'd try to convince you that I've never told a lie, but I do both as infrequently as possible. I distrust plot for 2 reasons: 1st, because our \textit{lives} are largely plotless, even when you add in all our reasonable precautions \& careful planning; \& 2nd, because I believe plotting \& the spontaneity of real creation aren't compatible. It's best that I be as clear about this as I can -- I want you to understand that my basic belief about the making of stories is that they pretty much make themselves. The job of the writer is to give them a place to grow (\& to transcribe them, of course). If you can see things this way (or at least try to), we can work together comfortably. If, on the other hand, you decide I'm crazy, that's fine. You won't be the 1st.

When, during the course of an interview for \textit{The New Yorker}, I told the interviewer (Mark Singer) that I believed stories are found things, like fossils in the ground, he said that he didn't believe me. I replied that that was fine, as long as he believed that \textit{I} believe it. \& I do. Stories aren't souvenir tee-shirts for GameBoys. Stories are relics, part of an undiscovered pre-existing world. The writer's job is to use the tools in his or her toolbox to get as much of each one out of the ground intact as possible. Sometimes the fossil you uncover is small; a seashell. Sometimes it's enormous, a \textit{Tyrannosaurus Rex} with all those gigantic ribs \& grinning teeth. Either way, short story or thousand-page whopper of a novel, the techniques of excavation remain basically the same.

No matter how good you are, no matter how much experience you have, it's probably impossible to get the entire fossil out of the ground without a few breaks \& losses. To get even \textit{most} of it, the shovel must give way to more delicate tools: airhose, palm-pick, perhaps even a toothbrush. Plot is a far bigger tool, the writer's jackhammer. You can liberate a fossil from hard ground with a jackhammer, no argument there, but you know as well as I do that the jackhammer is going to break almost as much stuff as it liberates. It's clumsy, mechanical, anticreative. Plot is, I think, the good writer's last resort \& the dullard's 1st choice. The story which results from it is apt to feel artificial \& labored.

I lean more heavily on intuition, \& have been able to do that because my books tend to be based on situation rather than story. Some of the ideas which have produced those books are more complex than others, but the majority start out with the stark simplicity of a department store window display or a waxwork tableau. I want to put a group of characters (perhaps a pair; perhaps even just 1) in some sort of predicament \& then watch them try to work themselves free. My job isn't to \textit{help} them work their way free, or manipulate them to safety -- those are jobs which require the noisy jackhammer of plot -- but to watch what happens \& then write it down.

The situation comes 1st. The characters -- always flat \& unfeatured, to begin with -- come next. Once these things are fixed in my mind, I begin to narrate. I often have an idea of what the outcome may be, but I have never demanded of a set of characters that they do things my way. On the contrary, I want them to do things \textit{their} way. In some instances, the outcome is what I visualized. In most, however, it's something I never expected. For a suspense novelist, this is a great thing. I am, after all, not just the novel's creator but its 1st reader. \& if \textit{I'm} not able to guess with any accuracy how the damned thing is going to turn out, even with my inside knowledge of coming events, I can be pretty sure of keeping the reader in a state of page-turning anxiety. \& why worry about the ending anyway? Why be such a control freak? Sooner or later every story comes out \textit{somewhere}.

In the early 1980s, my wife \& I went to London on a combined business{\tt/}pleasure trip. I fell asleep on the plane \& had a dream about a popular writer (it may or may not have been me, but it sure to God wasn't James Caan) who fell into the clutches of a psychotic fan living on a farm somewhere out in the back of the beyond. The fan was a woman isolated by her growing paranoia. She kept some livestock in the barn, including her pet pig, Misery. The pig was named after the continuing main character in the writer's best-selling bodice-rippers. My clearest memory of this dream upon waking was something the woman said to the writer, who had a broken leg \& was being kept prisoner in the back bedroom. I wrote it on an American Airlines cocktail napkin so I wouldn't forget it, then put it in my pocket. I lost it somewhere, but can remember most of what I wrote down:
\begin{quotation}
	\textit{She speaks earnestly but never quite makes eye contact. A big woman \& solid all through; she is an absence of hiatus.} (Whatever \textit{that} means; remember, I'd just woken up.) \textit{``I wasn't trying to be funny in a mean way when I named my pig Misery, no sir. Please don't think that. No, I named her in the spirit of fan love, which is the purest love there is. You should be flattered.''}
\end{quotation}
Tabby \& I stayed at Brown's Hotel in London, \& on our 1st night there I was unable to sleep. Some of it was what sounded like a trio of little-girl gymnasts in the room directly above ours, some of it was undoubtedly jet lag, but a lot of it was that airline cocktail napkin. Jotted on it was the seed of what I thought could be a really excellent story, one that might turn out funny \& satiric as well as scary. I thought it was just too rich not to write.

I got up, went downstairs, \& asked the concierge if there was a quiet place where I could work longhand for a bit. He led me to a gorgeous desk on the 2nd-floor stair landing. It had been Rudyard Kipling's desk, he told me with perhaps justifiable pride. I was a little intimidated by this intelligence, but the spot was quiet \& the desk seemed hospitable enough; it featured about an acre of cherrywood working surface, for 1 thing. Stoked on cup after cup of tea (I drank it by the gallon when I wrote $\ldots$ unless I was drinking beer, that is), I filled 16 pages of a steno notebook. I like to work longhand, actually; the only problem is that, once I get jazzed, I can't keep up with the lines forming in my head \& I get frazzled.

When I called it quits, I stopped in the lobby to thank the concierge again for letting me use Mr. Kipling's beautiful desk. ``I'm so glad you enjoyed it,'' he replied. He was wearing a misty, reminiscent little smile, as if he had known the writer himself. ``Kipling died there, actually. Of a stroke. While he was writing.''

I went back upstairs to catch a few hours' sleep, thinking of how often we are given information we really could have done without.

The working title of my story, which I thought would be a novella of about 30000 words, was ``The Annie Wilkes Edition.'' When I sat down at Mr. Kipling's beautiful desk I had the basic situation -- crippled writer, psycho fan -- firmly fixed in my mind. The actual \textit{story} did not as then exist (well, it did, but as a relic buried -- except for 16 handwritten pages, that is -- in the earth), but knowing the story wasn't necessary for me to begin work. I had located the fossil; the rest, I knew, would consist of careful excavation.

I'd suggest that what works for me may work equally well for you. If you are enslaved to (or intimidated by) the tiresome tyranny of the outline \& the notebook filled with ``Character Notes,'' it may liberate you. At the very least, it will turn your mind to something more interesting than Developing the Plot.

(An amusing sidelight: the century's greatest supporter of Developing the Plot may have been Edgar Wallace, a best-selling potboiler novelist of the 1920s. Wallace invented -- \& patented -- a device called the Edgar Wallace Plot Wheel. When you got stuck for the next Plot Development or needed an Amazing Turn of Events in a hurry, you simply spun the Plot Wheel \& read what came up in the window: \textit{a fortuitous arrival}, perhaps, or \textit{Heroine declares her love}. These gadgets apparently sold like hotcakes.)

By the time I had finished that 1st Brown's Hotel session, in which Paul Sheldon wakes up to find himself Annie Wilkes's prisoner, I thought I knew what was going to happen. Annie would demand that Paul write another novel about his plucky continuing character, Misery Chastain, one just for her. After 1st demurring, Paul would of course agree (a psychotic nurse, I thought, could be very persuasive). Annie would tell him she intended to sacrifice her beloved pig, Misery, to this project. \textit{Misery's Return} would, she'd say, consist of but 1 copy: a holographic manuscript bound in pigskin!

Here we'd fade out, I thought, \& return to Annie's remote Colorado retreat 6 or 8 months later for the surprise ending.

Paul is gone, his sickroom turned into a shrine to Misery Chastain, but Misery the pig is still very much in evidence, grunting serenely away in her sty beside the barn. On the walls of the ``Misery Room'' are book covers, stills from the Misery movies, pictures of Paul Sheldon, perhaps a newspaper headline reading \textsc{famed romance novelist still missing}. In the center of the room, carefully spotlighted, is a single book on a small table (a cherrywood table, of course, in honor or Mr. Kipling). It is the Annie Wilkes Edition of \textit{Misery's Return}. The binding is beautiful, \& it should be; it is the skin of Paul Sheldon. \& Paul himself? Hi bones might be buried behind the barn, but I thought it likely that the pig would have eaten the tasty parts.

Not bad, \& it would have made a pretty good story (not such a good novel, however; no one likes to root for a guy over the course of 300 pages only to discover that between chapters 16 \& 17 the pig ate him), but that wasn't the way things eventually went. Paul Sheldon turned out to be a good deal more resourceful than I initially thought, \& his efforts to play Scheherazade \& save his life gave me a chance to say some things about the redemptive power of writing that I had long felt but never articulated. Annie also turned out to be more complex than I'd 1st imagined her, \& she was great fun to write about -- here was a woman pretty much stuck with ``cockadoodie brat'' when it came to profanity, but who felt absolutely no qualms about chopping off her favorite writer's foot when he tried to get away from her. In the end, I felt that Annie was almost as much to be pitied as to be feared. \& none of the story's details \& incidents proceeded from plot; they were organic, each arising naturally from the initial situation, each an uncovered part of the fossil. \& I'm writing all this with a smile. As sick with drugs \& alcohol as I was much of the time, I had such fun with that one.

\textit{Gerald's Game} \& \textit{The Girl Who Loved Tom Gordon} are 2 other purely situational novels. If \textit{Misery} is ``2 characters in a house,'' then \textit{Gerald} is ``1 woman in a bedroom'' \& \textit{The Girl Who} is ``1 kid lost in the woods.'' As I told you, I \textit{have} written plotted novels, but the results, in books like \textit{Insomnia} \& \textit{Rose Madder}, have not been particularly inspiring. These are (much as I hate to admit it) stiff, trying-too-hard novels. The only plot-driven novel of mine which I really like is \textit{The Dead Zone} (\& in all fairness, I must say I like that one a great deal). 1 book which \textit{seems} plotted -- \textit{Bag of Bones} -- is actually another situation: ``widowed writer in a haunted house.'' The back story of \textit{Bag of Bones} is satisfying gothic (at least I think so) \& very complex, but none of the details were premeditated. The hisotry of TR-90 \& the story of what widowed writer Mike Noonan's wife was really up to during the last summer of her life arose spontaneously -- all those details were parts of the fossil, in other words.

A strong enough situation renders the whole question of plot moot, which is fine with me. The most interesting situations can usually be expressed as a \textit{What-if} question:

\textit{What if} vampires invaded a small New England village? (\textit{`Salem's Lot})

\textit{What if} a policeman in a remote Nevada town went berserk \& started killing everyone in sight? (\textit{Desperation})

\textit{What if} a cleaning woman suspected of a murder she got away with (her husband) fell under suspicion for a murder she did not commit (her employer)? (\textit{Dolores Claiborne})

\textit{What if} a young mother \& her son became trapped in their stalled car by a rabid dog? (\textit{Cujo})

These were all situations which occurred to me -- while showering, while driving, while taking my daily walk -- \& which I eventually turned into books. In no case were they plotted, not even to the extent of a single note jotted on a single piece of scrap paper, although some of the stories (e.g., \textit{Dolores Claiborne}) are almost as complex as those you find in murder mysteries. Please remember, however, that there is a huge difference between story \& plot. Story is honorable \& trustworthy; plot is shifty, \& best kept under house arrest.

Each of the novels summarized above was smoothed out \& detailed by the editorial process, of course, but most of the elements existed to begin with. ``A movie should be there in rough cut,'' the film editor Paul Hirsch once told me. The same is true of books. I think it's rare that incoherence or dull storytelling can be solved by something so minor as a 2nd draft.

This isn't a textbook, \& so there aren't a lot of exercises, but I want to offer you one now, in case you feel that all this talk about situation replacing plot is so much woolly-headed bullshit. I am going to show you the location of a fossil. Your job is to write 5 or 6 pages of unplotted narration concerning this fossil. Put another way, I want you to dig for the bones \& see what they look like. I think you may be quite surprised \& delighted with the results. Ready? Here we go.

Everyone is familiar with the basic details of the following story; with small variations, it seems to pop up in the Police Beat section of metropolitan daily papers every other week or so. A woman -- call her Jane -- marries a man who is bright, witty, \& pulsing with sexual magnetism. We'll call the guy Dick; it's the world's most Freudian name. Unfortunately, Dick has a dark side. He's short-tempered, a control freak, perhaps even (you'll find this out as he speaks \& acts) a paranoid. Jane tries mightily to overlook Dick's faults \& make the marriage work (why she tries so hard is something you will also find out; she will come onstage \& tell you). They have a child, \& for awhile things seem better. Then, when the little girls is 3 or so, the abuse \& the jealous tirades begin again. The abuse is verbal at 1st, then physical. Dick is convinced that Jane is sleeping with someone, perhaps someone from her job. Is it someone specific? I don't know \& don't care. Eventually Dick may tell you who he suspects. If he does, we'll both know, won't we?

At last poor Jane can't take it anymore. She divorces the schmuck \& gets custody of their daughter, Little Nell. Dick begins to stalk her. Jane responds by getting a restraining order, a document about as useful as a parasol in a hurricane, as many abused women will tell you. Finally, after an incident which you will write in vivid \& scary detail -- a public beating, perhaps -- Richard the Schmuck is arrested \& jailed. All of this is back story. How you work it in -- \& \textit{how much} of it you work in -- is up to you. In any case, it's not the situation. What follows is the situation.

1 day shortly after Dick's incarceration in the city jail, Jane picks up Little Nell at the daycare center \& ferries her to a friend's house for a birthday party. Jane then takes herself home, looking forward to 2 or 3 hours' unaccustomed peace \& quiet. Perhaps, she thinks, I'll take a nap. It's a house she's going to, even though she's a young working woman -- the situation sort of demands it. How she came by this house \& why she has the afternoon off are thins the story will tell you \& which will look neatly plotted if you come up with good reasons (perhaps the house belongs to her parents; perhaps she's house-sitting; perhaps another thing entirely).

Something pings at her, just below the level of consciousness, as she lets herself in, something that makes her uneasy. She can't isolate it \& tells herself it's just nerves, a little fallout from her 5 years of hell with Mr. Congeniality. What else \textit{could} it be? Dick is under lock \& key, after all.

Before taking her nap, Jane decides to have a cup of herbal tea \& watch the news. (Can you use that pot of boiling water on the stove later on? Perhaps, perhaps.) The lead item on \textit{Action News at 3} is a shocker: that morning, 3 men escaped from the city jail, killing a guard in the process. 2 of the 3 bad guys were recaptured almost at once, but the 3rd is still at large. None of the prisoners are identified by name (not in \textit{this} newscast, at least), but Jane, sitting in her empty house (which you will now have plausibly explained), knows beyond a shadow of a doubt that 1 of them was Dick. She knows because she has finally identified that ping of unease she felt in the foyer. It was the smell, faint \& fading, of Vitalis hair-tonic. \textit{Dick's} hair-tonic. Jane sits in her chair, her muscles lax with fright, unable to get up. \& as she hears Dick's footfalls begin to descend the stairs, she thinks: \textit{Only Dick would make sure he had hair-tonic, even in jail}. She must get up, must run, but she can't move $\ldots$

It's a pretty good story, yes? I think so, but not exactly unique. As I've already pointed out, \textsc{estranged hubby beats up} (or \textsc{murders) ex-wife} makes the paper every other week, sad but true. What I want you to do in this exercise is \textit{change the sexes of the antagonist \& protagonist} before beginning to work out the situation in your narrative -- make the ex-wife the stalker, in other words (perhaps it's a mental institution she's escaped from instead of the city jail), the husband the victim. Narrate this without plotting -- let the situation \& that 1 unexpected inversion carry you along. I predict you will succeed swimmingly $\ldots$ if, i.e., you are honest about how your characters speak \& behave. Honesty is storytelling makes up for a great many stylistic faults, as the work of wooden-prose writers like Theodore Dreister \& Ayn Rand shows, but lying is the great unrepairable fault. Liars prosper, no question about it, but only in the grand sweep of things, never down in the jungles of actual composition, where you must take your objective 1 bloody word at a time. If you begin to lie about what you know \& feel while you're down there, everything falls down.

When you finish your exercise, drop me a line at \url{www.stephenking.com} \& tell me how it worked for you. I can't promise to vet every reply, but I \textit{can} promise to read at least some of your adventures with great interest. I'm curious to know what kind of fossil you dig up, \& how much of it you are able to retrieve from the ground intact.

\fbox{\bf16} Description is what makes the reader a sensory participant in the story. Good description is a learned skill, 1 of the prime reasons why you cannot succeed unless you read a lot \& write a lot. It's not just a question of \textit{how-to}, you see; it's also a question of \textit{how much to}. Reading will help you answer how much, \& only reams of writing will help you with the how. You can learn only by doing.

Description begins with visualization of what it is you want the reader to experience. It ends with your translating what you see in your mind into words on the page. It's far from easy. As I've said, we've all heard someone say, ``Man, it was so great (or so horrible{\tt/}strange{\tt/}funny) $\ldots$ I just can't describe it!'' if you want to be a successful writer, you \textit{must} be able to describe it, \& in a way that will cause your reader to prickle with recognition. If you can do this, you will be paid for your labors, \& deservedly so. If you can't, you're going to collect a lot of rejection slips \& perhaps explore a career in the fascinating world of telemarketing.

Thin description leaves the reader feeling bewildered \& nearsighted. Overdescription buries him or her in details \& images. The trick is to find a happy medium. It's also important to know \textit{what} to describe \& what can be left alone while you get on with your main job, which is telling a story.

I'm not particularly keen on writing which exhaustively describes the physical characteristics of the people in the story \& what they're wearing (I find wardrobe inventory particularly irritating; if I want to read descriptions of clothes, I can always get a J. Crew catalogue). I can't remember many cases where I felt I had to describe what the people in a story of mine looked like -- I'd rather let the reader supply the faces, the builds, \& the clothing as well. If I tell you that Carrie White is a high school outcast with a bad complexion \& a fashion-victim wardrobe, I think you can do the rest, can't you? I don't need to give you a pimple-by-pimple, skirt-by-skirt rundown. We all remember 1 or more high school losers, after all; if I describe mine, it freezes out yours, \& I lose a little bit of the bond of understanding I want to forge between us. Description begins in the writer's imagination, but should finish in the readers'. When it comes to actually pulling this off, the writer is much more fortunate than the filmmaker, who is almost always doomed to show too much $\ldots$ including, in 9 cases out of 10, the zipper running up the monster's back.

I think locale \& texture are much more important to the reader's sense of actually being \textit{in} the story than any physical description of the players. Nor do I think that physical description should be a shortcut to character. So spare me, if you please, he hero's \textit{sharply intelligent blue eyes} \& \textit{outthrust determined chin}; likewise the heroine's \textit{arrogant cheekbones}. This sort of thing is bad technique \& lazy writing, the equivalent of all those tiresome adverbs.

For me, good description usually consists of a few well-chosen details that will stand for everything else. In most cases, these details will be the 1st ones that come to mind. Certainly they will do for a start. If you decide latter on that you'd like to change, add, or delete, you can do so -- it's what rewrite was invented for. But I think you will find that, in most cases, your 1st visualized details will be the truest \& best. You should remember (\& your reading will prove it over \& over again should you begin to doubt) that it's as easy to overdescribe as to underdescribe. Probably easier.

1 of my favorite restaurants in New York is the steakhouse Palm Too on 2nd Avenue. If I decide to set a scene in Palm Too, I'll certainly be writing about what I know, as I've been there on a number of occasions. Before beginning to write, I'll take a moment to call up an image of the place, drawing from my memory \& filling my mind's eye, an eye whose vision grows sharper the more it is used. I call it a mental eye because that's the phrase with which we're all familiar, but what I actually want to do is open \textit{all} my senses. This memory search will be brief but intense, a kind of hypnotic recall. \&, as with actual hypnosis, you'll find it easier to accomplish the more you attempt it.

The 1st 4 things which come to my mind when I think of Palm Too are: (a) the darkness of the bar \& the contrasting brightness of the backbar mirror, which catches \& reflects light from the street; (b) the sawdust on the floor; (c) the funky cartoon caricatures on the walls; (d) the smells of cooking steak \& fish.

If I think longer I can come up with more stuff (what I don't remember I'll make up -- during the visualization process, fact \& fiction become entwined), but there's no need for more. This isn't the Taj Mahal we're visiting, after all, \& I don't want to sell you the place. It's also important to remember it's not about the setting, anyway -- it's about the story, \& it's \textit{always} about the story. It will not behoove me (or you) to wander off into thickets of description just because it would be easy to do. We have other fish (\& steak) to fry.

Bearing that in mind, here's a sample bit of narration which takes a character into Palm Too:
\begin{quotation}
	The cab pulled up in front of Palm Too at 3:45 on a bright summer afternoon. Billy paid the driver, stepped out onto the sidewalk, \& took a quick look around for Martin. Not in sight. Satisfied, Billy went inside.
	
	After the hot clarity of 2nd Avenue, Palm Too was as dark as a cave. The backbar mirror picked up some of the street-glare \& glimmered in the gloom like a mirage. For a moment it was all Billy could see, \& then his eyes began to adjust. There were a few solitary drinkers at the bar. Beyond them, the ma\^ire d', his tie undone \& his shirt cuffs rolled back to show his hairy wrists, was talking with the bartender. There was still sawdust sprinkled on the floor, Billy noted, as if this were a 20s speakeasy instead of a millennium eatery where you couldn't smoke, let alone spit a job of tobacco between your feet. \& the cartoons dancing across the walls -- gossip-column caricatures of downtown political hustlers, newsmen who had long since retired or drunk themselves to death, celebrities you couldn't quite recognize -- still gamboled all the way to the ceiling. The air was redolent of steak \& fried onions. All of it the same as it ever was.
	
	The ma\^ire d' stepped forward. ``Can I help you, sir? We don't open for dinner until 6, but the bar --''
	
	``I'm looking for Richie Martin,'' Billy said.
\end{quotation}
Billy's arrival in the cab is narration -- action, if you like that word better. What follows after he steps through the door of the restaurant is pretty much straight description. I got in almost all of the details which 1st came to mind when I accessed my memories of the real Palm Too, \& I added a few other things, as well -- the ma\^ire d' between shifts is pretty good, I think; I love the undone tie \& the cuffs rolled up to expose the hairy wrists. It's like a photograph. The smell of fish is the only thing not here, \& that's because the smell of the onions was stronger.

We come back to actual storytelling with a bit of narration (the ma\^ire d' steps forward to center stage) \& then the dialogue. By now we see our location clearly. There are plenty of details I could have added -- the narrowness of the room, Tony Bennett on the sound system, the Yankees bumper-sticker on the cash register -- but what would be the point? When it comes to scene-setting \& all sorts of description, a meal is as good as a feast. We want to know if Billy has located Richie Martin -- that's the story we paid our 24 bucks to read. More about the restaurant would slow the pace of that story, perhaps annoying us enough to break the spell good fiction can weave. In many cases when a reader puts a story aside because it ``got boring,'' the boredom arose because the writer grew enchanted with his powers of description \& lost sight of his priority, which is to keep the ball rolling. If the reader wants to know more about Palm Too than can be found above, he or she can either visit the next time he or she is in New York, or send for a brochure. I've already spilled enough ink here for me to indicate Palm Too will be a major setting for my story. If it turns out not to be, I'd do well to revise the descriptive stuff down by a few lines in the next draft. Certainly I couldn't keep it in on the grounds that it's good; it \textit{should} be good, if I'm being paid to do it. What I'm not being paid to do is be self-indulgent.

There is straight description (``a few solitary drinkers at the bar'') \& a bit of rather more poetic description (``The backbar mirror $\ldots$ glimmered in the gloom like a mirage'') in my central descriptive paragraph about Palm Too. Both are okay, but I like the figurative stuff. The use of simile \& other figurative language is 1 of the chief delights of fiction -- reading it \& writing it, as well. When it's on target, a simile delights us in much the same way meeting an old friend in a crowd of strangers does. By comparing 2 seemingly unrelated objects -- a restaurant bar \& a cave, a mirror \& a mirage -- we are sometimes able to see an old thing in a new \& vivid way.\footnote{Although ``dark as a cave'' isn't all that riveting; certainly we've heard it before. It is, truth to tell, a bit lazy, not quite a clich\'e but certainly in the neighborhood.} Even if the result is mere clarity instead of beauty, I think writer \& reader are participating together in a kind of miracle. Maybe that's drawing it a little strong, but yeah -- it's what I believe.

When a simile or metaphor \textit{doesn't} work, the results are sometimes funny \& sometimes embarrassing. Recently I read this sentence in a forthcoming novel I prefer not to name: ``He sat stolidly beside the corpse, waiting for the medical examiner as patiently as a man waiting for a turkey sandwich.'' If there is a clarifying connection here, I wasn't able to make it. I consequently closed the book without reading further. If a writer knows what he or she is doing, I'll go along for the ride. If he or she doesn't $\ldots$ well, I'm in my 50s now, \& there are a lot of books out there. I don't have time to waste with the poorly written ones.

The Zen simile is only 1 potential pitfall of figurative language. The most common -- \& again, landing in this trap can usually be traced back to not enough reading -- is the use of clich\'e similes, metaphors, \& images. He ran \textit{like a madman}, she was pretty \textit{as a summer day}, the guy was \textit{a hot ticket}, Bob fought \textit{like a tiger} $\ldots$ don't waste my time (or anyone's) with such chestnuts. It makes you look either lazy or ignorant. Neither description will do your reputation as a writer much good.

My all-time favorite similes, by the way, come from the hardboiled-detective fiction of the 40s \& 50s, \& the literary descendants of the dime-dreadful writers. These favorites include ``It was darker than a carload of assholes'' (George V. Higgins) \& ``I lit a cigarette [that] tasted like a plumber's handkerchief'' (Raymond Chandler).

The key to good description begins with clear seeing \& ends with clear writing, the kind of writing that employs fresh images \& simple vocabulary. I began learning my lessons in this regard by reading Chandler, Hammett, \& Ross MacDonald; I gained perhaps even more respect for the power of compact, descriptive language from reading T.S. Eliot (those ragged claws scuttling across the ocean floor; those coffee spoons), \& William Carlos Williams (white chickens, red wheelbarrow, the plums that were in the ice box, so sweet \& so cold).

As with all other aspects of the narrative art, you will improve with practice, but practice will never make you perfect. Why should it? What fun would that be? \& the harder you try to be clear \& simple, the more you will learn about the complexity of our American dialect. It be slippery, precious; aye, it be very slippery, indeed. Practice the art, always reminding yourself that your job is to say what you see, \& then to get on with your story.

\fbox{\bf7} Let us now talk a little bit about dialogue, the audio portion of our programme. It's dialogue that gives your cast their voices, \& is crucial in defining their characters -- only what people do tells us more about what they're like, \& talk is sneaky: what people say often conveys their character to others in ways of which they -- the speakers -- are completely unaware.

You can tell me via straight narration that your main character, Mistuh Butts, never did well in school, never even \textit{went} much to school, but you can convey the same thing, \& much more vividly, by his speech $\ldots$ \& 1 of the cardinal rules of good fiction is never tell us a thing if you can show us, instead:
\begin{quotation}
	``What you reckon?'' the boy asked. He doodled a stick in the dirt without looking up. What he drew could have been a ball, or a planet, or nothing but a circle. ``You reckon the earth goes around the sun like they say?''
	
	``I don't know what they say,'' Mistuh Butts replied. ``I ain't never studied what thisun or thatun says, because eachun says a different thing until your head is finally achin \& you lose your aminite.''
	
	``What's aminite?'' the boy asked.
	
	``You don't never shut up the questions!'' Mistuh Butts cried. He seized the boy's stick \& snapped it. ``Aminite is in your belly when it's time to eat! Less you sick! \& folk says \textit{I'm} ignorant!''
	
	``Oh, \textit{appetite},'' the boy said placidly, \& began drawing again, this time with his finger.
\end{quotation}
Well-crafted dialogue will indicate if a character is smart or dumb (Mistuh Butts isn't necessarily a moron just because he can't say \textit{appetite}; we must listen to him awhile longer because making up our minds on that score), honest or dishonest, amusing or an old sobersides. Good dialogue, such as that written by George V. Higgins, Peter Straub, or Graham Greene, is a delight to read; bad dialogue is deadly.

Writers have different skill levels when it comes to dialogue. Your skills in this area can be improved, but, as a great man once said (actually it was Clint Eastwood), ``A man's got to know his limitations.'' H. P. Lovecraft was a genius when it came to tales of the macabre, but a terrible dialogue writer. He seems to have known it, too, because in the millions of words of fiction he wrote, fewer than \textit{5000} are dialogue. The following passage from ``The Color Out of Space,'' in which a dying farmer describes the alien presence which has invaded his well, showcases Lovecraft's dialogue problems. Folks, people just don't talk like this, even on their deathbeds:
\begin{quotation}
	``Nothin' $\ldots$ nothin' $\ldots$ the color $\ldots$ it burns $\ldots$ cold an' wet $\ldots$ but it burns $\ldots$ it lived in the well $\ldots$ I seen it $\ldots$ a kind o' smoke $\ldots$ jest like the flowers last spring $\ldots$ the well shone at night $\ldots$ everything alive $\ldots$ sucked the life out of everything $\ldots$ in the stone $\ldots$ it must a'come in that stone $\ldots$ pizened the whole place $\ldots$ dun't know what it wants $\ldots$ that round thing the men from the college dug out'n the stone $\ldots$ it was that same color $\ldots$ jest the same, like the flowers an' plants $\ldots$ seeds $\ldots$ I seen it the fust time this week $\ldots$ it beats down your mind an' then gets ye $\ldots$ burns ye up $\ldots$ It come from some place whar things ain't as they is here $\ldots$ 1 o' them professors said so $\ldots$''
\end{quotation}
\& so on \& so forth, in carefully constructed elliptical bursts of information. It's hard to say exactly what's wrong with Lovecraft's dialogue, other than the obvious: it's stilted \& lifeless, brimming with country corn-pone (``some place whar things ain't as they is here''). When dialogue is right, we know. When it's wrong we also know -- it jags on the ear like a badly tuned musical instrument.

Lovecraft was, by all accounts, both snobbish \& painfully shy (a galloping racist as well, his stories full of sinister Africans \& the sort of scheming Jews my Uncle Oren always worried about after 4 or 5 beers), the kind of writer who maintains a voluminous correspondence but gets along poorly with others in person -- were he alive today, he'd likely exist most vibrantly in various Internet chat-rooms. Dialogue is a skill best learned by people who enjoy talking \& listening to others -- particularly listening, picking up the accents, rhythms, dialect, \& slang of various groups. Loners such as Lovecraft often write it badly, or with the care of someone who is composing in a language other than his or her native tongue.

I don't know if contemporary novelist John Katzenbach is a loner or not, but his novel \textit{Hart's War} contains some memorably bad dialogue. Katzenbach is the sort of novelist who drives creative-writing teachers mad, a wonderful storyteller whose art is marred by self-repetition (a fault which is curable) \& an ear for talk that is pure tin (a fault which probably isn't). \textit{Hart's War} is a murder mystery set in a World War II POW camp -- a neat idea, but problematic in Katzenbach's hands once he really gets the plot boiling. Here is Wing Commander Phillip Pryce talking to his friends just before the Germans in charge of Stalag Luft 13 take him away, not to be repatriated as they claim, but probably to be shot in the woods.
\begin{quotation}
	Pryce grabbed at Tommy once again. ``Tommy,'' he whispered, ``this is not a coincidence! Nothing is what it seems! Dig deeper! Save him, lad, save him! For more than ever, now, I believe Scott is innocent! $\ldots$ You're on your own now, boys. \& remember, I'm counting on you to live through this! Survive! Whatever happens!''
	
	He turned back to the Germans. ``All right, \textit{Hauptmann},'' he said with a sudden, exceedingly calm determination. ``I'm ready now. Do with me what you will.''
\end{quotation}
Either Katzenbach does not realize what every line of the Wing Commander's dialogue is a clich\'e from a late=40s war movie or he's trying to use that similarity deliberately to awaken feelings of pity, sadness, \& perhaps nostalgia in his audience. Either way, it doesn't work. The only feeling the passage evokes is a kind of impatient incredulity. You wonder if any editor ever saw it, \& if so, what stayed his or he blue pencil. Given Katzenbach's considerable talents in other areas, his failure here tends to reinforce my idea that writing good dialogue is art as well as craft.

Many good dialogue writers simply seem to have been born with a well-tuned ear, just as some musicians \& singers have perfect or near-perfect pitch. Here's passage from Elmore Leonard's novel \textit{Be Cool}. You might compare it to the Lovecraft \& Katzenbach passages above, noting 1st of all that here we've got an honest-to-God exchange going on, \& not a stilted soliloquy:
\begin{quotation}
	Chili $\ldots$ looked up again as Tommy said, ``You doing okay?''
	
	``You want to know if I'm making out?''
	
	``I mean in your business. How's it going? I know you did okay with \textit{Get Leo}, a terrific picture, terrific. \& you know what else? It was good. But the sequel -- what was it called?''
	
	``\textit{Get Lost}.''
	
	``Yeah, well that's what happened before I got a chance to see it, it disappeared.''
	
	``It didn't open big so the studio walked away. I was against doing a sequel to begin with. But the guy running production at Tower says they're making the picture, with me or without me. I thought, well, if I can come up with a good story $\ldots$''
\end{quotation}
2 guys at lunch in Beverly Hills, \& right away we know they're both players. They may be phonies (\& maybe they're not), but they're an instant buy within the context of Leonard's story; in fact, we welcome them with open arms. Their talk is so real that part of what we feel is the guilty pleasure of anyone 1st tuning in \& then eavesdropping on an interesting conversation. We're  getting a sense of character, as well, although only in faint strokes. This is early on in the novel (page 2, actually), \& Leonard is an old pro. He knows he doesn't have to do it all at once. Still, don't we learn something about Tommy's character when he assures Chili that \textit{Geo Leo} is not only terrific, but also good?

We could ask ourselves if such dialogue is true to life or only to a certain \textit{idea} of life, a certain stereotyped image of Hollywood players, Hollywood lunches, Hollywood deals. This is a fair enough question, \& the answer is, perhaps not. Yet the dialogue \textit{does} ring true to our ear; at his best (\& although \textit{Be Cool} is quite entertaining, it is far from Leonard's best), Elmore Leonard is capable of a kind of street poetry. The skill necessary to write such dialogue comes from years of practice; the art comes from a creative imagination which is working hard \& having fun.

As with all other aspects of fiction, the key to writing good dialogue is honesty. \& if you \textit{are} honest about the words coming out of your characters' mouths, you'll find that you've let yourself in for a fair amount of criticism. Not a week goes by that I don't receive at least 1 pissed-off letter (most weeks there are more) accusing me of being foulmouthed, bigoted, homophobic, murderous, frivolous, or downright psychopathic. In the majority of cases what my correspondents are hot under the collar about relates to something in the dialogue: ``Let's get the fuck out of Dodge'' or ``We don't cotton much to niggers around here'' or ``What do you think you're doing, you fucking faggot?''

My mother, God rest her, didn't approve of profanity or any such talk; she called it ``the language of the ignorant.'' This did not, however, keep her from yelling ``Oh shit!'' if she burned the roast or nailed her thumb a good one while hammering a picture-hook in the wall. Nor does it prelude most people, Christian as well as heathen, from saying something similar (or even stronger) when the dog barfs on the shag carpet or the car slips off the jack. It's important to tell the truth; so much depends upon it, as William Carlos Williams almost said when he was writing about that red wheelbarrow. The Legion of Decency might not like the word \textit{shit}, \& you might not like it much, either, but sometimes you're just stuck with it -- no kid ever ran to his mother \& said that his little sister just \textit{defecated} in the tub. I suppose he might say \textit{pushed} or \textit{went woowoo}, but \textit{took a shit} is, I fear, very much the ballpark (little pitchers have big ears, after all).

You \textit{must} tell the truth if your dialogue is to have to resonance \& realism that \textit{Hart's War}, good story though it is, so sadly lacks -- \& that holds true all the way down to what folks say when they hit their thumb with the hammer. If you substitute ``Oh sugar!'' for ``Oh shit!'' because you're thinking about the Legion of Decency, you are breaking the unspoken contract that exists between the writer \& reader -- your promise to express the truth of how people act \& talk through the medium of a made-up story.

On the other hand, 1 of your characters (e.g., the protagonist's old maid aunt) really \textit{might} say \textit{Oh sugar} instead of \textit{Oh shit} after pounding her thumb with the hammer. You'll know which to use if you know your character, \& we'll learn something about the speaker that will make him or her move vivid \& interesting. The point is to let each character speak freely, without regard to what the Legion of Decency or the Christian Ladies' Reading Circle may approve of. To do otherwise would be cowardly as well as dishonest, \& believe me, writing fiction in America as we enter the 21st century is no job for intellectual cowards. There are lots of would-be censors out there, \& although they may have different agendas, they all want basically the same thing: for you to see the world they see $\ldots$ or to at least shut up about what you \textit{do} see that's different. They are agents of the status quo. Not necessarily bad guys, but dangerous guys if you happen to believe in intellectual freedom.

As it happens, I agree with my mother: profanity \& vulgarity \textit{is} the language of the ignorant \& the verbally challenged. \textit{Mostly}, that is; there are exceptions, including profane aphorisms of great color \& vitality. \textit{They always fuck you at the drive-thru; I'm busier than a 1-legged man in an ass-kicking contest; wish in 1 hand, shit in the other, see which one fills up 1st} -- these phrases \& others like them aren't for the drawing-room, but they \textit{are} striking \& pungent. Or consider this passage from \textit{Brain Storm}, by Richard Dooling, where vulgarity becomes poetry:
\begin{quotation}
	``Exhibit A: One loutish, headstrong penis, a barbarous cuntivore without a flyspeck of decency in him. The capscallion of all rapscallions. A scurvy, vermiform scug with a serpentine twinkle in his solitary eye. An orgulous Turk who strikes in the dark vaults of flesh like a penile thunderbolt. A greedy cur seeking shadows, slick crevices, tuna fish ecstasy, \& sleep $\ldots$''
\end{quotation}
Although not offered as dialogue, I want to reproduce another passage from Dooling here, because it speaks to the converse: that one can be quite admirably graphic resorting to vulgarity or profanity at all:
\begin{quotation}
	She straddled him \& prepared to make the necessary port connections, male \& female adapters ready, I{\tt/}O enabled, server{\tt/}client, master{\tt/}slave. Just a couple of high-end biological machines preparing to hot-dock with cable modems \& access each other's front-end processors.
\end{quotation}
If I were a Henry James or Jane Austen sort of guy, writing only about toffs or smart college folks, I'd hardly ever have to use a dirty word or a profane phrase; I might ever have had a book banned from America's school libraries or gotten a letter from some helpful fundamentalist fellow who wants me to know that I'm going to burn in hell, where all my millions of dollars won't buy me so much as a single drink of water. I did not, however, group up among folks of that sort. I grew up as a part of America's lower middle class, \& they're the people I can write about with the most honesty \& knowledge. It means that they say shit more often than sugar when they bang their thumbs, but I've made my peace with that. Was never much at war with it in the 1st place, as a matter of fact.

When I get 1 of Those Letters, or face another review that accuses me of being a vulgar lowbrow -- which to some extent I am -- I take comfort from the words of turn-of-the-century social realist Frank Norris, whose novels include \textit{The Octopus, The Pit}, \& \textit{McTeague}, an authentically great book. Norris wrote about working-class guys on ranches, in city laboring jobs, in factories. McTeague, the main character of Norris's finest work, is an unschooled dentist. Norris's books provoked a good deal of public outrage, to which Norris responded coolly \& disdainfully: ``What do I care for their opinions? I never truckled. I told them the truth.''

Some people don't want to hear the truth, of course, but that's not your problem. What would be is wanting to be a writer without wanting to shoot straight. Talk, whether ugly or beautiful, is an index of character; it can also be a breath of cool, refreshing air in a room some people would prefer to keep shut up. In the end, the important question has nothing to do with whether the talk in your story is sacred or profane; the only question is how it rings on the page \& in the ear. If you expect it to ring true, then you must talk yourself. Even more important, you must shut up \& listen to others talk.

\fbox{\bf8} Everything I've said about dialogue applies to building characters in fiction. The job boils down to 2 things: paying attention to how the real people around you behave \& then telling the truth about what you see. You may notice that your next-door neighbor picks his nose when he thinks no one is looking. This is a great detail, but noting it does you no good as a writer unless you're willing to dump it into a story at some point.

Are fictional characters drawn directly from life? Obviously not, at least on a 1-to-1 basis -- you'd \textit{better} not, unless you want to get sued or shot on your way to the mailbox some fine morning. In many cases, such as \textit{roman \`a clef} novels like \textit{Valley of the Dolls}, characters are drawn \textit{mostly} from life, but after readers get done playing the inevitable guessing game about who's who, these stories tend to be unsatisfying, stuffed with shadowbox celebrities who bonk each other \& then fade quickly from the reader's mind. I read \textit{Valley of the Dolls} shortly after it came out (I was a cook's boy at a western Maine resort that summer), gobbling it up as eagerly as everyone else who bought it, I suppose, but I can't remember much of what it was about. On the whole, I think I prefer the weekly codswallop served up the \textit{The National Enquirer}, where I can get recipes \& cheesecake photographs as well as scandal.

For me, what happens to characters as a story progresses depends solely on what I discover about them as I go along -- how they grow, in other words. Sometimes they grow a little. If they grow a lot, they begin to influence the course of the story instead of the other way around. I almost always start with something that's situational. I don't say that's right, only that it's the way I've always worked. If a story ends up that same way, however, I count it something of a failure no matter how interesting it may be to me or to others. I think the best stories always end up being about the people rather than the event, which is to say character-driven. Once you get beyond the short story, though (2000--4000 words, let's say), I'm not much of a believer in the so-called character study; I think that in the end, the story should always be the boss. Hey, if you want a character study, buy a biography or get season tickets to your local college's theater-lab productions. You'll get all the character you can stand.

It's also important to remember that no one is ``the bad guy'' or ``the best friend'' or ``the whore with a heart of gold'' in real life; in real life we each of us regard ourselves as the main character, the protagonist, the big cheese; the camera is on \textit{us}, baby. If you can bring this attitude into your fiction, you may not find it easier to create \textit{brilliant} characters, but it will be harder for you to create the sort of 1D dopes that populate so much pop fiction.

Annie Wilkes, the nurse who holds Paul Sheldon prisoner in \textit{Misery}, may seem psychopathic to us, but it's important to remember that she seems perfectly sane \& reasonable to herself -- heroic, in fact, a beleaguered woman trying to survive in a hostile world filled with cockadoodie brats. We see her go through dangerous mood-swings, but I tried never to come right out \& say ``Annie was depressed \& possibly suicidal that day'' or ``Annie seemed particularly happy that day.'' If I have to tell you, I lose. If, on the other hand, I can show you a silent, dirty-haired woman who compulsively gobbles cake \& candy, then have you draw the conclusion that Annie is in the depressive part of a manic-depressive cycle, I win. \& if I am able, even briefly, to give you a Wilkes'-eye-view of the world -- if I can make you understand her madness -- then perhaps I can make her someone you sympathize with or even identify with. The result? She's more frightening than ever, because she's close to real. If, on the other hand, I turn her into a cackling old crone, she's just another pop-up bogeylady. In that case I lose bigtime, \& so does the reader. Who would want to visit with such a stale shrew? That version of Annie was old when \textit{The Wizard of Oz} was in its 1st run.

It would be fair enough to ask, I suppose, if Paul Sheldon in \textit{Misery} is me. Certainly \textit{parts} of him are $\ldots$ but I think you will find that, if you continue to write fiction, every character you create is partly you. When you ask yourself what a certain character will do given a certain set of circumstances, you're making the decision based on what you yourself would (or, in the case of a bad guy, wouldn't) do. Added to these versions of yourself are the character traits, both lovely \& unlovely, which you observe in others (e.g., a guy who picks his nose when he thinks no one is looking). There is also a wonderful 3rd element: pure blue-sky imagination. This is the part which allowed me to be a psychotic nurse for a little while when I was writing \textit{Misery}. \& being Annie was not, by \& large, hard at all. In fact, it was sort of fun. I think being Paul was harder. He was sane, I'm sane, no 4 days at Disneyland there.

My novel \textit{The Dead Zone} arose from 2 questions: Can a political assassin ever be right? \& if he is, could you make him the protagonist of a novel? The good guy? These ideas called for a dangerously unstable politician, it seemed to me -- a fellow who could climb the political ladder by showing the world a jolly, jes'-folks face \& charming the voters by refusing to play the game in the usual way. (Greg Stillson's campaign tactics as I imagined them 20 years ago were very similar to the ones Jesse Ventura used in his successful campaign for the governor's seat in Minnesota. Thank goodness Ventura doesn't seem like Stillson in any other ways.)

\textit{The Dead Zone}'s protagonist, Johnny Smith, is also an everyday, jes'-folks sort of guy, only with Johnny it's no act. The 1 thing that sets him apart is a limited ability to see the future, gained as the result of a childhood accident. When Johnny shakes Greg Stillson's hand at a political rally, he has a vision of Stillson becoming the President of the United States \& subsequently starting World War III. Johnny comes to the conclusion that the only way he can keep this from happening -- the only way he can save the world, in other words -- is by putting a bullet in Stillson's head. Johnny is different from other violent, paranoid mystics in only 1 way: he really \textit{can} see the future. Only don't they all say that?

The situation had an edgy, outlaw feel to it that appealed to me. I thought the story would work if I could make Johnny a genuinely decent guy without turning him into a plaster saint. Same thing with Stillson, only backwards: I wanted him to be authentically nasty \& really scare the reader, not just because Stillson is always boiling with potential violence but because he is so goddam \textit{persuasive}. I wanted the reader to constantly be thinking: ``This guy is out of control -- how come somebody can't see through him?'' The fact that Johnny \textit{does} see through him would, I thought, put the reader even more firmly in Johnny's corner.

When we 1st meet the potential assassin, he's taking his girl to the county fair, riding the rides \& playing the games. What could be more normal or likable? The fact that he's on the verge of proposing to Sarah makes us like him even more. Later, when Sarah suggests they cap a perfect date by sleeping together for the 1st time, Johnny tells her he wants to wait until they're married. I felt I was walking a fine line on that one -- I wanted readers to see Johnny as sincere \& sincerely in love, a straight shooter but not a tight-assed prude. I was able to cut his principled behavior a bit by giving him a childish sense of humor; he greets Sarah wearing a glow-in-the-dark Halloween mask (the mask hopefully works in a symbolic way, too; certainly Johnny is perceived as a monster when he points a gun at candidate Stillson). ``Same old Johnny,'' Sarah says, laughing, \& by the time the 2 of them are headed back from the fair in Johnny's old Volkswagen Bug, I think Johnny Smith has become our friend, just an average American guy who's hoping to live happily ever after. The sort of guy who'd return your wallet with the money still in it if he found it on the street or stop \& help you change your flat tire if he came upon you broke down by the side of the road. Ever since John. F. Kennedy was shot in Dallas, the great American bogeyman has been the guy with the rifle in a high place. I wanted to make this guy into the reader's friend.

Johnny was hard. Taking an average guy \& making him vivid \& interesting always is. Greg Stillson (like most villains) was easier \& a lot more fun. I wanted to nail his dangerous, divided character in the 1st scene of the book. Here, several years before he runs for the U.S. House of Representatives in New Hampshire, Stillson is a young traveling salesman hawking Bibles to midwest country folk. When he stops at 1 farm, he is menaced by a snarling dog. Stillson remains friendly \& smiling -- Mr. Jes' Folks -- until he's positive no one's home at the farm. Then he sprays teargas into the dog's eyes \& kicks it to death.

If one is to measure success by reader response, the opening scene of \textit{The Dead Zone} (my 1st number-1 hardcover bestseller) was 1 of my most successful ever. Certainly it struck a raw nerve; I was deluged with letters, most of them protesting my outrageous cruelty to animals. I wrote back to these folks, pointing out the usual things: (a) Greg Stillson wasn't real; (b) the \textit{dog} wasn't real; (c) I myself had never in my life put the boot to 1 of my pets, or anyone else's. I also pointed out what might have been a little less obvious -- it was important to establish, right up front, that Gregory Ammas Stillson was an extremely dangerous man, \& very good at camouflage.

I continued to build the characters of Johnny \& Greg in alternating scenes until the confrontation at the end of the book, when things resolve themselves in what I hoped would be an unexpected way. The characters of my protagonist \& antagonist were determined by the story I had to tell -- by the fossil, the found object, in other words. My job (\& yours, if you decide this is a viable approach to storytelling) is to make sure these fictional folks behave in ways that will both help the story \& seem reasonable to us, given what we know about them (\& what we know about real life, of course). Sometimes villains feel self-doubt (as Greg Stillson does); sometimes they feel pity (as Annie Wilkes does). \& sometimes the good guy tries to turn away from doing the right thing, as Johnny Smith does $\ldots$ as Jesus Christ himself did, if you think about that prayer (``take this cup from my lips'') in the Garden of Gethsemane. \& if you do your job, your characters will come to life \& start doing stuff on their own. I know that sounds a little creepy if you haven't actually experienced it, but it's terrific fun when it happens. \& it will solve a lot of your problems, believe me.

'' -- \cite[pp. 108--107]{King2010}

%------------------------------------------------------------------------------%

\section{ON LIVING: A POSTSCRIPT}

%------------------------------------------------------------------------------%

\section{\& Furthermore, Part I: Door Shut, Door Open}

%------------------------------------------------------------------------------%

\section{\& Furthermore, Part II: A Booklist}

%------------------------------------------------------------------------------%

\section{Further to Furthermore, Part III}

%------------------------------------------------------------------------------%

\section{About Stephen King}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
	
\end{document}