\documentclass{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/hong/1_NQBH/reference/bib.bib}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\small \textsc{Sect.} ~\thesection}
\rhead{\small \nouppercase{\leftmark}}
\renewcommand{\sectionmark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Giver vs. Taker vs. Matcher}
\author{Nguyen Quan Ba Hong\footnote{Independent Researcher, Ben Tre City, Vietnam\\e-mail: \texttt{nguyenquanbahong@gmail.com}}}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
	This article is a subsection in the section \textit{Psychology} in \href{https://github.com/NQBH/hobby/blob/master/series_a_personal_journey_to_realms_of_knowledge/NQBH_series_a_personal_journey_to_realms_of_knowledge.pdf}{Series: A Personal Journey to Realms of Knowledge: The Promised Neverland}, i.e., in the manner of Linux directory format: Series: A Personal Journey to Realms of Knowledge: The Promised Neverland\texttt{/}psychology\texttt{/}giver vs. taker vs. matcher.
\end{abstract}

\textbf{Keywords.}

\textbf{MSC.}
\tableofcontents

%------------------------------------------------------------------------------%

\section{Introduction}
\textsc{Adam Grant.} \href{https://www.youtube.com/watch?v=YyXRYgjQXX0}{YouTube\texttt{/}Are you a giver or a taker?} \href{https://www.ted.com/talks/adam_grant_are_you_a_giver_or_a_taker}{TED\texttt{/}Are you a giver or a taker?}

\section{Future Works}

\section*{Acknowledgment}

\appendix

\section{Supplementary Material}

\printbibliography[heading=bibintoc]
	
\end{document}