\documentclass{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/hong/1_NQBH/reference/bib.bib}
\usepackage[english,vietnamese]{babel}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools,multicol}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\small \textsc{Sect.} ~\thesection}
\rhead{\small \nouppercase{\leftmark}}
\renewcommand{\sectionmark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Elementary Mathematics\texttt{/}Student}
\author{\selectlanguage{vietnamese} Nguyễn Quản Bá Hồng\footnote{Independent Researcher, Ben Tre City, Vietnam\\e-mail: \texttt{nguyenquanbahong@gmail.com}; website: \url{https://nqbh.github.io}.}}
\date{\today}

\begin{document}
\maketitle
\selectlanguage{english}
\begin{abstract}
	
\end{abstract}

\tableofcontents
\selectlanguage{vietnamese}

%------------------------------------------------------------------------------%

\section{Elementary Mathematics}

\subsection{Grade 6}
Jul 25, 27, 29; Aug 1, 3, 5, 8, 10, 12, 2022.
\begin{enumerate}
	\item \textsc{Nguyễn Thị Kim Anh.}
	\item \textsc{Trần Gia Bảo.}
	\item \textsc{Lê Duyên Minh Châu.}
	\item \textsc{Nguyễn Tâm Đắc.}
	\item \textsc{Nguyễn Hữu Hiếu.}
	\item \textsc{Nguyễn Hoàng Bảo Lam.}
	\item \textsc{Ngô Nguyễn Phương Nghi.}
	\item \textsc{Nguyễn Như Ngọc.}
	\begin{itemize}
		\item \textbf{Paid tutor fee.} Aug, 2022.
	\end{itemize}
	\item \textsc{Nguyễn Bình Nguyên.}
	\item \textsc{Lê Trọng Nhân.}
	\item \textsc{Trần Ngọc Nhi.}
	\item \textsc{Nguyễn Ngọc Thu Nhiên.}
	\item \textsc{Võ Lâm Quỳnh Như.}
	\item \textsc{Đào Minh Hồng Phát.}
	\item \textsc{Nguyễn Hữu Phát.}
	\item \textsc{Phan Tấn Quân.}
	\item \textsc{Trương Khánh Quỳnh.} Aug 3, 2022.
	\item \textsc{Nguyễn Thị Bảo Thi.}
	\item \textsc{Phan Nguyễn Anh Thơ.} Aug 5, 2022.
	\item \textsc{Đào Ngọc Quế Trân.}
	\item \textsc{Nguyễn Ngọc Thùy Trâm.}
	\item \textsc{Lê Phương Trinh.}
\end{enumerate}

\subsection{Grade 7}
\begin{enumerate}
	\item \textsc{Đàm Trung Tín.} Aug 2, 4, 6, 9, 11, 13, 2022.
\end{enumerate}

\subsection{Grade 8}
\begin{enumerate}
	\item \textsc{Trương Hồ Yến Quỳnh.} Aug 5, 2022.
\end{enumerate}

\subsection{Grade 9}

\subsection{Grade 10}

\subsection{Grade 11}
\begin{enumerate}
	\item \textsc{Nguyễn Thị Ngọc Nhi.} Aug 4, 6, 9, 11, 13, 2022.
\end{enumerate}

\subsection{Grade 12}

\section{Elementary Physics}

\subsection{Grade 11}
\begin{enumerate}
	\item \textsc{Nguyễn Thị Ngọc Nhi.} Aug 4, 6, 9, 11, 13, 2022.
\end{enumerate}

\section{Programming}

\subsection{Grade 11}
\begin{enumerate}
	\item \textsc{Nguyễn Thị Ngọc Nhi.} Aug 4, 6, 9, 11, 13, 2022.
\end{enumerate}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
	
\end{document}