\documentclass{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/hong/1_NQBH/reference/bib.bib}
\usepackage[utf8]{vietnam}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm,mathtools,float,graphicx,algpseudocode,algorithm,tcolorbox}
\usepackage[inline]{enumitem}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{hequa}{Hệ quả}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{dinhnghia}{Định nghĩa}[section]
\newtheorem{example}{Example}[section]
\newtheorem{vidu}{Ví dụ}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{baitoan}{Bài toán}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{cauhoi}{Câu hỏi}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{luuy}{Lưu ý}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{dinhly}{Định lý}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\small Sect.~\thesection}
\rhead{\small\nouppercase{\leftmark}}
\renewcommand{\sectionmark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Problems in English}
\author{Nguyễn Quản Bá Hồng\footnote{Independent Researcher, Ben Tre City, Vietnam\\e-mail: \texttt{nguyenquanbahong@gmail.com}; website: \url{https://nqbh.github.io}.}}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
	\textsc{[en]} This text is a collection of problems, from easy to advanced, for English. The latest version of this text has been stored \& downloadable at the following link: \href{https://github.com/NQBH/hobby/blob/master/language/english/problem/NQBH_english_problem.pdf}{GitHub\texttt{/}NQBH\texttt{/}hobby\texttt{/}language\texttt{/}english\texttt{/}problem}\footnote{\textsc{url}: \url{https://github.com/NQBH/hobby/blob/master/language/english/problem/NQBH_english_problem.pdf}.}.
	\vspace{2mm}
	
	\textsc{[vi]} Tài liệu này là 1 bộ sưu tập các bài tập Tiếng Anh chọn lọc từ cơ bản đến nâng cao. Phiên bản mới nhất của tài liệu này được lưu trữ ở link sau: \href{https://github.com/NQBH/hobby/blob/master/language/english/problem/NQBH_english_problem.pdf}{GitHub\texttt{/}NQBH\texttt{/}hobby\texttt{/}language\texttt{/}english\texttt{/}problem}.
\end{abstract}
\tableofcontents
\newpage

%------------------------------------------------------------------------------%

\section{Multiple Choice Tests}
Choose the best alternative to fill the gap in each of the following sentences.
\begin{enumerate}[leftmargin=8mm]
	\item I find the way she keeps playing with her hair really $\ldots$. {\sf A.} thrilled. {\sf B.} irritated. {\sf C.} thrilling. {\sf D.} irritating.
	\item She was $\ldots$ to discover that she had won 1st prize. {\sf A.} excited. {\sf B.} lucky. {\sf C.} astonished. {\sf D.} nervous.
	\item You must have left $\ldots$ when all your exams were over. {\sf A.} relieved. {\sf B.} upset. {\sf C.} irritated. {\sf D.} cross.
	\item \textit{The Silence of the Lambs} is the most $\ldots$ film I've ever seen. {\sf A.} terrified. {\sf B.} terrifying. {\sf C.} scared. {\sf D.} excited.
	\item I was $\ldots$ that Tom could come to the party after all. {\sf A.} irritating. {\sf B.} nervous. {\sf C.} furious. {\sf D.} glad.
	\item The manager really wanted Anna to join the company: he believed she would be a valuable $\ldots$ to his team of sales people. {\sf A.} addition. {\sf B.} accumulation. {\sf C.} supplement. {\sf D.} appendage.
	\item Everything included, the cost of our new living room came to a total of \pounds10,000. {\sf A.} full. {\sf B.} great. {\sf C.} gross. {\sf D.} grand.
	\item When the ship docked at Hamburg, they found a $\ldots$ in the hold. {\sf A.} gate-crasher. {\sf B.} stowaway. {\sf C.} interloper. {\sf D.} trespasser.
	\item If you know when you're coming back it would be better to buy a $\ldots$ ticket. {\sf A.} 1st-class. {\sf B.} single. {\sf C.} return. {\sf D.} 1-way.
	\item Cheese \& spinach pie is a traditional Greek $\ldots$. {\sf A.} cuisine. {\sf B.} food. {\sf C.} course. {\sf D.} dish.
	\item How long does it usually take to $\ldots$ for a domestic flight? {\sf A.} check in. {\sf B.} check out. {\sf C.} check over. {\sf D.} check on.
	\item The brochure stated the hotel was situated $\ldots$ beside the sea. {\sf A.} direct. {\sf B.} right. {\sf C.} precise. {\sf D.} exact.
	\item Sheila had butterflies in her $\ldots$ before the interview. {\sf A.} face. {\sf B.} heart. {\sf C.} palms. {\sf D.} stomach.
	\item I hadn't been riding before \& could barely $\ldots$ the horse at 1st. {\sf A.} get into. {\sf B.} get out of. {\sf C.} get on. {\sf D.} get off.
	\item I misread the time on the ticket \& we $\ldots$ the plane. {\sf A.} caught. {\sf B.} got on. {\sf C.} lost. {\sf D.} missed.
	\item Although Vicky looked pretty much the same after all those years, I noticed $\ldots$ changes which made her look even more beautiful than I remembered. {\sf A.} subtle. {\sf B.} sensitive. {\sf C.} fair. {\sf D.} joint.
	\item If they had been wearing $\ldots$ they might have survived when their plane crashed into the sea. {\sf A.} cabins. {\sf B.} seat-belts. {\sf C.} life-belts. {\sf D.} yachts.
	\item If you have a minor illness, it's usually better just to let nature take its $\ldots$. {\sf A.} course. {\sf B.} time. {\sf C.} path. {\sf D.} way.
	\item The scientist wanted to $\ldots$ himself from the controversy following the failure of the experiment. {\sf A.} drive. {\sf B.} space. {\sf C.} distance. {\sf D.} lengthen.
	\item They're starting their $\ldots$ by flying to Tokyo. {\sf A.} travel. {\sf B.} voyage. {\sf C.} journey. {\sf D.} trip.
	\item The shop assistant was most $\ldots$ when I told her that there was no price on the jar of olives. {\sf A.} helpless. {\sf B.} unhelpful. {\sf C.} thoughtful. {\sf D.} unhelpfully.
	\item He may not be good looking but he's 1 of the most $\ldots$ people I've ever met. {\sf A.} likely. {\sf B.} unlikely. {\sf C.} likable. {\sf D.} likeness.
	\item I don't know why exactly but I suspect him of being $\ldots$. {\sf A.} dishonesty. {\sf B.} dishonest. {\sf C.} dishonestly. {\sf D.} honesty.
	\item He hired a $\ldots$ to try to follow her everywhere she went. {\sf A.} private investigator. {\sf B.} caretaker. {\sf C.} lifeguard. {\sf D.} traffic warden.
	\item He runs a very $\ldots$ business selling spare parts for cars on the Internet. {\sf A.} profiting. {\sf B.} profitable. {\sf C.} profit. {\sf D.} profitably.
	\item The $\ldots$ in our building often falls asleep at the front desk. {\sf A.} caretaker. {\sf B.} stockbroker. {\sf C.} undertaker. {\sf D.} bookmaker.
	\item He took $\ldots$ with many of the points she made, claiming they were incorrect. {\sf A.} exception. {\sf B.} issue. {\sf C.} account. {\sf D.} trouble.
	\item The sink is completely blocked so we'll have to get a(n) $\ldots$. {\sf A.} miner. {\sf B.} undertaker. {\sf C.} surgeon. {\sf D.} plumber.
	\item The organizers $\ldots$ how many people would be attending the course \& there were not enough chairs. a. underestimated. {\sf B.} overestimated. {\sf C.} overrated. {\sf D.} estimated.
	\item $\ldots$ a minute! I can't find my keys. {\sf A.} Keep on. {\sf B.} Hold on. {\sf C.} Go on. {\sf D.} Carry on.
	\item You just never lift a $\ldots$ to help me -- I have to do everything myself! {\sf A.} shoulder. {\sf B.} hand. {\sf C.} finger. {\sf D.} wrist.
	\item Scientist seem to be trying to push the $\ldots$ of nature to the limit. {\sf A.} laws. {\sf B.} codes. {\sf C.} rules. {\sf D.} regulations.
	\item There was a $\ldots$ on the door who wouldn't let us into the disco. {\sf A.} referee. {\sf B.} lifeguard. {\sf C.} wrestler. {\sf D.} bouncer.
	\item I won't act in films that are too violent -- I $\ldots$ the line at that. {\sf A.} paint. {\sf B.} make. {\sf C.} draw. {\sf D.} write.
	\item I had lost my glasses in the water \& could barely $\ldots$ the edge of the river. {\sf A.} notice. {\sf B.} recognize. {\sf C.} remark. {\sf D.} distinguish.
	\item If you $\ldots$ to Berlin, call my friend Peter \& say ``hello'' from me. {\sf A.} will go. {\sf B.} won't go. {\sf C.} go. {\sf D.} going.
	\item She gets 15 per cent $\ldots$ on every insurance policy she sells. {\sf A.} salary. {\sf B.} commission. {\sf C.} bonus. {\sf D.} pension.
	\item My uncle is thinking of $\ldots$ next year \& spending the rest of his life traveling round the world. {\sf A.} retiring. {\sf B.} being made redundant. {\sf C.} being sacked. {\sf D.} being resigned.
	\item Before the opened the new factory a lot of the young people round here were $\ldots$ the dole. {\sf A.} in. {\sf B.} up. {\sf C.} at. {\sf D.} on.
	\item She has $\ldots$ a position as general manager. {\sf A.} applied. {\sf B.} applied for. {\sf C.} solicited. {\sf D.} asked.
	\item I worked all summer but I didn't $\ldots$ enough to save any money. {\sf A.} gain. {\sf B.} win. {\sf C.} pay. {\sf D.} earn.
	\item In the interview they asked if I had much $\ldots$ with computers. {\sf A.} experience. {\sf B.} experiment. {\sf C.} experiments. {\sf D.} experiences.
	\item Many governments have tried to ban $\ldots$. {\sf A.} syndicates. {\sf B.} employers. {\sf C.} trade unions. {\sf D.} employees.
	\item There is a special section of the newspaper entirely devoted to job $\ldots$. {\sf A.} advertisements. {\sf B.} announces. {\sf C.} propaganda. {\sf D.} publicity.
	\item I look forward to $\ldots$ from you \& receiving your brochure. {\sf A.} heard. {\sf B.} hearing. {\sf C.} hear. {\sf D.} hears.
	\item Are there any squash $\ldots$ in the area? {\sf A.} greens. {\sf B.} pitches. {\sf C.} rings. {\sf D.} courts.
	\item People sometimes use baseball $\ldots$ as weapons. {\sf A.} clubs. {\sf B.} sticks. {\sf C.} bats. {\sf D.} rackets.
	\item It doesn't matter which football team you support the $\ldots$ is always in the wrong. {\sf A.} linesman. {\sf B.} referee. {\sf C.} arbitrator. {\sf D.} umpire.
	\item I think the $\ldots$ is a bit too high, don't you? {\sf A.} net. {\sf B.} ring. {\sf C.} pitch. {\sf D.} set.
	\item $\ldots$, I want to explain why the team has not managed to get into the 1st division. {\sf A.} At the end. {\sf B.} In the end. {\sf C.} Finally. {\sf D.} Last.
	\item I'd like to buy the radio, but I haven't got any money on me at the moment. Could you $\ldots$ for me for a day or two? {\sf A.} bring it around. {\sf B.} lay it in. {\sf C.} take it in. {\sf D.} put it on 1 side.
	\item ``Were you told to get dinner ready?'' ``No, I did it of my own $\ldots$'' {\sf A.} desire. {\sf B.} accord. {\sf C.} idea. {\sf D.} will.
	\item Mr. Newrich invited his guests to an expensive restaurant. However, he felt $\ldots$ at ease. {\sf A.} sick. {\sf B.} painful. {\sf C.} ill. {\sf D.} unwell.
	\item Police blamed a small hooligan $\ldots$ in the crowd for the violence which occurred. {\sf A.} constituent. {\sf B.} element. {\sf C.} division. {\sf D.} portion.
	\item The police arrested the wrong man mainly because they $\ldots$ the names they had been given by the witness. {\sf A.} bewildered. {\sf B.} merged. {\sf C.} confused. {\sf D.} puzzled.
	\item The police car raced down the street with the $\ldots$ blaring. {\sf A.} siren. {\sf B.} bell. {\sf C.} gong. {\sf D.} alarm.
	\item His parents died when he was very young so he was $\ldots$ by an aunt. {\sf A.} grown up. {\sf B.} taken care. {\sf C.} taken after. {\sf D.} brought up.
	\item I have been eating honey so my fingers are $\ldots$. {\sf A.} dirty. {\sf B.} sticky. {\sf C.} wet. {\sf D.} yellow.
	\item Mrs. Proper always tells her child not to talk with his mouth $\ldots$. {\sf A.} wide. {\sf B.} open. {\sf C.} together. {\sf D.} full.
	\item I have always $\ldots$ my older brother for his courage \& honesty. {\sf A.} taken up. {\sf B.} taken after. {\sf C.} looked up to. {\sf D.} looked after.
	\item After the guests had left we were allowed to eat the $\ldots$ cakes. {\sf A.} remaining. {\sf B.} left. {\sf C.} missed. {\sf D.} additional.
	\item Don't be so $\ldots$! He was only joking. {\sf A.} sensible. {\sf B.} senseless. {\sf C.} sensitive. {\sf D.} insensitive.
	\item The conspirators were plotting the $\ldots$ of the government. {\sf A.} overthrow. {\sf B.} disaster. {\sf C.} demolition. {\sf D.} catastrophe.
	\item When I entered the room, the children were $\ldots$ over the last orange. {\sf A.} consulting. {\sf B.} discussing. {\sf C.} squabbling. {\sf D.} participating.
	\item My sister had just $\ldots$ 16. {\sf A.} completed. {\sf B.} turned. {\sf C.} become. {\sf D.} had.
	\item A good way of $\ldots$ good is keeping it in a fridge. {\sf A.} enduring. {\sf B.} extending. {\sf C.} prolonging. {\sf D.} preserving.
	\item Many Asian countries still rely on rice as the $\ldots$ food. {\sf A.} staple. {\sf B.} capital. {\sf C.} superior. {\sf D.} winning.
	\item The dish had a very interesting taste as it was $\ldots$ with lemon. {\sf A.} spiced. {\sf B.} pickled. {\sf C.} seasoned. {\sf D.} flavored.
	\item The sight of so many sweets made the children's mouth $\ldots$. {\sf A.} drip. {\sf B.} water. {\sf C.} moisten. {\sf D.} wet.
	\item Just think! Next month you'll be $\ldots$ \& it seems like only yesterday you were a baby. {\sf A.} in your teens. {\sf B.} in your teenage. {\sf B.} at your teens. {\sf D.} teenager.
	\item ``Another cup of coffee?'' -- ``No, but thanks $\ldots$.'' {\sf A.} not all all. {\sf B.} for all. {\sf C.} all the same. {\sf D.} you for all.
	\item It was very $\ldots$ of you to lend us your car for the whole week. {\sf A.} generous. {\sf B.} ambitious. {\sf C.} modest. {\sf D.} thoughtless.
	\item We'll have to find someone to $\ldots$ our plants while we're away on holiday. {\sf A.} bring up. {\sf B.} get on. {\sf C.} grow up. {\sf D.} look after.
	\item I don't think you should be so $\ldots$ with the children. Watching television for half an hour a day won't do them any harm. {\sf A.} stubborn. {\sf B.} strict. {\sf C.} reliable. {\sf D.} thoughtful.
	\item It was very $\ldots$ of you to eat the last slice of cake without asking. {\sf A.} loyal. {\sf B.} sensitive. {\sf C.} self-confident. {\sf D.} naughty.
	\item Even very active children often stop playing sport when they $\ldots$. {\sf A.} tell off. {\sf B.} bring up. {\sf C.} grow up. {\sf D.} look up.
	\item There is some controversy over who is responsible for the $\ldots$ of penicillin. {\sf A.} creation. {\sf B.} discovery. {\sf C.} invention. {\sf D.} revolution.
	\item I inferred $\ldots$ his remark that Mrs. Vegetarian doesn't eat meat. {\sf A.} from. {\sf B.} of. {\sf C.} to. {\sf D.} with.
	\item He said he would sue us, but I don't think he'll $\ldots$ his threat. {\sf A.} achieve. {\sf B.} bring about. {\sf C.} perform. {\sf D.} carry out.
	\item How about a glass of orange juice to $\ldots$ your thirst? {\sf A.} quash. {\sf B.} quench. {\sf C.} quell. {\sf D.} quieten.
	\item He was arrested for trying to pass $\ldots$ notes at the bank. {\sf A.} camouflaged. {\sf B.} fake. {\sf C.} counterfeit. {\sf D.} fraudulent.
	\item The safe deposit box $\ldots$ a high-pitched sound when it was moved. {\sf A.} ejected. {\sf B.} excluded. {\sf C.} expelled. {\sf D.} emitted.
	\item Her chain was made of $\ldots$ gold, unmixed with any other metal. {\sf A.} pure. {\sf B.} dishonest. {\sf C.} clean. {\sf D.} right.
	\item I haven't been $\ldots$ Pete lately. {\sf A.} getting up  to. {\sf B.} getting through to. {\sf C.} getting on with. {\sf D.} getting round to.
	\item Although it was only $\ldots$ jewellery, it looked real enough. {\sf A.} fictional. {\sf B.} untrue. {\sf C.} invented. {\sf D.} imitation.
	\item You are late again -- please try to be $\ldots$ in future. {\sf A.} accurate. {\sf B.} punctual. {\sf C.} efficient. {\sf D.} reliable.
	\item After he broke the window, the boy was $\ldots$ from school. {\sf A.} exiled. {\sf B.} excluded. {\sf C.} extracted. {\sf D.} expelled.
	\item A child's 1st 5 years are the most important as far as learning is $\ldots$. {\sf A.} concerned. {\sf B.} affected. {\sf C.} hit. {\sf D.} touched.
	\item I don't think they'll be taking $\ldots$ any more teachers this term because there aren't many students. {\sf A.} up. {\sf B.} in. {\sf C.} on. {\sf D.} over.
	\item Could you speak $\ldots$ a bit? We can't hear you at the back of the room. {\sf A.} out. {\sf B.} loud. {\sf C.} over. {\sf D.} up.
	\item I'm going to take a couple of days $\ldots$ next month to help my sister move house. {\sf A.} over. {\sf B.} off. {\sf C.} out. {\sf D.} up.
	\item The police $\ldots$ off the street where the bomb had gone off. {\sf A.} cordoned. {\sf B.} battened. {\sf C.} fastened. {\sf D.} shuttered.
	\item We all make mistakes; no-one is $\ldots$. {\sf A.} fallible. {\sf B.} infallible. mistaken. {\sf D.} unmistakable.
	\item Everyone laughed when he took $\ldots$ the teacher so well. {\sf A.} over. {\sf B.} up. {\sf C.} off. {\sf D.} out.
	\item He offered me \$500 to break my contract. That's $\ldots$. {\sf A.} bribery. {\sf B.} blackmail. {\sf C.} compensation. {\sf D.} reward.
	\item It was such a shock to receive a letter like that $\ldots$. {\sf A.} in the red. {\sf B.} out of the blue. {\sf C.} in the pink. {\sf D.} over the moon.
	\item The hooligan $\ldots$ the money out of my hand \& ran away. {\sf A.} snatched. {\sf B.} gripped. {\sf C.} clutched. {\sf D.} withdrew.
	\item As he was caught in $\ldots$ of a gun, he was immediately a suspect. {\sf A.} control. {\sf B.} possession. {\sf C.} ownership. {\sf D.} handling.
	\item He is doing a science $\ldots$ at the University of Edinburgh. {\sf A.} career. {\sf B.} certificate. {\sf C.} title. {\sf D.} degree.
	\item The teacher told us to learn the irregular verbs $\ldots$ hear. {\sf A.} to. {\sf B.} at. {\sf C.} by. {\sf D.} in.
	\item The conductor $\ldots$ the boys off for misbehaving on the bus. {\sf A.} told. {\sf B.} said. {\sf C.} shouted. {\sf D.} cried.
	\item The 1st thing he did was to $\ldots$ his shoelaces \& take his shoes off. {\sf A.} disconnect. {\sf B.} mislead. {\sf C.} undo. {\sf D.} unpack.
	\item After considering the case, the judge put the young offender $\ldots$ for 2 years. {\sf A.} in charge. {\sf B.} in control. {\sf C.} on trial. {\sf D.} on probation.
	\item An IQ test is supposed to measure the $\ldots$ of your intelligence. {\sf A.} level. {\sf B.} extent. {\sf C.} degree. {\sf D.} size.
	\item She was sent out of the examination room because the examiner caught her $\ldots$. {\sf A.} sneaking. {\sf B.} cheating. {\sf C.} lying. {\sf D.} copying.
	\item We used to $\ldots$ all sorts of things when our parents went out. {\sf A.} get through to. {\sf B.} get over. {\sf C.} get up to. {\sf D.} get on with.
	\item I new that something $\ldots$ was going on as soon as I walked into the room. {\sf A.} huge. {\sf B.} odd. {\sf C.} vast. {\sf D.} massive.
	\item If I don't write you a note to say you have a doctor's appointment, the teacher will think you are playing $\ldots$. {\sf A.} truant. {\sf B.} the fool. {\sf C.} for time. {\sf D.} hard to get.
	\item I've got so much spare time now I'm thinking of taking $\ldots$ a new hobby like stamp collecting. {\sf A.} in. {\sf B.} off. {\sf C.} up. {\sf D.} over.
	\item You are not very $\ldots$ today, Hugh. What's the matter? I've never known you so quiet. {\sf A.} chattering. {\sf B.} loud. {\sf C.} speaking. {\sf D.} talkative.
	\item Rita is not $\ldots$ of doing this work -- she should change her class. {\sf A.} fit. {\sf B.} capable. {\sf C.} possible. {\sf D.} suitable.
	\item It should be $\ldots$ that students are expected to attend classes regularly. {\sf A.} marked. {\sf B.} reminded. {\sf C.} perceived. {\sf D.} noted.
	\item The traffic policeman let her $\ldots$ with a warning though she was driving way over the speed limit. {\sf A.} up. {\sf B.} off. {\sf C.} on. {\sf D.} out.
	\item This is an exciting book which $\ldots$ new ground in educational research. {\sf A.} breaks. {\sf B.} reaches. {\sf C.} scratches. {\sf D.} turns.
	\item The tutorial system at Oxford \& Cambridge is the $\ldots$ of many universities. {\sf A.} sorrow. {\sf B.} jealousy. {\sf C.} regret. {\sf D.} envy.
	\item When the detectives finally trapped him, he had to $\ldots$ to lying. {\sf A.} recourse. {\sf B.} resource. {\sf C.} resort. {\sf D.} retort.
	\item He was very thrilled when he heard he had $\ldots$ the exam. {\sf A.} passed. {\sf B.} approved. {\sf C.} failed. {\sf D.} won.
	\item I wish he'd let us make some of the decisions instead of completely taking $\ldots$ all our meetings. {\sf A.} in. {\sf B.} over. {\sf C.} off. {\sf D.} up.
	\item $\ldots$ my opinion, students shouldn't have to repeat the year if they fail their exams. {\sf A.} In. {\sf B.} At. {\sf C.} On. {\sf D.} from.
	\item Prof. Rush was speaking so quickly I couldn't $\ldots$ what he said. {\sf A.} accept. {\sf B.} listen. {\sf C.} catch. {\sf D.} take.
	\item Jane said she could not $\ldots$ the information given in the lecture. {\sf A.} accumulate. {\sf B.} absorb. {\sf C.} admire. {\sf D.} listen.
	\item Mr. Wellbred went to a school which $\ldots$ good manners \& self-discipline. {\sf A.} blossomed. {\sf B.} planted. {\sf C.} harvested. {\sf D.} cultivated.
	\item The smell of the burnt cabbage was so $\ldots$ that it spread to every room. {\sf A.} pervasive. {\sf B.} effusive. {\sf C.} extensive. {\sf D.} diffuse.
	\item Suddenly I understood perfectly \& everything fell $\ldots$ place. {\sf A.} down. {\sf B.} for. {\sf C.} into. {\sf D.} out.
	\item Mr. Horrid was a terrible teacher \& obviously not $\ldots$ for teaching. {\sf A.} cut in. {\sf B.} cut on. {\sf C.} cut up. {\sf D.} cut out.
	\item Most of the $\ldots$ were unimpressed by his latest film. {\sf A.} reviews. {\sf B.} criticizing. {\sf C.} critics. {\sf D.} comments.
	\item It was so embarrassing when Romeo forgot his $\ldots$ in the second act. {\sf A.} paper. {\sf B.} lines. {\sf C.} part. {\sf D.} script.
	\item I hope the film has a happy $\ldots$. {\sf A.} finish. {\sf B.} end. {\sf C.} finale. {\sf D.} ending.
	\item It's surprising the performance went so well after only 3 $\ldots$. {\sf A.} rehearsals. {\sf B.} auditions. {\sf C.} applauses. {\sf D.} directions.
	\item High risk sports have really taken $\ldots$. {\sf A.} up. {\sf B.} off. {\sf C.} in. {\sf D.} on.
	\item The person $\ldots$ the orchestra was a young woman in her early 20s. {\sf A.} composing. {\sf B.} directing. {\sf C.} conducting. {\sf D.} leading.
	\item We're always playing tricks on Pete because he takes ages to $\ldots$ on. {\sf A.} catch. {\sf B.} keep. {\sf C.} pick. {\sf D.} get.
	\item Little Tom did not like his 1st $\ldots$ at school at all. {\sf A.} course. {\sf B.} period. {\sf C.} presence. {\sf D.} term.
	\item There was a(n) $\ldots$ against the College's new syllabuses. {\sf A.} bang. {\sf B.} outcry. {\sf C.} scream. {\sf D.} whist.
	\item New students must $\ldots$ for classes before term begins. {\sf A.} enrol. {\sf B.} enter. {\sf C.} join. {\sf D.} teach.
	\item When you listen to a lecture, it is useful to $\ldots$ the important points. {\sf A.} clear. {\sf B.} put down. {\sf C.} notify. {\sf D.} write on.
	\item Prof. Silver was a most effective speaker \& his audience seemed to $\ldots$ on his every word. {\sf A.} catch. {\sf B.} cling. {\sf C.} hang. {\sf D.} hold.
	\item My brother found it difficult to learn to write because he is $\ldots$. {\sf A.} left-handed. {\sf B.} right-handed. {\sf C.} single-handed. {\sf D.} 2-handed.
	\item In writing the account of his summer adventures, Neville chose not to $\ldots$ his experiences in the order in which they happened. {\sf A.} compose. {\sf B.} classify. {\sf C.} compare. {\sf D.} arrange.
	\item This cheese has gone $\ldots$. You'll have to throw it away. {\sf A.} mouldy. {\sf B.} decayed. {\sf C.} stale. {\sf D.} contaminated.
	\item The steak looked tender, but it was as tough as $\ldots$. {\sf A.} a belt. {\sf B.} a saddle. {\sf C.} old boots. {\sf D.} rubber.
	\item Dinner will be served $\ldots$ but we have time for a drink before then. {\sf A.} actually. {\sf B.} currently. {\sf C.} lately. {\sf D.} presently.
	\item Can you help me work $\ldots$ the answers to these maths problems? {\sf A.} in. {\sf B.} out. {\sf C.} up. {\sf D.} off.
	\item I can agree with you to a certain $\ldots$, Professor, but not entirely. {\sf A.} pan. {\sf B.} level. {\sf C.} extent. {\sf D.} way.
	\item The Examination Board have recently changed the $\ldots$ for the Diploma in History. {\sf A.} brochure. {\sf B.} syllabus. {\sf C.} programme. {\sf D.} compendium.
	\item Slow down a bit. I can't $\ldots$ up with you when you walk so fast. {\sf A.} get. {\sf B.} hold. {\sf C.} keep. {\sf D.} work.
	\item The lecturer spoke so fast that I found it hard to take $\ldots$ what he was saying. {\sf A.} away. {\sf B.} up. {\sf C.} over. {\sf D.} in.
	\item In the examination you may be asked for comments on various $\ldots$ of a topic. {\sf A.} angles. {\sf B.} features. {\sf C.} aspects. {\sf D.} qualities.
	\item Failing the final exam was a big $\ldots$ to my hopes. {\sf A.} band. {\sf B.} kick. {\sf C.} hit. {\sf D.} blow.
	\item Sign your name on the $\ldots$ line. {\sf A.} dotted. {\sf B.} broken. {\sf C.} drawn. {\sf D.} spotted.
	\item We shall have to use that glass as we are a cup $\ldots$. {\sf A.} few. {\sf B.} short. {\sf C.} minus. {\sf D.} less.
	\item Some language students reach a high $\ldots$ of competence in communication. {\sf A.} degree. {\sf B.} grade. {\sf C.} level. {\sf D.} mark.
	\item Unfortunately, state schools don't have much in the $\ldots$ of sports facilities compared to private schools. {\sf A.} part. {\sf B.} case. {\sf C.} means. {\sf D.} way.
	\item Your answers to the examination questions must $\ldots$ exactly the instructions given below. {\sf A.} accompany. {\sf B.} follow. {\sf C.} conform. {\sf D.} keep.
	\item It is very difficult to $\ldots$ the exact meaning of an idiom in a foreign language. {\sf A.} convert. {\sf B.} transfer. {\sf C.} exchange. {\sf D.} convey.
	\item On Sunday, Vivian studied for 7 hours $\ldots$. {\sf A.} on end. {\sf B.} at once. {\sf C.} in full. {\sf D.} at length.
	\item Tom handed in the test \& awaited the results $\ldots$. {\sf A.} in the same breath. {\sf B.} out of breath. {\sf C.} with bated breath. {\sf D.} under her breath.
	\item Don't tell me your name again. It's $\ldots$. {\sf A.} off my chest. {\sf B.} in my mouth. {\sf C.} at my fingertips. {\sf D.} on the tip of my tongue.
	\item She's had a $\ldots$ in her left shoulder for a week now. {\sf A.} sore. {\sf B.} ache. {\sf C.} pain. {\sf D.} hurt.
	\item Do you think parents should $\ldots$ their children from spending too much time online? {\sf A.} deny. {\sf B.} refuse. {\sf C.} disqualify. {\sf D.} prevent.
	\item He criticized her in the meeting \& I'm afraid she's really taken it to $\ldots$. {\sf A.} heart. {\sf B.} brain. {\sf C.} mind. {\sf D.} chest.
	\item He was so badly injured that they had to bring on a $\ldots$ \& carry him off the football pitch. {\sf A.} bed. {\sf B.} hammock. {\sf C.} stretcher. {\sf D.} bunk.
	\item In some countries you don't need a doctor's $\ldots$ to buy antibiotics. {\sf A.} recipe. {\sf B.} prescription. {\sf C.} receipt. {\sf D.} note.
	\item The nurse took his $\ldots$ \& checked that he was comfortable. {\sf A.} heat. {\sf B.} fever. {\sf C.} temperature. {\sf D.} thermal.
	\item You $\ldots$ be hungry again. We've just had lunch!
	\item Jane is so $\ldots$ in her work that it would be a pity to disturb her. {\sf A.} intent. {\sf B.} attentive. {\sf C.} consumed. {\sf D.} absorbed.
	\item It suddenly $\ldots$ on me what he really meant. {\sf A.} dawned. {\sf B.} came. {\sf C.} hit. {\sf D.} struck.
	\item You will need a pen \& some paper to $\ldots$ this problem. It is too difficult to do in your head. {\sf A.} discover. {\sf B.} work out. {\sf C.} realize. {\sf D.} find out.
	\item The little country hotel served them a meal fit for $\ldots$. {\sf A.} a king. {\sf B.} a lord. {\sf C.} an emperor. {\sf D.} a queen.
	\item Let me $\ldots$ the bill before we leave this restaurant. {\sf A.} arrange. {\sf B.} conclude. {\sf C.} settle. {\sf D.} pay up.
	\item Would you please $\ldots$ from smoking while the lecture is in progress? {\sf A.} refrain. {\sf B.} keep yourself. {\sf C.} avoid. {\sf D.} stop.
	\item This course $\ldots$ no previous knowledge of the subject. {\sf A.} assembles. {\sf B.} assumes. {\sf C.} assigns. {\sf D.} assures.
	\item The students paid $\ldots$ attention to their distinguished professor. {\sf A.} respectable. {\sf B.} respected. {\sf C.} respectful. {\sf D.} respective.
	\item Mr. Discontent wanted to complain to the waiter but was afraid of making a(n) $\ldots$. {\sf A.} act. {\sf B.} drama. {\sf C.} play. {\sf D.} scene.
	\item Have you $\ldots$ for any evening classes next term? {\sf A.} enrolled. {\sf B.} engaged. {\sf C.} inscribed. {\sf D.} signed.
	\item Can you pass me a knife? I want to $\ldots$ these onions. {\sf A.} fry. {\sf B.} roll. {\sf C.} chop. {\sf D.} grate.
	\item She had to $\ldots$ her dress because she had lost so much weight. {\sf A.} shift. {\sf B.} change. {\sf C.} loosen. {\sf D.} alter.
	\item Look at that beautiful brooch! I only paid \$3 for it at the sale. It was a real $\ldots$. {\sf A.} bargain. {\sf B.} affair. {\sf C.} benefit. {\sf D.} occasion.
	\item You'll be $\ldots$ your money if you buy that hat: you will never wear it. {\sf A.} giving away. {\sf B.} wasting. {\sf C.} saving. {\sf D.} losing.
	\item The cost of the material is in the $\ldots$ bill for the work. {\sf A.} included. {\sf B.} enclosed. {\sf C.} held. {\sf D.} combined.
	\item If you don't know how to wash that sweater, look at the $\ldots$ inside the collar. {\sf A.} badge. {\sf B.} label. {\sf C.} notice. {\sf D.} sign.
	\item I don't think I could eat any $\ldots$, but I'd like a coffee. {\sf A.} dessert. {\sf B.} starter. {\sf C.} main course. {\sf D.} second helping.
	\item Can you $\ldots$ the soup while I slice the bread? {\sf A.} sprinkle. {\sf B.} whisk. {\sf C.} roll. {\sf D.} stir.
	\item I find it really hard to walk in shoes with very high $\ldots$. {\sf A.} soles. {\sf B.} heels. {\sf C.} bottoms. {\sf D.} supports.
	\item Could you go down to the baker's \& get me a $\ldots$ of bread \& a couple of jam doughnuts. {\sf A.} loaf. {\sf B.} piece. {\sf C.} lump. {\sf D.} slice.
	\item He was charged with a(n) $\ldots$ of currency regulations. {\sf A.} break. {\sf B.} disrespect. {\sf C.} breach. {\sf D.} observance.
	\item Our insurance policy offers immediate $\ldots$ against the risk of burglary, accident or damage by fire. {\sf A.} cart. {\sf B.} security. {\sf C.} relief. {\sf D.} cover.
	\item The man jumped out of the window \& committed $\ldots$. {\sf A.} suicide. {\sf B.} homicide. {\sf C.} murder. {\sf D.} death.
	\item She is always wearing a strange red hat with a long $\ldots$ in it. {\sf A.} hair. {\sf B.} fur. {\sf C.} feather. {\sf D.} tail.
	\item ``How would you like your hair cut, sir?'' -- ``Just a $\ldots$ please.'' {\sf A.} cut. {\sf B.} trim. {\sf C.} shave. {\sf D.} snip.
	\item Corporal Lucky was hit by a bullet but he was only $\ldots$ wounded. {\sf A.} barely. {\sf B.} little. {\sf C.} slightly. {\sf D.} merely.
	\item My sister works on the $\ldots$ in the local supermarket on Saturday mornings. {\sf A.} check-through. {\sf B.} check-up. {\sf C.} check-in. {\sf D.} check-out.
	\item We had to wait in a $\ldots$ in the post office for over 10 minutes. {\sf A.} row. {\sf B.} line. {\sf C.} queue. {\sf D.} file.
	\item I lost the receipt \& the shop assistant refused to give me a $\ldots$. {\sf A.} discount. {\sf B.} refund. {\sf C.} guarantee. {\sf D.} exchange.
	\item They said they could order it for me because they didn't have my size in $\ldots$. {\sf A.} shop. {\sf B.} hold. {\sf C.} stock. {\sf D.} store.
	\item Our local supermarket charges a lot to $\ldots$ groceries. {\sf A.} deliver. {\sf B.} carry. {\sf C.} bring. {\sf D.} bear.
	\item The accused man was able to prove his innocence at the trial \& was $\ldots$. {\sf A.} absolved. {\sf B.} acquitted. {\sf C.} forgiven. {\sf D.} pardoned.
	\item If the terrorists are not sent to prison, there will be a public $\ldots$. {\sf A.} attack. {\sf B.} onslaught. {\sf C.} recrimination. {\sf D.} outcry.
	\item The prisoners had spent almost a month digging a $\ldots$ before the guards discovered it. {\sf A.} pipe. {\sf B.} tunnel. {\sf C.} subway. {\sf D.} pass.
	\item He was thrown into prison \& $\ldots$ of his property. {\sf A.} confiscated. {\sf B.} denied. {\sf C.} removed. {\sf D.} deprived.
	\item Don't take that road. There's always very $\ldots$ traffic along there. {\sf A.} heavy. {\sf B.} strong. {\sf C.} powerful. {\sf D.} hard.
	\item She's so $\ldots$ skinned that nothing offends her. {\sf A.} hard. {\sf B.} thick. {\sf C.} strong. {\sf D.} heavy.
	\item This tea is too $\ldots$ for me. Could you bring me some hot water? {\sf A.} strong. {\sf B.} thick. {\sf C.} hard. {\sf D.} heavy.
	\item The older children used to $\ldots$ on him in the playground at lunch time. {\sf A.} get. {\sf B.} hit. {\sf C.} pick. {\sf D.} beat.
	\item Put the lid back on the tin so that the biscuits $\ldots$ crisp \& fresh. {\sf A.} stay. {\sf B.} bring. {\sf C.} get. {\sf D.} become.
	\item I'm afraid that the herring we had for supper has given me $\ldots$. {\sf A.} sickness. {\sf B.} indisposition. {\sf C.} infection. {\sf D.} indigestion.
	\item Miss Diligent did 9 hours' $\ldots$ studying a day for her exam. {\sf A.} solid. {\sf B.} heavy. {\sf C.} powerful. {\sf D.} big.
	\item It was during my stay in India that I $\ldots$ taste for very hot curry. {\sf A.} gained. {\sf B.} acquired. {\sf C.} got. {\sf D.} received.
	\item I suppose I $\ldots$ lent the book to Simon, but I'm almost sure I didn't. {\sf A.} might. {\sf B.} could. {\sf C.} must have. {\sf D.} might have.
	\item You $\ldots$ that man our phone number. I didn't trust him at all. {\sf A.} must not have given. {\sf B.} might not have given. {\sf C.} should not have given. {\sf D.} could not have given.
	\item The hot weather has made the $\ldots$ even worse. {\sf A.} famine. {\sf B.} drought. {\sf C.} floods. {\sf D.} hail.
	\item Fortunately only 3 passengers suffered minor $\ldots$. {\sf A.} hurts. {\sf B.} pains. {\sf C.} injuries. {\sf D.} damages.
	\item The government have declared a state of $\ldots$ after yesterday's earthquake. {\sf A.} crisis. {\sf B.} emergency. {\sf C.} urgency. {\sf D.} danger.
	\item Scientists are working to find a cure for $\ldots$ like AIDS. {\sf A.} ills. {\sf B.} infirmities. {\sf C.} diseases. {\sf D.} complaints.
	\item Thousands of $\ldots$ are fleeing the war in the north. {\sf A.} emigrants. {\sf B.} employees. {\sf C.} escapees. {\sf D.} refugees.
	\item This lime juice needs $\ldots$ before you can drink it. {\sf A.} diluting. {\sf B.} dissolving. {\sf C.} softening. {\sf D.} watering.
	\item Put the orange $\ldots$ in the waste basket, not on the floor. {\sf A.} skin. {\sf B.} find. {\sf C.} shell. {\sf D.} peel.
	\item People were injured by $\ldots$ stones the size of tennis balls. {\sf A.} sleet. {\sf B.} gale. {\sf C.} hail. {\sf D.} snow.
	\item The food they serve here is excellent because they have a fine $\ldots$. {\sf A.} cooker. {\sf B.} chef. {\sf C.} headwaiter. {\sf D.} stove.
	\item I can't remember Mary's number. I'll have to $\ldots$ in the phone book. {\sf A.} look it up. {\sf B.} look it. {\sf C.} look for. {\sf D.} look it out.
	\item They arrived so late for the meal, that the food was $\ldots$. {\sf A.} dried. {\sf B.} hard. {\sf C.} spoilt. {\sf D.} lost.
	\item Don't bring those clothes in off the line yet. They're still a bit $\ldots$. {\sf A.} humid. {\sf B.} mild. {\sf C.} cool. {\sf D.} damp.
	\item She had just $\ldots$ the shell of the hard-boiled egg \& was starting to peel it. {\sf A.} burst. {\sf B.} cracked. {\sf C.} fractured. {\sf D.} shattered.
	\item A $\ldots$ of wind blew my hat off. {\sf A.} puff. {\sf B.} draft. {\sf C.} gust. {\sf D.} breeze.
	\item They have closed the center of the city to traffic because of $\ldots$. {\sf A.} overpopulation. {\sf B.} contamination. {\sf C.} famine. {\sf D.} pollution.
	\item I can get $\ldots$ on a lot less now that I don't have to pay rent. {\sf A.} over. {\sf B.} down. {\sf C.} by. {\sf D.} through.
	\item I am $\ldots$ furious with Charlotte. How could she forget our wedding anniversary? {\sf A.} absolutely. {\sf B.} extremely. {\sf C.} very. {\sf D.} terribly.
	\item Her sole $\ldots$ in the hotel kitchen was peeling potatoes. {\sf A.} role. {\sf B.} profession. {\sf C.} job. {\sf D.} work.
	\item Be careful not to $\ldots$ the tray too much or the glasses will slide off. {\sf A.} bend. {\sf B.} tilt. {\sf C.} incline. {\sf D.} twist.
	\item Can't I $\ldots$ you to another piece of cake? {\sf A.} convince. {\sf B.} persuade. {\sf C.} pull. {\sf D.} tempt.
	\item I couldn't open the wine as I didn't have a $\ldots$. {\sf A.} corkdriver. {\sf B.} corkscrew. {\sf C.} lever. {\sf D.} screwdriver.
	\item We've got 2 cats \& they get $\ldots$ very well together. {\sf A.} down. {\sf B.} through. {\sf C.} on. {\sf D.} by.
	\item I haven't had a chance to read anything apart from the $\ldots$ in today's paper. {\sf A.} headings. {\sf B.} titles. {\sf C.} captions. {\sf D.} headlines.
	\item How many television $\ldots$ do they have in your country? {\sf A.} canals. {\sf B.} channels. {\sf C.} ways. {\sf D.} numbers.
	\item I wish the papers weren't always so full of $\ldots$ about the royal family. {\sf A.} gossip. {\sf B.} rumor. {\sf C.} slander. {\sf D.} chat.
	\item My favorite radio $\ldots$ is Radio 3. {\sf A.} canal. {\sf B.} channel. {\sf C.} station. {\sf D.} way.
	\item A rude boy $\ldots$ his tongue out at me. {\sf A.} put. {\sf B.} stuck. {\sf C.} showed. {\sf D.} took.
	\item Do you know what $\ldots$ Radio 3 is on? {\sf A.} frequency. {\sf B.} line. {\sf C.} dial. {\sf D.} position.
	\item \textit{The Sun} is a newspaper with a $\ldots$ of over 3 million. {\sf A.} distribution. {\sf B.} coverage. {\sf C.} circulation. {\sf D.} sales.
	\item Can you pass me the $\ldots$ control? I want to see if my favorite soap has started. {\sf A.} far. {\sf B.} removed. {\sf C.} distant. {\sf D.} remote.
	\item He never really got $\ldots$ his relationship with Jenny. {\sf A.} through. {\sf B.} by. {\sf C.} over. {\sf D.} up.
	\item Waiting outside the examination room, I trembled with $\ldots$. {\sf A.} comprehension. {\sf B.} apprehension. {\sf C.} expectation. {\sf D.} tension.
	\item I'm terribly sorry but we haven't got any copies of the book in stock. They sold $\ldots$ early this morning. {\sf A.} out. {\sf B.} up. {\sf C.} through. {\sf D.} over.
	\item It is $\ldots$ that students will have doubled their vocabulary in 3 months. {\sf A.} worry. {\sf B.} foreseen. {\sf C.} hope. {\sf D.} anticipated.
	\item I wouldn't eat those gooseberries if I were you, they don't look $\ldots$ to me. {\sf A.} ripe. {\sf B.} ready. {\sf C.} formed. {\sf D.} underdone.
	\item In a multiple-choice exercise it's sometimes easier to $\ldots$ the wrong answers before choosing the right one.
	\item That fish smells terrible! Throw it in the $\ldots$. {\sf A.} bin. {\sf B.} rubbish. {\sf C.} sink. {\sf D.} waste-paper basket.
	\item I couldn't get $\ldots$ to Wendy so I left a message on her answering machine. {\sf A.} over. {\sf B.} through. {\sf C.} by. {\sf D.} up.
	\item I was completely $\ldots$ out after running so far. {\sf A.} exhausted. {\sf B.} run. {\sf C.} worn. {\sf D.} rubbed.
	\item She is hard-working \& efficient $\ldots$, she has had several years' experience of this kind of work. {\sf A.} As well as. {\sf B.} not only. {\sf C.} However. {\sf D.} Furthermore.
	\item Look $\ldots$! That shelf is about to fall down. {\sf A.} up. {\sf B.} out. {\sf C.} over. {\sf D.} down.
	\item So what have you been getting $\ldots$ while I've been away? {\sf A.} down to. {\sf B.} away with. {\sf C.} up to. {\sf D.} through to.
	\item $\ldots$ was she very rude to my mother, but she insulted my father too. {\sf A.} Not only. {\sf B.} As well as. {\sf C.} In addition to. {\sf D.} Furthermore.
	\item $\ldots$ passing all her examinations, she won a major tennis tournament this month. {\sf A.} In addition. {\sf B.} Moreover. {\sf C.} Not only. {\sf D.} As well as.
	\item Please give me 2 $\ldots$ of bread. {\sf A.} crumbs. {\sf B.} flakes. {\sf C.} slices. {\sf D.} lumps.
	\item She $\ldots$ her automatic cooker to cook the meat for 7 o'clock. {\sf A.} set. {\sf B.} put. {\sf C.} arranged. {\sf D.} timed.
	\item This knife is terribly $\ldots$. It won't even cut a piece of cheese. {\sf A.} sharp. {\sf B.} flat. {\sf C.} blunt. {\sf D.} thick.
	\item The apple was so $\ldots$ that he put some sugar on it. {\sf A.} bad. {\sf B.} juicy. {\sf C.} ripe. {\sf D.} sour.
	\item He rubbed his finger round the $\ldots$ of the glass. {\sf A.} ledge. {\sf B.} lip. {\sf C.} rim. {\sf D.} tip.
	\item A food mixer is a very useful $\ldots$ to have in the kitchen. {\sf A.} tool. {\sf B.} instrument. {\sf C.} scheme. {\sf D.} gadget.
	\item The refrigerator was $\ldots$ with food. {\sf A.} affluent. {\sf B.} crammed. {\sf C.} bloated. {\sf D.} full.
	\item Our new cat is black with white $\ldots$. {\sf A.} claws. {\sf B.} hooves. {\sf C.} wings. {\sf D.} paws.
	\item As the elephant walked along he swung his $\ldots$ from side to side. {\sf A.} beak. {\sf B.} fin. {\sf C.} trunk. {\sf D.} hump.
	\item The lion $\ldots$ when the keeper tried to make him go back into his cage. {\sf A.} roared. {\sf B.} neighed. {\sf C.} mooed. {\sf D.} barked.
	\item I think it's time you $\ldots$ \& told him you think things should be changed. {\sf A.} killed 2 birds with 1 stone. {\sf B.} put the cat among the pigeons. {\sf C.} took the bull by the horns. {\sf D.} got out of the rat race.
	\item Why did you $\ldots$ \& mention the party to Roger? It was supposed to be a surprise. {\sf A.} let the cat out of the bag. {\sf B.} put the cat among the pigeons. {\sf C.} have kittens. {\sf D.} kill 2 birds with 1 stone.
	\item It must be true. I heard it straight from the $\ldots$ mouth. {\sf A.} dog's. {\sf B.} horse's. {\sf C.} camel's. {\sf D.} cat's.
	\item I know you're upset about breaking up with Tony but there are plenty more $\ldots$. {\sf A.} horses in the stable. {\sf B.} cows in the field. {\sf C.} tigers in the zoo. {\sf D.} fish in the sea.
	\item Half way up the mountain there was a $\ldots$ where climbers could stay over night. {\sf A.} hut. {\sf B.} bungalow. {\sf C.} cottage. {\sf D.} bed-sit.
	\item It took them over an hour to put up their $\ldots$. {\sf A.} hut. {\sf B.} tent. {\sf C.} caravan. {\sf D.} bungalow.
	\item He lives in a tiny $\ldots$ up in the mountains. {\sf A.} suburb. {\sf B.} city. {\sf C.} village. {\sf D.} town.
	\item He likes to spend his holidays lying on a beach \& she likes to spend hers walking in $\ldots$. {\sf A.} a country. {\sf B.} the country. {\sf C.} country. {\sf D.} some country.
	\item Have you seen the exhibition at that new $\ldots$ in the high street? {\sf A.} stadium. {\sf B.} theater. {\sf C.} cathedral. {\sf D.} gallery.
	\item Competition from goods produced with cheaper labor has meant that many $\ldots$ have closed. {\sf A.} markets. {\sf B.} factories. {\sf C.} stadiums. {\sf D.} galleries.
	\item The royal wedding was celebrated in the magnificent 16th-century $\ldots$. {\sf A.} cathedral. {\sf B.} theater. {\sf C.} market. {\sf D.} gallery.
	\item The puppy was $\ldots$ so much I nearly dropped him. {\sf A.} sniggering. {\sf B.} wriggling. {\sf C.} chuckling. {\sf D.} giggling.
	\item We were $\ldots$ that the flight would be delayed for 2 hours. {\sf A.} invited. {\sf B.} ordered. {\sf C.} informed. {\sf D.} instructed.
	\item He managed to $\ldots$ the branch of a tree \& avoid being swept away by the flood waters. {\sf A.} grasp. {\sf B.} clutch. {\sf C.} hug. {\sf D.} munch.
	\item That man next to me on the bus must have been a $\ldots$. My wallet has gone. {\sf A.} shoplifter. {\sf B.} kidnapper. {\sf C.} hijacker. {\sf D.} pickpocket.
	\item The murder took place in the main street in the middle of the day, but there were no $\ldots$. {\sf A.} judges. {\sf B.} prosecutions. {\sf C.} witnesses. {\sf D.} juries.
	\item He was $\ldots$ when he was caught trying to catch a plane to France. {\sf A.} arrested. {\sf B.} put on probation. {\sf C.} executed. {\sf D.} acquitted.
	\item Skilled $\ldots$ have produced perfect copies of many of the world's famous paintings. {\sf A.} muggers. {\sf B.} smugglers. {\sf C.} rapists. {\sf D.} forgers.
	\item The judge released him on $\ldots$ but he'll have to appear in court again next week. {\sf A.} sentence. {\sf B.} bail. {\sf C.} punishment. {\sf D.} fine.
	\item We made $\ldots$ the nearest port when we heard the storm warning. {\sf A.} out. {\sf B.} of. {\sf C.} up. {\sf D.} for.
	\item The judge's ruling $\ldots$ a wave of protest campaigns across the country. {\sf A.} instigated. {\sf B.} launched. {\sf C.} provoked. {\sf D.} commenced.
	\item Mrs. Baker rolled out the pastry \& cut it into pretty $\ldots$ before baking it. {\sf A.} shapes. {\sf B.} pictures. {\sf C.} measures. {\sf D.} sizes.
	\item If you don't lift that saucepan carefully, it will spill \& you may $\ldots$ yourself. {\sf A.} cut. {\sf B.} peel. {\sf C.} skin. {\sf D.} scald.
	\item I don't like my steak cooked too long. I like it $\ldots$. {\sf A.} burnt. {\sf B.} underdone. {\sf C.} raw. {\sf D.} well done.
	\item Frozen food should always be $\ldots$ before it is cooked. {\sf A.} defrosted. {\sf B.} dissolved. {\sf C.} melted. {\sf D.} softened.
	\item When you have cracked the eggs, put the $\ldots$ in the rubbish bin. {\sf A.} coverings. {\sf B.} shells. {\sf C.} covers. {\sf D.} skins.
	\item She $\ldots$ the tablecloth carefully \& put it away in a drawer. {\sf A.} bent. {\sf B.} creased. {\sf C.} folded. {\sf D.} pleated.
	\item A fat lady sat on my shopping bag in the bus \& $\ldots$ my tomatoes. {\sf A.} squashed. {\sf B.} crushed. {\sf C.} broke. {\sf D.} squeezed.
	\item There was a sharp $\ldots$ \& the plate broke into 2 pieces. {\sf A.} bang. {\sf B.} crack. {\sf C.} crash. {\sf D.} noise.
	\item Can you give me a teaspoon to $\ldots$ my tea? {\sf A.} beat. {\sf B.} spin. {\sf C.} turn. {\sf D.} stir.
	\item Instead of telling her where he'd really been, he made $\ldots$ a complicated story. {\sf A.} out. {\sf B.} for. {\sf C.} up. {\sf D.} of.
	\item We'll make $\ldots$ for lost time if we keep working for a couple more hours. {\sf A.} out. {\sf B.} up. {\sf C.} for. {\sf D.} of.
	\item It's 15 years since Timothy worked in Holland \& his Dutch is pretty now $\ldots$. {\sf A.} stale. {\sf B.} scratchy. {\sf C.} sloppy. {\sf D.} rusty.
	\item The person who $\ldots$ an examination is supposed to see that nobody tries to cheat. {\sf A.} dominates. {\sf B.} governs. {\sf C.} supervises. {\sf D.} leads.
	\item Add the essence $\ldots$ until the teaspoonful is beaten into the mixture. {\sf A.} drop by drop. {\sf B.} inch by inch. {\sf C.} leaf by leaf. {\sf D.} spot by spot.
	\item Please don't talk in the $\ldots$ because there is an examination in the lecture hall. {\sf A.} promenade. {\sf B.} lane. {\sf C.} corridor. {\sf D.} way.
	\item I know you have a good voice \& have ambitions to be an opera singer but don't give up your day $\ldots$ yet! {\sf A.} job. {\sf B.} work. {\sf C.} situation. {\sf D.} place.
	\item I am in the $\ldots$ as to where Chris was last night. {\sf A.} shade. {\sf B.} pink. {\sf C.} dark. {\sf D.} black.
	\item Breaking his leg dealt a $\ldots$ to his chance of becoming a professional footballer. {\sf A.} thump. {\sf B.} strike. {\sf C.} hit. {\sf D.} blow.
	\item Jane is a sympathetic listener. She lent me a(n) $\ldots$ when I lost my job. {\sf A.} mind. {\sf B.} ear. {\sf C.} mouth. {\sf D.} eye.
	\item Let's make a dash for the train now as the rain seems to be $\ldots$ off. {\sf A.} reducing. {\sf B.} slowing. {\sf C.} easing. {\sf D.} running.
	\item Our firm is so successful because it is at the cutting $\ldots$ of computer technology. {\sf A.} limit. {\sf B.} edge. {\sf C.} verge. {\sf D.} fringe.
	\item He was $\ldots$ intensively for 2 weeks before the tennis tournament. {\sf A.} coached. {\sf B.} practiced. {\sf C.} learned. {\sf D.} taught.
	\item Our journey was $\ldots$ by the icy roads, which forced us to drive very slowly. {\sf A.} obstructed. {\sf B.} impeded. {\sf C.} barred. {\sf D.} blocked.
	\item Before the parachute jump he was carefully $\ldots$ in safety procedures. {\sf A.} shown. {\sf B.} familiarized. {\sf C.} presented. {\sf D.} instructed.
	\item She $\ldots$ her daughter to the care of a babysitter for the evening. {\sf A.} confided. {\sf B.} consigned. {\sf C.} entrusted. {\sf D.} resigned.
	\item The engineer $\ldots$ the machine with a hammer \&, miraculously, it roared back to life. {\sf A.} slapped. {\sf B.} smacked. {\sf C.} punched. {\sf D.} whacked.
	\item Fighting among rebel soldiers $\ldots$ last night \& a curfew has now been imposed on the city. {\sf A.} enhanced. {\sf B.} intensified. {\sf C.} heightened. {\sf D.} aggravated.
	\item While other companies collapsed in the economic recession, Cartwright Ltd. $\ldots$ \& share prices rose. {\sf A.} earned. {\sf B.} exploited. {\sf C.} profited. {\sf D.} gained.
	\item Rebecca $\ldots$ her fianc\'e across the face during an argument \& walked out of the restaurant. {\sf A.} slapped. {\sf B.} smashed. {\sf C.} struck. {\sf D.} punched.
	\item The yearly procession to the temple in honor of the local deity was an important religious $\ldots$ in that city. {\sf A.} activity. {\sf B.} festival. {\sf C.} fast. {\sf D.} habit.
	\item Although she was able to walk with the aid of crutches, having a broken ankle $\ldots$ her movements considerably. {\sf A.} barred. {\sf B.} blocked. {\sf C.} intervened. {\sf D.} hindered.
	\item The Oscar winning actress simply $\ldots$ charm \& professionalism in her acceptance speech. {\sf A.} exuded. {\sf B.} excluded. {\sf C.} expunged. {\sf D.} extricated.
	\item After buying an expensive new penthouse Marianna was flat $\ldots$. {\sf A.} shattered. {\sf B.} torn. {\sf C.} smashed. {\sf D.} broke.
	\item The new political party came to the $\ldots$ after the general election. {\sf A.} fore. {\sf B.} back. {\sf C.} side. {\sf D.} front.
	\item On  the 8th day of the strike the Minister $\ldots$ fit to make a statement. {\sf A.} appeared. {\sf B.} showed. {\sf C.} looked. {\sf D.} saw.
	\item Harry blew a $\ldots$ when his holiday was canceled. {\sf A.} fuse. {\sf B.} plug. {\sf C.} switch. {\sf D.} socket.
	\item Despite her poor exam results, Alice put a $\ldots$ face on the situation. {\sf A.} tough. {\sf B.} brave. {\sf C.} courageous. {\sf D.} bold.
	\item I was thrilled to meet Paul McCartney in the $\ldots$ when I sat next to him at the theater. {\sf A.} meat. {\sf B.} blood. {\sf C.} flesh. {\sf D.} vein.
	\item I'm very sorry, but these $\ldots$ are out of stock at the moment. {\sf A.} wares. {\sf B.} goods. {\sf C.} commodities. {\sf D.} supplies.
	\item Each $\ldots$ of the house must pay his own tax. {\sf A.} dweller. {\sf B.} settler. {\sf C.} resident. {\sf D.} inhabitant.
	\item Once at the skating rink, Ivan was allowed to skate to his heart's $\ldots$. {\sf A.} happiness. {\sf B.} delight. {\sf C.} content. {\sf D.} contentment.
	\item Floods have completely $\ldots$ the farmer's crops. {\sf A.} ruined. {\sf B.} damaged. {\sf C.} harmed. {\sf D.} injured.
	\item I $\ldots$ the notice-board quickly, looking for second-hand cars for sale. {\sf A.} peered at. {\sf B.} scanned. {\sf C.} scrutinized. {\sf D.} inspected.
	\item The farmer makes money by $\ldots$ pedigree horses. {\sf A.} bringing up. B nurturing. {\sf C.} growing. {\sf D.} breeding.
	\item For months I sat with my binoculars watching a bird $\ldots$ its young. {\sf A.} breed. {\sf B.} rear. {\sf C.} bring up. {\sf D.} grow.
	\item ``He $\ldots$ my rubber, miss!'' shouted the boy. {\sf A.} swiped. {\sf B.} ripped off. {\sf C.} abducted. {\sf D.} acquired.
	\item Many forest-$\ldots$ animals were killed in the fire. {\sf A.} settling. {\sf B.} residing. {\sf C.} inhabiting. {\sf D.} dwelling.
	\item This city has 4 million $\ldots$. {\sf A.} residents. {\sf B.} inhabitants. {\sf C.} dwellers. {\sf D.} settlers.
	\item During pioneer days a lot of land in the United States was up for $\ldots$. {\sf A.} gain. {\sf B.} promotion. {\sf C.} grabs. {\sf D.} taking.
	\item This lovely new dress fits like a $\ldots$. {\sf A.} glove. {\sf B.} gauntlet. {\sf C.} gown. {\sf D.} treat.
	\item Investors have taken $\ldots$ from the improving economic situation. {\sf A.} consolation. {\sf B.} courage. {\sf C.} heart. {\sf D.} meaning.
	\item We at \textit{Buyrite} throw down the $\ldots$ to competitors to match us for price, quality, \& service. {\sf A.} gauntlet. {\sf B.} mitten. {\sf C.} sword. {\sf D.} hat.
	\item Every time the government meets their demands, the union leaders move the $\ldots$. {\sf A.} lamp-posts. {\sf B.} goalposts. {\sf C.} bus stops. {\sf D.} roadblocks.
	\item The designer refuses to gild the $\ldots$, preferring clean, simple lines for his creations. {\sf A.} lily. {\sf B.} flower. {\sf C.} rose. {\sf D.} daisy.
	\item With fuel in short supply, machinery in the factory slowly ground to a(n) $\ldots$. {\sf A.} stop. {\sf B.} halt. {\sf C.} pause. {\sf D.} end.
	\item The teacher was adamant \& stuck to his $\ldots$ about the date of the final exam. {\sf A.} weapons. {\sf B.} neck. {\sf C.} thumb. {\sf D.} guns.
	\item David decided that smoking was ruining his health \& so gave it up for $\ldots$. {\sf A.} good. {\sf B.} always. {\sf C.} all. {\sf D.} once.
	\item For busy people in today's society, lifestyle management in gaining $\ldots$. {\sf A.} points. {\sf B.} speed. {\sf C.} terrain. {\sf D.} ground.
	\item The government is making little $\ldots$ in its fight to beat inflation. {\sf A.} forward. {\sf B.} advance. {\sf C.} headway. {\sf D.} improvement.
	\item The art teacher gave the children a fee $\ldots$ in their creative compositions. {\sf A.} offer. {\sf B.} gift. {\sf C.} kick. {\sf D.} hand.
	\item The forensic expert $\ldots$ the evidence looking for particles of skin. {\sf A.} scanned. {\sf B.} glimpsed at. {\sf C.} scrutinized. {\sf D.} peered at.
	\item My brother is always $\ldots$ me about my hooked nose. {\sf A.} teasing. {\sf B.} pestering. {\sf C.} getting on at. {\sf D.} harassing.
	\item He managed to $\ldots$ the flow of blood by tightly bandaging the wound. {\sf A.} restrain. {\sf B.} check. {\sf C.} curb. {\sf D.} inhibit.
	\item We intend this to be a(n) $\ldots$ project, taking us into the next decade. {\sf A.} constant. {\sf B.} incessant. {\sf C.} steady. {\sf D.} ongoing.
	\item It was only when he had been unemployed for 6 months that Neil's situation hit $\ldots$. {\sf A.} base. {\sf B.} down. {\sf C.} home. {\sf D.} back.
	\item My father has decided to $\ldots$ a beard to cover a small scar he has on his chin. {\sf A.} rear. {\sf B.} bring up. {\sf C.} breed. {\sf D.} grow.
	\item My grandmother was a lovely person who $\ldots$ pleasure from helping others. {\sf A.} derived. {\sf B.} gathered. {\sf C.} deduced. {\sf D.} collected.
	\item You were really $\ldots$ when you paid \$100 for those shoes. They're not even leather! {\sf A.} swiped. {\sf B.} ripped off. {\sf C.} pinched. {\sf D.} pilfered.
	\item I can't tell you $\ldots$ the population of Prague, but there's an encyclopedia in the cupboard. {\sf A.} off-hand. {\sf B.} in hand. {\sf C.} at hand. {\sf D.} on hand.
	\item After making several bad business deals the company was losing money hand over $\ldots$. {\sf A.} finger. {\sf B.} wrist. {\sf C.} fist. {\sf D.} thumb.
	\item Maggie is so moody \& unpredictable. She's apt to fly off the $\ldots$ without any real cause. {\sf A.} strap. {\sf B.} handle. {\sf C.} catch. {\sf D.} belt.
	\item A(n) $\ldots$ proportion of the population did not vote in the last elections. {\sf A.} essential. {\sf B.} fundamental. {\sf C.} grave. {\sf D.} significant.
	\item When a former secret agent tried to publish his memoirs, the government had certain parts of the book $\ldots$. {\sf A.} forbidden. {\sf B.} prohibited. {\sf C.} censored. {\sf D.} disallowed.
	\item Benjamin Britten, the composer, is probably most $\ldots$ for his opera ``Peter Grimes''. {\sf A.} conspicuous. {\sf B.} famous. {\sf C.} remarkable. {\sf D.} distinguished.
	\item $\ldots$ scientists from around the world met in London to discuss a revolutionary new drug. {\sf A.} Noteworthy. {\sf B.} Elevated. {\sf C.} Prestigious. {\sf D.} Eminent.
	\item Greg has, to all intents \& $\ldots$, finished his degree course, with the exception of his final dissertation. {\sf A.} purposes. {\sf B.} reasons. {\sf C.} aims. {\sf D.} proposals.
	\item It was $\ldots$ justice for Ted to receive a parking fine when he was at the doctor's. {\sf A.} poetic. {\sf B.} only. {\sf C.} rough. {\sf D.} small.
	\item There are more people employed, by \& $\ldots$ in the service sector than in manufacturing nowadays. {\sf A.} large. {\sf B.} all. {\sf C.} by. {\sf D.} through.
	\item We giggled at the sight of Mrs. Brown $\ldots$ down the road in her 6-inch stiletto heels. {\sf A.} staggering. {\sf B.} tottering. {\sf C.} reeling. {\sf D.} stumbling.
	\item He slammed my hand in the car door \&, to add insult to $\ldots$, didn't stop to apologize. {\sf A.} damage. {\sf B.} harm. {\sf C.} injury. {\sf D.} infamy.
	\item We were all on a $\ldots$-edge until the very end of the Hitchcock film. {\sf A.} chair. {\sf B.} knife. {\sf C.} cliff. {\sf D.} razor.
	\item Those stones have been here since time $\ldots$. {\sf A.} immemorable. {\sf B.} immortal. {\sf C.} innumerable. {\sf D.} immemorial.
	\item Enjoy your ``Sunwoy'' cruise, safe in the $\ldots$ that everything has been taken care of. {\sf A.} awareness. {\sf B.} recognition. {\sf C.} knowledge. {\sf D.} fact.
	\item This shoe repairer is so quick that he can sole \& heel your shoes in a(n) $\ldots$. {\sf A.} moment. {\sf B.} hurry. {\sf C.} instant. {\sf D.} jiffy.
	\item Sue spent weeks in hospital after being seriously $\ldots$ in a car crash. {\sf A.} ruined. {\sf B.} injured. {\sf C.} impaired. {\sf D.} damaged.
	\item When she lost her temper, she $\ldots$ the vase her ex-husband \& brought her. {\sf A.} snapped. {\sf B.} clipped. {\sf C.} smashed. {\sf D.} broke off.
	\item My neighbor has threatened to $\ldots$ over our dispute about property boundaries. {\sf A.} take legal action. {\sf B.} take legal aid. {\sf C.} try me. {\sf D.} bring me to trial.
	\item Mr. Wright is being $\ldots$ for fraud at the Old Bailey courthouse, this afternoon. {\sf A.} taken to court. {\sf B.} charged. {\sf C.} summoned. {\sf D.} tried.
	\item As long as you have the most $\ldots$ ingredients for the recipe, you can make do without the other things. {\sf A.} critical. {\sf B.} essential. {\sf C.} crucial. {\sf D.} vital.
	\item The heavy rain lashed down $\ldots$ throughout the night without letting up. {\sf A.} continually. {\sf B.} perpetually. {\sf C.} continuously. {\sf D.} eternally.
	\item The border guards have been accused of $\ldots$ refugees. {\sf A.} pestering. {\sf B.} getting on at. {\sf C.} teasing. {\sf D.} harassing.
	\item My young nieces kept $\ldots$ me to buy them sweets. {\sf A.} pestering. {\sf B.} teasing. {\sf C.} getting on at. {\sf D.} harassing.
	\item I $\ldots$ a plate when I was washing up, but it can still be used. {\sf A.} snapped. {\sf B.} smashed. {\sf C.} chipped. {\sf D.} shattered.
	\item The windscreen $\ldots$ into a thousand pieces when the car hit a lamp-post. {\sf A.} smashed. {\sf B.} shattered. {\sf C.} snapped. {\sf D.} cracked.
	\item The optician says you have to wear glasses, like it or $\ldots$ it. {\sf A.} jump. {\sf B.} loathe. {\sf C.} dislike. {\sf D.} lump.
	\item You will be putting your life on the $\ldots$ if you take up skydiving. {\sf A.} line. {\sf B.} way. {\sf C.} lane. {\sf D.} ground.
	\item His friends \& family left him in the $\ldots$ when he went bankrupt. {\sf A.} church. {\sf B.} lurch. {\sf C.} end. {\sf D.} street.
	\item I'll reserve $\ldots$ on Ben Shipley's latest novel until I've read it. {\sf A.} condemnation. {\sf B.} thought. {\sf C.} judgment. {\sf D.} criticism.
	\item Tara was really laying it $\ldots$ on about her accident at work. {\sf A.} fine. {\sf B.} broad. {\sf C.} thin. {\sf D.} thick.
	\item Were you on the $\ldots$ when you said you had resigned from work? {\sf A.} wagon. {\sf B.} flat. {\sf C.} level. {\sf D.} town.
	\item Stars of the Hollywood silver screen tended to be larger than $\ldots$. {\sf A.} life. {\sf B.} drama. {\sf C.} reality. {\sf D.} ever.
	\item Judging from the noise it is making, the washing machine is on its last $\ldots$. {\sf A.} gasp. {\sf B.} legs. {\sf C.} resort. {\sf D.} breath.
	\item I left the house in a hurry \& my bedroom was $\ldots$ with clothes. {\sf A.} scattered. {\sf B.} dispersed. {\sf C.} strewn. {\sf D.} sprinkled.
	\item Always having had $\ldots$ feet, Delia is off again, backpacking round India. {\sf A.} itchy. {\sf B.} scratchy. {\sf C.} sore. {\sf D.} light.
	\item The skaters $\ldots$ gracefully across the ice. {\sf A.} slid. {\sf B.} glided. {\sf C.} slipped. {\sf D.} skidded.
	\item The cat slept peacefully $\ldots$ in the long grass. {\sf A.} nestled. {\sf B.} huddled. {\sf C.} snuggled. {\sf D.} cuddled.
	\item Mr. Wright $\ldots$ his vegetable garden carefully. {\sf A.} attends. {\sf B.} grows. {\sf C.} maintains. {\sf D.} tends.
	\item Employees of the company are forbidden to $\ldots$ information about the secret formula. {\sf A.} betray. {\sf B.} divulge. {\sf C.} portray. {\sf D.} unveil.
	\item The soldier didn't $\ldots$ his comrades even when he was pressured to do so. {\sf A.} reveal. {\sf B.} divulge. {\sf C.} betray. {\sf D.} disclose.
	\item The removal men $\ldots$ the heavy piano up the stairs with great difficulty. {\sf A.} toted. {\sf B.} shoved. {\sf C.} thrust. {\sf D.} heaved.
	\item She $\ldots$ a few clothes into a case \& hurried to the airport. {\sf A.} scattered. {\sf B.} piled. {\sf C.} shoved. {\sf D.} heaved.
	\item My fear of wasps $\ldots$ from the fact that I was severely stung as a child. {\sf A.} begets. {\sf B.} stems. {\sf C.} generates. {\sf D.} commences.
	\item This year's series of open-air plays $\ldots$ with a performance of ``Electra''. {\sf A.} stems. {\sf B.} originates. {\sf C.} derives. d. commences.
	\item We could hear the monkeys $\ldots$ long before we reached their cage. {\sf A.} chatting. {\sf B.} prattling. {\sf C.} chattering. {\sf D.} babbling.
	\item Jane $\ldots$ about the surprise party for Sheila \& now the whole idea is ruined. {\sf A.} blabbed. {\sf B.} gossiped. {\sf C.} prattled. {\sf D.} chatted.
	\item I think you $\ldots$ a nerve when you mentioned Ralph's forthcoming retirement. {\sf A.} touched. {\sf B.} had. {\sf C.} drilled. {\sf D.} hit.
	\item Yes, I saw the last episode of the series, but I'm none the $\ldots$ about what really happened. {\sf A.} better. {\sf B.} surer. {\sf C.} cleverer. {\sf D.} wiser.
	\item Thousands of refugees are camping at the $\ldots$ between the 2 countries, hoping to find asylum. {\sf A.} boundary. {\sf B.} brim. {\sf C.} border. {\sf D.} rim.
	\item After 6 months of convalescence in a nursing home, Simon is finally on the $\ldots$. {\sf A.} mend. {\sf B.} top. {\sf C.} go. {\sf D.} run.
	\item For stamps \& coins to be of value, they need to be in mint $\ldots$. {\sf A.} state. {\sf B.} situation. {\sf C.} condition. {\sf D.} appearance.
	\item The professor looked as if he were $\ldots$ away in a world of his own. {\sf A.} yards. {\sf B.} kilometers. {\sf C.} streets. {\sf D.} miles.
	\item Dave gave the office party a $\ldots$ that year. {\sf A.} miss. {\sf B.} hit. {\sf C.} chance. {\sf D.} break.
	\item The cup is full to the $\ldots$ so be careful when you carry it. {\sf A.} brim. {\sf B.} boundary. {\sf C.} verge. {\sf D.} border.
	\item Young children are often $\ldots$ to illnesses such as measles. {\sf A.} liable. {\sf B.} sensitive. {\sf C.} susceptible. {\sf D.} apt.
	\item Some people like to $\ldots$ their neck doing dangerous sports. {\sf A.} endanger. {\sf B.} break. {\sf C.} twist. {\sf D.} risk.
	\item After losing my job, I was on the $\ldots$ of a nervous breakdown. {\sf A.} border. {\sf B.} brim. {\sf C.} verge. {\sf D.} bounds.
	\item The school playing fields are out of $\ldots$ while equipment is being set up for the cricket match.
	\item Children should keep away from the river $\ldots$ in case they fall in. {\sf A.} bounds. {\sf B.} bank. {\sf C.} brim. {\sf D.} border.
	\item We're spending our holidays on the $\ldots$ this summer. {\sf A.} beach. {\sf B.} bank. {\sf C.} coast. {\sf D.} shore.
	\item The old lady $\ldots$ her bag for fear she might lose it. {\sf A.} seized. {\sf B.} grabbed. {\sf C.} grasped. {\sf D.} clutched.
	\item Isn't packing 3 suitcases for a weekend away going a bit $\ldots$? {\sf A.} overload. {\sf B.} overhead. {\sf C.} overboard. {\sf D.} overskill.
	\item In spite of his stammer, Gerald rose to the $\ldots$ \& delivered a fascinating speech. {\sf A.} top. {\sf B.} situation. {\sf C.} bait. {\sf D.} occasion.
	\item ``How often do you visit your parents?'' ``$\ldots$ so often, even though they live some distance away.'' {\sf A.} Even. {\sf B.} Ever. {\sf B.} Every. {\sf D.} Never.
	\item My puppy was very expensive because he's a special $\ldots$. {\sf A.} breed. {\sf B.} sect. {\sf C.} tribe. {\sf D.} caste.
	\item ``Do you think I should ask Andy?'' ``I'm sure he'll be only $\ldots$ willing to help you with the project.'' {\sf A.} just. {\sf B.} too. {\sf C.} that. {\sf D.} so.
	\item It is possible to $\ldots$ out of the pension scheme if you do not wish to participate. {\sf A.} back. {\sf B.} charge. {\sf B.} opt. {\sf D.} break.
	\item You say you need new clothes but your wardrobe is full to $\ldots$ with dresses. {\sf A.} overfilling. {\sf B.} overflowing. {\sf C.} overlaying. {\sf D.} overstepping.
	\item Frank was informed at the $\ldots$ that the assignment would be no easy matter. {\sf A.} offset. {\sf B.} upset. {\sf C.} reset. {\sf D.} outset.
	\item it will be overstepping the $\ldots$ if you address the managing director by his 1st name. {\sf A.} mark. {\sf B.} line. {\sf C.} grade. {\sf D.} limit.
	\item I was $\ldots$ due to a personality clash with my boss. {\sf A.} resigned. {\sf B.} sacked. {\sf C.} retired. {\sf D.} made redundant.
	\item The $\ldots$ of the failure of the bank resulted in many people losing their life-savings. {\sf A.} catastrophe. {\sf B.} problem. {\sf C.} trouble. {\sf D.} cause.
	\item The play was such a $\ldots$ that it came off after only 2 nights. {\sf A.} cataclysm. {\sf B.} calamity. {\sf C.} catastrophe. {\sf D.} disaster.
	\item Hundreds of workers have been $\ldots$ due to financial problems at the factory. {\sf A.} sacked. {\sf B.} resigned. {\sf C.} made redundant. {\sf D.} retired.
	\item The $\ldots$ in India resembles the class system structure in Britain. {\sf A.} breed. {\sf B.} caste. {\sf C.} tribe. {\sf D.} sect.
	\item Anthropologists have spent years studying the social system of this $\ldots$. {\sf A.} breed. {\sf B.} caste. {\sf C.} tribe. {\sf D.} sect.
	\item Countries often become $\ldots$ to promote trade \& industry between them. {\sf A.} allies. {\sf B.} colleagues. {\sf C.} acquaintances. {\sf D.} accomplices.
	\item Most people have few friends but many $\ldots$ who they only see from time to time. {\sf A.} allies. {\sf B.} accomplices. {\sf C.} associates. {\sf D.} acquaintances.
	\item Many people are feeling the $\ldots$ now that there is an economic recession. {\sf A.} strain. {\sf B.} pinch. {\sf C.} prod. {\sf D.} pain.
	\item Victoria went to great $\ldots$ to complete a high quality representation. {\sf A.} efforts. {\sf B.} torment. {\sf C.} pains. {\sf D.} difficulty.
	\item Julia's work is not a $\ldots$ on Natalie's. {\sf A.} patch. {\sf B.} spot. {\sf C.} scratch. {\sf D.} scrap.
	\item Somebody as conceited as Ron needs bringing down a $\ldots$ or two. {\sf A.} step. {\sf B.} notch. {\sf C.} rung. {\sf D.} peg.
	\item New technology is being $\ldots$ in at work. {\sf A.} phased. {\sf B.} shown. {\sf C.} installed. {\sf D.} filtered.
	\item The foreman came to the building site to pay the workmen's $\ldots$. {\sf A.} fees. {\sf B.} allowance. {\sf C.} wages. {\sf D.} salaries.
	\item I have no appetite \& I am lethargic. I've been feeling under $\ldots$ for ages. {\sf A.} pair. {\sf B.} threat. {\sf C.} stress. {\sf D.} par.
	\item The economic situation makes many people unwilling to take the $\ldots$ \& open their own businesses. {\sf A.} plunge. {\sf B.} bull. {\sf C.} initiative. {\sf D.} opportunity.
	\item I don't know how I can $\ldots$ up the courage to tell him the awful news. {\sf A.} pick. {\sf B.} set. {\sf C.} store. {\sf D.} pluck.
	\item The new soap opera on Channel 3. $\ldots$ new depths in terms of tastelessness. {\sf A.} fills. {\sf B.} reaches. {\sf C.} plumbs. {\sf D.} achieves.
	\item We'll keep you $\ldots$ of any further changes in the examination specifications. {\sf A.} noticed. {\sf B.} announced. {\sf C.} mailed. {\sf D.} posted.
	\item Relief workers $\ldots$ hundreds of crates of second-hand clothing which had been donated. {\sf A.} revealed. {\sf B.} unpacked. {\sf C.} unwrapped. {\sf D.} unveiled.
	\item The Queen was present at the official $\ldots$ of the statue. {\sf A.} unveiling. {\sf B.} unpacking. {\sf C.} unwrapping. {\sf D.} revealing.
	\item The hut finally fell down, unable to $\ldots$ days of violent winds \& rain. {\sf A.} put up with. {\sf B.} withstand. {\sf C.} support. {\sf D.} afford.
	\item As a single parent, it's difficult for her to $\ldots$ a family. {\sf A.} put up with. {\sf B.} afford. {\sf C.} support. {\sf D.} withstand.
	\item I phoned the company, who $\ldots$ me that my goods had been despatched. {\sf A.} confirmed. {\sf B.} assured. {\sf C.} guaranteed. {\sf D.} reassured.
	\item It was extremely $\ldots$ of you to drive home in that blizzard. {\sf A.} reckless. {\sf B.} adventurous. {\sf C.} impervious. {\sf D.} instinctive.
	\item Please send a reply to the wedding invitation in order to $\ldots$ your attendance. {\sf A.} guarantee. {\sf B.} confirm. {\sf C.} assure. {\sf D.} reassure.
	\item I really admire the hero of the film. He's so $\ldots$. {\sf A.} reckless. {\sf B.} adventurous. {\sf C.} foolhardy. {\sf D.} instinctive.
	\item A modern cash register keeps a record of every $\ldots$. {\sf A.} contract. {\sf B.} deal. {\sf C.} purchaser. {\sf D.} transaction.
	\item Looking after a house, 4 children, a lazy husband \& 2 dogs is real $\ldots$. {\sf A.} drudgery. {\sf B.} labor. {\sf C.} toil. {\sf D.} grind.
	\item The experimental play was only a $\ldots$ success, which disappointed the playwright. {\sf A.} local. {\sf B.} reserved. {\sf C.} qualified. {\sf D.} cautious.
	\item It takes time to get a financial system up \& $\ldots$ after the introduction of a new currency. {\sf A.} running. {\sf B.} proceeding. {\sf C.} walking. {\sf D.} going.
	\item Life is so full of both good fortune \& misfortunes that you have to learn to take the rough with the $\ldots$. {\sf A.} tough. {\sf B.} ready. {\sf C.} calm. {\sf D.} smooth.
	\item Stephen really lost his $\ldots$ when his dental appointment was canceled yet again. {\sf A.} rag. {\sf B.} voice. {\sf C.} calm. {\sf D.} head.
	\item In today's competitive world, everyone needs a regular holiday to $\ldots$ their batteries. {\sf A.} refill. {\sf B.} recast. {\sf C.} recharge. {\sf D.} rectify.
	\item When my new motor kept breaking down, I knew I'd been taken for a $\ldots$ by the second-hand car salesman. {\sf A.} drive. {\sf B.} ride. {\sf C.} walk. d. stroll.
	\item It has been medically proven that a poor diet leads to $\ldots$ growth. {\sf A.} controlled. {\sf B.} tempered. {\sf C.} curbed. {\sf D.} stunted.
	\item Don't take it as $\ldots$ that you'll be promoted in your job; other colleagues stand a good chance too. {\sf A.} read. {\sf B.} standard. {\sf C.} fixed. {\sf D.} word.
	\item My husband \& I are not rich but we are $\ldots$. {\sf A.} wealthy. {\sf B.} tight. {\sf C.} comfortable. {\sf D.} hard-up.
	\item Believing that the firm had given him a raw $\ldots$, he took his case to an industrial tribunal. {\sf A.} deal. {\sf B.} meal. {\sf C.} apple. {\sf D.} nerve.
	\item It was a hot summer day \& ice cream salesmen were doing a $\ldots$ trade. {\sf A.} busy. {\sf B.} roaring. {\sf C.} bustling. {\sf D.} lucrative.
	\item Bob is so short-tempered; he should try to $\ldots$ his anger. {\sf A.} stunt. {\sf B.} control. {\sf C.} temper. {\sf D.} curb.
	\item If you $\ldots$ your demands, they may be accepted by others. {\sf A.} temper. {\sf B.} control. {\sf C.} stunt. {\sf D.} curb.
	\item The teacher told us to $\ldots$ the poem by the next lesson. {\sf A.} remember. {\sf B.} recollect. {\sf C.} memorize. {\sf D.} remind.
	\item I don't $\ldots$ her name, as we met only once. {\sf A.} memorize. {\sf B.} recollect. {\sf C.} reminisce. {\sf D.} remind.
	\item Don't expect Alan to buy you a drink. He earns plenty of money but he's very $\ldots$. {\sf A.} hard-up. {\sf B.} wealthy. {\sf C.} tight. {\sf D.} comfortable.
	\item Could I borrow a few pounds? I'm a bit $\ldots$ at the moment. {\sf A.} comfortable. {\sf B.} wealthy. {\sf C.} tight. {\sf D.} hard-up.
	\item The audience were in $\ldots$ as they watched the latest Aykebourne comedy. {\sf A.} finger. {\sf B.} place. {\sf C.} point. {\sf D.} thumb.
	\item Don't mention work to Ray, as it's a sore $\ldots$ with him at the moment. {\sf A.} finger. {\sf B.} place. {\sf C.} point. {\sf D.} thumb.
	\item The trouble with socializing with colleagues is that they usually end up talking $\ldots$. {\sf A.} sense. {\sf B.} rot. {\sf C.} back. {\sf D.} shop.
	\item The dog $\ldots$ under the table when I shouted at him for eating my steak. {\sf A.} slithered. {\sf B.} slunk. {\sf C.} scampered. {\sf D.} strutted.
	\item The opposition will be elected into government at the next election, without a $\ldots$ of a doubt. {\sf A.} shadow. {\sf B.} shade. {\sf C.} benefit. {\sf D.} hue.
	\item Take your mobile phone with you just to be on the $\ldots$ side. {\sf A.} sunny. {\sf B.} safe. {\sf C.} straight. {\sf D.} secure.
	\item It's hard to get back into the $\ldots$ of things after a long holiday. {\sf A.} pace. {\sf B.} way. {\sf C.} rhythm. {\sf D.} swing.
	\item Noisy parties are really not my $\ldots$. {\sf A.} scene. {\sf B.} liking. {\sf C.} idea. {\sf D.} preference.
	\item I hate the way Tony $\ldots$ around looking so self-important. {\sf A.} slithers. {\sf B.} scampers. {\sf C.} struts. {\sf D.} slinks.
	\item Rosie the kitten $\ldots$ playfully around with a ball. {\sf A.} scampered. {\sf B.} strutted. C slithered. {\sf D.} slunk.
	\item A long, green snake $\ldots$ through the grass \& disappeared. {\sf A.} strutted. {\sf B.} slithered. {\sf C.} slunk. {\sf D.} scampered.
	\item She tried to $\ldots$ Tom's importance to the company in order to gain a promotion for herself. {\sf A.} reduce. {\sf B.} dwindle. {\sf C.} shrink. {\sf D.} diminish.
	\item Due to inflation, my savings have gradually $\ldots$ to practically nothing. {\sf A.} dwindled. {\sf B.} shrank. {\sf C.} reduced. {\sf D.} diminished.
	\item My new pullover $\ldots$ to half its previous size when I washed it. {\sf A.} reduced. {\sf B.} shrank. {\sf C.} diminished. {\sf D.} dwindled.
	\item I need to $\ldots$ my weight by about 10 kilos for health reasons. {\sf A.} shrink. {\sf B.} diminish. {\sf C.} dwindle. {\sf D.} reduce.
	\item I don't think you're suitably $\ldots$ for this party. {\sf A.} dressed. {\sf B.} clad. {\sf C.} wearing. {\sf D.} clothed.
	\item Jack was disappointed not to be promoted as he was given to $\ldots$ that the job would be his. {\sf A.} know. {\sf B.} realize. {\sf C.} understand. {\sf D.} say.
	\item The Prime Minister has $\ldots$ the controversial statement he made about nuclear arms. {\sf A.} extracted. {\sf B.} retracted. {\sf C.} pulled out. {\sf D.} renounced.
	\item The King was forced to $\ldots$ his right to the throne when he married a divorcee. {\sf A.} extract. {\sf B.} pullout. {\sf C.} retract. {\sf D.} renounce.
	\item The team had to $\ldots$ of the competition because of injuries. {\sf A.} renounce. {\sf B.} extract. {\sf C.} pull out. {\sf D.} retract.
	\item The text was complicated \& therefore very difficult to $\ldots$ information from. {\sf A.} extract. {\sf B.} renounce. {\sf C.} retract. {\sf D.} pullout.
	\item She $\ldots$ her finger on a sharp rose thorn. {\sf A.} stung. {\sf B.} bit. {\sf C.} pricked. {\sf D.} stabbed.
	\item The baby was $\ldots$ by a bee whilst playing in the garden. {\sf A.} bitten. {\sf B.} stung. {\sf C.} pricked. {\sf D.} stabbed.
	\item The film star categorically $\ldots$ any connection with the scandal. {\sf A.} refuses. {\sf B.} revokes. {\sf C.} rejects. {\sf D.} denies.
	\item Due to his reckless driving his driving license was $\ldots$ by the court. {\sf A.} revoked. {\sf B.} denied. {\sf C.} refused. {\sf D.} rejected.
	\item Sue begged Tom to help her but he $\ldots$. {\sf A.} denied. {\sf B.} revoked. {\sf C.} rejected. {\sf D.} refused.
	\item The mountains were $\ldots$ in thick cloud. {\sf A.} clothed. {\sf B.} clothing. {\sf C.} dressing. {\sf D.} wearing.
	\item That loud heavy metal music from next door is $\ldots$ me up the wall. {\sf A.} sending. {\sf B.} driving. {\sf C.} bringing. {\sf D.} pushing.
	\item Bright children who are $\ldots$ on the uptake may get bored easily if they are not stimulated enough at school. {\sf A.} swift. {\sf B.} fast. {\sf C.} finest. {\sf D.} best.
	\item This new model of car is the $\ldots$ in driving luxury. {\sf A.} penultimate. {\sf B.} ultimate. {\sf C.} quick. {\sf D.} rapid.
	\item Shift work does have its $\ldots$ sometimes. {\sf A.} uses. {\sf B.} conveniences. {\sf C.} reasons. {\sf D.} purposes.
	\item My husband told me in no $\ldots$ terms that I would have to economize on household expenses. {\sf A.} unsure. {\sf B.} uncertain. {\sf C.} vague. {\sf D.} unclear.
	\item We had a $\ldots$ of a time at Jason's party yesterday. {\sf A.} week. {\sf B.} period. {\sf C.} whole. {\sf D.} whale.
	\item In $\ldots$ did I knock on the huge oak door, for nobody answered. {\sf A.} vane. {\sf B.} mane. {\sf C.} vain. {\sf D.} vein.
	\item Investors were caught $\ldots$ by the sharp drop in share prices. {\sf A.} undecided. {\sf B.} unawares. {\sf C.} unsuspecting. {\sf D.} unconscious.
	\item They were planning, $\ldots$ to Hilary, to throw a surprise party for her. {\sf A.} unaware. {\sf B.} oblivious. {\sf C.} unbeknown. {\sf D.} unknown.
	\item The economic situation in the country is going from bad to $\ldots$ after the stock market crash. {\sf A.} worse. {\sf B.} best. {\sf C.} better. {\sf D.} worst.
	\item Tom always tries to $\ldots$ perfection in everything he does. {\sf A.} accomplish. {\sf B.} attain. {\sf C.} gain. {\sf D.} manage.
	\item The firm will go bankrupt if it cannot meet its $\ldots$. {\sf A.} liabilities. {\sf B.} duties. {\sf C.} charges. {\sf D.} promises.
	\item The charity managed to $\ldots$ a great deal during its most recent project in Africa. {\sf A.} attain. {\sf B.} reach. {\sf C.} gain. {\sf D.} accomplish.
	\item Thank you very much, but I am afraid that, due to the political situation, the President must $\ldots$ your invitation to tour your country. {\sf A.} reject. {\sf B.} renounce. {\sf C.} snub. {\sf D.} decline.
	\item Peter was heartbroken when Sue $\ldots$ his offer of marriage so cruelty. {\sf A.} spurned. {\sf B.} disclaimed. {\sf C.} denied. {\sf D.} renounced.
	\item I wasn't $\ldots$ to see a queue outside the new sports center. {\sf A.} taken aback. {\sf B.} astounded. {\sf C.} amazed. {\sf D.} surprised.
	\item She was a little $\ldots$ by this strange coincidence. {\sf A.} astounded. {\sf B.} amazed. {\sf C.} taken back. {\sf D.} flabbergasted.
	\item We $\ldots$ to the manager about the assistant's behavior. {\sf A.} objected. {\sf B.} complained. {\sf C.} criticized. {\sf D.} nagged.
	\item Instead of $\ldots$ to yourself, why don't you make a formal complaint? {\sf A.} complaining. {\sf B.} whining. {\sf C.} objecting. {\sf D.} grumbling.
	\item The latest advertising $\ldots$ for facial tissues is a free gift with every purchase. {\sf A.} scam. {\sf B.} snare. {\sf C.} gimmick. {\sf D.} plot.
	\item Jewellery worth over \pounds1 million was $\ldots$ from a house in Wimbledon last night. {\sf A.} stolen. {\sf B.} burgled. {\sf C.} robbed. {\sf D.} mugged.
	\item Because it was the 1st time he'd been in trouble with the police, he was let $\ldots$ with a warning. {\sf A.} down. {\sf B.} off. {\sf C.} out. {\sf D.} away.
	\item The whole class will stay here until the person who broke the chair $\ldots$ up. {\sf A.} owns. {\sf B.} confesses. {\sf C.} tells. {\sf D.} admits.
	\item There will also be $\ldots$ to moderate winds blowing from the north. {\sf A.} calm. {\sf B.} fine. {\sf C.} thin. {\sf D.} light.
	\item My mother told me that, as a child, I used to $\ldots$ continually for sweets. {\sf A.} grumble. {\sf B.} whine. {\sf C.} complain. {\sf D.} pester.
	\item You'll need to put more effort $\ldots$ your work if you want to pass the exam. {\sf A.} on. {\sf B.} through. {\sf C.} into. {\sf D.} at.
	\item The $\ldots$ ordered the pilot of the plane to fly to Venezuela. {\sf A.} hijacker. {\sf B.} kidnapper. {\sf C.} mugger. {\sf D.} blackmailer.
	\item Tomorrow we can expect to see an end to the $\ldots$ rain in the region, but it will remain cloudy. {\sf A.} rough. {\sf B.} hard. {\sf C.} strong. {\sf D.} heavy.
	\item Sea levels are expected to $\ldots$ considerably in the next few decades. {\sf A.} rise. {\sf B.} raise. {\sf C.} arise. {\sf D.} lift.
	\item I was wondering if you could $\ldots$ me up at your place for the night. {\sf A.} give. {\sf B.} live. {\sf C.} take. {\sf D.} put.
	\item He put his own life at $\ldots$ in an attempt to save his dying friend. {\sf A.} danger. {\sf B.} risk. {\sf C.} hazard. {\sf D.} threat.
	\item Heavier fines should be imposed on those people who $\ldots$ litter. {\sf A.} waste. {\sf B.} dump. {\sf C.} throw. {\sf D.} drop.
	\item You only need to put a $\ldots$ of salt in the mixture, not a whole teaspoonful! {\sf A.} piece. {\sf B.} drop. {\sf C.} pinch. {\sf D.} little.
	\item He was $\ldots$ an injection in his arm. {\sf A.} put. {\sf B.} given. {\sf C.} had. {\sf D.} treated.
	\item You have very high blood $\ldots$, Mr. Woolgar. {\sf A.} tension. {\sf B.} infection. {\sf C.} level. {\sf D.} pressure.
	\item I should like to point out that 2 months have $\ldots$ since you promised to come \& repair my television. {\sf A.} elapsed. {\sf B.} expired. {\sf C.} lapsed. {\sf D.} transpired.
	\item He $\ldots$ me to do more exercise. {\sf A.} suggested. {\sf B.} offered. {\sf C.} recommended. {\sf D.} said.
	\item I wish you'd get $\ldots$ of these old books -- you never read them any more. {\sf A.} away. {\sf B.} out. {\sf C.} lost. {\sf D.} rid.
	\item My mum's just knitted me a lovely $\ldots$ sweater. {\sf A.} high-heeled. {\sf B.} second-hand. {\sf C.} long-sleeved. {\sf D.} ankle-length.
	\item $\ldots$ should not open their exam papers until told to do so by the invigilator. {\sf A.} Participants. {\sf B.} Candidates. {\sf C.} Competitors. {\sf D.} Takers.
	\item Most of the people in the $\ldots$ were friends or family of the actors \& actresses. {\sf A.} public. {\sf B.} spectators. {\sf C.} audience. {\sf D.} viewers.
	\item By the time we $\ldots$ to the station the train had already left. {\sf A.} got. {\sf B.} reached. {\sf C.} arrived. {\sf D.} found.
	\item These jeans don't $\ldots$ me any more; they're too tight round the waist. {\sf A.} suit. {\sf B.} match. {\sf C.} fit. {\sf D.} go with.
	\item Do you mind if I $\ldots$ that track again? I really like it. {\sf A.} put. {\sf B.} listen. {\sf C.} touch. {\sf D.} play.
	\item Johnny Strummer's the $\ldots$ vocalist \& the bass guitarist sings the backing vocals. {\sf A.} 1st. {\sf B.} lead. {\sf C.} front. {\sf D.} head.
	\item The hotel is situated next to an 18-hold gold $\ldots$. {\sf A.} course. {\sf B.} court. {\sf C.} field. {\sf D.} pitch.
	\item Share prices on the Stock Exchange plunged sharply in the morning but $\ldots$ slightly in the afternoon. {\sf A.} regained. {\sf B.} recovered. {\sf C.} restored. {\sf D.} retrieved.
	\item I couldn't $\ldots$ laughing when he fell in the pool with all his clothes on; it was so funny! {\sf A.} help. {\sf B.} afford. {\sf C.} stand. {\sf D.} give up.
	\item The sales assistants are so $\ldots$; they spend more time chatting to each other than serving customers. {\sf A.} helpless. {\sf B.} unhelpful. {\sf C.} helping. {\sf D.} helped.
	\item ``How dare you!'' she screamed, \&, picking up her bag, she $\ldots$ of the room. {\sf A.} made a bolt. {\sf B.} dashed. {\sf C.} stormed out. {\sf D.} sprinted.
	\item The device is conveniently small, but $\ldots$ it isn't particularly easy to use. {\sf A.} on the other hand. {\sf B.} on the other side. {\sf C.} on the contrary. {\sf D.} in addition.
	\item Chris was $\ldots$ between buying a new house \& going on a round-the-world cruise. {\sf A.} pulled. {\sf B.} moving. {\sf C.} torn. {\sf D.} leaning.
	\item She hasn't spoken German for over a year \& I think she wants to brush $\ldots$ a few things. {\sf A.} out in. {\sf B.} on with. {\sf C.} over on. {\sf D.} up on.
	\item Where do you think his new enthusiasm $\lfloor$ from? {\sf A.} springs. {\sf B.} leaps. {\sf C.} sources. {\sf D.} traces.
	\item I wish you would stop sitting on the $\ldots$ \& decide whose side you're on. {\sf A.} wall. {\sf B.} crossroads. {\sf C.} fence. {\sf D.} middle.
	\item My brother's $\ldots$ taller than me. {\sf A.} bit. {\sf B.} something. {\sf C.} more. {\sf D.} slightly.
	\item She's even broken the door! Look, the handle was $\ldots$ off. {\sf A.} parted. {\sf B.} come. {\sf C.} gone. {\sf D.} split.
	\item A(n) $\ldots$ attempt by the army was quickly foiled. {\sf A.} mutiny. {\sf B.} overthrow. {\sf C.} overtake. {\sf D.} coup.
	\item Take your umbrella, just in case -- it's better to be $\ldots$ than sorry. {\sf A.} safe. {\sf B.} secure. {\sf C.} sure. {\sf D.} sound.
	\item You can leave the vegetables but eat as $\ldots$ of the meat as you can. {\sf A.} most. {\sf B.} more. {\sf C.} much. {\sf D.} many.
	\item This research $\ldots$ the importance of motivating students. {\sf A.} lightens. {\sf B.} highlights. {\sf C.} lights up. {\sf D.} enlightens.
	\item I've won \pounds100 on the Lottery -- what a $\ldots$ of luck! {\sf A.} touch. {\sf B.} point. {\sf C.} stroke. {\sf D.} patch.
	\item It will be easier to wash your muddy trousers if you leave them to $\ldots$ in water 1st. {\sf A.} drown. {\sf B.} bathe. {\sf C.} moisten. {\sf D.} soak.
	\item It was not until Sue left home that she realized the importance of close family $\ldots$. {\sf A.} cords. {\sf B.} ropes. {\sf C.} ties. {\sf D.} strings.
	\item Glen was $\ldots$ of the company's financial problems, otherwise he wouldn't have agreed to take the job with them. {\sf A.} innocent. {\sf B.} untrained. {\sf C.} untaught. {\sf D.} ignorant.
	\item It's funny you should say that. I've just had the $\ldots$ thought. {\sf A.} identical. {\sf B.} like. {\sf C.} alike. {\sf D.} similar.
	\item ``You are not comparing $\ldots$ with like,'' said Margaret. {\sf A.} same. {\sf B.} like. {\sf C.} both. {\sf D.} each.
	\item I was $\ldots$ by officials at the airport, as I was carrying fresh food. They made me throw it all away. {\sf A.} restrained. {\sf B.} confined. {\sf C.} detained. {\sf D.} impeded.
	\item Job seekers are required to visit the $\ldots$ center at least once a week to be eligible for benefits. {\sf A.} job. {\sf B.} unemployment. {\sf C.} employ. {\sf D.} work.
	\item It's time the government $\ldots$ a stand against tax evaders \& began prosecuting them. {\sf A.} had. {\sf B.} got. {\sf C.} took. {\sf D.} gave.
	\item John's $\ldots$ to my news was very positive; he really seemed pleased for me. {\sf A.} answer. {\sf B.} reply. {\sf C.} feedback. {\sf D.} reaction.
	\item He $\ldots$ his bag \& ran out of the house. {\sf A.} caught. {\sf B.} held. {\sf C.} grabbed. {\sf D.} held up.
	\item How many times has Venus Williams $\ldots$ her sister? a. won. {\sf B.} defeated. {\sf C.} lost. {\sf D.} beaten.
	\item Some people ran onto the $\ldots$ after the 2nd goal. {\sf A.} pitch. {\sf B.} field. {\sf C.} court. {\sf D.} stadium.
	\item I enjoy eating hot $\ldots$ food but it gives me indigestion. {\sf A.} creamy. {\sf B.} spicy. {\sf C.} greasy. {\sf D.} savory.
	\item Helen offered to give me the $\ldots$ for her delicious pizza. {\sf A.} recipe. {\sf B.} instructions. {\sf C.} receipt. {\sf D.} prescription.
	\item The crowds $\ldots$ the streets trying to catch sight of their hero. {\sf A.} flocked. {\sf B.} overpopulated. {\sf C.} thronged. {\sf D.} huddled.
	\item I hope that this punishment will $\ldots$ you a lesson. {\sf A.} learn. {\sf B.} teach. {\sf C.} know. {\sf D.} give.
	\item Because she doesn't behave in a very $\ldots$ way, people think that she's rather strange. {\sf A.} usual. {\sf B.} standard. {\sf C.} conventional. {\sf D.} average.
	\item Global $\ldots$ is causing climate changes. {\sf A.} warming. {\sf B.} pollution. {\sf C.} heating. {\sf D.} destruction.
	\item I would love to go to the exhibition with you, but I'm afraid I can't $\ldots$ the time. {\sf A.} leave. {\sf B.} spare. {\sf C.} save. {\sf D.} lose.
	\item We were alerted to the fire when we saw thick, heavy smoke $\ldots$ out of the window. {\sf A.} billowing. {\sf B.} clouding. {\sf C.} choking. {\sf D.} pushing.
	\item The $\ldots$ came on board \& inspected all the sailors on the ship. {\sf A.} ally. {\sf B.} pacifist. {\sf C.} private. {\sf D.} admiral.
	\item Now there's only 1 wage coming in, we'll all have to tighten our $\ldots$. {\sf A.} waists. {\sf B.} pockets. {\sf C.} belts. {\sf D.} hats.
	\item We all wish you luck as you $\ldots$ on a new career. {\sf A.} move. {\sf B.} embark. {\sf C.} alight. {\sf D.} board.
	\item The Sales Manager spent several months $\ldots$ up his team. {\sf A.} drawing. {\sf B.} sending. {\sf C.} keeping. building.
	\item Many people would find the $\ldots$ of a human clone a frightening one. {\sf A.} scene. {\sf B.} vision. {\sf C.} sight. {\sf D.} shape.
	\item You really should buy a DVD player. You have to move with the $\ldots$. {\sf A.} times. {\sf B.} time. {\sf C.} timing. {\sf D.} timer.
	\item Sewage causes a serious health hazard when it is $\ldots$ directly into rivers \& seas, as it is in this country. {\sf A.} terminated. {\sf B.} blown. {\sf C.} flushed. {\sf D.} eschewed.
	\item When I took over the business I got more than I $\ldots$. {\sf A.} asked after. {\sf B.} came in for. {\sf C.} drew up. {\sf D.} bargained for.
	\item Your son has the $\ldots$ of a fine musician. {\sf A.} makings. {\sf B.} beginnings. {\sf C.} looks. {\sf D.} talents.
	\item He found that the test was child's $\ldots$. {\sf A.} matches. {\sf B.} games. {\sf C.} play. {\sf D.} delight.
	\item ``I don't believe you! You're having me $\ldots$!'' said Jack. {\sf A.} out. {\sf B.} on. {\sf C.} over. {\sf D.} up.
	\item I saw that film everyone's been raving $\ldots$. Maybe I'm missing something, but I thought it was fairly ordinary. {\sf A.} on. {\sf B.} for. {\sf C.} up. {\sf D.} about.
	\item I was in a $\ldots$ as to what to do. If I told the truth, he would get into trouble, but if I said nothing I would be in more trouble. {\sf A.} doubt. {\sf B.} quandary. {\sf C.} hitch. {\sf D.} complexity.
	\item The indiscriminate use of pesticides has $\ldots$ many rare species. {\sf A.} wiped out. {\sf B.} devastated. {\sf C.} extincted. {\sf D.} canceled.
	\item The business was set up purely as a front for $\ldots$ money. {\sf A.} washing. {\sf B.} laundering. {\sf C.} cleaning. {\sf D.} rinsing.
	\item In the afternoons we would go for a $\ldots$ along the seafront. {\sf A.} strut. {\sf B.} stretch. {\sf C.} stride. {\sf D.} stroll.
	\item ``I was in Oxford last weekend \& you'll never guess who I bumped $\ldots$,'' he said. {\sf A.} for. {\sf B.} into. {\sf C.} along. {\sf D.} through.
	\item Polynesia. The very name conjures $\ldots$ images of sun drenched beaches \& warm turquoise seas. {\sf A.} out. {\sf B.} in. {\sf C.} over. {\sf D.} up.
	\item Pulling the contract out of the envelope, she ripped it to $\ldots$. {\sf A.} smithereens. {\sf B.} tiny pieces. {\sf C.} shreds. {\sf D.} half.
	\item Our legs ached as we $\ldots$ wearily through the snow. {\sf A.} strutted. {\sf B.} scampered. {\sf C.} galloped. {\sf D.} trudged.
	\item Having $\ldots$ my vote, I left the polling station. {\sf A.} cast. {\sf B.} passed. {\sf C.} placed. {\sf D.} bid.
	\item It's time we bought some new furniture. Look at these chairs. They're $\ldots$ to pieces. {\sf A.} battered. {\sf B.} falling. {\sf C.} crumbling. {\sf D.} going.
	\item With unemployment at a record level, retraining programmes would $\ldots$ off in the long run. {\sf A.} pay. {\sf B.} go. {\sf C.} wear. {\sf D.} pass.
	\item James never shows his emotions; no matter what happens, he always keeps a stiff upper $\ldots$. {\sf A.} mouth. {\sf B.} kip. {\sf C.} head. {\sf D.} eye.
	\item I picked up the letter. She darted forward $\ldots$ $\ldots$ it. ``That's mine!'' she said. {\sf A.} gripped. {\sf B.} wrenched. {\sf C.} yanked. {\sf D.} grabbed.
	\item Of course you'll pass. You write well \& you have an excellent $\ldots$ of the subject. {\sf A.} grip. {\sf B.} seizure. {\sf C.} grasp. {\sf D.} embrace.
	\item After 3 hours of walking in the hot summer sun, we were $\ldots$ for a drink. {\sf A.} sighing. {\sf B.} gulping. {\sf C.} panting. {\sf D.} gasping.
	\item It was a(n) $\ldots$ conversation. We were both talking about different things without realizing it. {\sf A.} thick. {\sf B.} unwise. {\sf C.} absurd. {\sf D.} foolish.
	\item It is people who are of voting age who make up the $\ldots$ of a country. {\sf A.} electorate. {\sf B.} constituency. {\sf C.} ballot papers. {\sf D.} electors.
	\item It's only a small lamp, so it doesn't $\ldots$ off much light. {\sf A.} cast. {\sf B.} give. {\sf C.} shed. {\sf D.} spend.
	\item The fire $\ldots$ for days until the monsoon rains put it out. {\sf A.} ignited. {\sf B.} scorched. {\sf C.} flashed. {\sf D.} raged.
	\item He prefers $\ldots$ water to tap water. {\sf A.} canned. {\sf B.} mineral. {\sf C.} well. {\sf D.} fresh.
	\item Fresh fruit juice is better for you than $\ldots$ drinks. {\sf A.} fuzzy. {\sf B.} saccharine. {\sf C.} fizzy. {\sf D.} bubbly.
	\item We had to pay for our food but not for the drinks -- they were on the $\ldots$. {\sf A.} cafe. {\sf B.} house. {\sf C.} compliment. {\sf D.} consumption.
	\item They don't sell tea or coffee. They only sell $\ldots$ drinks. {\sf A.} soft. {\sf B.} sweet. {\sf C.} plain. {\sf D.} mild.
	\item We're over the $\ldots$! Who wouldn't be? We've just won \pounds1 million! {\sf A.} planet. {\sf B.} clouds. {\sf C.} stars. {\sf D.} moon.
	\item You dropped it down the stairs? You're lucky it is still in one $\ldots$! {\sf A.} whole. {\sf B.} piece. {\sf C.} entirely. {\sf D.} unit.
	\item Grabbing the twig, he $\ldots$ it in two. {\sf A.} snapped. {\sf B.} crushed. {\sf C.} crumbled. {\sf D.} tore.
	\item The fumes were so thick that he was $\ldots$ for breath. {\sf A.} suffocating. {\sf B.} inhaling. {\sf C.} gasping. {\sf D.} wheezing.
	\item We knew the concert was sold out, but we still went to the stadium $\ldots$ the off-chance that someone might want to sell us their tickets. {\sf A.} with. {\sf B.} on. {\sf C.} by. {\sf D.} in.
	\item The whole country is up in $\ldots$ about the new tax the government has put on books. {\sf A.} arms. {\sf B.} fists. {\sf C.} anger. {\sf D.} rage.
	\item I couldn't $\ldots$ him because he spoke far too quickly. {\sf A.} gather. {\sf B.} catch. {\sf C.} see. {\sf D.} follow.
	\item At the beginning of the school year, every teacher is $\ldots$ a classroom. {\sf A.} allocated. {\sf B.} distributed. {\sf C.} registered. {\sf D.} sorted.
	\item There have been no $\ldots$ in our department for over a year. {\sf A.} pensions. {\sf B.} perks. {\sf C.} promotions. {\sf D.} permits.
	\item He took the pills \& a week later he was as right as $\ldots$. {\sf A.} dawn. {\sf B.} water. {\sf C.} sunshine. {\sf D.} rain.
	\item I loved him at 1st, but when I got to know him better, I went $\ldots$ him. {\sf A.} out of. {\sf B.} from. {\sf C.} off. {\sf D.} ahead of.
	\item The new student found the informality at school $\ldots$ at 1st. {\sf A.} off-putting. {\sf B.} foreign. {\sf C.} mysterious. {\sf D.} blinding.
	\item Marge's knowledge of the subject was only $\ldots$. {\sf A.} external. {\sf B.} superficial. {\sf C.} outward. {\sf D.} outer.
	\item John is a most $\ldots$ young man; he can do a lot of different jobs well. {\sf A.} capable. {\sf B.} cunning. {\sf C.} industrious. {\sf D.} laborious.
	\item I can make $\ldots$ a bed in the spare room if you'd like to stay the night. {\sf A.} out. {\sf B.} of. {\sf C.} for. {\sf D.} up.
	\item Could you stand $\ldots$ for me \& teach my history class tomorrow, Maggie? {\sf A.} down. {\sf B.} out. {\sf C.} in. {\sf D.} up.
	\item Some people are against informality at lectures but, personally, I $\ldots$ the idea. {\sf A.} cheer. {\sf B.} applaud. {\sf C.} clap. {\sf D.} shout.
	\item My mother had to take private pupils in order to $\ldots$ her salary as a teacher. {\sf A.} augment. {\sf B.} expand. {\sf C.} complete. {\sf D.} inflate.
	\item Liz deliberately told Andy some lies about his girlfriend, just to make him feel angry \& sad. What a $\ldots$ thing to do! {\sf A.} tough. {\sf B.} daft. {\sf C.} fussy. {\sf D.} wicked.
	\item Deborah is going to take extra lessons to $\ldots$ what she missed while she was away. {\sf A.} put up with. {\sf B.} cut down on. {\sf C.} catch up on. {\sf D.} take up with.
	\item It was a great $\ldots$ to study under such an outstanding teacher. {\sf A.} privilege. {\sf B.} fortune. {\sf C.} favor. {\sf D.} value.
	\item We had a terrible row a few years ago \& we still haven't made it $\ldots$. {\sf A.} out. {\sf B.} up. {\sf C.} for. {\sf D.} of.
	\item The examiners often $\ldots$ extremely difficult questions for the written exams. {\sf A.} set. {\sf B.} make. {\sf C.} create. {\sf D.} write.
	\item A university professor's view is rarely that of the man in the $\ldots$. {\sf A.} bus. {\sf B.} street. {\sf C.} queue. {\sf D.} factory.
	\item He made $\ldots$ that it was his own work, but it was easy to see that he had copied it from a book. {\sf A.} up. {\sf B.} off. {\sf C.} out. {\sf D.} for.
	\item I expect all of you to be here 10 minutes before the examination begins, without $\ldots$. {\sf A.} fail. {\sf B.} failure. {\sf C.} fault. {\sf D.} miss.
	\item After the serious talk with his tutor, Hilary $\ldots$ himself more conscientiously to his studies. {\sf A.} engaged. {\sf B.} converted. {\sf C.} applied. {\sf D.} exerted.
	\item You completely misunderstood my instructions; you got hold of the wrong end of the $\ldots$. {\sf A.} line. {\sf B.} rope. {\sf C.} story. {\sf D.} stick.
	\item Margaret is very $\ldots$ up \& thinks she is superior to her classmates. {\sf A.} fed. {\sf B.} looked. {\sf C.} stuck. {\sf D.} turned.
	\item I haven't had a very $\ldots$ week. I seem to have done nothing at all. {\sf A.} economic. {\sf B.} enthusiastic. {\sf C.} extensive. {\sf D.} productive.
	\item Mr. Snob $\ldots$ that only his club sells real English ale. {\sf A.} exults. {\sf B.} crows. {\sf C.} brags. {\sf D.} rejoices.
	\item It is $\ldots$ impossible to find a good educational computer program. {\sf A.} almost. {\sf B.} barely. {\sf C.} hardly. {\sf D.} merely.
	\item Try our delicious 4-course meal, offering you superb $\ldots$ for money. {\sf A.} cost. {\sf B.} value. {\sf C.} price. {\sf D.} worth.
	\item At many restaurants a service $\ldots$ is included in the bill. {\sf A.} tip. {\sf B.} gratuity. {\sf C.} tax. {\sf D.} charge.
	\item I'm afraid I've only had time to $\ldots$ the article you recommended. {\sf A.} glance. {\sf B.} scan. {\sf C.} peruse. {\sf D.} look.
	\item A new restaurant has opened here which specializes in French $\ldots$. {\sf A.} cuisine. {\sf B.} cook. {\sf C.} feeding. {\sf D.} kitchen.
	\item Turn the book round, you've got it $\ldots$. {\sf A.} upside-down. {\sf B.} inside out. {\sf C.} downside up. {\sf D.} outside in.
	\item I can't make anything $\ldots$ his writing. {\sf A.} from. {\sf B.} in. {\sf C.} out. {\sf D.} of.
	\item Judging by the $\ldots$ Sean has put into his essay, he should do well. {\sf A.} effort. {\sf B.} exercise. {\sf C.} labor. {\sf D.} toil.
	\item Tom was so $\ldots$ at maths at school that he became the youngest student ever to be accepted by a college. {\sf A.} keen. {\sf B.} hopeful. {\sf C.} brilliant. {\sf D.} proud.
	\item Mr. Careful kept the $\ldots$ round his garden neatly clipped. {\sf A.} fence. {\sf B.} flowers. {\sf C.} plants. {\sf D.} hedge.
	\item I'm relying on you, gentlemen, so please don't $\ldots$. {\sf A.} allow me off. {\sf B.} drop me off. {\sf C.} let me down. {\sf D.} drop me down.
	\item Mary looked as if she hadn't a $\ldots$ what Prof. Brown was talking about. {\sf A.} guess. {\sf B.} clue. {\sf C.} point. {\sf D.} thought.
	\item ``Get out of my garden, you nasty creatures!'' the man shouted, shaking his $\ldots$ at the kitchen. {\sf A.} ear. {\sf B.} head. {\sf C.} hand. {\sf D.} fist.
	\item I was startled by the sound of a twig $\ldots$ on the window. {\sf A.} tapping. {\sf B.} pattering. {\sf C.} rustling. {\sf D.} creaking.
	\item Someone from the Ministry of Education is coming to $\ldots$ our classes. {\sf A.} control. {\sf B.} inspect. {\sf C.} look on. {\sf D.} overlook.
	\item Before joining a course of study you must fill in a long $\ldots$ form. a. enrollment. {\sf B.} induction. {\sf C.} inscription. {\sf D.} personal.
	\item The new experimental system of enrollment didn't $\ldots$ expectations. {\sf A.} climb up to. {\sf B.} reach. {\sf C.} come up to. {\sf D.} rise to.
	\item Can you hold the ladder $\ldots$ while I climb p \& pick the apples? {\sf A.} steady. {\sf B.} rigid. {\sf C.} solid. {\sf D.} firm.
	\item The main disadvantage to our house is that the only $\ldots$ to the garden is through the bathroom. {\sf A.} passage. {\sf B.} access. {\sf C.} doorway. {\sf D.} communication.
	\item It's raining hard now. Let's stand in that shed to $\ldots$. {\sf A.} dry. {\sf B.} protect. {\sf C.} shelter. {\sf D.} save.
	\item He had to $\ldots$ the mud off his shoes before coming into the house. {\sf A.} clear away. {\sf B.} kick. {\sf C.} scratch. {\sf D.} scrape.
	\item Some schools have very $\ldots$ rules of behavior which must be obeyed. {\sf A.} solid. {\sf B.} strict. {\sf C.} straight. {\sf D.} strong.
	\item In some countries, students are selected $\ldots$ to their current level academic attainment.
	\item Your progress will be $\ldots$ in 3 months' time. {\sf A.} counted. {\sf B.} enumerated. {\sf C.} priced. {\sf D.} evaluated.
	\item Ms. Stickyfinger was charged with $\ldots$ the funds of the organization. {\sf A.} mislaying. {\sf B.} misappropriating. {\sf C.} mistaking. {\sf D.} misplacing.
	\item The woman $\ldots$ for her husband's life when he was found guilty of murder. {\sf A.} hid. {\sf B.} debated. {\sf C.} disputed. {\sf D.} pleaded.
	\item It was a reasonable $\ldots$ to draw in the light of the evidence. {\sf A.} assessment. {\sf B.} conclusion. {\sf C.} interpretation. {\sf D.} verdict.
	\item Vivian is studying to become a member of the medical $\ldots$. {\sf A.} employment. {\sf B.} position. {\sf C.} profession. {\sf D.} post.
	\item The whole story was a $\ldots$ of her imagination. {\sf A.} figment. B fantasy. {\sf C.} fabrication. {\sf D.} figure.
	\item The witness $\ldots$ the statements made by the accused man. {\sf A.} agreed. {\sf B.} confessed. {\sf C.} corroborated. {\sf D.} testified.
	\item A prominent local figure was $\ldots$ as co-respondent in a divorce case. {\sf A.} accused. {\sf B.} cited. {\sf C.} nominated. {\sf D.} quoted.
	\item The community was angered by the $\ldots$ punishment given their friend. {\sf A.} severe. {\sf B.} austere. {\sf C.} august. {\sf D.} vigorous.
	\item Those acting for the defendant propose to appeal $\ldots$ the sentence. {\sf A.} to. {\sf B.} for. {\sf C.} out. {\sf D.} against.
	\item I must buy a new leather $\ldots$ to go round my suitcase. {\sf A.} rope. {\sf B.} strap. {\sf C.} string. {\sf D.} thread.
	\item There is a bomb in the garden, it could $\ldots$ at any moment. {\sf A.} explode. {\sf B.} crack. {\sf C.} crash. {\sf D.} bang.
	\item I wish you'd let me speak for myself \& not $\ldots$ the words out of my mouth. {\sf A.} grab. {\sf B.} pull. {\sf C.} take. {\sf D.} snatch.
	\item All $\ldots$ barristers are expected to study at the Inns of Court. {\sf A.} hopeful. {\sf B.} wishful. {\sf C.} willing. {\sf D.} prospective.
	\item It is a criminal offence to $\ldots$ the facts. {\sf A.} express. {\sf B.} oppress. {\sf C.} suppress. {\sf D.} repress.
	\item It has been decided to hold a public $\ldots$ into the cause of the accident. {\sf A.} inquiry. {\sf B.} examination. {\sf C.} autopsy. {\sf D.} interrogation.
	\item You should only make serious accusations like that if they have a sound $\ldots$ in fact. {\sf A.} foothold. {\sf B.} basis. {\sf C.} framework. {\sf D.} principle.
	\item The suspect is $\ldots$ to have been in the neighborhood at the time of the crime. {\sf A.} accused. {\sf B.} affirmed. {\sf C.} alleged. {\sf D.} announced.
	\item The $\ldots$ question in this case is whether the accused had a motive for this crime or not. {\sf A.} valuable. {\sf B.} forcible. {\sf C.} supreme. {\sf D.} crucial.
	\item After a close cross-examination, the barrister was $\ldots$ his client was telling the truth. {\sf A.} content. {\sf B.} satisfied. {\sf C.} happy. {\sf D.} glad.
	\item Would it be possible to make an early $\ldots$ to have my hair cut tomorrow? {\sf A.} appointment. {\sf B.} arrangement. {\sf C.} date. {\sf D.} engagement.
	\item The tweed for this suit was $\ldots$ in Scotland. {\sf A.} knitted. {\sf B.} faced. {\sf C.} threaded. {\sf D.} woven.
	\item His comments $\ldots$ little or no relation to the facts of the case. {\sf A.} give. {\sf B.} bear. {\sf C.} possess. {\sf D.} reflect.
	\item The police were $\ldots$ in their examination of the murder site. {\sf A.} thorough. {\sf B.} exhausting. {\sf C.} intense. {\sf D.} concentrated.
	\item This was 1 of the few crimes he didn't $\ldots$. {\sf A.} achieve. {\sf B.} make. {\sf C.} commit. {\sf D.} perform.
	\item The army $\ldots$ defeat at the hands of its enemy. {\sf A.} bore. {\sf B.} tolerated. {\sf C.} supported. {\sf D.} suffered.
	\item A sergeant in the army wears 3 $\ldots$ on his arm. {\sf A.} scratches. {\sf B.} captions. {\sf C.} stripes. {\sf D.} medal.s
	\item Making colors vanish is 1 of my machine's $\ldots$ features. {\sf A.} certain. {\sf B.} unique. {\sf C.} solitary. {\sf D.} only.
	\item The encouraging news enabled the troops to go on with $\ldots$ hope. {\sf A.} renewed. {\sf B.} recurrent. {\sf C.} novel. {\sf D.} renovated.
	\item She tied her $\ldots$ round her head to protect her hair from the rain. {\sf A.} belt. {\sf B.} scarf. {\sf C.} glove. {\sf D.} umbrella.
	\item She bought some green shoes to $\ldots$ her green dress \& handbag. {\sf A.} equal. {\sf B.} mate. {\sf C.} match. {\sf D.} pair.
	\item She often wears a $\ldots$ hat to protect her head when the sun is very hot. {\sf A.} felt. {\sf B.} fur. {\sf C.} top. {\sf D.} straw.
	\item A feeling of tiredness almost $\ldots$ the solder on duty \& he had to struggle hard to keep awake. {\sf A.} conquered. {\sf B.} defeated. {\sf C.} overcame. {\sf D.} put out.
	\item I was informed by the police constable that he would be forced to take me into $\ldots$. {\sf A.} custody. {\sf B.} confinement. {\sf C.} detection. {\sf D.} guardianship.
	\item You are surely not suggesting that these $\ldots$ young children could have planned such an evil deed. {\sf A.} lovely. {\sf B.} innocent. {\sf C.} natural. {\sf D.} pure.
	\item Sergeant Proud can hardly $\ldots$ having to accept orders from the colonel who is much younger than himself. {\sf A.} resist. {\sf B.} mind. {\sf C.} bear. {\sf D.} suffer.
	\item Your shoes are a(n) $\ldots$! You haven't cleaned them for weeks. {\sf A.} accident. {\sf B.} disgrace. {\sf C.} blow. {\sf D.} shock.
	\item Her shoes were so old that her $\ldots$ were sticking out of them. {\sf A.} fingers. {\sf B.} thumbs. {\sf C.} toes. {\sf D.} tips.
	\item If I could $\ldots$ these shoes a little, they would be much more comfortable. {\sf A.} expand. {\sf B.} stretch. {\sf C.} increase. {\sf D.} grow.
	\item The washing is not quite dry, it needs $\ldots$. {\sf A.} airing. {\sf B.} heating. {\sf C.} rinsing. {\sf D.} warming.
	\item Unless you pack your trousers more carefully, they will get $\ldots$. {\sf A.} folded. {\sf B.} creased. {\sf C.} spoilt. {\sf D.} wrinkled.
	\item I gave Mary a black silk scarf with pink $\ldots$. {\sf A.} balls. {\sf B.} marks. {\sf C.} stains. {\sf D.} spots.
	\item You should iron out the $\ldots$ in that dress. {\sf A.} creases. {\sf B.} crumples. {\sf C.} folds. {\sf D.} wrinkles.
	\item When it is very hot, you may $\ldots$ the top button of your shirt.
	\item His jeans were so tight, they $\ldots$ when he sat down. {\sf A.} broke. {\sf B.} cracked. {\sf C.} split. {\sf D.} snapped.
	\item My wife has excellence $\ldots$ in clothes. {\sf A.} taste. {\sf B.} flavor. {\sf C.} gusto. {\sf D.} choice.
	\item Why don't you wear a(n) $\ldots$ when you do the washing-up? {\sf A.} towel. {\sf B.} cloth. {\sf C.} duster. {\sf D.} apron.
	\item My brother is always wearing the same $\ldots$ shirt. {\sf A.} lined. {\sf B.} spotted. {\sf C.} striped. {\sf D.} tie.
	\item The boxer in the dark $\ldots$ is sure to win. He's much better than the other one. {\sf A.} costume. {\sf B.} pants. {\sf C.} trousers. {\sf D.} trunks.
	\item Those new dresses may be striking but that sort of $\ldots$ does not suit everyone. {\sf A.} fashion. {\sf B.} form. {\sf C.} model. {\sf D.} mood.
	\item Come here! You've got a $\ldots$ of cotton on your coat. I'll take it off. {\sf A.} cord. {\sf B.} fibre. {\sf C.} string. {\sf D.} thread.
	\item Wearing odd socks would be classed as $\ldots$ behavior. {\sf A.} erotic. {\sf B.} ecstatic. {\sf C.} eccentric. {\sf D.} exotic.
	\item Mr. Smith arrived wearing a suit, but he put on his $\ldots$ before going into the workshop. {\sf A.} underclothes. {\sf B.} overcoat. {\sf C.} overwork. {\sf D.} overalls.
	\item He turned up his $\ldots$ to protect his neck from the cold wind. {\sf A.} cap. {\sf B.} collar. {\sf C.} scarf. {\sf D.} sleeve.
	\item The girl's dress was the 1st thing that $\ldots$ him to her. {\sf A.} attracted. {\sf B.} fetched. {\sf C.} lured. {\sf D.} tempted.
	\item Hearing a knock at the front door, the woman threw a(n) $\ldots$ over her pyjamas \& went downstairs to open the door. {\sf A.} bra. {\sf B.} nightdress. {\sf C.} evening dress. {\sf D.} dressing-gown.
	\item I'm the only one at this party in a long dress -- I feel quite out of $\ldots$. {\sf A.} order. {\sf B.} practice. {\sf C.} place. {\sf D.} turn.
	\item He couldn't find two that matched, so he was forced to wear $\ldots$ socks. {\sf A.} odd. {\sf B.} mixed. {\sf C.} uneven. {\sf D.} unlike.
	\item It is a good idea to be $\ldots$ dressed when you go for an interview. {\sf A.} boldly. {\sf B.} clearly. {\sf C.} smartly. {\sf D.} finely.
	\item There's a fantastic $\ldots$ from the top of the Empire State Building! {\sf A.} appearance. {\sf B.} view. {\sf C.} sight. {\sf D.} look.
	\item We usually do go by train, even though the car $\ldots$ is a lot quicker. {\sf A.} travel. {\sf B.} voyage. {\sf C.} trip. {\sf D.} journey.
	\item I'm going to ask for directions because I think we've $\ldots$ our way. {\sf A.} missed. {\sf B.} lost. {\sf C.} misplaced. {\sf D.} mistaken.
	\item 3 people were captured at the $\ldots$ today trying to get into the country. {\sf A.} line. {\sf B.} equator. c. border. {\sf D.} edge.
	\item Once we get to the hotel, let's just $\ldots$ quickly \& then do a bit of sightseeing. {\sf A.} set down. {\sf B.} check in. {\sf C.} make up. {\sf D.} turn up.
	\item Passengers requiring a special meal during the flight should inform the airline in $\ldots$. {\sf A.} advance. {\sf B.} ahead. {\sf C.} front. {\sf D.} forward.
	\item The $\ldots$ in the stadium all cheered the athletes. {\sf A.} viewers. {\sf B.} spectators. {\sf C.} witnesses. {\sf D.} onlookers.
	\item I got a new baseball $\ldots$ yesterday. {\sf A.} stick. {\sf B.} club. {\sf C.} racket. {\sf D.} bat.
	\item After a hard match. United managed to $\ldots$ with City, 1--1. {\sf A.} draw. {\sf B.} equal. {\sf C.} exact. {\sf D.} score.
	\item In this game, the players $\ldots$ it in turns to throw the dice. {\sf A.} have. {\sf B.} do. {\sf C.} make. {\sf D.} take.
	\item The police asked to see the man's driving $\ldots$. {\sf A.} degree. {\sf B.} certificate. {\sf C.} licence. {\sf D.} pass.
	\item Just ask them if you can play \& I'm sure they'll let you $\ldots$. {\sf A.} take up. {\sf B.} join in. {\sf C.} go off. {\sf D.} take to.
	\item Everyone expects Johnson to $\ldots$ Smith in today's final. {\sf A.} win. {\sf B.} score. {\sf C.} beat. {\sf D.} champion.
	\item The brochure says that the hotel has a great $\ldots$ of the sea. {\sf A.} appearance. {\sf B.} view. {\sf C.} look. {\sf D.} sight.
	\item I don't think people should be allowed to perform experiments $\ldots$ animals. {\sf A.} for. {\sf B.} over. {\sf C.} to. {\sf D.} on.
	\item Sending text $\ldots$ from your mobile is a great way to keep in touch with friends. {\sf A.} letters. {\sf B.} compositions. {\sf C.} messages. {\sf D.} notes.
	\item Let's face $\ldots$ -- we are destroying the environment \& we need to do something now. A.facts. {\sf B.} truth. {\sf C.} things. {\sf D.} information.
	\item My dad's really angry with my mum because she was caught $\ldots$ last week when she was driving to work. {\sf A.} running. {\sf B.} accelerating. {\sf C.} speeding. {\sf D.} chasing.
	\item Researchers have $\ldots$ to the conclusion that your personality is affected by your genes. {\sf A.} got. {\sf B.} come. {\sf C.} reached. {\sf D.} arrived.
	\item It's amazing that creatures survive $\ldots$ the bottom of the ocean. {\sf A.} in. {\sf B.} by. {\sf C.} to. {\sf D.} at.
	\item Fox Talbot $\ldots$ the 1st photograph in 1835. {\sf A.} took. {\sf B.} did. {\sf C.} drew. {\sf D.} gave.
	\item Do you mind if I just $\ldots$ a quick phone call from here? {\sf A.} do. {\sf B.} take. {\sf C.} have. {\sf D.} make.
	\item Many lives were saved $\ldots$ the introduction of antibiotics. {\sf A.} with. {\sf B.} at. {\sf C.} into. {\sf D.} in.
	\item The $\ldots$ of the nuclear accident is still unknown. {\sf A.} reason. {\sf B.} base. {\sf C.} cause. {\sf D.} motive.
	\item My father works at the university, doing research $\ldots$ weather control. {\sf A.} to. {\sf B.} on. {\sf C.} of. {\sf D.} from.
	\item The telescope will photograph distant galaxies, $\ldots$ an attempt to understand the universe. {\sf A.} on. {\sf B.} for. {\sf C.} in. {\sf D.} with.
	\item Do you know who $\ldots$ the fact that sound travels in waves? {\sf A.} discovered. {\sf B.} developed. {\sf C.} invented. {\sf D.} found.
	\item I'm thinking of getting a fax $\ldots$. {\sf A.} motor. {\sf B.} engine. {\sf C.} machine. {\sf D.} instrument.
	\item The $\ldots$ of the outbreak of the disease is still unknown. {\sf A.} aim. {\sf B.} reason. {\sf C.} motive. {\sf D.} cause.
	\item 10 million text messages are sent on $\ldots$ every minute. {\sf A.} normal. {\sf B.} average. {\sf C.} general. {\sf D.} common.
	\item Once they've finished the extension to the $\ldots$, there'll be room for them to take on a lot more workers. {\sf A.} factory. {\sf B.} trade. {\sf C.} industry. {\sf D.} manufacturing.
	\item Shannen doesn't dye her hair; it's $\ldots$ blonde. {\sf A.} physically. {\sf B.} organically. {\sf C.} logically. {\sf D.} naturally.
	\item Watch $\ldots$ for words like ``so-called'' in articles as they express the writer's bias. {\sf A.} about. {\sf B.} out. {\sf C.} over. {\sf D.} around.
	\item They $\ldots$ a description of the robber on \textit{Crime Time} \& it sounded like you! {\sf A.} made. {\sf B.} told. {\sf C.} gave. {\sf D.} said.
	\item I've $\ldots$ it in at the mains but it still doesn't seem to work. {\sf A.} worked. {\sf B.} plugged. {\sf C.} narrowed. {\sf D.} turned.
	\item It's not always easy to $\ldots$ the difference between fact \& opinion. {\sf A.} make. {\sf B.} do. {\sf C.} say. {\sf D.} tell.
	\item The debate will $\ldots$ place tonight. {\sf A.} be. {\sf B.} have. {\sf C.} take. {\sf D.} make.
	\item In my $\ldots$, freedom of the press must be maintained. {\sf A.} mind. {\sf B.} thought. {\sf C.} sight. {\sf D.} view.
	\item Media mogul Ronald Morduck has $\ldots$ control of another tabloid. {\sf A.} made. {\sf B.} taken. {\sf C.} given. {\sf D.} found.
	\item J. K. Rowling has $\ldots$ an enormous influence on children's literature. {\sf A.} had. {\sf B.} given. {\sf C.} done. {\sf D.} set.
	\item I could spend hours $\ldots$ the Internet! {\sf A.} swimming. {\sf B.} diving. {\sf C.} sailing. {\sf D.} surfing.
	\item They said $\ldots$ the news that the price of petrol is going up again. {\sf A.} from. {\sf B.} on. {\sf C.} in. {\sf D.} at.
	\item They were $\ldots$ out free tickets to the concert to everyone in the shopping center. {\sf A.} handing. {\sf B.} holding. {\sf C.} carrying. {\sf D.} bringing.
	\item We had the volume turned down, so I couldn't make $\ldots$ what they were talking about. {\sf A.} up. {\sf B.} in. {\sf C.} out. {\sf D.} off.
	\item I saw the interview while I was $\ldots$ through a magazine at the hairdresser's. {\sf A.} flicking. {\sf B.} clicking. {\sf C.} picking. {\sf D.} ticking.
	\item Why did you have to bring that subject $\ldots$ at dinner? {\sf A.} off. {\sf B.} out. {\sf C.} over. {\sf D.} up.
	\item People can become very $\ldots$ when they are stuck in traffic for a long time. {\sf A.} nervous. {\sf B.} stressful. {\sf C.} bad-tempered. {\sf D.} pressed.
	\item The tabloids completely $\ldots$ that story about Bruce Willis. It's not true at all! {\sf A.} made up. {\sf B.} filled in. {\sf C.} stood out. {\sf D.} turned over.
	\item The issue $\ldots$ question is more complex than you think. {\sf A.} from. {\sf B.} at. {\sf C.} on. {\sf D.} in.
	\item The programme usually $\ldots$ on at half past eight. {\sf A.} comes. {\sf B.} does. {\sf C.} turns. {\sf D.} looks.
	\item Everyone said they had $\ldots$ themselves at the wedding. {\sf A.} pleased. {\sf B.} impressed. {\sf C.} enjoyed. {\sf D.} excited.
	\item Mary seems to go out with a different $\ldots$ of friends almost every night. {\sf A.} collection. {\sf B.} band. {\sf C.} company. {\sf D.} group.
	\item They mentioned it in the article, but they didn't $\ldots$ into detail. {\sf A.} make. {\sf B.} go. {\sf C.} take. {\sf D.} come.
	\item More \& more people are living into $\ldots$ age \& it's a serious social problem. {\sf A.} high. {\sf B.} ancient. {\sf C.} old. {\sf D.} far.
	\item Stephanie seems to be very $\ldots$ with her classmates. {\sf A.} popular. {\sf B.} famous. {\sf C.} known. {\sf D.} likeable.
	\item Most $\ldots$ people have no idea what it's like to be famous. {\sf A.} usual. {\sf B.} ordinary. {\sf C.} typical. {\sf D.} medium.
	\item When my parents got divorced, my best friend was very $\ldots$ \& listened to all my problems. {\sf A.} likeable. {\sf B.} amusing. {\sf C.} enjoyable. d. sympathetic.
	\item Rita's very $\ldots$ \& easily gets upset when people criticize her. {\sf A.} sensitive. {\sf B.} level-headed. {\sf C.} sensible. {\sf D.} open-minded.
	\item Police were called in when the $\ldots$ of people began to get violent. {\sf A.} audience. {\sf B.} company. {\sf C.} crowd. {\sf D.} herd.
	\item My dad says he once met Robbie Williams when he was still $\ldots$. {\sf A.} hidden. {\sf B.} unknown. {\sf C.} infamous. {\sf D.} unrelated.
	\item The government is trying to convince people $\ldots$ the need for higher taxes. {\sf A.} of. {\sf B.} with. {\sf C.} that. {\sf D.} for.
	\item The MP asked $\ldots$ the prime minister was aware of the growing social problem. {\sf A.} that. {\sf B.} him. {\sf C.} what. {\sf D.} if.
	\item I believe that judges should be independent $\ldots$ the government. {\sf A.} to. {\sf B.} from. {\sf C.} with. {\sf D.} on.
	\item Most people seem to agree $\ldots$ the newspaper's criticism of the government. {\sf A.} with. {\sf B.} on. {\sf C.} to. {\sf D.} that.
	\item I don't think people should rely $\ldots$ the state, even if they are unemployed. {\sf A.} to. {\sf B.} by. {\sf C.} with. {\sf D.} on.
	\item Local residents object $\ldots$ the new power station in their area. {\sf A.} to have. {\sf B.} having. {\sf C.} to having. {\sf D.} of having.
	\item Do you think they should ban people $\ldots$ smoking in public places? {\sf A.} that. {\sf B.} from. {\sf C.} in. {\sf D.} of.
	\item The mayor was attacked $\ldots$ public money. {\sf A.} for wasting. {\sf B.} to waste. {\sf C.} he waste. {\sf D.} of wasting.
	\item This fascinating book covers some of the most $\ldots$ crimes of the 20th century. {\sf A.} unknown. {\sf B.} hidden. {\sf C.} infamous. {\sf D.} covered.
	\item The government should do more for $\ldots$ people. {\sf A.} ordinary. {\sf B.} usual. c. everyday. {\sf D.} typical.
	\item Ivan tells me he really $\ldots$ himself at your barbecue last week. {\sf A.} pleased. {\sf B.} interested. {\sf C.} played. {\sf D.} enjoyed.
	\item I know we had an argument but now I'd quite like to $\ldots$. {\sf A.} took down. {\sf B.} fall out. {\sf C.} make up. {\sf D.} bring up.
	\item Harry \& Sam both denied that the fight was their $\ldots$. {\sf A.} blame. {\sf B.} fault. {\sf C.} criticism. {\sf D.} cause.
\end{enumerate}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
	
\end{document}