\documentclass{article}
\usepackage[backend=biber,natbib=true,style=authoryear]{biblatex}
\addbibresource{/home/hong/1_NQBH/reference/bib.bib}
\usepackage[utf8]{inputenc}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{float}
\usepackage{graphicx}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{ques}{Question}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\usepackage[left=0.5in,right=0.5in,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\small \textsc{Sect.} ~\thesection}
\rhead{\small \nouppercase{\leftmark}}
\renewcommand{\sectionmark}[1]{\markboth{#1}{}}
\cfoot{\thepage}
\def\labelitemii{$\circ$}

\title{Journal}
\author{Collector: Nguyen Quan Ba Hong}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

%------------------------------------------------------------------------------%

\section{Computer Science Journal}

\section{Engineering Journal}

\subsection{\href{https://www.tandfonline.com/journals/unhb20}{Numerical Heat Transfer, Part B: Fundamentals}}
\begin{itemize}
	\item IF 2020: 1.586.
	\item 5-Year IF 2020: 1.506.
	\item SJR 2020: Q2, 0.547.
\end{itemize}

\section{Mathematical Journal}

\subsection{\href{https://www.springer.com/journal/10444}{Advances in Computational Mathematics}}
\begin{itemize}
	\item H-index: 61.
	\item IF 2020: 1.910.
	\item 5-Year IF 2020: 1.995.
	\item SJR 2020: Q1, 1.
\end{itemize}
\textit{Advances in Computational Mathematics} publishes high quality, accessible, and original articles at the forefront of computational and applied mathematics, with a clear potential for impact across the sciences.

The journal emphasizes 3 core areas:
\begin{enumerate}
	\item approximation theory and computational geometry,
	\item numerical analysis, modeling and simulation,
	\item imaging, signal processing and data analysis.
\end{enumerate}
This journal welcomes papers that are accessible to a broad audience in the mathematical sciences and that show either an advance in computational methodology or a novel scientific application area, or both. Methods papers should rely on rigorous analysis and\texttt{/}or convincing numerical studies.

\subsection{\href{https://www.degruyter.com/journal/key/ans/html}{Advanced Nonlinear Studies} (ANS)}
\begin{itemize}
	\item H-index: 36.
	\item IF: 1.446.
	\item SJR 2020: Q1, 1.31.
\end{itemize}
\textit{Advanced Nonlinear Studies} (ANS) publishes cutting-edge research on nonlinear problems, particularly those involving Integral and Partial Differential Equations, Dynamical Systems, Calculus of Variations, Geometric Analysis, and related areas. ANS also publishes novel and interesting applications of these areas to problems in biology, engineering, materials sciences, physics, and other sciences.

\paragraph{NQBH's contributions.} \cite{Dao_Diaz_Nguyen2020}.

\subsection{\href{https://msp.org/apde}{Analysis \& PDE} (APDE)}
\begin{itemize}
	\item H-index: 39.
	\item SJR 2020: 3.11.
\end{itemize}
\textit{Analysis \& PDE} (APDE) aims to be the leading specialized scholarly publication in mathematical analysis. The purpose of APDE is the advancement of mathematics.

\paragraph{NQBH's favorite articles.} \cite{Tao2013}.

\subsection{\href{https://www.esaim-cocv.org/}{ESAIM. Control, Optimisation and Calculus of Variations} (ESAIM: COCV)}
\begin{itemize}
	\item IF: 1.619.
	\item 5-Year IF: 1.594.
	\item SJR 2020: Q1, 1.02.
\end{itemize}

\subsection{\href{https://onlinelibrary.wiley.com/journal/10970363}{International Journal for Numerical Methods in Fluids}}
\begin{itemize}
	\item IF: 2.107.
	\item SJR 2020: Q1, 0.94.
\end{itemize}

\subsection{\href{https://www.springer.com/journal/21}{Journal of Mathematical Fluid Mechanics}}
\begin{itemize}
	\item IF 2020: 1.298.
	\item 5-Year IF 2020: 1.453.
	\item SJR 2020: Q1, 1.
\end{itemize}

\subsection{\href{https://www.ams.org/publications/journals/journalsframework/mcom}{Mathematics of Computation} (MCOM)}
\begin{itemize}
	\item Journal Citation Reports' IF: 2.417.
	\item SJR 2020: Q1, 1.95.
\end{itemize}

\subsection{\href{https://www.siam.org/publications/journals/siam-journal-on-control-and-optimization-sicon}{SIAM Journal on Control \& Optimization} (SICON)}
\begin{itemize}
	\item SJR 2020: Q1, 1.49.
\end{itemize}

\section{Physical Journal}

\subsection{\href{https://www.journals.elsevier.com/journal-of-computational-physics}{Journal of Computational Physics}}
\begin{itemize}
	\item IF: 3.553.
	\item SJR 2020: Q1, 1.88.
\end{itemize}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
\end{document}