\documentclass{article}
\usepackage[backend=biber,natbib=true,style=alphabetic,maxbibnames=50]{biblatex}
\addbibresource{/home/nqbh/reference/bib.bib}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=1cm,right=1cm,top=5mm,bottom=5mm,footskip=4mm]{geometry}
\def\labelitemii{$\circ$}

\title{Can't Hurt Me: Master Your Mind \& Defy the Odds}
\author{David Goggins}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

%------------------------------------------------------------------------------%

\section{Reviews}

\subsection{Amazon's Review}
``For David Goggins, childhood was a nightmare -- poverty, prejudice, \& physical abuse colored his days \& haunted his nights. But through self-discipline, mental toughness, \& hard work, Goggins transformed himself from a depressed, overweight young man with no future into a U.S. Armed Forces icon \& one of the world's top endurance athletes. The only man in history to complete elite training as a Navy SEAL, Army Ranger, \& Air Force Tactical Air Controller, he went on to set records in numerous endurance events, inspiring Outside magazine to name him ``The Fittest (Real) Man in America.''

In Can't Hurt Me, he shares his astonishing life story \& reveals that most of us tap into only 40\% of our capabilities. Goggins calls this The 40\% Rule, \& his story illuminates a path that anyone can follow to push past pain, demolish fear, \& reach their full potential.''

\subsection{Editorial Reviews}

\begin{itemize}
	\item ``David Goggins is a being of pure will \& inspiration. Just listening to this guy talk makes you want to run up a mountain. I firmly believe people like him can change the course of the world just by inspiring us to push harder \& dig deeper in everything we do. His goal to be `uncommon amongst uncommon people' is something we can all use to propel ourselves to fulfill our true potential. I'm a better man having met him.'' -- Joe Rogan, Standup Comedian \& Host of the \textit{Joe Rogan Experience Podcast}
	\item ``David Goggins lives out every goal, every dream no matter what. PERIOD. He's unstoppable. There's no limit to him because he doesn't live in a comfort zone. His mental \& physical capacity are equal. Goggins proves that your body can handle anything if you let your mind keep up. There's no way to stop something or someone that doesn't understand the concept of being beat.'' -- Marcus Luttrell, Retired Navy SEAL, Author of New York Times Best Seller \textit{Lone Survivor}
	\item ``Modern neuroscience is teaching us that the path to courage \& success arrives through embracing pain \& fear, not by avoiding them. If ever there was a real-life example of this, it is the story of David Goggins. In his unrelenting pursuit to self-conquer, Goggins taught himself how to tap into that elusive holy grail of human existence: the ability to rewire one's own brain in order to continually do better \& actually become better, regardless of feelings, external conditions, or motivational state. \textit{Can't Hurt Me} is the remarkable description of that journey \& the capacity to leverage \& better the mind. More importantly, it also teaches you how.'' -- Andrew D. Huberman, PhD, Professor of Neurobiology, Stanford University School of Medicine
	\item ``David Goggins throws the door open on pain, evil, darkness, the worst \& yes, the best of humanity, \& the strength of the human soul $\ldots$ \& that's just in chapter 1. If you are looking for a book that will heal, stretch, inspire, \& dig into the corners of what it takes to persevere \& overcome in a messed-up world, this is your book.'' -- Taya Kyle, Widow of American Sniper Chris Kyle, Author of \textit{New York Times} Best Seller \textit{American Wife}
	\item ``By the time you finish David Goggins's new book, you'll have kicked your victim mentality in the butt. Where you go from there is entirely up to you -- as Goggins makes clear in this entertaining \& poignant memoir cum inspirational how-to. As the man with a hole in his heart tells you, there are no excuses in life, only reasons to try harder.'' -- Jim DeFelice, Author of \textit{American Sniper}
	\item ``David Goggins's book is not the 1st about overcoming severe hardships to achieve success, but it is certainly 1 of the most compelling. His story of beating the odds, of achieving athlete greatness, of serving his country \& his charities, \& of mastering his own destiny will inspire all of us to reach a little higher \& give a little more. `I will never quit' is a tenet of the Navy SEAL ethos, \& one that David Goggins applies to everything he does.'' -- Admiral Eric Olson, U.S. Navy (Retired); Former Commander, United States Special Operations Command; Chairman, Special Operations Warrior Foundation
	\item ``I'm inspired that people like this guy exist. Not everyone will live a life like David Goggins, but he is proof that anyone could if given the right headspace within.'' -- Kelly Slater, Eleven-Time World Champion Surfer
	\item ``Guaranteed to galvanize more than a few couch potatoes into action.'' -- Kirkus Reviews
\end{itemize}

%------------------------------------------------------------------------------%

\section{Quotes}
See \cite{Goggins2018}.
\begin{enumerate}
	\item ``Everything in life is a mind game! Whenever we get swept under by life's dramas, large \& small, we are forgetting that no matter how bad the pain gets, no matter how harrowing the torture, all bad things end.''
	\item ``If you want to master the mind \& remove your governor, you'll have to become addicted to hard work. Because passion \& obsession, even talent, are only useful tools if you have the work ethic to back them up.''
	\item ``The reason it's important to push hardest when you want to quit the most is because it helps you callous your mind. It's the same reason why you have to do your best work when you are the least motivated. That's why I loved PT in BUD\texttt{/}S \& why I still love it today. Physical challenges strengthen my mind so I'm ready for whatever life throws at me, \& it will do the same for you.''
	\item ``By the time I graduated, I knew that the confidence I'd managed to develop didn't come from a perfect family or God-given talent. It came from personal accountability which brought me self respect, \& self-respect will always light a way forward.''
\end{enumerate}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
	
\end{document}