\documentclass{article}
\usepackage[backend=biber,natbib=true,style=alphabetic,maxbibnames=50]{biblatex}
\addbibresource{/home/nqbh/reference/bib.bib}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=red,citecolor=magenta]{hyperref}
\usepackage{algorithm,algpseudocode,amsmath,amssymb,amsthm,float,graphicx,mathtools}
\allowdisplaybreaks
\numberwithin{equation}{section}
\newtheorem{assumption}{Assumption}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{notation}{Notation}[section]
\newtheorem{principle}{Principle}[section]
\newtheorem{problem}{Problem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{question}{Question}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{theorem}{Theorem}[section]
\usepackage[left=1cm,right=1cm,top=5mm,bottom=5mm,footskip=4mm]{geometry}
\def\labelitemii{$\circ$}

\title{Undisputed Truth}
\author{Mike Tyson, Larry Sloman}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

%------------------------------------------------------------------------------%

\section{Quotes}
See \cite{Tyson_Sloman2013}.
\begin{enumerate}
	\item ``Fear is the greatest obstacle to learning. But fear is your best friend. Fear is like fire. If you learn to control it, you let it work for you. If you don't learn to control it, it'll destroy you \& everything around you. Like a snowball on a hill, you can pick it up \& throw it or do anything you want with it before it starts rolling down, but once it rolls down \& gets so big, it'll crush you to death. So one must never allow fear to develop \& built up without having control over it, because it you don't you won't be able to achieve your objective or save your life.''
	\item ``Man is not mean to be humble, he's meant to be humbled.''
	\item ``a man who was friendly with everyone was an enemy to himself.''
	\item ``You think you know the difference between a hero \& a coward, Mike? Well, there is no difference between a hero \& a coward in what they feel. It's what they do that makes them different. The hero \& the coward feel exactly the same but you have to have the discipline to do what a hero does \& to keep yourself from doing what the coward does.''
	\item ``The ability to apply the discipline, the ability to do what needs to be done no matter how he feels inside, in my opinion, is the definition of a true professional.''
	\item ``Your mind is not your friend, Mike. I hope you know that. You have to fight with your mind, control it, put it in its place. You have to control your emotions. Fatigue in the ring is 90\% psychological. It's just the excuse of a man who wants to quit.''
	\item ``The boxing's the easy part. When you get into the ring to fight, that's the vacation. But when you get in the gym, you have to do things over \& over till you're sore \& deep in your mind you say, `I don't want to do this anymore,' I push that out of my mind.''
	\item ``My style was impetuous, my defenses were impregnable, \& I was ferocious.''
	\item ``God is not going to send someone to hell for my mistakes. So God \& I have to deal with my own salvation.''
	\item ``1 day Cus came to me. ``You know, Camille really wants you to do your chores. I could care less if you did, but you should do them because it will make you a better boxer.'' ``How's taking out the trash going to make me a better boxer?'' I scoffed. ``Because doing something you hate to do like you love it is good conditioning for someone aspiring towards greatness.''
	\item ``MIKE WOULD LIKE TO THANK: Cus D'Amato, my mentor, friend, \& general. Because of you, my life has reached heights I could never have imagined. Without you, I don't know where I would be today. My gratitude to you is immeasurable.''
	\item ``He was very vindictive \& always out for revenge. He couldn't live without enemies, so he created them.''
	\item ``Confidence was everything. But in order to possess that confidence, you had to test yourself \& put yourself on the line. It doesn't come from osmosis, out of the air. It comes from consistently going over the visualization in your mind to help you develop the confidence that you want to possess.''
	\item ``The one who lost \& quit, well their demons will follow them to their grave because they had a chance to face them \& they didn't. You have to face your demons, Mike, or they will follow you to eternity.''
	\item ``I was a titan, the reincarnation of Alexander the Great. My style was impetuous, my defenses were impregnable, \& I was ferocious.''
	\item ``What if I failed \& lost? I had been in a million fights on the streets.''
	\item ``Some students would make fun of me, but nobody fucked with me.''
	\item ``In her book, Robin implied that we hadn't slept together, but I actually nailed her the 1st or 2nd night when she came to my hotel. Instead she claimed that we strolled through the mall \& played with puppies at pet shops for hours. Can you see me in a motherfucking mall, the heavyweight champ of the world? What the fuck am I doing in a mall?''
	\item ``Freedom is a very dangerous thing. We ration it very closely.''
	\item ``Burning cocaine is the worst smell in the world. It smells like burning plastic \& rat poison combined. A friend of mine once told me, that when you want to know something about anything, put some fire under it. The fire brings out everything. You want to know something about a motherfucker? Put some fire under his ass. Well, when you put some fire under that cocaine, you know what it's made out of.''
	\item ``I had been reading Nietzsche \& thought I was a Superman. I could barely spell my name but I was a Superman.''
	\item ``nadir.''
	\item ``You can't give kids a fighting chance if you don't give them a fighting mind.''
	\item ``What kind of bird don't fly? Jailbird! Jailbird!''
	\item ``But, sir, please let me finish this round \& go 1 more round. That's what we normally do,'' I pleaded. I want to impress Cus. I guess I had. When we got out of the ring, Cus's 1st words to Bobby were, ``That's the heavyweight champion of the world.''
	\item ``We sat down \& Cus told me he couldn't believe I was only 13 years old. \& then he told me what my future would be. He had seen me spar for not even 6  minutes, but he said it in a way that was like law. ``You looked splendid,'' he said. ``You're a great fighter.'' It was compliment after compliment. ``If you listen to me, I can make you the youngest heavyweight champion of all time.''
	\item ``used to watch these comic book characters on TV, the X-Men \& 1 of my favorites, Apocalypse. Apocalypse would say, ``I'm not malevolent, I just am.'' Cayton \& Jacobs wanted me to be friendly with everybody, sociable, but I knew a man who was friendly with everyone was an enemy to himself.''
	\item ``There are animals disguised as human beings out here, you are not sophisticated enough to decipher the two.''
	\item ``if you don't have the spiritual warrior in you, you'll never be a fighter.''
	\item ``The moment the bell rings, \& you come in contact with each other, suddenly your opponent seems like everybody else, because now your imagination has dissipated. The fight itself is the only reality that matters. You have to learn to impose your will \& take control over that reality.''
	\item ``Cus told me that to be a great fighter you had to get out of your head. He would have me sit down \& he'd say, ``Transcend. Focus. Relax until you see yourself looking at yourself. Tell me when you get there.''''
	\item ``Cus was a strong believer that in your mind you had to be the entity that you wanted to be. If you wanted to be heavyweight champion of the world, you had to.''
	\item ``Sometimes I'd give my mother a hundred bucks \& she wouldn't pay me back. She didn't respect me like that. I'd say, ``You own me some money, Ma.'' \& she'd just say, ``You owe me your life, boy. I'm not paying you back.''''
	\item ``Later in that tape, the dispatcher kept telling Desiree that she was the victim, she was the victim, \& then Desiree slipped something very interesting in.''
	\item ``Hammerle began to ask her about the improbabilities that he saw about the case. ``Look, you met Mike Tyson, \& you saw him making all of these passes.''
	\item ``My mom's funeral was pathetic. She had saved up some money for a plot in Linden, New Jersey. There were only 8 of us there -- me, my brother \& sister, my father Jimmy, her boyfriend Eddie, \& 3 of my mother's friends. I wore a suit that I had bought with some of the money that I had stolen. She only had a thin cardboard casket \& there wasn't enough money for a headstone. Before we left the grave, I said, ``Mom, I promise I'm going to be a good guy. I'm going to be the best fighter ever \& everybody is going to know my name. When they think of Tyson, they're not going to think of Tyson Foods or Cicely Tyson, they're going to think of Mike Tyson.'' I said this to her because this was what Cus had been telling me about the Tyson name. Up until then, our family's only claim to fame was that we shared the same last name as Cicely. My mom loved Cicely Tyson.''
	\item ``Tyson came in with a lot of excess baggage. The press has vilified him. Not a day goes by that the press doesn't bring up his faults. This is not the Tyson I know. The Tyson I knew is a sensitive, thoughtful, caring man. He may be terrifying in the ring, but that ends when he leaves the ring.''
	\item ``I was never taught how to handle my celebrity status. I don't tell kids it's right to be Mike Tyson. Parents serve as better role models.''
	\item ``Don't use that word `can't.' You can't say `can't.''
	\item ``I came home 1 day \& there was a white baby in the house. What the fuck is this shit? I thought.''
	\item ``No matter what anyone says, no matter the excuse or explanation, whatever a person does in the end is what he intended to do all along.''
	\item ``If I hide the cuffs, then I'm a bitch. Jim thought that hiding my cuffs would stop me from experiencing shame, but that would have been the shame. I had to be seen with that steel on me. Fuck everybody else, the people who understand, they have got to see me with that steel on. I was going to warrior school.''
	\item ``Because doing something you hate to do like you love it is good conditioning for someone aspiring towards greatness.''
	\item ``Damn, why couldn't I have been with a nonconfrontational kind of white guy?''
	\item ``Well, I know what I am, \& that's what counts more than anything else, because the people don't know what I go through. They think I'm born this way. They don't know what it took to get this way.''
	\item ``As long as I am breathing, your legacy will continue to live on.''
	\item ````People who are only interested in having fun cannot accomplish anything.'' I thought I was stronger than people who were weak \& partying.''
	\item ``Cus asked Steve to leave the room so he could talk to me in private. That's when he told me he was dying from pneumonia. I couldn't believe what he was telling me. He didn't look morbidly ill. He was buffing. He had energy \& zest. he was eating ice cream. He was chilling out, but I started freaking out. ``I don't want to do this shit without you,'' I said, choking back tears. ``I'm not going to do it.'' ``Well, if you don't fight, you'll realize that people can come back from the grave, because I'm going to haunt you for the rest of your life.'' I told him ``Okay,'' \& then he took my hand. ``The world has to see you, Mike. You're going to be champ of the world, the greatest out there.'''
	\item ``He was certain that I didn't suffer from manic depression. When Halpern called their shrink to see why he had diagnosed me as a manic-depressive, he started backpedaling. He said I wasn't a full-fledged manic-depressive, I merely had a mood disorder, something he called ``Boxer Syndrome.'' That was a new one for Freud.''
	\item ``may have withheld critical evidence, a borderline incompetent trial defense attorney whose bumbling defense may have been more responsible for the guilty verdict than everything else combined \& a jury that paid more attention to Tyson's bad-boy public image than to the incompleteness of the case's facts. Garrison's best achievement at trial was his strategy in successfully presenting Desiree Washington as the shy, inexperienced, naive, prim \& proper college student that they in fact knew her not to be. Utilizing the full protection.''
	\item ``I spent most of the six weeks.''
	\item ``I couldn't even spell fucking Adidas. But I knew how they made me feel.''
	\item ``Woooo! I loved that shit.''
\end{enumerate}

%------------------------------------------------------------------------------%

\printbibliography[heading=bibintoc]
	
\end{document}